### 2023-12-19 doc/manuscript.Rmd
4 new alliances

---
title: "Anthropogenic plant communities of the Iberian Atlantic ecoregion"
output: github_document
bibliography: refs.bib
csl: journal-of-vegetation-science.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

# Methods

## Study ecoregion

We studied the anthropogenic plant communities of the Iberian Atlantic ecoregion, i.e. the territories with a temperate climate in the north-western Iberian Peninsula. Our study ecoregion broadly corresponds with the Cantabrian Mixed Forests ecoregion *sensu* (Olson et al. 2001), with the Iberian part of the European Atlantic province *sensu* Rivas-Martínez et al. 2017b, and with the Iberian section of the Atlantic biogeographical region of the European Environmental Agency. To define the study region we followed a fine-scale revision of its geographic borders at high spatial resolution using the most recent biogeographical map of the Iberian Atlantic territories (Fernández Prieto et al. 2020). Our revised Iberian Atlantic ecoregion encompasses 79,891 km^2^ of Portugal, Spain and France. Its climate is temperate with average monthly temperatures ranging from 6 to 20°C, and with total annual precipitation between 1,100 and 2,600 mm, with some areas having warm-dry summers but without the distinctive 2-month drought of the Mediterranean climate (Moreno et al. 1990, Costa et al. 1997, Rivas-Martínez et al. 2017a, 2017b). Nonetheless, the Iberian Atlantic ecoregion is located at the transition between the Mediterranean and the temperate climatic zones in SW Europe.

## Vegetation data

To define our target vegetation units, we followed the definitions of anthropogenic vegetation in the recent classification of the vegetation of Europe by Mucina et al. In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Bidentetea*. To these we added the class *Cymbalario-Parietarietea diffusae* included by Mucina et al. in the vegetation of rock crevices and screes, since this class includes the vegetation of human-made walls in our ecoregion. For the nomenclature of sintaxa we follow Mucina et al. for alliances and higher ranks, and Rivas-Martínez et al. for associations.

As vegetation data for our analyses, we obtained 28,775 vegetation relevés (i.e. records of plant species coexisting in a given sampling plot at a given date) from the Iberian and Macaronesian Vegetation Information System (SIVIM, http://sivim.info/sivi; Font et al. 2010, 2012) representing all vegetation types in our study area (i.e. anthropogenic, semi-natural and natural). Since these plots were sampled by different authors throughout the decades, there was no taxonomical consistency among them, so we homogenised all taxon names and unified the nomenclature for all plots. We followed the methodology of Jiménez-Alfaro et al. (2021b) for the Cantabrian Mountains, which includes c. 80% of the flora in the study area, to create a new checklist of the ecoregion based on the Euro+Med Plantbase (Euro+Med 2022). Other vascular plants that were not included in Euro+Med (e.g. several *Quercus* or *Salix* hybrids, non-native species, etc.), as well as the bryophytes, were classified according to Plants of the World Online (POWO 2022). Fungi and algae were removed. We merged taxa into aggregates (i) when there were sets of very close species with a difficult identification in the field (e.g. *Viola riviniana*, *V. reichenbachiana* and *V. sylvestris* were merged into *Viola riviniana* aggr.), (ii) when the taxon was recently separated in different taxa (e.g. *Hedera helix*, *Hedera hibernica* and *Hedera canariensis* were merged into *Hedera helix* aggr.) and (iii) when it was assumed that not all authors were able to identify or recognize infraspecific taxa (e.g. all *Narcissus pseudonarcissus* subspecies and varieties were merged into *Narcissus pseudonarcissus* aggr.).

We improved the geographical information of the SIVIM relevés using a downscaling procedure. The original position of vegetation relevés in SIVIM was recorded as 10 km x 10 km (52.2%) and 1 km x 1 km (47.8%) grid cells. We assigned more precise geographic coordinates to these plots using the elevation recorded originally by the authors and a Digital Elevation Model (DEM) of the Iberian Peninsula at 200 m grid resolution. In the original 10 km x 10 km or 1 km x 1 km grid cell of each relevé, we (i) kept only those 200 m cells that agreed with the relevé’s elevation (± 50 m); (ii) randomly selected one of these 200 m cells; and (iii) assigned to the relevé the coordinate of the 200 m cell’s centroid, in decimal degrees. By doing this, the plots were downscaled at higher resolution within the 10 km x 10 km cells, assuming that 200 m cells at similar elevations within the same grid cell also have similar climatic conditions.

To identify anthropogenic relevés in the SIVIM database we followed two complementary approaches. Fist, we used the expert system created to classify vegetation plots into EUNIS pan-European habitat classification types (Chytrý et al. 2020). The expert system uses the floristic composition of plots to assign them to habitat types depending on the presence and abundance of diagnostic species. We used the habitats assigned by the expert system to retrieve plots assigned to any habitat related to the vegetation classes we had defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation). Second, we also retrieved relevés if they had been classified by the original authors into any sintaxa belonging to our target anthropogenic vegetation classes (**Supplementary 1**). This produced an initial pool of 3,160 relevés from SIVIM, to which we added 89 relevés of urban or peri-urban plant communities conducted by us or extracted by us from the local literature. After initial exploration of this dataset using Two-Way Indicator Species Analysis (TWINSPAN, Roleček et al. 2009) and DCA ordination, we identified 652 relevés as either not belonging to the target anthropogenic classes or to the geographical area of the ecoregion. We removed these, leaving a dataset of 2,508 relevés for further analysis.

## Definition of target alliances and training dataset

As a previous step to classification, we prepared a revised checklist of potential anthropogenic alliances that could be present in our study ecoregion, based on regional sintaxonomical checklists (Galicia, Asturias, Cantabria) and the recent revision at the European level (Mucina, maps, FloraVeg). The provisional list included 38 alliances. Then, we filtered our dataset to include only those relevés which had been assigned by the authors to any of these 38 alliances (n = 2,201 relevés).

We used sequential TWINSPAN to do a preliminary re-classification of the 2,201 filtered relevés, attempting to re-match the relevés into the *a priori* defined checklist of 38 alliances. We used 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity, and the R package 'twinspanR' to do the TWINSPAN. To start, we performed TWINSPAN followed by pairwise PERMANOVA fitted with the R package 'vegan' to detect the maximum number of mutually different clusters in the original dataset. The number was 15, so we kept a TWINSPAN classification of the whole dataset into 15 clusters, followed by separate TWINSPANS within each of the 15 clusters. This iterative process allowed us re-assign  relevés to 29 alliances, leaving 789 relevés unclassified.

## Semi-supervised classification

The next step was to apply a semi-supervised classification to assign the so far unclassified relevés to one of the resulting 29 alliances. Semi-supervised classification uses previously defined existing groups of data to assign the unclassified relevés to them (as in supervised classification) but also creates new groups to place the data points that do not match the already existing groups (as in unsupervised classification), allowing to detect new groups and to update the previous classification (De Cáceres et al. 2010, Tichý et al. 2014). We used the 1,719 relevés that we had already re-classified into the 29 alliances types as a training set for the semi-supervised classification. Attempts by the classification to create new groups resulted on new groups with no ecological significance, and thus we kept the results of the classification into the 29 *a priori* alliances. For the classification, we applied a supervised classification with the noise clustering fuzzy algorithm, which allows the relevés to be classified either into the *a priori* groups or into a "noise" group which includes outliers and transitional plots (Wiser & De Caceres 2013). We set the fuzziness coefficient to a low value, m = 1, because we expected there would be many plots of transitional species composition and the distance to noise class was set to d = 1. The analysis was performed using the R package ‘vegclust’ (https://cran.r-project.org/web/packages/vegclust). This analysis resulted in the final classification of 2,085 relevés into 29 alliances and 423 relevés left in the noise group. We used pair-wise PERMANOVA as implemented in the package 'RVAideMemoire' to test the significance of the 29 resulting groups (with 9,999 iterations, euclidean distances and Holm's p-value correction).

## Characteristic species

We calculated sets of characteristic species for each of the 29 alliances. We defined dominant species as species with more than 25% cover in at least 5% of the plots of the alliance; constant species as species with a frequency higher than 50% in the alliance; and diagnostic species as species with an IndVal value with a p-value lower than 0.01 calculated with the package 'labdsv' with 1.000.000 iterations.

## Community traits and ecological requirements

To describe the resulting 29 alliances we compiled a dataset of species traits and ecological preferences for the species pool in the 2,508 classified relevés. For each species, we extracted from Flora iberica its average height, its median month of flowering, the length of the flowering period, and whether it was a neophyte or not. From Floraveg we extracted the species's life forms, and kept 'therophyte' and 'geophyte' for further analysis since these categories had the largest contribution to variances as per an exploratory PCA. We also extracted the ecological indicator values and the disturbance indicator values of each species. In the case of the indicators, we assigned values to species lacking values by calculating weighted average values of the species with values for each plot (weighting by species cover), and then assigning to the missing species the weighted avergae of the plots where it was present (weighting by species cover). We explored the main axes of variation using PCA in FactoMineR, separately for species traits and ecological preferences.

### 2023-12-19 src/1-import-releves.R

@@ -1,16 +1,15 @@
 library(tidyverse)
 
-openxlsx::read.xlsx("data/urban-sintaxa.xlsx", sheet = 2) %>%
-  select(Manmade, Alliance, Class) %>%
+read.csv("results/sintaxonomy/original-sintaxonomy.csv") %>%
+  select(Anthropogenic, Alliance, Class, Original) %>%
   unique %>%
-  merge(openxlsx::read.xlsx("data/urban-sintaxa.xlsx", sheet = 1)) %>% 
-  arrange(Sintaxon) -> sintaxa 
+  arrange(Original) -> sintaxa 
 
 read.csv("../canteunis/results/SIVIM/header2.3.csv", fileEncoding = "latin1") %>%
   mutate(EUNISL1 = substring(EUNIS2.3, 1, 1)) %>%
   filter(EUNISL1 == "V" | # manmade EUNIS habitats
            EUNIS2.3 %in% c("R55", "R57") | # Epilobietea fringes not in EUNIS V
-           Sintaxon %in% pull(filter(sintaxa, Manmade == "Yes"), Sintaxon)) %>% # SIVIM sintaxa belonging to anthropogenic classes
+           Sintaxon %in% pull(filter(sintaxa, Anthropogenic == "Yes"), Original)) %>% # SIVIM sintaxa belonging to anthropogenic classes
   filter(!SIVIMID %in% pull(read.csv("data/new_releves/not-manmade-cantabrian.csv", fileEncoding = "latin1"), SIVIMID)) %>%
   data.frame %>%
   select(SIVIMID, Sintaxon, EUNIS2.3, Area, Year, DEG_LON, DEG_LAT, Accuracy, ALTITUDE..M., Aspect, Slope) %>%
@@ -19,6 +18,7 @@ read.csv("../canteunis/results/SIVIM/header2.3.csv", fileEncoding = "latin1") %>
          Longitude = DEG_LON,
          Latitude = DEG_LAT) %>%
   rbind(read.csv("data/new_releves/header-additions.csv", fileEncoding = "latin1")) %>%
+  rename(Original = Sintaxon) %>%
   merge(sintaxa, all.x = TRUE) %>%
   filter(! Class %in% c("Trifolio-Geranietea sanguinei",
                         "Mulgedio-Aconitetea",
@@ -30,7 +30,7 @@ read.csv("../canteunis/results/SIVIM/header2.3.csv", fileEncoding = "latin1") %>
[…]

### 2023-12-19 src/10-maps.R

@@ -22,31 +22,35 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
                             "Papaveretea" = "Papaveretea rhoeadis",
                             "Polygono-Poetea" = "Polygono-Poetea annuae"))%>%
   mutate(Alliance = fct_relevel(Alliance, 
-                             "Galio valantiae-Parietarion judaicae",
-                             "Cymbalario-Asplenion",
-                             "Polycarpion tetraphylli",
-                             "Polygono-Coronopodion",
-                             "Saginion procumbentis",
-                             "Scleranthion annui",
-                             "Oxalidion europeae",
-                             "Spergulo arvensis-Erodion cicutariae",
-                             "Allion triquetri",
-                             "Geranio pusilli-Anthriscion caucalidis",
-                             "Echio-Galactition tomentosae",
-                             "Linario polygalifoliae-Vulpion alopecuri",
-                             "Sisymbrion officinalis",
-                             "Carduo carpetani-Cirsion odontolepidis",
-                             "Cirsion richterano-chodati",
-                             "Dauco-Melilotion",
-                             "Geo urbani-Alliarion officinalis",
-                             "Arction lappae",
-                             "Balloto-Conion maculati",
-                             "Aegopodion podagrariae",
-                             "Epilobion angustifolii",
-                             "Cynancho-Convolvulion sepium",
-                             "Senecionion fluviatilis",
-                             "Bidention tripartitae",
-                             "Paspalo-Agrostion semiverticillati")) %>%
+                               "Galio valantiae-Parietarion judaicae",
+                               "Cymbalario-Asplenion",
+                               "Polycarpion tetraphylli",
+                               "Polygono-Coronopodion",
+                               "Saginion procumbentis",
+                               "Caucalidion lappulae",
+                               "Scleranthion annui",
+                               "Oxalidion europeae",
+                               "Spergulo arvensis-Erodion cicutariae",
+                               "Allion triquetri",
+                               "Geranio pusilli-Anthriscion caucalidis",
+                               "Chenopodion muralis",
+                               "Echio-Galactition tomentosae",
+                               "Linario polygalifoliae-Vulpion alopecuri",
+                               "Sisymbrion officinalis",
+                               "Convolvulo arvensis-Agropyrion repentis",
[…]

### 2023-12-19 src/2-twinspan.R

@@ -2,16 +2,17 @@ library(tidyverse); library(vegan); library(labdsv)
 
 rm(list = ls())
 
-openxlsx::read.xlsx("data/urban-sintaxa.xlsx", sheet = 2) %>%
-  filter(Manmade == "Yes") %>%
-  filter(Assigned != "Not Cantabrian") %>%
+read.csv("results/sintaxonomy/original-sintaxonomy.csv", fileEncoding = "latin1") %>%
+  filter(Anthropogenic == "Yes") %>%
+  #filter(Assigned != "Not Cantabrian") %>%
   pull(Alliance) %>%
   unique -> alliances
 
 read.csv("data/urban-header-5.0.csv", fileEncoding = "latin1") %>%
   filter(Alliance %in% alliances) -> header
 
-read.csv("data/urban-species-5.2.csv", fileEncoding = "latin1") -> species
+read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") %>%
+  filter(SIVIMID %in% header$SIVIMID) -> species
 
 species %>%
   select(SIVIMID, Analysis.Names, Cover.percent) %>%
@@ -124,7 +125,7 @@ data.frame(Community = indicators$maxcls, Indicator = indicators$indcls,
 
 header2 %>% filter(Cluster == 1) %>% select(-Cluster) -> header3
 
-read.csv("data/urban-species-5.2.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% header3$SIVIMID) -> species
+read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% header3$SIVIMID) -> species
 
 species %>%
   select(SIVIMID, Analysis.Names, Cover.percent) %>%
@@ -138,7 +139,7 @@ twinspanR::twinspan(
   modif = TRUE,
   cut.levels = c(0, 15, 25),
   min.group.size = 10,
-  clusters = 2,
+  clusters = 10,
   diss = "multi.sorensen",
   mean.median = "mean",
   show.output.on.console = FALSE,
@@ -170,12 +171,6 @@ df2 %>%
   geom_point(data = cent, shape = 21, size = 5, aes(fill = Cluster), show.legend = T)
 
 ### Sintaxa
-
-header4 %>%
-  group_by(Cluster, Class) %>%
-  tally %>%
-  arrange(Cluster, -n) %>%
-  data.frame
 
 header4 %>%
   group_by(Cluster, Alliance) %>%
[…]

### 2023-12-19 src/3-vegclust.R

@@ -16,7 +16,7 @@ read.csv("data/urban-header-5.1.csv", fileEncoding = "latin1") %>%
 
 ### Prepare community composition matrices
 
-read.csv("data/urban-species-5.2.csv", fileEncoding = "latin1") %>% 
+read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") %>% 
   filter(SIVIMID %in% headerClassified$SIVIMID) %>%
   select(SIVIMID, Analysis.Names, Cover.percent) %>%
   spread(Analysis.Names, Cover.percent, fill = 0) %>%
@@ -24,7 +24,7 @@ read.csv("data/urban-species-5.2.csv", fileEncoding = "latin1") %>%
   decostand("normalize") -> # Chord transformation of community data
   dfClassified
 
-read.csv("data/urban-species-5.2.csv", fileEncoding = "latin1") %>% 
+read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") %>% 
   filter(SIVIMID %in% headerUnclassified$SIVIMID) %>%
   select(SIVIMID, Analysis.Names, Cover.percent) %>%
   spread(Analysis.Names, Cover.percent, fill = 0) %>%
@@ -103,14 +103,14 @@ header2 %>%
   print(n = 500)
 
 header2 %>%
-  group_by(Sintaxon, Cluster) %>%
+  group_by(Original, Cluster) %>%
   tally %>%
   arrange(Cluster, -n) %>%
   print(n = 500)
 
 ### DCA of new groups
 
-read.csv("data/urban-species-5.2.csv", fileEncoding = "latin1") -> species
+read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") -> species
 
 species %>%
   merge(header2) %>%
@@ -200,7 +200,7 @@ header2 %>%
   filter(Cluster %in% c("M1")) -> 
   headerUnclassifiedM
 
-read.csv("data/urban-species-5.2.csv", fileEncoding = "latin1") %>% 
+read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") %>% 
   filter(SIVIMID %in% headerClassifiedM$SIVIMID) %>%
   select(SIVIMID, Analysis.Names, Cover.percent) %>%
   spread(Analysis.Names, Cover.percent, fill = 0) %>%
@@ -208,7 +208,7 @@ read.csv("data/urban-species-5.2.csv", fileEncoding = "latin1") %>%
[…]

### 2023-12-19 src/4-traits.R

@@ -1,36 +1,40 @@
 library(tidyverse)
 
-read.csv("data/urban-species-5.2.csv", fileEncoding = "latin1") -> species
+read.csv("data/traits.csv", fileEncoding = "latin1") -> traits
 
-read.csv("data/urban-traits-5.2.csv", fileEncoding = "latin1") -> traits
- 
-read.csv("data/indicators/EIVE_Paper_1.0_SM_13_v.2022-11-06.xlsx - mainTable.csv", dec = ",") %>%
-  select(TaxonConcept, 
-         EIVE1.M,
-         EIVE1.N,
-         EIVE1.R,
-         EIVE1.L,
-         EIVE1.T) %>%
-  rename(link = TaxonConcept) %>%
-  unique -> eivs
+species %>% filter(! Analysis.Names %in% traits$Analysis.Names)
 
-read.csv("data/indicators/disturbance_indicator_values.csv") %>%
-  select(species, Disturbance.Severity:Soil.Disturbance) %>%
-  rename(link = species) %>%
-  unique -> deivs
-
-traits %>%
-  mutate(link = gsub(" aggr.", "", Analysis.Names)) %>%
-  merge(eivs, all.x = TRUE) %>%
-  merge(deivs, all.x = TRUE) %>%
-  select(-link) -> traits2
-
-filter(species, ! Analysis.Names %in% traits2$Analysis.Names)
-filter(traits2, ! Analysis.Names %in% species$Analysis.Names)
-
-traits2 %>% write.csv("data/urban-traits-5.2.csv", fileEncoding = "latin1", row.names = FALSE)
-
-read.csv("data/urban-traits-5.2.csv", fileEncoding = "latin1") %>%
-  select(-c(heightmean, repetido)) %>%
-  rename(neophyte = alien) %>%
-  write.csv("data/traits.csv", fileEncoding = "latin1", row.names = FALSE)
+# read.csv("data/urban-species-5.2.csv", fileEncoding = "latin1") -> species
+# 
+# read.csv("data/urban-traits-5.2.csv", fileEncoding = "latin1") -> traits
+#  
+# read.csv("data/indicators/EIVE_Paper_1.0_SM_13_v.2022-11-06.xlsx - mainTable.csv", dec = ",") %>%
+#   select(TaxonConcept, 
+#          EIVE1.M,
+#          EIVE1.N,
+#          EIVE1.R,
+#          EIVE1.L,
+#          EIVE1.T) %>%
[…]

### 2023-12-19 src/6-characteristic-spp.R

@@ -15,33 +15,6 @@ species %>%
   spread(Analysis.Names, Cover.percent, fill = 0) %>%
   column_to_rownames(var = "SIVIMID") -> df1
 
-### Indicator spp
-
-# species %>%
-#   merge(header, by = "SIVIMID") %>%
-#   select(Semisupervised, SIVIMID, Analysis.Names, Cover.percent) %>%
-#   spread(Analysis.Names, Cover.percent, fill = 0) -> plots
-# 
-# plots %>%
-#   pull(Semisupervised) -> groups
-# 
-# levels(groups) -> group.labels
-# 
-# plots %>%
-#   select(-c(SIVIMID, Semisupervised)) %>%
-#   indval(groups, numitr = 1000000) -> indicators
-# 
-# save(indicators, file = "results/characteristic-spp/indval.RData")
-
-load(file = "results/characteristic-spp/indval.RData")
-
-data.frame(Community = indicators$maxcls, Indicator = indicators$indcls,
-           p = indicators$pval, p_adj = p.adjust(indicators$pval, "holm")) %>%
-  rownames_to_column(var = "Species") %>%
-  group_by(Community) %>%
-  slice_max(Indicator, n = 4) %>%
-  data.frame
-
 ### Phidelity (not working in R, extracted with JUICE)
 
 # ### Get groups
@@ -69,6 +42,111 @@ data.frame(Community = indicators$maxcls, Indicator = indicators$indcls,
 # save(phi, file = "results/characteristic-spp/phidelity.RData")
 # 
 # load(file = "results/characteristic-spp/phidelity.RData"); summary(phi)
+
+### Indicator spp
+
+# species %>%
+#   merge(header, by = "SIVIMID") %>%
+#   select(Semisupervised, SIVIMID, Analysis.Names, Cover.percent) %>%
+#   spread(Analysis.Names, Cover.percent, fill = 0) -> plots
+# 
+# plots %>%
+#   pull(Semisupervised) -> groups
+# 
+# levels(groups) -> group.labels
+# 
+# plots %>%
+#   select(-c(SIVIMID, Semisupervised)) %>%
+#   indval(groups, numitr = 1000000) -> indicators
+# 
+# save(indicators, file = "results/characteristic-spp/indval.RData")
+
+load(file = "results/characteristic-spp/indval.RData")
+
+header %>%
+  select(Alliance) %>%
[…]

### 2023-12-19 src/7-dca.R

@@ -88,15 +88,19 @@ vegan::scores(dca1) %>%
                                 "Polycarpion tetraphylli",
                                 "Polygono-Coronopodion",
                                 "Saginion procumbentis",
+                                "Caucalidion lappulae",
                                 "Scleranthion annui",
                                 "Oxalidion europeae",
                                 "Spergulo arvensis-Erodion cicutariae",
                                 "Allion triquetri",
                                 "Geranio pusilli-Anthriscion caucalidis",
+                                "Chenopodion muralis",
                                 "Echio-Galactition tomentosae",
                                 "Linario polygalifoliae-Vulpion alopecuri",
                                 "Sisymbrion officinalis",
+                                "Convolvulo arvensis-Agropyrion repentis",
                                 "Carduo carpetani-Cirsion odontolepidis",
+                                "Silybo mariani-Urticion piluliferae", 
                                 "Cirsion richterano-chodati",
                                 "Dauco-Melilotion",
                                 "Geo urbani-Alliarion officinalis",
@@ -111,12 +115,12 @@ vegan::scores(dca1) %>%
   mutate(Cluster = as.numeric(Cluster)) -> df2
 
 aggregate(cbind(DCA1, DCA2) ~ Cluster, data = df2, FUN = mean) %>%
-  mutate(DCA1 = ifelse(Cluster == 20, DCA1-.3, DCA1)) %>%
+  mutate(DCA1 = ifelse(Cluster == 26, DCA1-.3, DCA1)) %>%
   mutate(DCA1 = ifelse(Cluster == 19, DCA1+.00, DCA1))%>%
-  mutate(DCA1 = ifelse(Cluster == 22, DCA1-.1, DCA1)) %>%
-  mutate(DCA1 = ifelse(Cluster == 15, DCA1+.03, DCA1))%>%
-  mutate(DCA2 = ifelse(Cluster == 24, DCA2+.03, DCA2))%>%
-  mutate(DCA2 = ifelse(Cluster == 25, DCA2-.03, DCA2)) -> cent
+  mutate(DCA1 = ifelse(Cluster == 18, DCA1-.15, DCA1)) %>%
+  mutate(DCA1 = ifelse(Cluster == 20, DCA1+.03, DCA1))%>%
+  mutate(DCA1 = ifelse(Cluster == 24, DCA1+.2, DCA1))%>%
[…]

### 2023-12-19 src/8-cwm-pca.R

@@ -12,13 +12,13 @@ library(tidyverse)
 #   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
 #   group_by(SIVIMID, Trait) %>%
 #   summarise(Value = weighted.mean(Value, Cover.percent)) -> plotmeans
-#
+# 
 # read.csv("data/species.csv", fileEncoding = "latin1") %>%
 #   merge(plotmeans) %>%
 #   group_by(Analysis.Names, Trait) %>%
 #   summarise(Value = weighted.mean(Value, Cover.percent)) %>%
 #   rename(ravalue = Value) -> sppmeans
-#
+# 
 # read.csv("data/traits.csv", fileEncoding = "latin1") %>%
 #   select(Analysis.Names,
 #          EIVE1.M:Soil.Disturbance) %>%
@@ -27,13 +27,13 @@ library(tidyverse)
 #   mutate(value = ifelse(is.na(Value), ravalue, Value)) %>%
 #   select(-c(Value, ravalue)) %>%
 #   spread(Trait, value) -> raspp
-#
+# 
 # read.csv("data/traits.csv", fileEncoding = "latin1") %>%
 #   select(-c(EIVE1.M:Soil.Disturbance)) %>%
 #   merge(raspp) -> traits
-#
+# 
 # ### CwMs
-#
+# 
 # traits %>%
 #   gather(Trait, Value, -Analysis.Names) %>%
 #   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
@@ -41,7 +41,7 @@ library(tidyverse)
 #   summarise(Value = weighted.mean(Value, Cover.percent)) %>%
 #   spread(Trait, Value) %>%
 #   merge(read.csv("data/header.csv", fileEncoding = "latin1")) -> cwms
-#
+# 
 # save(cwms, file = "results/cwms/cwms.RData")
 
 rm(list = ls())
@@ -209,8 +209,8 @@ ggplot(vars, aes(Dim.1, Dim.2)) +
   geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
   geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
   facet_wrap(~ PCA, scales = "free") +
-  geom_segment(aes(x = 0, y = 0, xend = Dim.1, yend = Dim.2), arrow = arrow(length = unit(1/2, "picas"))) +
[…]

### 2023-12-19 src/9-cwm-boxplots.R

@@ -62,31 +62,35 @@ cwms %>%
                             "Mowing frequency (indicator)" = "Mowing.Frequency",
                             "Grazing pressure (indicator)" = "Grazing.Pressure")) %>%
   mutate(Alliance = fct_relevel(Alliance, 
-                                "Galio valantiae-Parietarion judaicae",
-                                "Cymbalario-Asplenion",
-                                "Polycarpion tetraphylli",
-                                "Polygono-Coronopodion",
-                                "Saginion procumbentis",
-                                "Scleranthion annui",
-                                "Oxalidion europeae",
-                                "Spergulo arvensis-Erodion cicutariae",
-                                "Allion triquetri",
-                                "Geranio pusilli-Anthriscion caucalidis",
-                                "Echio-Galactition tomentosae",
-                                "Linario polygalifoliae-Vulpion alopecuri",
-                                "Sisymbrion officinalis",
-                                "Carduo carpetani-Cirsion odontolepidis",
-                                "Cirsion richterano-chodati",
-                                "Dauco-Melilotion",
-                                "Geo urbani-Alliarion officinalis",
-                                "Arction lappae",
-                                "Balloto-Conion maculati",
-                                "Aegopodion podagrariae",
-                                "Epilobion angustifolii",
-                                "Cynancho-Convolvulion sepium",
-                                "Senecionion fluviatilis",
-                                "Bidention tripartitae",
-                                "Paspalo-Agrostion semiverticillati")) %>%
+                               "Galio valantiae-Parietarion judaicae",
+                               "Cymbalario-Asplenion",
+                               "Polycarpion tetraphylli",
+                               "Polygono-Coronopodion",
+                               "Saginion procumbentis",
+                               "Caucalidion lappulae",
+                               "Scleranthion annui",
+                               "Oxalidion europeae",
+                               "Spergulo arvensis-Erodion cicutariae",
+                               "Allion triquetri",
+                               "Geranio pusilli-Anthriscion caucalidis",
+                               "Chenopodion muralis",
+                               "Echio-Galactition tomentosae",
+                               "Linario polygalifoliae-Vulpion alopecuri",
+                               "Sisymbrion officinalis",
+                               "Convolvulo arvensis-Agropyrion repentis",
+                               "Carduo carpetani-Cirsion odontolepidis",
+                               "Silybo mariani-Urticion piluliferae", 
[…]

### 2023-12-19 src/run.R

@@ -1,0 +1,70 @@
+### Test significance of classification
+
+rm(list = ls())
+
+library(tidyverse); library(vegan)
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>%
+  filter(! Alliance %in% "Noise") -> 
+  header
+
+read.csv("data/species.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% header$SIVIMID) -> species
+
+species %>%
+  select(SIVIMID, Analysis.Names, Cover.percent) %>%
+  spread(Analysis.Names, Cover.percent, fill = 0) %>%
+  column_to_rownames(var = "SIVIMID") -> df1
+
+### Permanova
+
+df1 %>% rownames_to_column(var = "SIVIMID") %>% merge(header, by = "SIVIMID") -> df1b
+RVAideMemoire::pairwise.perm.manova(dist(df1, "euclidian"), df1b$Semisupervised, nperm = 9999, p.method = "holm") -> pairperma
+adonis2(df1 ~ Semisupervised, data = df1b, permutations = 9999) -> perma
+anova(betadisper(vegdist(df1, method = "euclidean"), df1b$Semisupervised)) -> betadis
+
+### Export results
+
+save(pairperma, file = "results/permanova/pairwise.RData")
+save(perma, file = "results/permanova/permanova.RData")
+save(betadis, file = "results/permanova/beta-dispersion.RData")
+
+load(file = "results/permanova/pairwise.RData"); pairperma
+load(file = "results/permanova/permanova.RData"); perma
+load(file = "results/permanova/beta-dispersion.RData"); betadis
+
+
+### https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html
+
+rm(list = ls())
+
+library(tidyverse); library(indicspecies); library(labdsv)
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>%
+  filter(! Alliance %in% "Noise")  -> 
+  header
+
+read.csv("data/species.csv", fileEncoding = "latin1") %>% 
+  filter(SIVIMID %in% header$SIVIMID) -> species
+
+species %>%
+  select(SIVIMID, Analysis.Names, Cover.percent) %>%
+  spread(Analysis.Names, Cover.percent, fill = 0) %>%
+  column_to_rownames(var = "SIVIMID") -> df1
+
+### Indicator spp
+
+species %>%
[…]

### 2023-12-19 doc/manuscript.Rmd
methods section

@@
 
+We performed all data analyses with R version 4.3.1 [@RN5387], using the R package 'tidyverse' [@RN4662] for data processing and visualization.
+The original datasets, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
+
 ## Study ecoregion
@@
 the territories with a temperate climate in the north-western Iberian Peninsula.
-Our study ecoregion broadly corresponds with the Cantabrian Mixed Forests ecoregion *sensu* (Olson et al.
-2001), with the Iberian part of the European Atlantic province *sensu* Rivas-Martínez et al.
-2017b, and with the Iberian section of the Atlantic biogeographical region of the European Environmental Agency.
-To define the study region we followed a fine-scale revision of its geographic borders at high spatial resolution using the most recent biogeographical map of the Iberian Atlantic territories (Fernández Prieto et al.
-2020).
-Our revised Iberian Atlantic ecoregion encompasses 79,891 km^2^ of Portugal, Spain and France.
-Its climate is temperate with average monthly temperatures ranging from 6 to 20°C, and with total annual precipitation between 1,100 and 2,600 mm, with some areas having warm-dry summers but without the distinctive 2-month drought of the Mediterranean climate (Moreno et al.
-1990, Costa et al.
-1997, Rivas-Martínez et al.
-2017a, 2017b).
+Our study ecoregion broadly corresponds with the Cantabrian Mixed Forests ecoregion *sensu* Olson *et al.* [@RN4667], with the Iberian part of the European Atlantic province *sensu* Rivas-Martínez *et al.* [@RN4699] and with the Iberian section of the Atlantic biogeographical region of the European Environmental Agency (https://www.eea.europa.eu/data-and-maps/figures/biogeographical-regions-in-europe-2).
+To define in detail the study ecoregion we did a fine-scale revision of its geographic borders at high spatial resolution using the most recent biogeographical map of the Iberian Atlantic territories [@RN5372].
+The revised Iberian Atlantic ecoregion encompasses 79,891 km^2^ of Portugal, Spain and France.
+Its climate is temperate with average monthly temperatures ranging from 6 to 20°C, and with total annual precipitation between 1,100 and 2,600 mm, with some areas having warm-dry summers but without the distinctive 2-month drought of the Mediterranean climate [@RN5373].
 Nonetheless, the Iberian Atlantic ecoregion is located at the transition between the Mediterranean and the temperate climatic zones in SW Europe.
+
+## Definition of anthropogenic vegetation
+
+To define our target vegetation, we followed the definitions of anthropogenic vegetation in the recent classification of the vegetation of Europe by Mucina *et al.* [@RN4698].
+In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Bidentetea*.
+To these, we added the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls in our ecoregion.
+For the nomenclature of sintaxa we follow Mucina *et al.* [@RN4698] for alliances and higher ranks, and Rivas-Martínez *et al.* [@RN4697] for associations.
 
@@
 
-To define our target vegetation units, we followed the definitions of anthropogenic vegetation in the recent classification of the vegetation of Europe by Mucina et al.
-In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Bidentetea*.
-To these we added the class *Cymbalario-Parietarietea diffusae* included by Mucina et al.
-in the vegetation of rock crevices and screes, since this class includes the vegetation of human-made walls in our ecoregion.
-For the nomenclature of sintaxa we follow Mucina et al.
-for alliances and higher ranks, and Rivas-Martínez et al.
-for associations.
+As vegetation data for our analyses, we obtained all vegetation relevés (i.e.
+records of plant species coexisting in a given sampling plot at a given date) from our study ecoregion that were available in SIVIM, the Iberian and Macarnformation System [@RN5374].
+These were 28,775 relevés representing all  vegetation types (i.e.
+anthropogenic, semi-natural and natural).
+Since these relevés were sampled by different authors throughout the decades, there was no taxonomical consistency among them, so we homogenised all taxon names for all dataset.
+We followed the methodology applied by Jiménez-Alfaro *et al.* [@RN5375] to the Cantabrian Mountains, an area within the study ecoregion which includes c.
+80% of the ecoregional flora, to create a new taxonomical checklist for the ecoregion based on the Euro+Med Plantbase (Euro+Med 2022).
+For vascular plants not included in Euro+Med (e.g.
+several *Quercus* or *Salix* hybrids, non-native species, etc.), as well as for bryophytes, we followed Plants of the World Online [@RN5376].
+We removed Fungi and algae.
+For analytical purposes, we merged taxa into aggregates (1) when there were sets of very close species with a difficult identification in the field (e.g.
+*Viola riviniana* + *V.
+reichenbachiana* + *V.
+sylvestris* = *Viola riviniana* aggr.), (2) when the taxon was recently separated in different taxa (e.g.
+*Hedera helix* + *Hedera hibernica* + *Hedera canariensis* = *Hedera helix* aggr.) and (3) when it was assumed that not all authors were able to identify or recognize infraspecific taxa (e.g.
+all *Narcissus pseudonarcissus* subspecies and varieties = *Narcissus pseudonarcissus* aggr.).
 
-As vegetation data for our analyses, we obtained 28,775 vegetation relevés (i.e.
-records of plant species coexisting in a given sampling plot at a given date) from the Iberian and Macaronesian Vegetation Information System (SIVIM, http://sivim.info/sivi; Font et al.
-2010, 2012) representing all vegetation types in our study area (i.e.
-anthropogenic, semi-natural and natural).
-Since these plots were sampled by different authors throughout the decades, there was no taxonomical consistency among them, so we homogenised all taxon names and unified the nomenclature for all plots.
-We followed the methodology of Jiménez-Alfaro et al.
-(2021b) for the Cantabrian Mountains, which includes c.
-80% of the flora in the study area, to create a new checklist of the ecoregion based on the Euro+Med Plantbase (Euro+Med 2022).
-Other vascular plants that were not included in Euro+Med (e.g.
-several *Quercus* or *Salix* hybrids, non-native species, etc.), as well as the bryophytes, were classified according to Plants of the World Online (POWO 2022).
-Fungi and algae were removed.
-We merged taxa into aggregates (i) when there were sets of very close species with a difficult identification in the field (e.g.
-*Viola riviniana*, *V.
-reichenbachiana* and *V.
-sylvestris* were merged into *Viola riviniana* aggr.), (ii) when the taxon was recently separated in different taxa (e.g.
-*Hedera helix*, *Hedera hibernica* and *Hedera canariensis* were merged into *Hedera helix* aggr.) and (iii) when it was assumed that not all authors were able to identify or recognize infraspecific taxa (e.g.
-all *Narcissus pseudonarcissus* subspecies and varieties were merged into *Narcissus pseudonarcissus* aggr.).
-
-We improved the geographical information of the SIVIM relevés using a downscaling procedure.
+For mapping purposes we improved the geographical information of the SIVIM relevés using a downscaling procedure.
 The original position of vegetation relevés in SIVIM was recorded as 10 km x 10 km (52.2%) and 1 km x 1 km (47.8%) grid cells.
-We assigned more precise geographic coordinates to these plots using the elevation recorded originally by the authors and a Digital Elevation Model (DEM) of the Iberian Peninsula at 200 m grid resolution.
-In the original 10 km x 10 km or 1 km x 1 km grid cell of each relevé, we (i) kept only those 200 m cells that agreed with the relevé’s elevation (± 50 m); (ii) randomly selected one of these 200 m cells; and (iii) assigned to the relevé the coordinate of the 200 m cell’s centroid, in decimal degrees.
-By doing this, the plots were downscaled at higher resolution within the 10 km x 10 km cells, assuming that 200 m cells at similar elevations within the same grid cell also have similar climatic conditions.
+We assigned more precise geographic coordinates to these plots using the elevation recorded originally by the authors and a digital elevation model (DEM) of the Iberian Peninsula at 200 m grid resolution.
+From the original 10 km x 10 km or 1 km x 1 km grid cell of each relevé, we (1) kept only those 200 m cells that agreed with the relevé's elevation (± 50 m); (2) randomly selected one of these 200 m cells; and (3) assigned to the relevé the coordinate of the 200 m cell's centroid, in decimal degrees.
+By doing this, the plots were downscaled at higher resolution within their cells, assuming that 200 m cells at similar elevations within the same grid cell also have similar climatic conditions.
 
 To identify anthropogenic relevés in the SIVIM database we followed two complementary approaches.
-Fist, we used the expert system created to classify vegetation plots into EUNIS pan-European habitat classification types (Chytrý et al.
-2020).
-The expert system uses the floristic composition of plots to assign them to habitat types depending on the presence and abundance of diagnostic species.
-We used the habitats assigned by the expert system to retrieve plots assigned to any habitat related to the vegetation classes we had defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation).
-Second, we also retrieved relevés if they had been classified by the original authors into any sintaxa belonging to our target anthropogenic vegetation classes (**Supplementary 1**).
-This produced an initial pool of 3,160 relevés from SIVIM, to which we added 89 relevés of urban or peri-urban plant communities conducted by us or extracted by us from the local literature.
-After initial exploration of this dataset using Two-Way Indicator Species Analysis (TWINSPAN, Roleček et al.
-2009) and DCA ordination, we identified 652 relevés as either not belonging to the target anthropogenic classes or to the geographical area of the ecoregion.
-We removed these, leaving a dataset of 2,508 relevés for further analysis.
+Fist, we used the expert system created to classify vegetation plots into EUNIS pan-European habitat classification types [@RN5061].
+The expert system uses the floristic composition of plots to assign them to habitat types depending on the presence and abundance of characteristic species.
+We used the habitats assigned by the expert system to retrieve plots that had been assigned to any habitat related to the vegetation classes we had previously defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation).
+Second, we also retrieved relevés if they had been classified by the original authors of the relevé into any sintaxa belonging to our target anthropogenic vegetation classes (**Supplementary 1**).
+This produced an initial pool of 3,160 relevés from SIVIM, to which we added 89 relevés of urban or peri-urban plant communities conducted by us or extracted by us from the local literature [@RN5377; @RN5378].
+After initial exploration of this dataset using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379], we identified 652  outlier relevés which either did not belong to the target anthropogenic classes or to the geographical area of the ecoregion.
+We removed these, leaving a dataset of 2,508 relevés for further analysi: 2,419 relevés from SIVIM, originally recorded in 89 publications see (**Supplementary 2**); plus 89 relevés added by us.
 
@@
 
-As a previous step to classification, we prepared a revised checklist of potential anthropogenic alliances that could be present in our study ecoregion, based on regional sintaxonomical checklists (Galicia, Asturias, Cantabria) and the recent revision at the European level (Mucina, maps, FloraVeg).
+As a previous step before classification, we prepared a revised checklist of potential anthropogenic alliances that could be present in our study ecoregion, based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
 The provisional list included 38 alliances.
+
 Then, we filtered our dataset to include only those relevés which had been assigned by the authors to any of these 38 alliances (n = 2,201 relevés).
-
-We used sequential TWINSPAN to do a preliminary re-classification of the 2,201 filtered relevés, attempting to re-match the relevés into the *a priori* defined checklist of 38 alliances.
-We used 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity, and the R package 'twinspanR' to do the TWINSPAN.
-To start, we performed TWINSPAN followed by pairwise PERMANOVA fitted with the R package 'vegan' to detect the maximum number of mutually different clusters in the original dataset.
-The number was 15, so we kept a TWINSPAN classification of the whole dataset into 15 clusters, followed by separate TWINSPANS within each of the 15 clusters.
-This iterative process allowed us re-assign  relevés to 29 alliances, leaving 789 relevés unclassified.
+We used sequential TWINSPANs to do a preliminary re-classification of the 2,201 filtered relevés, attempting to re-match the relevés into the *a priori* defined checklist of 38 alliances.
+We used 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity, and the R package 'twinspanR' [@RN5380] to do the TWINSPANs.
+To start, we performed TWINSPAN followed by pairwise PERMANOVA (with 9,999 iterations, euclidean distances and Holm's p-value correction) fitted with the R package 'RVAideMemoire' [@RN5381] to detect the maximum number of mutually different clusters in the filtered dataset.
+The number was 15, so we kept a TWINSPAN classification into 15 clusters, followed by separate TWINSPANS within each clusters.
+This iterative process allowed us to validate 29 alliances and assign 1,719 relevés to them, leaving 789 relevés as unclassified.
 
@@
 
-The next step was to apply a semi-supervised classification to assign the so far unclassified relevés to one of the resulting 29 alliances.
+Next we applied a semi-supervised classification to assign the so far unclassified relevés to one of the validated 29 alliances.
 Semi-supervised classification uses previously defined existing groups of data to assign the unclassified relevés to them (as in supervised classification) but also creates new groups to place the data points that do not match the already existing groups (as in unsupervised classification), allowing to detect new groups and to update the previous classification (De Cáceres et al.
@@
 2014).
-We used the 1,719 relevés that we had already re-classified into the 29 alliances types as a training set for the semi-supervised classification.
+We used the 1,719 relevés that we had already re-classified into the 29 alliances as a training set for the semi-supervised classification.
 Attempts by the classification to create new groups resulted on new groups with no ecological significance, and thus we kept the results of the classification into the 29 *a priori* alliances.
-For the classification, we applied a supervised classification with the noise clustering fuzzy algorithm, which allows the relevés to be classified either into the *a priori* groups or into a "noise" group which includes outliers and transitional plots (Wiser & De Caceres 2013).
-We set the fuzziness coefficient to a low value, m = 1, because we expected there would be many plots of transitional species composition and the distance to noise class was set to d = 1.
-The analysis was performed using the R package ‘vegclust’ (https://cran.r-project.org/web/packages/vegclust).
-This analysis resulted in the final classification of 2,085 relevés into 29 alliances and 423 relevés left in the noise group.
-We used pair-wise PERMANOVA as implemented in the package 'RVAideMemoire' to test the significance of the 29 resulting groups (with 9,999 iterations, euclidean distances and Holm's p-value correction).
+For the classification, we applied a noise clustering fuzzy algorithm, which allows the relevés to be classified either into the *a priori* groups or into a 'noise' group which includes outliers and transitional relevés [@RN5382].
+We set the fuzziness coefficient to a low value (m = 1) to accommodate a high number of transitional relevés; and we set the distance to the noise class to d = 1.
+We performed this analysis with the R package 'vegclust' [@RN5383].
+The semi-supervised classification with noise clustering resulted in the final classification of 2,085 relevés into 29 alliances and 423 relevés left in the noise group.
+We used pair-wise PERMANOVA in the same conditions as reported above to test the significance of the 29 resulting groups; and Detrended Correspondence Analysis (DCA) as implemented in the R package 'vegan' [@RN5386] to visualize the relationships between the vegetation groups.
 
@@
 
-We calculated sets of characteristic species for each of the 29 alliances.
-We defined dominant species as species with more than 25% cover in at least 5% of the plots of the alliance; constant species as species with a frequency higher than 50% in the alliance; and diagnostic species as species with an IndVal value with a p-value lower than 0.01 calculated with the package 'labdsv' with 1.000.000 iterations.
+We calculated sets of characteristic species [@RN5061] for each of the 29 alliances.
+We defined dominant species as species with more than 25% cover in at least 5% of the relevés of the alliance; constant species as species with a frequency higher than 50% in the alliance; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.01, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations.
 
@@
 
-To describe the resulting 29 alliances we compiled a dataset of species traits and ecological preferences for the species pool in the 2,508 classified relevés.
-For each species, we extracted from Flora iberica its average height, its median month of flowering, the length of the flowering period, and whether it was a neophyte or not.
-From Floraveg we extracted the species's life forms, and kept 'therophyte' and 'geophyte' for further analysis since these categories had the largest contribution to variances as per an exploratory PCA.
-We also extracted the ecological indicator values and the disturbance indicator values of each species.
-In the case of the indicators, we assigned values to species lacking values by calculating weighted average values of the species with values for each plot (weighting by species cover), and then assigning to the missing species the weighted avergae of the plots where it was present (weighting by species cover).
-We explored the main axes of variation using PCA in FactoMineR, separately for species traits and ecological preferences.
+To describe the resulting 29 alliances we compiled a dataset of species traits and ecological preferences for the species pool of the 2,508 classified relevés.
+For each species, we extracted from *Flora iberica* [-@RN3243] its maximum height, its median month of flowering, the length of the flowering period, and whether it was a neophyte or not.
+From FloraVeg.EU [-@RN5385] we extracted the species's life forms, and kept 'therophyte' and 'geophyte' for further analysis since these categories had the largest contribution to variance as per an exploratory Principal Component Analysis (PCA) performed using the R package 'FactoMineR' [@RN3166].
+We also extracted the ecological indicator values [@RN5384] and the disturbance indicator values [@RN5101] of each species.
+In the case of the ecological and disturbance indicators, we assigned values to species lacking them by (1) calculating weighted average values of the species with values for each plot (weighting by species cover), and (2) assigning to the missing species the weighted average values of the plots where they were present (weighting by species cover).
+Then, for each relevé we calculated its community-weighted mean (weighting by species cover) for each trait and ecological preference, and explored the main axes of variation using PCA, separately for species traits and ecological preferences.
+
+# Data availability
+
+The original datasets, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
+Upon publication, a version of record of the repository will be deposited in Zenodo.
+
+# References
+
+::: {#refs}
+:::
+
+# Figures
+
+```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "."}
+knitr::include_graphics("../results/figures/dca.png")
+```
+
+```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "."}
+knitr::include_graphics("../results/figures/cwms.png")
+```
+
+```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "."}
+knitr::include_graphics("../results/figures/boxplots.png")
+```
+
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "."}
+knitr::include_graphics("../results/figures/maps.png")
+```

### 2023-12-20 doc/manuscript.Rmd
revised fake Silybo-Urticion

@@
 We performed all data analyses with R version 4.3.1 [@RN5387], using the R package 'tidyverse' [@RN4662] for data processing and visualization.
-The original datasets, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
+The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
 
@@
 the territories with a temperate climate in the north-western Iberian Peninsula.
-Our study ecoregion broadly corresponds with the Cantabrian Mixed Forests ecoregion *sensu* Olson *et al.* [@RN4667], with the Iberian part of the European Atlantic province *sensu* Rivas-Martínez *et al.* [@RN4699] and with the Iberian section of the Atlantic biogeographical region of the European Environmental Agency (https://www.eea.europa.eu/data-and-maps/figures/biogeographical-regions-in-europe-2).
+Our study ecoregion broadly corresponds with the Cantabrian Mixed Forests ecoregion *sensu* Olson *et al.* [-@RN4667], with the Iberian part of the European Atlantic province *sensu* Rivas-Martínez *et al.* [-@RN4699] and with the Iberian section of the Atlantic biogeographical region of the European Environmental Agency (https://www.eea.europa.eu/data-and-maps/figures/biogeographical-regions-in-europe-2).
 To define in detail the study ecoregion we did a fine-scale revision of its geographic borders at high spatial resolution using the most recent biogeographical map of the Iberian Atlantic territories [@RN5372].
@@
 Its climate is temperate with average monthly temperatures ranging from 6 to 20°C, and with total annual precipitation between 1,100 and 2,600 mm, with some areas having warm-dry summers but without the distinctive 2-month drought of the Mediterranean climate [@RN5373].
-Nonetheless, the Iberian Atlantic ecoregion is located at the transition between the Mediterranean and the temperate climatic zones in SW Europe.
+Nonetheless, the Iberian Atlantic ecoregion is located at the transition between the temperate and the Mediterranean climates in southern Europe.
 
@@
 
-To define our target vegetation, we followed the definitions of anthropogenic vegetation in the recent classification of the vegetation of Europe by Mucina *et al.* [@RN4698].
+To define our target vegetation, we followed the definitions of anthropogenic vegetation in the recent classification of the vegetation of Europe by Mucina *et al.* [-@RN4698].
 In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Bidentetea*.
-To these, we added the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls in our ecoregion.
-For the nomenclature of sintaxa we follow Mucina *et al.* [@RN4698] for alliances and higher ranks, and Rivas-Martínez *et al.* [@RN4697] for associations.
+For a complete synthesis, we added the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [-@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls in our ecoregion.
+For the nomenclature of sintaxa we follow Mucina *et al.* [-@RN4698] for alliances and higher ranks, and Rivas-Martínez *et al.* [-@RN4697] for associations.
 
@@
 As vegetation data for our analyses, we obtained all vegetation relevés (i.e.
-records of plant species coexisting in a given sampling plot at a given date) from our study ecoregion that were available in SIVIM, the Iberian and Macarnformation System [@RN5374].
+records of plant species coexisting in a given sampling plot at a given date) from our study ecoregion that were available in SIVIM, the Iberian and Macaronesian Vegetation Information System [@RN5374].
 These were 28,775 relevés representing all vegetation types (i.e.
 anthropogenic, semi-natural and natural).
-Since these relevés were sampled by different authors throughout the decades, there was no taxonomical consistency among them, so we homogenised all taxon names for all dataset.
-We followed the methodology applied by Jiménez-Alfaro *et al.* [@RN5375] to the Cantabrian Mountains, an area within the study ecoregion which includes c.
-80% of the ecoregional flora, to create a new taxonomical checklist for the ecoregion based on the Euro+Med Plantbase (Euro+Med 2022).
+Since these relevés were sampled by different authors throughout the decades, there was no taxonomical consistency among them, so we homogenized all taxon names for the whole dataset.
+We followed the methodology applied by Jiménez-Alfaro *et al.* [-@RN5375] to the Cantabrian Mountains, an area within the study ecoregion which includes c.
+80% of the ecoregional flora, to create a new taxonomical checklist for the ecoregion based on the Euro+Med Plantbase [@RN5044].
 For vascular plants not included in Euro+Med (e.g.
@@
 
-For mapping purposes we improved the geographical information of the SIVIM relevés using a downscaling procedure.
+For mapping purposes, we improved the geographical information of the SIVIM relevés using a downscaling procedure.
 The original position of vegetation relevés in SIVIM was recorded as 10 km x 10 km (52.2%) and 1 km x 1 km (47.8%) grid cells.
@@
 After initial exploration of this dataset using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379], we identified 652 outlier relevés which either did not belong to the target anthropogenic classes or to the geographical area of the ecoregion.
-We removed these, leaving a dataset of 2,508 relevés for further analysi: 2,419 relevés from SIVIM, originally recorded in 89 publications see (**Supplementary 2**); plus 89 relevés added by us.
+We removed these, leaving a dataset of 2,508 relevés for further analysis: 2,419 relevés from SIVIM, originally recorded in 89 publications see (**Supplementary 2**); plus 89 relevés added by us.
 
@@
 We used 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity, and the R package 'twinspanR' [@RN5380] to do the TWINSPANs.
-To start, we performed TWINSPAN followed by pairwise PERMANOVA (with 9,999 iterations, euclidean distances and Holm's p-value correction) fitted with the R package 'RVAideMemoire' [@RN5381] to detect the maximum number of mutually different clusters in the filtered dataset.
-The number was 15, so we kept a TWINSPAN classification into 15 clusters, followed by separate TWINSPANS within each clusters.
-This iterative process allowed us to validate 29 alliances and assign 1,719 relevés to them, leaving 789 relevés as unclassified.
+To start, we did a TWINSPAN classification into 15 clusters, followed by separate TWINSPANS within each cluster.
+This iterative process allowed us to validate 28 alliances and assign 1,725 relevés to them, leaving 783 relevés as unclassified.
 
@@
 
-Next we applied a semi-supervised classification to assign the so far unclassified relevés to one of the validated 29 alliances.
-Semi-supervised classification uses previously defined existing groups of data to assign the unclassified relevés to them (as in supervised classification) but also creates new groups to place the data points that do not match the already existing groups (as in unsupervised classification), allowing to detect new groups and to update the previous classification (De Cáceres et al.
-2010, Tichý et al.
-2014).
-We used the 1,719 relevés that we had already re-classified into the 29 alliances as a training set for the semi-supervised classification.
-Attempts by the classification to create new groups resulted on new groups with no ecological significance, and thus we kept the results of the classification into the 29 *a priori* alliances.
+Next, we applied a semi-supervised classification to assign the so far unclassified relevés to one of the validated 28 alliances.
+Semi-supervised classification uses previously defined groups of data to assign the unclassified relevés to them (as in supervised classification) but also creates new groups to place the data points that do not match the already existing groups (as in unsupervised classification), allowing to detect new groups and to update the previous classification [@RN5383].
+We used the 1,719 relevés that we had already re-classified into the 28 alliances as a training set for the semi-supervised classification.
+Attempts by the classification to create new groups resulted on new groups with no ecological significance, and thus we kept the results of the classification into the 28 *a priori* alliances.
 For the classification, we applied a noise clustering fuzzy algorithm, which allows the relevés to be classified either into the *a priori* groups or into a 'noise' group which includes outliers and transitional relevés [@RN5382].
@@
 We performed this analysis with the R package 'vegclust' [@RN5383].
-The semi-supervised classification with noise clustering resulted in the final classification of 2,085 relevés into 29 alliances and 423 relevés left in the noise group.
-We used pair-wise PERMANOVA in the same conditions as reported above to test the significance of the 29 resulting groups; and Detrended Correspondence Analysis (DCA) as implemented in the R package 'vegan' [@RN5386] to visualize the relationships between the vegetation groups.
+The semi-supervised classification with noise clustering resulted in the final classification of 2,081 relevés into 28 alliances and 427 relevés left in the noise group.
+We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package 'RVAideMemoire' [@RN5381] to detect the maximum number of mutually different clusters in the filtered dataset; and Detrended Correspondence Analysis (DCA) as implemented in the R package 'vegan' [@RN5386] to visualize the relationships between the vegetation groups.
 
@@
 
-We calculated sets of characteristic species [@RN5061] for each of the 29 alliances.
-We defined dominant species as species with more than 25% cover in at least 5% of the relevés of the alliance; constant species as species with a frequency higher than 50% in the alliance; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.01, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations.
+We calculated sets of characteristic species [@RN5061] for each of the 28 alliances.
+We defined dominant species as species with more than 25% cover in at least 5% of the relevés of the alliance; constant species as species with a frequency higher than 50% in the alliance; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.05, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations.
 
@@
 
-To describe the resulting 29 alliances we compiled a dataset of species traits and ecological preferences for the species pool of the 2,508 classified relevés.
+To describe the resulting 28 alliances, we compiled a dataset of species traits and ecological preferences for the species pool of the 2,508 classified relevés.
 For each species, we extracted from *Flora iberica* [-@RN3243] its maximum height, its median month of flowering, the length of the flowering period, and whether it was a neophyte or not.
-From FloraVeg.EU [-@RN5385] we extracted the species's life forms, and kept 'therophyte' and 'geophyte' for further analysis since these categories had the largest contribution to variance as per an exploratory Principal Component Analysis (PCA) performed using the R package 'FactoMineR' [@RN3166].
+From FloraVeg.EU [-@RN5385] we extracted the species’ life forms and kept 'therophyte' and 'geophyte' for further analysis since these two categories had the largest contribution to variance as per an exploratory Principal Component Analysis (PCA) performed using the R package 'FactoMineR' [@RN3166].
 We also extracted the ecological indicator values [@RN5384] and the disturbance indicator values [@RN5101] of each species.
@@
 
-The original datasets, R code for the analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
+The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
 Upon publication, a version of record of the repository will be deposited in Zenodo.

### 2023-12-20 src/10-maps.R

@@ -87,7 +87,7 @@ rgdal::readOGR(dsn = "data/map", layer = "IberoAtlantic") -> Ecoregions # Map fi
 
 header %>%
   ggplot(aes(Longitude, Latitude)) +
-  facet_wrap( ~ Alliance, ncol = 5) +
+  facet_wrap( ~ Alliance, ncol = 4) +
   geom_polygon(data = Ecoregions, aes(x = long, y = lat, group = group), 
                color = "black", fill = "gainsboro", size = 0.25, show.legend = FALSE) +
   geom_point(aes(color = Class), show.legend = TRUE, alpha = 0.4) +
@@ -124,5 +124,5 @@ header %>%
   guides(colour = guide_legend(override.aes = list(alpha = 1))) -> Fig2; Fig2
 
 ggsave(Fig2, file = "results/figures/maps.png", bg = "white", 
-       path = NULL, scale = 1, width = 179, height = 160, units = "mm", dpi = 600)
+       path = NULL, scale = 1, width = 179, height = 215, units = "mm", dpi = 600)
 

### 2023-12-20 src/2-twinspan.R

@@ -1133,6 +1133,10 @@ data.frame(Community = indicators$maxcls, Indicator = indicators$indcls,
 
 ### Save
 
+# header3 %>%
+#   mutate(L1 = "Epilobion angustifolii") %>%
+#   write.csv("data/urban-header-5.1I.csv", fileEncoding = "latin1")
+
 header4 %>%
   mutate(L1 = fct_recode(as.factor(Cluster),
                          "Allion triquetri" = "1",
@@ -1253,7 +1257,7 @@ header4 %>%
                          "Balloto-Conion maculati" = "6",
                          "Echio-Galactition tomentosae" = "7",
                          "Arction lappae" = "8",
-                         "Silybo mariani-Urticion piluliferae" = "9",
+                         "Dauco-Melilotion" = "9",
                          "Arction lappae" = "10",
                          "Dauco-Melilotion" = "11",
                          "Convolvulo arvensis-Agropyrion repentis" = "12")) %>%
@@ -1278,7 +1282,7 @@ twinspanR::twinspan(
   modif = TRUE,
   cut.levels = c(0, 15, 25),
   min.group.size = 10,
-  clusters = 4,
+  clusters = 5,
   diss = "multi.sorensen",
   mean.median = "mean",
   show.output.on.console = FALSE,
@@ -1367,7 +1371,8 @@ header4 %>%
                          "Cirsion richterano-chodati" = "1",
                          "Cirsion richterano-chodati" = "2",
                          "Carduo carpetani-Cirsion odontolepidis" = "3",
-                         "Cirsion richterano-chodati" = "4")) %>%
+                         "Cirsion richterano-chodati" = "4",
+                         "Cirsion richterano-chodati" = "5")) %>%
   select(-Cluster) %>%
   write.csv("data/urban-header-5.1K.csv", fileEncoding = "latin1")
 
@@ -1608,7 +1613,7 @@ twinspanR::twinspan(
   modif = TRUE,
   cut.levels = c(0, 15, 25),
   min.group.size = 10,
-  clusters = 3,
+  clusters = 5,
   diss = "multi.sorensen",
   mean.median = "mean",
   show.output.on.console = FALSE,
@@ -1643,6 +1648,12 @@ df2 %>%
 
 header4 %>%
   group_by(Cluster, Alliance) %>%
+  tally %>%
+  arrange(Cluster, -n) %>%
+  data.frame
+
+header4 %>%
+  group_by(Cluster, Original) %>%
   tally %>%
[…]

### 2023-12-20 src/3-vegclust.R

@@ -385,10 +385,9 @@ header3 %>%
                          "Saginion procumbentis" = "F25",
                          "Scleranthion annui" = "F26",
                          "Senecionion fluviatilis" = "F27",
-                         "Silybo mariani-Urticion piluliferae" = "F28",
-                         "Sisymbrion officinalis" = "F29",
+                         "Sisymbrion officinalis" = "F28",
                          "Allion triquetri" = "F3",
-                         "Spergulo arvensis-Erodion cicutariae" = "F30",
+                         "Spergulo arvensis-Erodion cicutariae" = "F29",
                          "Arction lappae" = "F4",
                          "Balloto-Conion maculati" = "F5",
                          "Bidention tripartitae" = "F6",
@@ -418,6 +417,9 @@ read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") %>%
 header %>% group_by(SIVIMID) %>% filter(!SIVIMID %in% species$SIVIMID)
 species %>% group_by(SIVIMID) %>% filter(!SIVIMID %in% header$SIVIMID)
   
+header %>% filter(Alliance == "Noise") %>% group_by(SIVIMID)
+header %>% filter(Alliance != "Noise") %>% group_by(SIVIMID)
+
 # openxlsx::read.xlsx("data/urban-sintaxa.xlsx", sheet = 1) %>%
 #   rename(Original = Sintaxon) %>%
 #   select(Original, Alliance) %>%

### 2023-12-20 src/5-permanova.R

@@ -4,28 +4,28 @@ rm(list = ls())
 
 library(tidyverse); library(vegan)
 
-read.csv("data/header.csv", fileEncoding = "latin1") -> 
-  header
-
-read.csv("data/species.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% header$SIVIMID) -> species
-
-species %>%
-  select(SIVIMID, Analysis.Names, Cover.percent) %>%
-  spread(Analysis.Names, Cover.percent, fill = 0) %>%
-  column_to_rownames(var = "SIVIMID") -> df1
-
-### Permanova
-
-df1 %>% rownames_to_column(var = "SIVIMID") %>% merge(header, by = "SIVIMID") -> df1b
-RVAideMemoire::pairwise.perm.manova(dist(df1, "euclidian"), df1b$Semisupervised, nperm = 9999, p.method = "holm") -> pairperma
-adonis2(df1 ~ Semisupervised, data = df1b, permutations = 9999) -> perma
-anova(betadisper(vegdist(df1, method = "euclidean"), df1b$Semisupervised)) -> betadis
-
-### Export results
-
-save(pairperma, file = "results/permanova/pairwise.RData")
-save(perma, file = "results/permanova/permanova.RData")
-save(betadis, file = "results/permanova/beta-dispersion.RData")
+# read.csv("data/header.csv", fileEncoding = "latin1") %>% filter(! Alliance %in% "Noise") -> 
+#   header
+# 
+# read.csv("data/species.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% header$SIVIMID) -> species
+# 
+# species %>%
+#   select(SIVIMID, Analysis.Names, Cover.percent) %>%
+#   spread(Analysis.Names, Cover.percent, fill = 0) %>%
+#   column_to_rownames(var = "SIVIMID") -> df1
+# 
+# ### Permanova
+# 
+# df1 %>% rownames_to_column(var = "SIVIMID") %>% merge(header, by = "SIVIMID") -> df1b
+# RVAideMemoire::pairwise.perm.manova(dist(df1, "euclidian"), df1b$Alliance, nperm = 100000, p.method = "holm") -> pairperma
+# adonis2(df1 ~ Alliance, data = df1b, permutations = 9999) -> perma
+# anova(betadisper(vegdist(df1, method = "euclidean"), df1b$Alliance)) -> betadis
+# 
+# ### Export results
+# 
+# save(pairperma, file = "results/permanova/pairwise.RData")
[…]

### 2023-12-20 src/6-characteristic-spp.R

@@ -4,7 +4,7 @@ rm(list = ls())
 
 library(tidyverse); library(indicspecies); library(labdsv)
 
-read.csv("data/header.csv", fileEncoding = "latin1") -> 
+read.csv("data/header.csv", fileEncoding = "latin1") %>% filter(! Alliance %in% "Noise")-> 
   header
 
 read.csv("data/species.csv", fileEncoding = "latin1") %>% 
@@ -22,7 +22,7 @@ species %>%
 # df1 %>% 
 #   rownames_to_column(var = "SIVIMID") %>% 
 #   merge(header, by = "SIVIMID") %>% 
-#   pull(Semisupervised) -> groups
+#   pull(Alliance) -> groups
 # 
 # ### Transform relevés to presence/absence
 # 
@@ -47,16 +47,16 @@ species %>%
 
 # species %>%
 #   merge(header, by = "SIVIMID") %>%
-#   select(Semisupervised, SIVIMID, Analysis.Names, Cover.percent) %>%
+#   select(Alliance, SIVIMID, Analysis.Names, Cover.percent) %>%
 #   spread(Analysis.Names, Cover.percent, fill = 0) -> plots
 # 
 # plots %>%
-#   pull(Semisupervised) -> groups
+#   pull(Alliance) -> groups
 # 
 # levels(groups) -> group.labels
 # 
 # plots %>%
-#   select(-c(SIVIMID, Semisupervised)) %>%
+#   select(-c(SIVIMID, Alliance)) %>%
 #   indval(groups, numitr = 1000000) -> indicators
 # 
 # save(indicators, file = "results/characteristic-spp/indval.RData")
@@ -64,6 +64,7 @@ species %>%
 load(file = "results/characteristic-spp/indval.RData")
 
 header %>%
+  filter(!Alliance %in% "Noise") %>%
   select(Alliance) %>%
   unique %>%
   mutate(Community = as.numeric(as.factor(Alliance))) -> communities
@@ -71,7 +72,7 @@ header %>%
 data.frame(Community = indicators$maxcls, Indicator = indicators$indcls,
            p = indicators$pval, p_adj = p.adjust(indicators$pval, "holm")) %>%
   rownames_to_column(var = "Species") %>%
-  filter(p_adj < 0.01) %>%
+  filter(p_adj < 0.05) %>%
   merge(communities) %>%
   select(-c(Community, p)) %>%
   arrange(Alliance, -Indicator) %>%

### 2023-12-20 src/7-dca.R

@@ -115,12 +115,10 @@ vegan::scores(dca1) %>%
   mutate(Cluster = as.numeric(Cluster)) -> df2
 
 aggregate(cbind(DCA1, DCA2) ~ Cluster, data = df2, FUN = mean) %>%
-  mutate(DCA1 = ifelse(Cluster == 26, DCA1-.3, DCA1)) %>%
-  mutate(DCA1 = ifelse(Cluster == 19, DCA1+.00, DCA1))%>%
-  mutate(DCA1 = ifelse(Cluster == 18, DCA1-.15, DCA1)) %>%
-  mutate(DCA1 = ifelse(Cluster == 20, DCA1+.03, DCA1))%>%
-  mutate(DCA1 = ifelse(Cluster == 24, DCA1+.2, DCA1))%>%
-  mutate(DCA1 = ifelse(Cluster == 23, DCA1-.2, DCA1)) -> cent
+  mutate(DCA1 = ifelse(Cluster == 18, DCA1+.05, DCA1)) %>%
+  mutate(DCA1 = ifelse(Cluster == 25, DCA1-.05, DCA1))%>%
+  mutate(DCA1 = ifelse(Cluster == 23, DCA1+.1, DCA1))%>%
+  mutate(DCA1 = ifelse(Cluster == 22, DCA1-.25, DCA1)) -> cent
 merge(df2, setNames(cent, c("Cluster", "oDCA1", "oDCA2")), by = "Cluster", sort = FALSE) %>%
   mutate(Group = paste(Cluster, Alliance)) %>%
   mutate(Group = fct_reorder(Group, Cluster)) -> segs
@@ -168,7 +166,6 @@ df2 %>%
                                 "darkmagenta", 
                                 "darkmagenta",  
                                 "darkmagenta",  
-                                "darkmagenta",
                                 "darkmagenta",
                                 "limegreen",
                                 "limegreen",

### 2023-12-20 src/8-cwm-pca.R

@@ -4,45 +4,45 @@ library(tidyverse)
 
 ### Reciprocal averaging
 
-# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
-#   select(Analysis.Names,
-#          EIVE1.M:Soil.Disturbance) %>%
-#   gather(Trait, Value, -Analysis.Names) %>%
-#   na.omit %>%
-#   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-#   group_by(SIVIMID, Trait) %>%
-#   summarise(Value = weighted.mean(Value, Cover.percent)) -> plotmeans
-# 
-# read.csv("data/species.csv", fileEncoding = "latin1") %>%
-#   merge(plotmeans) %>%
-#   group_by(Analysis.Names, Trait) %>%
-#   summarise(Value = weighted.mean(Value, Cover.percent)) %>%
-#   rename(ravalue = Value) -> sppmeans
-# 
-# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
-#   select(Analysis.Names,
-#          EIVE1.M:Soil.Disturbance) %>%
-#   gather(Trait, Value, -Analysis.Names) %>%
-#   merge(sppmeans, all.x = TRUE) %>%
-#   mutate(value = ifelse(is.na(Value), ravalue, Value)) %>%
-#   select(-c(Value, ravalue)) %>%
-#   spread(Trait, value) -> raspp
-# 
-# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
-#   select(-c(EIVE1.M:Soil.Disturbance)) %>%
-#   merge(raspp) -> traits
-# 
-# ### CwMs
-# 
-# traits %>%
-#   gather(Trait, Value, -Analysis.Names) %>%
-#   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-#   group_by(SIVIMID, Trait) %>%
-#   summarise(Value = weighted.mean(Value, Cover.percent)) %>%
-#   spread(Trait, Value) %>%
-#   merge(read.csv("data/header.csv", fileEncoding = "latin1")) -> cwms
-# 
-# save(cwms, file = "results/cwms/cwms.RData")
+read.csv("data/traits.csv", fileEncoding = "latin1") %>%
+  select(Analysis.Names,
+         EIVE1.M:Soil.Disturbance) %>%
+  gather(Trait, Value, -Analysis.Names) %>%
+  na.omit %>%
[…]

### 2023-12-20 src/run.R

@@ -4,8 +4,7 @@ rm(list = ls())
 
 library(tidyverse); library(vegan)
 
-read.csv("data/header.csv", fileEncoding = "latin1") %>%
-  filter(! Alliance %in% "Noise") -> 
+read.csv("data/header.csv", fileEncoding = "latin1") %>% filter(! Alliance %in% "Noise") ->
   header
 
 read.csv("data/species.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% header$SIVIMID) -> species
@@ -18,9 +17,9 @@ species %>%
 ### Permanova
 
 df1 %>% rownames_to_column(var = "SIVIMID") %>% merge(header, by = "SIVIMID") -> df1b
-RVAideMemoire::pairwise.perm.manova(dist(df1, "euclidian"), df1b$Semisupervised, nperm = 9999, p.method = "holm") -> pairperma
-adonis2(df1 ~ Semisupervised, data = df1b, permutations = 9999) -> perma
-anova(betadisper(vegdist(df1, method = "euclidean"), df1b$Semisupervised)) -> betadis
+RVAideMemoire::pairwise.perm.manova(dist(df1, "euclidian"), df1b$Alliance, nperm = 100000, p.method = "holm") -> pairperma
+adonis2(df1 ~ Alliance, data = df1b, permutations = 9999) -> perma
+anova(betadisper(vegdist(df1, method = "euclidean"), df1b$Alliance)) -> betadis
 
 ### Export results
 
@@ -32,15 +31,13 @@ load(file = "results/permanova/pairwise.RData"); pairperma
 load(file = "results/permanova/permanova.RData"); perma
 load(file = "results/permanova/beta-dispersion.RData"); betadis
 
-
 ### https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html
 
 rm(list = ls())
 
 library(tidyverse); library(indicspecies); library(labdsv)
 
-read.csv("data/header.csv", fileEncoding = "latin1") %>%
-  filter(! Alliance %in% "Noise")  -> 
+read.csv("data/header.csv", fileEncoding = "latin1") -> 
   header
 
 read.csv("data/species.csv", fileEncoding = "latin1") %>% 
@@ -51,20 +48,139 @@ species %>%
   spread(Analysis.Names, Cover.percent, fill = 0) %>%
   column_to_rownames(var = "SIVIMID") -> df1
 
[…]

### 2023-12-20 src/6-characteristic-spp.R
permanova & characteristic

@@ -63,15 +63,14 @@ species %>%
 
 load(file = "results/characteristic-spp/indval.RData")
 
-header %>%
-  filter(!Alliance %in% "Noise") %>%
+read.csv("data/header.csv", fileEncoding = "latin1") %>%
   select(Alliance) %>%
   unique %>%
   mutate(Community = as.numeric(as.factor(Alliance))) -> communities
 
 data.frame(Community = indicators$maxcls, Indicator = indicators$indcls,
            p = indicators$pval, p_adj = p.adjust(indicators$pval, "holm")) %>%
-  rownames_to_column(var = "Species") %>%
+  rownames_to_column(var = "Species") %>% 
   filter(p_adj < 0.05) %>%
   merge(communities) %>%
   select(-c(Community, p)) %>%
@@ -127,11 +126,12 @@ species %>%
 
 species %>%
   merge(header, by = "SIVIMID") %>%
+  filter(Cover.percent > 25) %>%
   group_by(Alliance, Analysis.Names) %>%
   summarise(F = length(Analysis.Names)) %>%
   merge(ns) %>%
   mutate(Frequency = F/n) %>%
-  filter(Frequency >= .50) %>%
+  filter(Frequency >= .05) %>%
   mutate(Species = Analysis.Names,
          Type = "Dominant",
          Value = Frequency*100) %>%

### 2023-12-20 src/run.R

@@ -1,186 +1,0 @@
-### Test significance of classification
-
-rm(list = ls())
-
-library(tidyverse); library(vegan)
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>% filter(! Alliance %in% "Noise") ->
-  header
-
-read.csv("data/species.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% header$SIVIMID) -> species
-
-species %>%
-  select(SIVIMID, Analysis.Names, Cover.percent) %>%
-  spread(Analysis.Names, Cover.percent, fill = 0) %>%
-  column_to_rownames(var = "SIVIMID") -> df1
-
-### Permanova
-
-df1 %>% rownames_to_column(var = "SIVIMID") %>% merge(header, by = "SIVIMID") -> df1b
-RVAideMemoire::pairwise.perm.manova(dist(df1, "euclidian"), df1b$Alliance, nperm = 100000, p.method = "holm") -> pairperma
-adonis2(df1 ~ Alliance, data = df1b, permutations = 9999) -> perma
-anova(betadisper(vegdist(df1, method = "euclidean"), df1b$Alliance)) -> betadis
-
-### Export results
-
-save(pairperma, file = "results/permanova/pairwise.RData")
-save(perma, file = "results/permanova/permanova.RData")
-save(betadis, file = "results/permanova/beta-dispersion.RData")
-
-load(file = "results/permanova/pairwise.RData"); pairperma
-load(file = "results/permanova/permanova.RData"); perma
-load(file = "results/permanova/beta-dispersion.RData"); betadis
-
-### https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html
-
-rm(list = ls())
-
-library(tidyverse); library(indicspecies); library(labdsv)
-
-read.csv("data/header.csv", fileEncoding = "latin1") -> 
-  header
-
-read.csv("data/species.csv", fileEncoding = "latin1") %>% 
-  filter(SIVIMID %in% header$SIVIMID) -> species
-
-species %>%
-  select(SIVIMID, Analysis.Names, Cover.percent) %>%
-  spread(Analysis.Names, Cover.percent, fill = 0) %>%
-  column_to_rownames(var = "SIVIMID") -> df1
-
-### Phidelity (not working in R, extracted with JUICE)
-
[…]

### 2023-12-21 doc/manuscript.Rmd
results section

@@
 After initial exploration of this dataset using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379], we identified 652 outlier relevés which either did not belong to the target anthropogenic classes or to the geographical area of the ecoregion.
-We removed these, leaving a dataset of 2,508 relevés for further analysis: 2,419 relevés from SIVIM, originally recorded in 89 publications see (**Supplementary 2**); plus 89 relevés added by us.
+We removed these, leaving a dataset of 2,508 relevés for the analyses of this article: 2,419 relevés from SIVIM, originally recorded in 89 publications see (**Supplementary 2**); plus 89 relevés added by us.
 
-## Definition of target alliances and training dataset
+## Statistical analyses
 
-As a previous step before classification, we prepared a revised checklist of potential anthropogenic alliances that could be present in our study ecoregion, based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
-The provisional list included 38 alliances.
-
-Then, we filtered our dataset to include only those relevés which had been assigned by the authors to any of these 38 alliances (n = 2,201 relevés).
-We used sequential TWINSPANs to do a preliminary re-classification of the 2,201 filtered relevés, attempting to re-match the relevés into the *a priori* defined checklist of 38 alliances.
+[Definition of target alliances and training dataset.]{.underline} As a previous step before classification, we prepared a revised checklist of potential anthropogenic alliances that could be present in our study ecoregion, based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
+Then, we filtered our dataset to include only those relevés which had been assigned by the authors to any of these alliances, and used sequential TWINSPANs to do a preliminary re-classification of these filtered relevés, attempting to re-match the relevés into the *a priori* defined checklist of alliances.
 We used 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity, and the R package 'twinspanR' [@RN5380] to do the TWINSPANs.
 To start, we did a TWINSPAN classification into 15 clusters, followed by separate TWINSPANS within each cluster.
-This iterative process allowed us to validate 28 alliances and assign 1,725 relevés to them, leaving 783 relevés as unclassified.
+This preliminary TWINSPAN allowed us to (1) validate which alliances were present in the study ecoregion; and (2) prepare a training dataset for semi-supervised classification (see below).
 
-## Semi-supervised classification
-
-Next, we applied a semi-supervised classification to assign the so far unclassified relevés to one of the validated 28 alliances.
-Semi-supervised classification uses previously defined groups of data to assign the unclassified relevés to them (as in supervised classification) but also creates new groups to place the data points that do not match the already existing groups (as in unsupervised classification), allowing to detect new groups and to update the previous classification [@RN5383].
-We used the 1,719 relevés that we had already re-classified into the 28 alliances as a training set for the semi-supervised classification.
-Attempts by the classification to create new groups resulted on new groups with no ecological significance, and thus we kept the results of the classification into the 28 *a priori* alliances.
-For the classification, we applied a noise clustering fuzzy algorithm, which allows the relevés to be classified either into the *a priori* groups or into a 'noise' group which includes outliers and transitional relevés [@RN5382].
+[Semi-supervised classification.]{.underline} Next, we applied a semi-supervised classification to the whole dataset.
+Semi-supervised classification uses a training subset of *a priori* classified relevés to classify a secondary subset of unclassified relevés [@RN5383].
+Our training subset included all the relevés that (1) had been classified by the authors of the relevé into any of the alliances that we had validated during the previous TWINSPANs, and (2) had also been classified by us into these alliances during our TWINSPANs.
+The secondary subset included all the relevés which did not meet these two criteria.
+Our goal was to refine and not to simply extend the classification, so we applied the semi-supervised classification to the whole dataset, i.e.
+we allowed relevés from the training subset to be re-assigned to other groups during the classification.
+In addition, semi-supervised classification can create new groups to place data points that do not match the already *a priori* groups.
+Furthermore, we applied a noise clustering fuzzy algorithm, which allows the relevés to be classified either into the *a priori* groups or into a 'noise' group which includes outliers and transitional relevés [@RN5382].
 We set the fuzziness coefficient to a low value (m = 1) to accommodate a high number of transitional relevés; and we set the distance to the noise class to d = 1.
 We performed this analysis with the R package 'vegclust' [@RN5383].
-The semi-supervised classification with noise clustering resulted in the final classification of 2,081 relevés into 28 alliances and 427 relevés left in the noise group.
-We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package 'RVAideMemoire' [@RN5381] to detect the maximum number of mutually different clusters in the filtered dataset; and Detrended Correspondence Analysis (DCA) as implemented in the R package 'vegan' [@RN5386] to visualize the relationships between the vegetation groups.
+We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package 'RVAideMemoire' [@RN5381] to test the significant of the final vegetation groups; and Detrended Correspondence Analysis (DCA) as implemented in the R package 'vegan' [@RN5386] to visualize the relationships between the groups.
 
-## Characteristic species
+[Characteristic species and EUNIS habitat regionalization.]{.underline} We calculated sets of characteristic species [@RN5061] for each of the final vegetation groups.
+We defined dominant species as species with more than 25% cover in at least 5% of the relevés of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.05, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations.
+Additionally, we assigned to each group a regionalized level 4 EUNIS habitat code, using already codes when they existed, and creating new codes when not.
 
-We calculated sets of characteristic species [@RN5061] for each of the 28 alliances.
-We defined dominant species as species with more than 25% cover in at least 5% of the relevés of the alliance; constant species as species with a frequency higher than 50% in the alliance; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.05, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations.
-
-## Community traits and ecological requirements
-
-To describe the resulting 28 alliances, we compiled a dataset of species traits and ecological preferences for the species pool of the 2,508 classified relevés.
+[Community traits and ecological requirements.]{.underline} To describe the resulting vegetation groups, we compiled a dataset of species traits and ecological preferences for the species pool of the classified relevés.
 For each species, we extracted from *Flora iberica* [-@RN3243] its maximum height, its median month of flowering, the length of the flowering period, and whether it was a neophyte or not.
@@
 Then, for each relevé we calculated its community-weighted mean (weighting by species cover) for each trait and ecological preference, and explored the main axes of variation using PCA, separately for species traits and ecological preferences.
+
+# Results
+
+## Overview of the classification
+
+Our revision of the literature produced a provisional list of 38 anthropogenic vegetation alliances that could be present in the Iberian Atlantic ecoregion (**Supplementary 3**).
+In our study dataset (n = 2,508 relevés), there were 2,201 relevés that had been assigned by the authors to any of these alliances.
+We used these 2,201 relevés for a preliminary re-classification based on TWINSPAN, attempting to re-match the relevés into the *a priori* defined checklist of 38 alliances.
+This procedure allowed us to validate 28 alliances as present in the ecoregion.
+A total of 1,725 relevés had been classified by both the original authors and our TWINSPAN into these alliances, and these became our training subset for semi-supervised classification.
+Attempts by the semi-supervised classification to create new groups resulted on new groups with no ecological significance, and thus we kept the the 28 alliances as final vegetation groups (**Table 1**).
+Most of the alliances had a wide distribution in the ecoregion according to our dataset (**Figure 1**), although five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis*, *Paspalo-Agrostion semiverticillati*) had only isolated occurrences in a few points towards the southern limit of the Iberian Atlantic territories.
+The regional literature included in our dataset recognized 72 associations (**Supplementary 4**) within these alliances.
+The semi-supervised classification with noise clustering resulted in the final classification of 2,081 relevés into 28 alliances and 427 relevés left in the noise group.
+
+Focusing on the classified plots, the dataset included 1,167 taxa or taxa aggregates.
+Of these, 101 were neophytes (i.e.
+9% of the species pool).
+Regarding life forms, the species pool was dominated by hemicryptophytes (43%) and therophytes (41%).
+The 10 most frequent species were *Ochlopoa annua* (621 occurences), *Urtica dioica* (592), *Sonchus oleraceus* (540), *Stellaria media* (516), *Capsella bursa-pastoris* (390), *Polygonum aviculare* (375), *Dactylis glomerata* (363), *Senecio vulgaris* (332) and *Anisantha sterilis* (301).
+Considering only those relevés with the most frequent plot size (20 m^2^, n = 256 relevés), the average species richness per plot was 16 (minimum = 3, maximum = 37).
+
+DCA ordination of the floristic composition of the classified relevés (**Fig.
+2**) suggested a high consistency of the vegetation classes *Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Sisymbrietea* and *Bidentetea*.
+The alliances belonging to the two perennial ruderal vegetation classes (*Artemisietea vulgaris* and *Epilobietea angustifolii*) were mixed together.
+The winter-annual class *Chenopodietea* also showed high heterogeneity, specially because of the isolation of the altered dune alliance *Linario polygalifoliae-Vulpion alopecuri*.
+
+## Community-level traits and ecological preferences
+
+PCA ordination of the community-weighted means of the species traits (**Fig.
+3A**) indicated that the first two axes explained 65% of the variability.
+The major contributors to axis 1 (41% variance explained) were community height, proportion of geophytes, mean month of flowering and proportion of therophytes; it separated taller communities with more geophytes and later flowering (classes *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Bidentetea*) from shorter communities with more therophytes and earlier flowering (classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis* and *Sisymbrietea*).
+The major contributors to axis 2 (24% variance explained) were proportion of neophytes and width of the flowering season; it separated communities with more neophytes and longer flowering seasons (classes *Bidentetea*, *Digitario sanguinalis-Eragrostietea minoris* and *Polygono-Poetea annuae*) from communities with less neophytes and shorter flowering seasons (classes *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Chenopodietea*).
+
+PCA ordination of the community-weighted means of the species ecological and disturbance preferences (**Fig.
+3B**) indicated that the first two axes explained 56% of the variability.
+The major contributors to axis 1 (35% variance explained) were diturbance frequency, soil disturbance, mowing frequency and light requirements; it separated communities thriving with frequent disturbances in open situations (classes *Polygono-Poetea annuae*, *Sisymbrietea*, *Digitario-Eragrostietea* and *Papaveretea rhoeadis*) from communities prefering less frequent disturbances and more shade (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii* and *Bidentetea*).
+The major contributors to axis 2 (21% variance explained) were nutrient requirements and disturbance severity; it separated more nitrophilous communities adapted to more severe disturbances (classes *Digitario-Eragrostietea*, *Sisymbrietea* and *Epilobietea angustifolii*) from less nitrophilous communities adapted to lower intensity disturbances (classes *Cymbalario-Parietarietea diffusae*, *Artemisietea vulgaris* and *Chenopodietea*).
+
+## Classified anthropogenic plant communities
+
+***Cymbalario-Parietarietea diffusae*** Chasmophytic vegetation of human-made walls.
+
+1.
+*Galio valantiae-Parietarion judaicae* Vegetation of dry walls.
+Occupies warmer, drier and more nitriphicated situations than the other alliance in the class, has a shorter flowering season and is more adept to disturbances (**Fig.
+4**).
+Wide distrubution in the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Parietaria judaica* aggr., *Centranthus ruber*, *Asplenium trichomanes*.
+
+2.
+*Cymbalario-Asplenion* Vegetation of mesic walls.
+Occupies colder, wetter and less nitriphicated situations than the other alliance in the class, has a longer flowering season and is less adept to disturbances (**Fig.
+4**).
+Wide distrubution in the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Cymbalaria muralis*.
+
+***Polygono-Poetea annuae*** Dwarf vegetation of heavily-trampled sites.
+
+3.
+*Polycarpion tetraphylli* Dwarf-annual trampled vegetation of warm and sunny sites.
+Occupies warmer, sunnier and less nitriphicated situations than the other alliances in the class and has an earlier flowering season (**Fig.
+4**).
+Wide distrubution in the ecoregion, specially along the coast and in the more Mediterranean valleys of the interior (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Sagina apetala*, *Plantago coronopus*, *Crassula tillaea*, *Spergularia marina*.
+
+4.
+*Polygono-Coronopodion* Dwarf-annual trampled vegetation of dry sites.
+Occupies colder, drier and more nitriphicated situations than the other alliances in the class, and has a higher proportion of neophytes (**Fig.
+4**).
+Wide distrubution in the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Matricaria suaveolens*, *Polygonum aviculare*.
+
+5.
+*Saginion procumbentis* Dwarf trampled vegetation of mesic sites.
+Occupies colder, wetter and more nitriphicated situations than the other alliance in the class, and is less dominated by therophytes (**Fig.
+4**).
+Distribution concentrated in the central part of the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Sagina procumbens*, *Bryum argenteum*.
+
+***Papaveretea rhoeadis*** Annual segetal vegetation of crops and gardens.
+
+6.
+*Caucalidion lappulae* Annual segetal vegetation of cereal crops on base-rich soils.
+Adapted to more frequent and severe disturbances than the other alliances in the class  (**Fig.
+4**).
+Distribution mostly limited to the more Mediterranean valleys south of the Cantabrian Mountains (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Cyanus segetum*, *Papaver rhoeas*, *Trifolium arvense*, *Anacyclus clavatus*, *Valerianella eriocarpa*.
+
+7.
+*Scleranthion annui* Annual segetal vegetation of cereal crops on base-poor soils.
+Occupies more acidic soils than the other alliances in the class, and flowers earlier (**Fig.
+4**).
+Wide distribution in the ecoregion, but specially in the west, where acidic bedrocks dominate (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Mibora minima*, *Rumex acetosella*.
+
+8.
+*Oxalidion europeae* Annual segetal vegetation of gardens and root crops.
+Occupies wetter, shadier and more nitriphicated situations than the other alliances in the class,  has a longer flowering season and a higher proportion of neophytes (**Fig.
+4**).
+Wide distribution in the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Stellaria media*, *Veronica persica*.
+
+***Digitario sanguinalis-Eragrostietea minoris*** Summer-annual C4 segetal vegetation.
+
+9.
+*Spergulo arvensis-Erodion cicutariae* Summer-annual C4 segetal vegetation.
+Late-flowering alliance rich in neophytes and grasses, occurying in highly-distubed and nitriphicated soils (**Fig.
+4**).
+Wide distribution in the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Digitaria sanguinalis*, *Echinochloa crus-galli*, *Chenopodium album*, *Amaranthus hybridus* aggr.
+
+***Chenopodietea*** Winter-annual ruderal vegetation.
+
+10.
+*Allion triquetri* Geophyte-rich fringe vegetation.
+Occupies warmer and shadier situations than the other alliances in the class, has a higher proportion of geophytes and a lower proportion of therophytes, and is adapted to less frequent disturbances (**Fig.
+4**).
+Wide distribution in the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Urtica membranacea*, *Smyrnium olusatrum*, *Tradescantia fluminensis*.
+
+11.
+*Geranio pusilli-Anthriscion caucalidis* Winter-annual fringe vegetation.
+Occupies situations similar to the previous alliance, but in colder sites, and is dominated by therophytes instead of geophytes (**Fig.
+4**).
+Mostly distributed in the more Mediterranean valleys south of the Cantabrian Mountains (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Geranium lucidum*, *Anthriscus caucalis*.
+
+12.
+*Chenopodion muralis* Low-growth winter-annual ruderal vegetation.
+Occupies colder situations than the other alliances in the class and more often disturbed sites, and has a longer and later flowering season (**Fig.
+4**).
+Sparse occurencce throughout the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Malva neglecta*, *Urtica urens*.
+
+13.
+*Echio-Galactition tomentosae* Tall-herb winter-annual ruderal vegetation.
+Occupies drier, sunnier and less nutrient-rich situations than the other alliances in the class (**Fig.
+4**).
+Mostly distributed in the west of the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Galactites tomentosus*, *Coleostephus myconis*, *Anisantha rigida*.
+
+14.
+*Linario polygalifoliae-Vulpion alopecuri* Ephemeral annual vegetation of disturbed coastal dunes.
+Occupies drier and more nutrient-poor situations than the other alliances in the class (**Fig.
+4**) and has a high number of coastal dune species.
+Isolated occurrence in the southern coasts of the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Vulpia alopecuros*, *Malcolmia littorea*, *Crucianella maritima*, *Artemisia campestris* aggr.
+
+***Sisymbrietea*** Summer-annual ruderal vegetation.
+
+15.
+*Sisymbrion officinalis* Summer-annual ruderal vegetation.
+Rich in therophytes and adapted to severe disturbances (**Fig.
+4**).
+Wide distribution in the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Hordeum murinum*, *Sisymbrium officinale*.
+
+***Artemisietea vulgaris*** Perennial ruderal vegetation of dry sites.
+
+16.
+*Convolvulo arvensis-Agropyrion repentis* Semiruderal grasslands.
+Occupies wetter and shadier situations than the other alliances in the class, and has a higher proportion of geophytes and grasses (**Fig.
+4**).
+Isolated occurencce in the south of the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Poa compressa*, *Elytrigia repens*, *Potentilla reptans*, *Valerianella locusta*.
+
+17.
+*Carduo carpetani-Cirsion odontolepidis* Thistle ruderal vegetation of warm sites.
+Occupies warmer situations than the other thistle in the class (**Fig.
+4**).
+Mostly distributed in the Cantabrian Mountains (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Onopordum acanthium*.
+
+18.
+*Cirsion richterano-chodati* Thistle ruderal vegetation of cold sites.
+Occupies colder situations than the other thistle alliance in the class (**Fig.
+4**).
+Mostly distributed in the Cantabrian Mountains (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Cirsium eriophorum*, *Carduus carpetanus*, *Carduus nutans* aggr.
+
+19.
+*Dauco-Melilotion* Biennial ruderal vegetation.
+Occupies wetter situations than the two thistle alliances in the class, and has a higher proportion of therophytes (**Fig.
+4**).
+Sparse occurencce throughout the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Helminthotheca echioides*, *Daucus carota*, *Melilotus albus*, *Foeniculum vulgare*.
+
+***Epilobietea angustifolii*** Perennial ruderal vegetation of mesic to wet sites.
+
+20.
+*Geo urbani-Alliarion officinalis* Low-herb short-lived semiruderal and fringe vegetation.
+Occupies shadier and less disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants (**Fig.
+4**).
+Wide distribution in the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Urtica dioica*.
+
+21.
+*Arction lappae* Low-herb short-lived ruderal vegetation.
+Occupies sunnier and more frequently-disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants (**Fig.
+4**).
+Distribution concentrated in the Cantabrian Mountains (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Oxybasis rubra*, *Senecio duriaei*.
+
+22.
+*Balloto-Conion maculati* Tall-herb perennial ruderal vegetation.
+Occupies warmer situations than the other alliances in the class (**Fig.
+4**).
+Distribution concentrated in the Cantabrian Mountains and the east of the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Sambucus ebulus*.
+
+23.
+*Aegopodion podagrariae* Tall-herb perennial vegetation of forest margins and clearings.
+Occupies shadier and less disturbed situations than the other alliances in the class (**Fig.
+4**).
+Sparse occurencce throughout the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Anthriscus sylvestris*, *Heracleum sphondylium*, *Galium aparine*.
+
+24.
+*Epilobion angustifolii* Tall-herb perennial vegetation of forest margins and clearings in acidic soils.
+Occupies colder, nutrient-poorer, more acidic and less disturbed situations than the other alliances in the class (**Fig.
+4**).
+Distributed towards the west of the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Epilobium angustifolium*, *Digitalis purpurea*, *Luzula lactea*.
+
+25.
+*Cynancho-Convolvulion sepium* Tall-herb vegetation of nutrient-rich riparian habitats.
+Occupies wetter situations and less disturbed situations than the other alliances in the class, and has a later flowering season (**Fig.
+4**).
+Dispersed throughout the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Eupatorium cannabinum*, *Angelica sylvestris*, *Picris hieracioides*.
+
+26.
+*Senecionion fluviatilis* Tall-herb vegetation of nutrient-rich river banks and ditches.
+Occupies warmer and wetter situations situations than the other alliances in the class, flowers later, has taller plants and is highly dominated by neophytes and geophytes (**Fig.
+4**).
+Isolated occurence in the western coast of the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Arundo donax*, *Silene latifolia*.
+
+***Bidentetea*** Summer-annual pioneer vegetation of temporarily-flooded sites.
+
+27.
+*Bidention tripartitae* Summer-annual pioneer vegetation of temporarily-flooded sites.
+Occupies nutrient-richer sites than the other alliance in the class and has a higher proportion of therophytes (**Fig.
+4**).
+Dispersed throughout the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Persicaria hydropiper*, *Bidens frondosus*, *Lythrum salicaria*.
+
+28.
+*Paspalo-Agrostion semiverticillati* Summer-annual pioneer vegetation of temporarily-flooded and warm sites.
+Occupies warmer and wetter situations than the other alliance in the class, has a higher proportion of neophytes and has a longer flowering season (**Fig.
+4**).
+Isolated occurrences in the south of the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Paspalum distichum*, *Cyperus longus*, *Schoenoplectus lacustris*.
 
@@
 
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "."}
+knitr::include_graphics("../results/figures/maps.png")
+```
+
 ```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "."}
@@
 ```
-
-```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "."}
-knitr::include_graphics("../results/figures/maps.png")
-```

### 2023-12-21 src/11-spp-numbers.R

@@ -1,0 +1,88 @@
+library(tidyverse)
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise") %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+  group_by(Analysis.Names)
+
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise") %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+  group_by(Analysis.Names) %>% tally %>% arrange(-n)
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise") %>%
+  group_by(Area) %>% tally %>% arrange(-n)
+
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise" & Area == "20") %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+  select(SIVIMID, Analysis.Names) %>%
+  unique %>%
+  group_by(SIVIMID) %>% tally %>% arrange(-n)
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise" & Area == "20") %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+  select(SIVIMID, Analysis.Names) %>%
+  unique %>%
+  group_by(SIVIMID) %>% tally %>% arrange(n)
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise" & Area == "20") %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+  select(SIVIMID, Analysis.Names) %>%
+  unique %>%
+  group_by(SIVIMID) %>% tally %>% pull(n) %>% mean
+
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise") %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+  select(SIVIMID, Analysis.Names) %>%
+  unique %>%
+  group_by(SIVIMID) %>% tally %>% arrange(-n)
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
[…]

### 2023-12-21 src/8-cwm-pca.R

@@ -4,45 +4,45 @@ library(tidyverse)
 
 ### Reciprocal averaging
 
-read.csv("data/traits.csv", fileEncoding = "latin1") %>%
-  select(Analysis.Names,
-         EIVE1.M:Soil.Disturbance) %>%
-  gather(Trait, Value, -Analysis.Names) %>%
-  na.omit %>%
-  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  group_by(SIVIMID, Trait) %>%
-  summarise(Value = weighted.mean(Value, Cover.percent)) -> plotmeans
-
-read.csv("data/species.csv", fileEncoding = "latin1") %>%
-  merge(plotmeans) %>%
-  group_by(Analysis.Names, Trait) %>%
-  summarise(Value = weighted.mean(Value, Cover.percent)) %>%
-  rename(ravalue = Value) -> sppmeans
-
-read.csv("data/traits.csv", fileEncoding = "latin1") %>%
-  select(Analysis.Names,
-         EIVE1.M:Soil.Disturbance) %>%
-  gather(Trait, Value, -Analysis.Names) %>%
-  merge(sppmeans, all.x = TRUE) %>%
-  mutate(value = ifelse(is.na(Value), ravalue, Value)) %>%
-  select(-c(Value, ravalue)) %>%
-  spread(Trait, value) -> raspp
-
-read.csv("data/traits.csv", fileEncoding = "latin1") %>%
-  select(-c(EIVE1.M:Soil.Disturbance)) %>%
-  merge(raspp) -> traits
-
-### CwMs
-
-traits %>%
-  gather(Trait, Value, -Analysis.Names) %>%
-  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  group_by(SIVIMID, Trait) %>%
-  summarise(Value = weighted.mean(Value, Cover.percent)) %>%
-  spread(Trait, Value) %>%
-  merge(read.csv("data/header.csv", fileEncoding = "latin1")) -> cwms
-
-save(cwms, file = "results/cwms/cwms.RData")
+# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
+#   select(Analysis.Names,
+#          EIVE1.M:Soil.Disturbance) %>%
+#   gather(Trait, Value, -Analysis.Names) %>%
+#   na.omit %>%
[…]

### 2024-01-16 doc/manuscript.Rmd
intro

@@
 title: "Anthropogenic plant communities of the Iberian Atlantic ecoregion"
-output: github_document
+output: 
+  bookdown::word_document2:
+    reference_docx: template.docx
+    number_sections: false
 bibliography: refs.bib
@@
 ```
+
+# Introduction
+
+Anthropogenic vegetation is the set of plant communities that occur as a direct consequence of human activities, and is chiefly composed by the weed vegetation of arable fields and the ruderal vegetation of human settlements and their surroundings [@RN5441].
+The formation of anthropogenic vegetation in temperate Europe has been characterized by a constant enrichment produced by the arrival of alien species: first archaeophytes starting in the Neolithic and peaking in the Bronze Age; and then neophytes from the Modern Age to the present day [@RN5439].
+Synanthropic archaeophytes include many Mediterranean taxa that started their expansion into temperate areas during the Mid-Holocene warm period, but it is difficult to disentangle to what extent this expansion was favored by human activities and/or ongoing climatic changes [@RN5432; @RN5443], especially in southern European areas close to the transition between the Mediterranean and the temperate climates [@RN5444].
+Meanwhile, synanthropic neophytes include many American, African and Asian taxa that arrived as a consequence of the global trade networks established by the Western European colonial powers [@RN5451].
+In the last century, anthropogenic vegetation has shown a decrease in species richness and diversity [@RN5423], as rare synanthropic species have become rarer, and neophytes have increased their abundance at the expense of natives and archaeophytes [@RN5441].
+
+The anthropogenic communities of human-made habitats make up an important part of the supply of natural elements in urban and peri-urban habitats [@RN5445].
+In densely-populated city landscapes, the plant communities that colonize urban vacant lots [@RN2300] are a useful biodiversity resource with high ecological and societal potential [@RN5446].
+Anthropogenic vegetation cover can also impact regional-level extreme climatic events [@RN5440].
+Furthermore, many synanthropic plants can accumulate heavy metals in their biomass [@RN5433], making them valuable candidates for nature-based solutions such as bioremediation and restoration of degraded landscapes [@RN5447].
+
+The Iberian Atlantic ecoregion has had a long history of human habitation dating back to the Cantabrian Upper Paleolithic [@RN5450].
+This, together with its transitional position at the border between the temperate and Mediterranean climatic zones of Europe [@RN4700], suggests a long regional history for archaeophyte-rich anthropogenic plant communities, and supports a current high diversity of anthropogenic plant communities combining temperate and Mediterranean floristic elements [@RN5008].
+Historical trade links with America and Asia, together with the warm temperate climate, have made the region a hotspot for plant invasions by neophytes [@Checklist].
+Recent processes of postindustrial land-use change have left large areas of brownfields whose management and restoration requires an understanding of anthropogenic communities able to colonize or restore them [@RN5448; @RN5449].
+Although there is a long tradition of studying weed and ruderal communities in the ecoregion [@RN5364; @RN5361; @RN5359; @RN5360] there is a current need for a regional-level synthesis that revises the classification and updates it in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
+In this article we have performed such a synthesis with two main objectives: (1) provide an updated classification of anthropogenic plant communities in the Iberian Atlantic ecoregion, and (2) describe the main patterns of variation in terms of species traits and ecological and disturbance preferences.
 
@@
 We defined dominant species as species with more than 25% cover in at least 5% of the relevés of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.05, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations.
-Additionally, we assigned to each group a regionalized level 4 EUNIS habitat code, using already codes when they existed, and creating new codes when not.
+Additionally, we assigned to each group a regionalized level 4 EUNIS habitat code, using pre-existing codes when they existed, and creating new codes when not.
 
@@
 A total of 1,725 relevés had been classified by both the original authors and our TWINSPAN into these alliances, and these became our training subset for semi-supervised classification.
-Attempts by the semi-supervised classification to create new groups resulted on new groups with no ecological significance, and thus we kept the the 28 alliances as final vegetation groups (**Table 1**).
+Attempts by the semi-supervised classification to create new groups resulted on new groups with no ecological significance, and thus we kept the 28 alliances as final vegetation groups (**Table 1**).
 Most of the alliances had a wide distribution in the ecoregion according to our dataset (**Figure 1**), although five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis*, *Paspalo-Agrostion semiverticillati*) had only isolated occurrences in a few points towards the southern limit of the Iberian Atlantic territories.
@@
 Regarding life forms, the species pool was dominated by hemicryptophytes (43%) and therophytes (41%).
-The 10 most frequent species were *Ochlopoa annua* (621 occurences), *Urtica dioica* (592), *Sonchus oleraceus* (540), *Stellaria media* (516), *Capsella bursa-pastoris* (390), *Polygonum aviculare* (375), *Dactylis glomerata* (363), *Senecio vulgaris* (332) and *Anisantha sterilis* (301).
+The 10 most frequent species were *Ochlopoa annua* (621 occurrences), *Urtica dioica* (592), *Sonchus oleraceus* (540), *Stellaria media* (516), *Capsella bursa-pastoris* (390), *Polygonum aviculare* (375), *Dactylis glomerata* (363), *Senecio vulgaris* (332) and *Anisantha sterilis* (301).
 Considering only those relevés with the most frequent plot size (20 m^2^, n = 256 relevés), the average species richness per plot was 16 (minimum = 3, maximum = 37).
@@
 3B**) indicated that the first two axes explained 56% of the variability.
-The major contributors to axis 1 (35% variance explained) were diturbance frequency, soil disturbance, mowing frequency and light requirements; it separated communities thriving with frequent disturbances in open situations (classes *Polygono-Poetea annuae*, *Sisymbrietea*, *Digitario-Eragrostietea* and *Papaveretea rhoeadis*) from communities prefering less frequent disturbances and more shade (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii* and *Bidentetea*).
+The major contributors to axis 1 (35% variance explained) were disturbance frequency, soil disturbance, mowing frequency and light requirements; it separated communities thriving with frequent disturbances in open situations (classes *Polygono-Poetea annuae*, *Sisymbrietea*, *Digitario-Eragrostietea* and *Papaveretea rhoeadis*) from communities preferring less frequent disturbances and more shade (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii* and *Bidentetea*).
 The major contributors to axis 2 (21% variance explained) were nutrient requirements and disturbance severity; it separated more nitrophilous communities adapted to more severe disturbances (classes *Digitario-Eragrostietea*, *Sisymbrietea* and *Epilobietea angustifolii*) from less nitrophilous communities adapted to lower intensity disturbances (classes *Cymbalario-Parietarietea diffusae*, *Artemisietea vulgaris* and *Chenopodietea*).
@@
 
-***Cymbalario-Parietarietea diffusae*** Chasmophytic vegetation of human-made walls.
+***Cymbalario-Parietarietea diffusae*** Vegetation of human-made walls.
 
@@
 *Galio valantiae-Parietarion judaicae* Vegetation of dry walls.
-Occupies warmer, drier and more nitriphicated situations than the other alliance in the class, has a shorter flowering season and is more adept to disturbances (**Fig.
+Occupies warmer, drier and more nutrient-rich situations than the other alliance in the class, has a shorter flowering season and is more adapted to disturbances (**Fig.
 4**).
-Wide distrubution in the ecoregion (**Fig.
+Wide distribution in the ecoregion (**Fig.
 1**).
@@
 *Cymbalario-Asplenion* Vegetation of mesic walls.
-Occupies colder, wetter and less nitriphicated situations than the other alliance in the class, has a longer flowering season and is less adept to disturbances (**Fig.
+Occupies colder, wetter and less nutrient-rich situations than the other alliance in the class, has a longer flowering season and is less adapted to disturbances (**Fig.
 4**).
-Wide distrubution in the ecoregion (**Fig.
+Wide distribution in the ecoregion (**Fig.
 1**).
@@
 *Polycarpion tetraphylli* Dwarf-annual trampled vegetation of warm and sunny sites.
-Occupies warmer, sunnier and less nitriphicated situations than the other alliances in the class and has an earlier flowering season (**Fig.
+Occupies warmer, sunnier and less nutrient-rich situations than the other alliances in the class and has an earlier flowering season (**Fig.
 4**).
-Wide distrubution in the ecoregion, specially along the coast and in the more Mediterranean valleys of the interior (**Fig.
+Wide distribution in the ecoregion, specially along the coast and in the more Mediterranean valleys of the interior (**Fig.
 1**).
@@
 *Polygono-Coronopodion* Dwarf-annual trampled vegetation of dry sites.
-Occupies colder, drier and more nitriphicated situations than the other alliances in the class, and has a higher proportion of neophytes (**Fig.
+Occupies colder, drier and more nutrient-rich situations than the other alliances in the class, and has a higher proportion of neophytes (**Fig.
 4**).
-Wide distrubution in the ecoregion (**Fig.
+Wide distribution in the ecoregion (**Fig.
 1**).
@@
 *Saginion procumbentis* Dwarf trampled vegetation of mesic sites.
-Occupies colder, wetter and more nitriphicated situations than the other alliance in the class, and is less dominated by therophytes (**Fig.
+Occupies colder, wetter and more nutrient-rich situations than the other alliance in the class, and is less dominated by therophytes (**Fig.
 4**).
@@
 
-***Papaveretea rhoeadis*** Annual segetal vegetation of crops and gardens.
+***Papaveretea rhoeadis*** Annual weed vegetation of crops and gardens.
 
 6.
-*Caucalidion lappulae* Annual segetal vegetation of cereal crops on base-rich soils.
+*Caucalidion lappulae* Annual weed vegetation of cereal crops on base-rich soils.
 Adapted to more frequent and severe disturbances than the other alliances in the class  (**Fig.
@@
 7.
-*Scleranthion annui* Annual segetal vegetation of cereal crops on base-poor soils.
+*Scleranthion annui* Annual weed vegetation of cereal crops on base-poor soils.
 Occupies more acidic soils than the other alliances in the class, and flowers earlier (**Fig.
@@
 8.
-*Oxalidion europeae* Annual segetal vegetation of gardens and root crops.
-Occupies wetter, shadier and more nitriphicated situations than the other alliances in the class,  has a longer flowering season and a higher proportion of neophytes (**Fig.
+*Oxalidion europeae* Annual weed vegetation of gardens and root crops.
+Occupies wetter, shadier and more nutrient-rich situations than the other alliances in the class, has a longer flowering season and a higher proportion of neophytes (**Fig.
 4**).
@@
 
-***Digitario sanguinalis-Eragrostietea minoris*** Summer-annual C4 segetal vegetation.
+***Digitario sanguinalis-Eragrostietea minoris*** Summer-annual C4 weed vegetation.
 
 9.
-*Spergulo arvensis-Erodion cicutariae* Summer-annual C4 segetal vegetation.
-Late-flowering alliance rich in neophytes and grasses, occurying in highly-distubed and nitriphicated soils (**Fig.
+*Spergulo arvensis-Erodion cicutariae* Summer-annual C4 weed vegetation.
+Late-flowering alliance rich in neophytes and grasses, occurring in highly-disturbed and nutrient-rich soils (**Fig.
 4**).
@@
 4**).
-Sparse occurencce throughout the ecoregion (**Fig.
+Sparse occurrence throughout the ecoregion (**Fig.
 1**).
@@
 4**).
-Isolated occurencce in the south of the ecoregion (**Fig.
+Isolated occurrence in the south of the ecoregion (**Fig.
 1**).
@@
 *Carduo carpetani-Cirsion odontolepidis* Thistle ruderal vegetation of warm sites.
-Occupies warmer situations than the other thistle in the class (**Fig.
+Occupies warmer situations than the other thistle alliance in the class (**Fig.
 4**).
@@
 *Senecionion fluviatilis* Tall-herb vegetation of nutrient-rich river banks and ditches.
-Occupies warmer and wetter situations situations than the other alliances in the class, flowers later, has taller plants and is highly dominated by neophytes and geophytes (**Fig.
+Occupies warmer and wetter situations than the other alliances in the class, flowers later, has taller plants and is highly dominated by neophytes and geophytes (**Fig.
 4**).
-Isolated occurence in the western coast of the ecoregion (**Fig.
+Isolated occurrence in the western coast of the ecoregion (**Fig.
 1**).
@@
 
+# Discussion
+
+The major driver of variation in trampled community composition is soil moisture [@RN5434].
+
+In medieval Switzerland, summer and winter weeds communities could not be distinguished [@RN5431].
+
+The class *Digitario sanguinalis-Eragrostietea minoris* includes thermophilic vegetation dominated by C4 plants that occupies arable land as well as trampled and ruderal sites [@RN5426].
+
+Anthropogenic vegetation composition is sensitive to micro-scale management decisions, such as the ownership of domestic dogs [@RN5438].
+
+Neophytes were more represented in early successional anthropogenic communities [@RN5423].
+
+Arable land plants were mainly therophytes and alien species; while ruderal habitats were richer in perennials and C strategists [@RN5436].
+
+Compare proportion of neophytes to other areas [@RN5429; @@RN5435]
+
+Neophytes have no special preferences for local habitat conditions and their highest proportion was found mainly in disturbed habitats at low elevations. [@RN5435]
+
+Therophytes and hemicryptophytes were most abundant in the life form spectra [@RN5428]
+
 # Data availability
@@
 
-```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "."}
+```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Figure 1.
+Distribution of the anthropogenic alliances in the Iberian Atlantic ecoregion.
+Each dot represents a relevé.
+Dots are colored by vegetation classes."}
 knitr::include_graphics("../results/figures/maps.png")
@@
 
-```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "."}
+```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Figure 2.
+Ordination of the floristic composition of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+The ordination is the results of a Detrended Correspondence Analysis grouped by vegetation classes (A, left) and alliances (B, right)."}
 knitr::include_graphics("../results/figures/dca.png")
@@
 
-```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "."}
+```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Figure 3.
+Ordination of the community-level species traits and ecological preferences.
+The ordinations are the results of a Principal Component Analysis calculated for the community-weighted means for species traits (A, left) and ecological and disturbance indicator values (B, right)."}
 knitr::include_graphics("../results/figures/cwms.png")
@@
 
-```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "."}
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
+Community-level species traits and ecological preferences by alliance.
+Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
 knitr::include_graphics("../results/figures/boxplots.png")

### 2024-04-10 doc/manuscript.Rmd
revised figures

@@
 Anthropogenic vegetation is the set of plant communities that occur as a direct consequence of human activities, and is chiefly composed by the weed vegetation of arable fields and the ruderal vegetation of human settlements and their surroundings [@RN5441].
-The formation of anthropogenic vegetation in temperate Europe has been characterized by a constant enrichment produced by the arrival of alien species: first archaeophytes starting in the Neolithic and peaking in the Bronze Age; and then neophytes from the Modern Age to the present day [@RN5439].
-Synanthropic archaeophytes include many Mediterranean taxa that started their expansion into temperate areas during the Mid-Holocene warm period, but it is difficult to disentangle to what extent this expansion was favored by human activities and/or ongoing climatic changes [@RN5432; @RN5443], especially in southern European areas close to the transition between the Mediterranean and the temperate climates [@RN5444].
-Meanwhile, synanthropic neophytes include many American, African and Asian taxa that arrived as a consequence of the global trade networks established by the Western European colonial powers [@RN5451].
-In the last century, anthropogenic vegetation has shown a decrease in species richness and diversity [@RN5423], as rare synanthropic species have become rarer, and neophytes have increased their abundance at the expense of natives and archaeophytes [@RN5441].
-
 The anthropogenic communities of human-made habitats make up an important part of the supply of natural elements in urban and peri-urban habitats [@RN5445].
@@
 
-The Iberian Atlantic ecoregion has had a long history of human habitation dating back to the Cantabrian Upper Paleolithic [@RN5450].
-This, together with its transitional position at the border between the temperate and Mediterranean climatic zones of Europe [@RN4700], suggests a long regional history for archaeophyte-rich anthropogenic plant communities, and supports a current high diversity of anthropogenic plant communities combining temperate and Mediterranean floristic elements [@RN5008].
-Historical trade links with America and Asia, together with the warm temperate climate, have made the region a hotspot for plant invasions by neophytes [@Checklist].
-Recent processes of postindustrial land-use change have left large areas of brownfields whose management and restoration requires an understanding of anthropogenic communities able to colonize or restore them [@RN5448; @RN5449].
-Although there is a long tradition of studying weed and ruderal communities in the ecoregion [@RN5364; @RN5361; @RN5359; @RN5360] there is a current need for a regional-level synthesis that revises the classification and updates it in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
-In this article we have performed such a synthesis with two main objectives: (1) provide an updated classification of anthropogenic plant communities in the Iberian Atlantic ecoregion, and (2) describe the main patterns of variation in terms of species traits and ecological and disturbance preferences.
+The anthropogenic vegetation of temperate Europe is formed by an original pool of native species favored by humans, but this pool has been constantly enriched by the arrival of alien species: first archaeophytes starting in the Neolithic and peaking in the Bronze Age; and then neophytes from the Modern Age (after 1.500 A.D.) to the present day [@RN5439].
+Synanthropic archaeophytes include many Mediterranean taxa that started their expansion into temperate areas during the Mid-Holocene warm period, but it is difficult to disentangle to what extent this expansion was favored by human activities and/or ongoing climatic changes [@RN5432; @RN5443], especially in southern European areas close to the transition between the Mediterranean and the temperate climates [@RN5444].
+Meanwhile, synanthropic neophytes include many American, African and Asian taxa that arrived as a consequence of the global trade networks established by the Western European colonial powers [@RN5451].
+In the last century, anthropogenic vegetation has shown a decrease in species richness and diversity [@RN5423], as rare synanthropic species have become rarer, and neophytes have increased their abundance at the expense of natives and archaeophytes [@RN5441].
+
+Antropogenic vegetation is likely to be similar within ecological regions (ecoregions, hereafter), since they are assumed to encompass areas with a similar biogeographic history and recurrent local ecosystems [@RN5491].
+However, ecoregional-scale studies on the diversity of antropogenic vegetation are scarce in the literature.
+The Iberian Atlantic ecoregion (a.k.a.
+Cantabrian Mixed Forests ecoregion; NW Iberian Peninsula) has a long history of human habitation dating back to the Cantabrian Upper Paleolithic [@RN5450].
+This, together with its transitional position at the border between the temperate and Mediterranean climatic zones of Europe [@RN4700], suggests a long regional history for archaeophyte-rich anthropogenic plant communities, likely supporting a current high diversity of anthropogenic plant communities with both temperate and Mediterranean floristic elements [@RN5008].
+Historical trade links with America and Asia, together with the warm and humid temperate climate, have made the region a hotspot for neophyte invasions [@RN5492; @RN5493].
+Recent post-industrial land-use changes have left large areas with abandoned industrial sites (i.e.
+brownfields) whose management and restoration requires an understanding of anthropogenic communities able to colonize or restore them [@RN5448; @RN5449].
+Although there is a long tradition of studying weed and ruderal communities in the ecoregion [@RN5364; @RN5361; @RN5359; @RN5360] there is a current need for an ecoregional-level synthesis that revises and updates the classification and updates in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
+In this article we have performed such a synthesis with two main objectives: (1) provide an updated classification of anthropogenic plant communities in the Iberian Atlantic ecoregion, and (2) describe the main patterns of vegetation diversity in terms of floristic origin (natives, archaeophytes, neophytes); life forms; functional traits (plant height and flowering time); and ecological indicator values (temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity).
 
 # Methods
-
-We performed all data analyses with R version 4.3.1 [@RN5387], using the R package 'tidyverse' [@RN4662] for data processing and visualization.
-The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
 
@@
 
-As vegetation data for our analyses, we obtained all vegetation relevés (i.e.
-records of plant species coexisting in a given sampling plot at a given date) from our study ecoregion that were available in SIVIM, the Iberian and Macaronesian Vegetation Information System [@RN5374].
-These were 28,775 relevés representing all vegetation types (i.e.
-anthropogenic, semi-natural and natural).
-Since these relevés were sampled by different authors throughout the decades, there was no taxonomical consistency among them, so we homogenized all taxon names for the whole dataset.
-We followed the methodology applied by Jiménez-Alfaro *et al.* [-@RN5375] to the Cantabrian Mountains, an area within the study ecoregion which includes c.
-80% of the ecoregional flora, to create a new taxonomical checklist for the ecoregion based on the Euro+Med Plantbase [@RN5044].
-For vascular plants not included in Euro+Med (e.g.
-several *Quercus* or *Salix* hybrids, non-native species, etc.), as well as for bryophytes, we followed Plants of the World Online [@RN5376].
-We removed Fungi and algae.
-For analytical purposes, we merged taxa into aggregates (1) when there were sets of very close species with a difficult identification in the field (e.g.
-*Viola riviniana* + *V.
-reichenbachiana* + *V.
-sylvestris* = *Viola riviniana* aggr.), (2) when the taxon was recently separated in different taxa (e.g.
-*Hedera helix* + *Hedera hibernica* + *Hedera canariensis* = *Hedera helix* aggr.) and (3) when it was assumed that not all authors were able to identify or recognize infraspecific taxa (e.g.
-all *Narcissus pseudonarcissus* subspecies and varieties = *Narcissus pseudonarcissus* aggr.).
-
-For mapping purposes, we improved the geographical information of the SIVIM relevés using a downscaling procedure.
-The original position of vegetation relevés in SIVIM was recorded as 10 km x 10 km (52.2%) and 1 km x 1 km (47.8%) grid cells.
-We assigned more precise geographic coordinates to these plots using the elevation recorded originally by the authors and a digital elevation model (DEM) of the Iberian Peninsula at 200 m grid resolution.
-From the original 10 km x 10 km or 1 km x 1 km grid cell of each relevé, we (1) kept only those 200 m cells that agreed with the relevé's elevation (± 50 m); (2) randomly selected one of these 200 m cells; and (3) assigned to the relevé the coordinate of the 200 m cell's centroid, in decimal degrees.
-By doing this, the plots were downscaled at higher resolution within their cells, assuming that 200 m cells at similar elevations within the same grid cell also have similar climatic conditions.
-
-To identify anthropogenic relevés in the SIVIM database we followed two complementary approaches.
-Fist, we used the expert system created to classify vegetation plots into EUNIS pan-European habitat classification types [@RN5061].
-The expert system uses the floristic composition of plots to assign them to habitat types depending on the presence and abundance of characteristic species.
-We used the habitats assigned by the expert system to retrieve plots that had been assigned to any habitat related to the vegetation classes we had previously defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation).
-Second, we also retrieved relevés if they had been classified by the original authors of the relevé into any sintaxa belonging to our target anthropogenic vegetation classes (**Supplementary 1**).
-This produced an initial pool of 3,160 relevés from SIVIM, to which we added 89 relevés of urban or peri-urban plant communities conducted by us or extracted by us from the local literature [@RN5377; @RN5378].
-After initial exploration of this dataset using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379], we identified 652 outlier relevés which either did not belong to the target anthropogenic classes or to the geographical area of the ecoregion.
-We removed these, leaving a dataset of 2,508 relevés for the analyses of this article: 2,419 relevés from SIVIM, originally recorded in 89 publications see (**Supplementary 2**); plus 89 relevés added by us.
+We retrieved anthropogenic vegetation plots from SIVIM, the Iberian and Macaronesian Vegetation Information System [@RN5374].
+To identify plots as anthropogenic we followed two criteria: (1) We used the expert system created by Chytrý *et al.* [-@RN5061] to classify vegetation plots into EUNIS pan-European habitat types.
+This expert system uses the floristic composition of plots to assign them to habitat types depending on the presence and abundance of characteristic species.
+We used the habitats assigned by the expert system to retrieve plots that had been assigned to any habitat related to the vegetation classes we had previously defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation) (R55 and R57 include communities that are classified as "grasslands" by EUNIS but as "anthropogenic" by by Mucina *et al.* [-@RN4698]).
+(2) We also retrieved vegetation plots from SIVIM if they had been classified by the original authors of the relevé into any sintaxa belonging to our target anthropogenic vegetation classes (**Supplementary 1**).
+This produced an initial pool of 3,160 vegetation plots from SIVIM, to which we added 89 vegetation plots of urban or peri-urban plant communities sampled by us or extracted from local literature not in SIVIM [@RN5377; @RN5378].
+After initial exploration of this dataset using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379], we identified 652 outlier plots which either did not belong to the target anthropogenic classes or to the geographical area of the ecoregion.
+The majority of these plots corresponded to coastal plant communities and had been misclassified by the expert system.
+We removed these, leaving a dataset of 2,508 vegetation  plots for data analyses: 2,419 plots from SIVIM, originally recorded in 89 publications (**Supplementary 2**); plus 89 plots added by us.
 
@@
 
+We performed all data analyses with R version 4.3.1 [@RN5387], using the R package 'tidyverse' [@RN4662] for data processing and visualization.
+The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
+We homogenized all taxon names in the analyses and the article using Euro+Med [@RN5044] or Plants of the World Online for taxa not in Euro+Med [@RN5376].
+
 [Definition of target alliances and training dataset.]{.underline} As a previous step before classification, we prepared a revised checklist of potential anthropogenic alliances that could be present in our study ecoregion, based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
-Then, we filtered our dataset to include only those relevés which had been assigned by the authors to any of these alliances, and used sequential TWINSPANs to do a preliminary re-classification of these filtered relevés, attempting to re-match the relevés into the *a priori* defined checklist of alliances.
+Then, we filtered our dataset to extract a subset with the vegetation plots which had been assigned by the authors to any of these alliances.
+We used TWINSPAN to conduct a preliminary numerical classification of this subset, attempting to match the plots into the *a priori* defined checklist of alliances.
 We used 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity, and the R package 'twinspanR' [@RN5380] to do the TWINSPANs.
@@
 [Semi-supervised classification.]{.underline} Next, we applied a semi-supervised classification to the whole dataset.
-Semi-supervised classification uses a training subset of *a priori* classified relevés to classify a secondary subset of unclassified relevés [@RN5383].
-Our training subset included all the relevés that (1) had been classified by the authors of the relevé into any of the alliances that we had validated during the previous TWINSPANs, and (2) had also been classified by us into these alliances during our TWINSPANs.
-The secondary subset included all the relevés which did not meet these two criteria.
+Semi-supervised classification uses a training subset of *a priori* classified vegetation plots to classify a secondary subset of unclassified plots [@RN5383].
+Our training subset included all the plots that (1) had been classified by the authors of the relevé into any of the alliances that we had validated during the previous TWINSPANs, and (2) had also been matched by us into these alliances during our TWINSPANs.
+The secondary subset included all the plots which did not meet these two criteria.
 Our goal was to refine and not to simply extend the classification, so we applied the semi-supervised classification to the whole dataset, i.e.
-we allowed relevés from the training subset to be re-assigned to other groups during the classification.
-In addition, semi-supervised classification can create new groups to place data points that do not match the already *a priori* groups.
-Furthermore, we applied a noise clustering fuzzy algorithm, which allows the relevés to be classified either into the *a priori* groups or into a 'noise' group which includes outliers and transitional relevés [@RN5382].
-We set the fuzziness coefficient to a low value (m = 1) to accommodate a high number of transitional relevés; and we set the distance to the noise class to d = 1.
+we allowed plots from the training subset to be re-assigned to other groups during the classification.
+In addition, semi-supervised classification can create new groups to place data points that do not match the already existing *a priori* groups.
+Furthermore, we applied a noise clustering fuzzy algorithm, which allows the plots to be classified either into the *a priori* groups or into a 'noise' group which includes outliers and transitional plots [@RN5382].
+We set the fuzziness coefficient to a low value (m = 1) to accommodate a high number of transitional plots; and we set the distance to the noise class to d = 1.
 We performed this analysis with the R package 'vegclust' [@RN5383].
-We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package 'RVAideMemoire' [@RN5381] to test the significant of the final vegetation groups; and Detrended Correspondence Analysis (DCA) as implemented in the R package 'vegan' [@RN5386] to visualize the relationships between the groups.
+We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package 'RVAideMemoire' [@RN5381] to test the significance of the final vegetation groups; along with Detrended Correspondence Analysis (DCA) as implemented in the R package 'vegan' [@RN5386] to visualize the relationships between the groups.
 
 [Characteristic species and EUNIS habitat regionalization.]{.underline} We calculated sets of characteristic species [@RN5061] for each of the final vegetation groups.
-We defined dominant species as species with more than 25% cover in at least 5% of the relevés of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.05, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations.
-Additionally, we assigned to each group a regionalized level 4 EUNIS habitat code, using pre-existing codes when they existed, and creating new codes when not.
+We defined dominant species as species with more than 25% cover in at least 5% of the vegetation plots of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.05, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations.
+Additionally, we assigned to each group a regionalized level 4 EUNIS habitat code, using pre-existing codes when they existed (https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp), and creating new codes when not.
 
-[Community traits and ecological requirements.]{.underline} To describe the resulting vegetation groups, we compiled a dataset of species traits and ecological preferences for the species pool of the classified relevés.
+[Community traits and ecological requirements.]{.underline} To describe the resulting vegetation groups, we compiled a dataset of species traits and ecological preferences for the species pool of the classified vegetation plots.
 For each species, we extracted from *Flora iberica* [-@RN3243] its maximum height, its median month of flowering, the length of the flowering period, and whether it was a neophyte or not.
 From FloraVeg.EU [-@RN5385] we extracted the species’ life forms and kept 'therophyte' and 'geophyte' for further analysis since these two categories had the largest contribution to variance as per an exploratory Principal Component Analysis (PCA) performed using the R package 'FactoMineR' [@RN3166].
-We also extracted the ecological indicator values [@RN5384] and the disturbance indicator values [@RN5101] of each species.
+We also collected the ecological indicator values [@RN5384] and the disturbance indicator values [@RN5101] of each species.
 In the case of the ecological and disturbance indicators, we assigned values to species lacking them by (1) calculating weighted average values of the species with values for each plot (weighting by species cover), and (2) assigning to the missing species the weighted average values of the plots where they were present (weighting by species cover).
-Then, for each relevé we calculated its community-weighted mean (weighting by species cover) for each trait and ecological preference, and explored the main axes of variation using PCA, separately for species traits and ecological preferences.
+Then, we calculated the plot-level community-weighted mean (weighting by species cover) for each trait and ecological preference, and explored the main axes of variation using PCA, separately for species traits and ecological preferences.
 
@@
 Our revision of the literature produced a provisional list of 38 anthropogenic vegetation alliances that could be present in the Iberian Atlantic ecoregion (**Supplementary 3**).
-In our study dataset (n = 2,508 relevés), there were 2,201 relevés that had been assigned by the authors to any of these alliances.
-We used these 2,201 relevés for a preliminary re-classification based on TWINSPAN, attempting to re-match the relevés into the *a priori* defined checklist of 38 alliances.
+In our study dataset (n = 2,508 vegetation plots), there were 2,201 plots that had been assigned by the original authors to one of these alliances.
+We used these 2,201 plots for a preliminary re-classification based on TWINSPAN, attempting to re-match the plots into the *a priori* defined checklist of 38 alliances.
 This procedure allowed us to validate 28 alliances as present in the ecoregion.
-A total of 1,725 relevés had been classified by both the original authors and our TWINSPAN into these alliances, and these became our training subset for semi-supervised classification.
-Attempts by the semi-supervised classification to create new groups resulted on new groups with no ecological significance, and thus we kept the 28 alliances as final vegetation groups (**Table 1**).
+A total of 1,725 plots had been classified by both the original authors and our TWINSPAN into these alliances, and these became our training subset for the semi-supervised classification.
+Attempts by the semi-supervised classification to create new groups resulted in new groups with no ecological significance, and thus we kept the original 28 alliances as final vegetation groups (**Table 1**).
 Most of the alliances had a wide distribution in the ecoregion according to our dataset (**Figure 1**), although five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis*, *Paspalo-Agrostion semiverticillati*) had only isolated occurrences in a few points towards the southern limit of the Iberian Atlantic territories.
 The regional literature included in our dataset recognized 72 associations (**Supplementary 4**) within these alliances.
-The semi-supervised classification with noise clustering resulted in the final classification of 2,081 relevés into 28 alliances and 427 relevés left in the noise group.
+The semi-supervised classification with noise clustering resulted in the final classification of 2,081 plots into 28 alliances and 427 plots left out in the noise group.
 
@@
 The 10 most frequent species were *Ochlopoa annua* (621 occurrences), *Urtica dioica* (592), *Sonchus oleraceus* (540), *Stellaria media* (516), *Capsella bursa-pastoris* (390), *Polygonum aviculare* (375), *Dactylis glomerata* (363), *Senecio vulgaris* (332) and *Anisantha sterilis* (301).
-Considering only those relevés with the most frequent plot size (20 m^2^, n = 256 relevés), the average species richness per plot was 16 (minimum = 3, maximum = 37).
+Considering only those plots with the most frequent plot size (20 m^2^, n = 256 plots), the average species richness per plot was 16 (minimum = 3, maximum = 37).
 
-DCA ordination of the floristic composition of the classified relevés (**Fig.
+DCA ordination of the floristic composition of the classified plots (**Fig.
 2**) suggested a high consistency of the vegetation classes *Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Sisymbrietea* and *Bidentetea*.
 The alliances belonging to the two perennial ruderal vegetation classes (*Artemisietea vulgaris* and *Epilobietea angustifolii*) were mixed together.
-The winter-annual class *Chenopodietea* also showed high heterogeneity, specially because of the isolation of the altered dune alliance *Linario polygalifoliae-Vulpion alopecuri*.
+The winter-annual class *Chenopodietea* also showed high heterogeneity, especially because of the isolation of the altered dune alliance *Linario polygalifoliae-Vulpion alopecuri*.
 
@@
 3B**) indicated that the first two axes explained 56% of the variability.
-The major contributors to axis 1 (35% variance explained) were disturbance frequency, soil disturbance, mowing frequency and light requirements; it separated communities thriving with frequent disturbances in open situations (classes *Polygono-Poetea annuae*, *Sisymbrietea*, *Digitario-Eragrostietea* and *Papaveretea rhoeadis*) from communities preferring less frequent disturbances and more shade (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii* and *Bidentetea*).
+The major contributors to axis 1 (35% variance explained) were disturbance frequency, soil disturbance, mowing frequency and light requirements; it separated communities thriving with frequent disturbances in open situations (classes *Polygono-Poetea annuae*, *Sisymbrietea*, *Digitario-Eragrostietea* and *Papaveretea rhoeadis*) from communities preferring less frequent disturbances and shadier areas (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii* and *Bidentetea*).
 The major contributors to axis 2 (21% variance explained) were nutrient requirements and disturbance severity; it separated more nitrophilous communities adapted to more severe disturbances (classes *Digitario-Eragrostietea*, *Sisymbrietea* and *Epilobietea angustifolii*) from less nitrophilous communities adapted to lower intensity disturbances (classes *Cymbalario-Parietarietea diffusae*, *Artemisietea vulgaris* and *Chenopodietea*).
@@
 
+Segetal vegetation of Palestine [@RN5494].
+
 # Data availability
@@
 Dots are colored by vegetation classes."}
-knitr::include_graphics("../results/figures/maps.png")
+knitr::include_graphics("../results/figures/F1 - maps.png")
 ```
@@
 The ordination is the results of a Detrended Correspondence Analysis grouped by vegetation classes (A, left) and alliances (B, right)."}
-knitr::include_graphics("../results/figures/dca.png")
+knitr::include_graphics("../results/figures/F2 - pca.png")
 ```
@@
 The ordinations are the results of a Principal Component Analysis calculated for the community-weighted means for species traits (A, left) and ecological and disturbance indicator values (B, right)."}
-knitr::include_graphics("../results/figures/cwms.png")
+knitr::include_graphics("../results/figures/F3 - origins.png")
 ```
@@
 Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
-knitr::include_graphics("../results/figures/boxplots.png")
+knitr::include_graphics("../results/figures/F4 - lifeforms.png")
 ```
+
+```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
+Community-level species traits and ecological preferences by alliance.
+Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
+knitr::include_graphics("../results/figures/F5 - height-flowering.png")
+```
+
+```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
+Community-level species traits and ecological preferences by alliance.
+Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
+knitr::include_graphics("../results/figures/F6 - indicators.png")
+```
+
+```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
+Community-level species traits and ecological preferences by alliance.
+Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
+knitr::include_graphics("../results/figures/F7 - pca.png")
+```

### 2024-04-10 src/10-maps.R

@@ -1,128 +1,0 @@
-rm(list = ls())
-
-library(tidyverse); library(vegan)
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>%
-  filter(! Alliance %in% "Noise") %>%
-  mutate(Class = fct_relevel(Class,
-                             "Cymbalario-Parietarietea diffusae",
-                             "Polygono-Poetea annuae",
-                             "Papaveretea rhoeadis",
-                             "Digitario sanguinalis-Eragrostietea minoris",
-                             "Chenopodietea",
-                             "Sisymbrietea",
-                             "Artemisietea vulgaris",
-                             "Epilobietea angustifolii",
-                             "Bidentetea")) %>%
-  mutate(Class = fct_recode(Class, 
-                            "Artemisietea" = "Artemisietea vulgaris",
-                            "Parietarietea" = "Cymbalario-Parietarietea diffusae",
-                            "Digitario-Eragrostietea" = "Digitario sanguinalis-Eragrostietea minoris",
-                            "Epilobietea" = "Epilobietea angustifolii",
-                            "Papaveretea" = "Papaveretea rhoeadis",
-                            "Polygono-Poetea" = "Polygono-Poetea annuae"))%>%
-  mutate(Alliance = fct_relevel(Alliance, 
-                               "Galio valantiae-Parietarion judaicae",
-                               "Cymbalario-Asplenion",
-                               "Polycarpion tetraphylli",
-                               "Polygono-Coronopodion",
-                               "Saginion procumbentis",
-                               "Caucalidion lappulae",
-                               "Scleranthion annui",
-                               "Oxalidion europeae",
-                               "Spergulo arvensis-Erodion cicutariae",
-                               "Allion triquetri",
-                               "Geranio pusilli-Anthriscion caucalidis",
-                               "Chenopodion muralis",
-                               "Echio-Galactition tomentosae",
-                               "Linario polygalifoliae-Vulpion alopecuri",
-                               "Sisymbrion officinalis",
-                               "Convolvulo arvensis-Agropyrion repentis",
-                               "Carduo carpetani-Cirsion odontolepidis",
-                               "Silybo mariani-Urticion piluliferae", 
-                               "Cirsion richterano-chodati",
-                               "Dauco-Melilotion",
-                               "Geo urbani-Alliarion officinalis",
[…]

### 2024-04-10 src/10-origins.R

@@ -1,0 +1,113 @@
+rm(list = ls())
+
+library(tidyverse)
+
+### Reciprocal averaging
+
+read.csv("data/traits.csv", fileEncoding = "latin1") %>%
+  select(Analysis.Names,
+         origin) %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1"), all.y = TRUE) %>%
+  group_by(SIVIMID, origin) %>%
+  summarise(y = length(Analysis.Names)) %>%
+  merge(read.csv("data/header.csv", fileEncoding = "latin1"), all.y = TRUE) %>%
+  group_by(Class, Alliance, origin) %>%
+  summarise(n = sum(y)) %>%
+  group_by(Alliance) %>% 
+  mutate(proportion = n / sum(n)) %>%
+  group_by() %>% 
+  filter(Alliance != "Noise") %>%
+  mutate(origin = fct_relevel(origin,
+                              "native",
+                             "archaeophyte", 
+                             "neophyte")) %>%
+  mutate(origin = fct_recode(origin, 
+                            "Natives" = "native",
+                            "Archaeophytes" = "archaeophyte",
+                            "Neophytes" = "neophyte")) %>%
+  mutate(Alliance = fct_relevel(Alliance,
+                                "Paspalo-Agrostion semiverticillati",
+                                "Bidention tripartitae",
+                                "Senecionion fluviatilis",
+                                "Cynancho-Convolvulion sepium",
+                                "Epilobion angustifolii",
+                                "Aegopodion podagrariae",
+                                "Balloto-Conion maculati",
+                                "Arction lappae",
+                                "Geo urbani-Alliarion officinalis",
+                                "Dauco-Melilotion", 
+                                "Cirsion richterano-chodati",
+                                "Silybo mariani-Urticion piluliferae",
+                                "Carduo carpetani-Cirsion odontolepidis",
+                                "Convolvulo arvensis-Agropyrion repentis",
+                                "Sisymbrion officinalis",
+                                "Linario polygalifoliae-Vulpion alopecuri",
+                                "Echio-Galactition tomentosae",
+                                "Chenopodion muralis",
+                                "Geranio pusilli-Anthriscion caucalidis",
+                                "Allion triquetri",
+                                "Spergulo arvensis-Erodion cicutariae",
+                                "Oxalidion europeae",
+                                "Scleranthion annui",
+                                "Caucalidion lappulae",
[…]

### 2024-04-10 src/11-lifeforms.R

@@ -1,0 +1,119 @@
+rm(list = ls())
+
+library(tidyverse)
+
+### Reciprocal averaging
+
+read.csv("data/traits.csv", fileEncoding = "latin1") %>%
+  select(Analysis.Names,
+         bryophyte:phanerophyte) %>%
+  gather(Trait, Value,
+         bryophyte:phanerophyte) %>%
+  filter(Value == 1) %>%
+  mutate(origin = ifelse(Trait == "therophyte", "Therophyte", "Other")) %>%
+  mutate(origin = ifelse(Trait == "geophyte", "Geophyte", origin)) %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1"), all.y = TRUE) %>%
+  group_by(SIVIMID, origin) %>%
+  summarise(y = length(Analysis.Names)) %>%
+  merge(read.csv("data/header.csv", fileEncoding = "latin1"), all.y = TRUE) %>%
+  group_by(Class, Alliance, origin) %>%
+  summarise(n = sum(y)) %>%
+  na.omit() %>%
+  group_by(Alliance) %>% 
+  mutate(proportion = n / sum(n)) %>%
+  group_by() %>% 
+  filter(Alliance != "Noise") %>%
+  mutate(origin = fct_relevel(origin, 
+                              "Other",
+                              "Geophyte",
+                              "Therophyte")) %>%
+  mutate(origin = fct_recode(origin, 
+                             "Natives" = "native",
+                             "Archaeophytes" = "archaeophyte",
+                             "Neophytes" = "neophyte")) %>%
+  mutate(Alliance = fct_relevel(Alliance,
+                                "Paspalo-Agrostion semiverticillati",
+                                "Bidention tripartitae",
+                                "Senecionion fluviatilis",
+                                "Cynancho-Convolvulion sepium",
+                                "Epilobion angustifolii",
+                                "Aegopodion podagrariae",
+                                "Balloto-Conion maculati",
+                                "Arction lappae",
+                                "Geo urbani-Alliarion officinalis",
+                                "Dauco-Melilotion", 
+                                "Cirsion richterano-chodati",
+                                "Silybo mariani-Urticion piluliferae",
+                                "Carduo carpetani-Cirsion odontolepidis",
+                                "Convolvulo arvensis-Agropyrion repentis",
+                                "Sisymbrion officinalis",
+                                "Linario polygalifoliae-Vulpion alopecuri",
+                                "Echio-Galactition tomentosae",
+                                "Chenopodion muralis",
[…]

### 2024-04-10 src/11-publications.R

@@ -1,16 +1,0 @@
-library(tidyverse)
-
-read.csv("../canteunis/data/sivim/invATL.csv", fileEncoding = "latin1") %>%
-  select(codi_xtr2, autors, anny, titol, publicacio) %>%
-  rename(SIVIMID = codi_xtr2) %>%
-  unique %>%
-  filter(SIVIMID %in% header$SIVIMID) %>%
-  select(-SIVIMID) %>%
-  unique %>%
-  filter(autors != "") %>%
-  arrange(autors, anny, titol) %>%
-  rename(Authors = autors,
-         Year = anny,
-         Title = titol,
-         Publication = publicacio) %>%
-  write.csv("results/publications/publications.csv", fileEncoding = "latin1", row.names = FALSE)

### 2024-04-10 src/11-spp-numbers.R

@@ -1,88 +1,0 @@
-library(tidyverse)
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>% 
-  filter(Alliance != "Noise") %>%
-  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  group_by(Analysis.Names)
-
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>% 
-  filter(Alliance != "Noise") %>%
-  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  group_by(Analysis.Names) %>% tally %>% arrange(-n)
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>% 
-  filter(Alliance != "Noise") %>%
-  group_by(Area) %>% tally %>% arrange(-n)
-
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>% 
-  filter(Alliance != "Noise" & Area == "20") %>%
-  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  select(SIVIMID, Analysis.Names) %>%
-  unique %>%
-  group_by(SIVIMID) %>% tally %>% arrange(-n)
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>% 
-  filter(Alliance != "Noise" & Area == "20") %>%
-  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  select(SIVIMID, Analysis.Names) %>%
-  unique %>%
-  group_by(SIVIMID) %>% tally %>% arrange(n)
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>% 
-  filter(Alliance != "Noise" & Area == "20") %>%
-  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  select(SIVIMID, Analysis.Names) %>%
-  unique %>%
-  group_by(SIVIMID) %>% tally %>% pull(n) %>% mean
-
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>% 
-  filter(Alliance != "Noise") %>%
-  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  select(SIVIMID, Analysis.Names) %>%
-  unique %>%
-  group_by(SIVIMID) %>% tally %>% arrange(-n)
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>% 
[…]

### 2024-04-10 src/12-height-flowering.R

@@ -1,0 +1,194 @@
+rm(list = ls())
+
+library(tidyverse)
+
+### Reciprocal averaging
+
+# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
+#   select(Analysis.Names,
+#          EIVE1.M:Soil.Disturbance) %>%
+#   gather(Trait, Value, -Analysis.Names) %>%
+#   na.omit %>%
+#   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+#   group_by(SIVIMID, Trait) %>%
+#   summarise(Value = weighted.mean(Value, Cover.percent)) -> plotmeans
+# 
+# read.csv("data/species.csv", fileEncoding = "latin1") %>%
+#   merge(plotmeans) %>%
+#   group_by(Analysis.Names, Trait) %>%
+#   summarise(Value = weighted.mean(Value, Cover.percent)) %>%
+#   rename(ravalue = Value) -> sppmeans
+# 
+# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
+#   select(Analysis.Names,
+#          EIVE1.M:Soil.Disturbance) %>%
+#   gather(Trait, Value, -Analysis.Names) %>%
+#   merge(sppmeans, all.x = TRUE) %>%
+#   mutate(value = ifelse(is.na(Value), ravalue, Value)) %>%
+#   select(-c(Value, ravalue)) %>%
+#   spread(Trait, value) -> raspp
+# 
+# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
+#   select(-c(EIVE1.M:Soil.Disturbance)) %>%
+#   merge(raspp) -> traits
+# 
+# ### CwMs
+# 
+# traits %>%
+#   gather(Trait, Value, -Analysis.Names) %>%
+#   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+#   group_by(SIVIMID, Trait) %>%
+#   summarise(Value = weighted.mean(Value, Cover.percent)) %>%
+#   spread(Trait, Value) %>%
+#   merge(read.csv("data/header.csv", fileEncoding = "latin1")) -> cwms
+# 
+# save(cwms, file = "results/cwms/cwms.RData")
+
+load(file = "results/cwms/cwms.RData")
+
+cwms %>%
+  filter(heightmax < 300) %>%
+  filter(Alliance != "Noise") %>%
+  select(SIVIMID,
+         Class,
+         Alliance,
[…]

### 2024-04-10 src/13-indicators.R

@@ -1,0 +1,303 @@
+rm(list = ls())
+
+library(tidyverse)
+
+### Reciprocal averaging
+
+# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
+#   select(Analysis.Names,
+#          EIVE1.M:Soil.Disturbance) %>%
+#   gather(Trait, Value, -Analysis.Names) %>%
+#   na.omit %>%
+#   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+#   group_by(SIVIMID, Trait) %>%
+#   summarise(Value = weighted.mean(Value, Cover.percent)) -> plotmeans
+# 
+# read.csv("data/species.csv", fileEncoding = "latin1") %>%
+#   merge(plotmeans) %>%
+#   group_by(Analysis.Names, Trait) %>%
+#   summarise(Value = weighted.mean(Value, Cover.percent)) %>%
+#   rename(ravalue = Value) -> sppmeans
+# 
+# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
+#   select(Analysis.Names,
+#          EIVE1.M:Soil.Disturbance) %>%
+#   gather(Trait, Value, -Analysis.Names) %>%
+#   merge(sppmeans, all.x = TRUE) %>%
+#   mutate(value = ifelse(is.na(Value), ravalue, Value)) %>%
+#   select(-c(Value, ravalue)) %>%
+#   spread(Trait, value) -> raspp
+# 
+# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
+#   select(-c(EIVE1.M:Soil.Disturbance)) %>%
+#   merge(raspp) -> traits
+# 
+# ### Plot means
+# 
+# traits %>%
+#   select(Analysis.Names, Disturbance.Frequency, Disturbance.Severity, 
+#          EIVE1.T, EIVE1.M, EIVE1.L, EIVE1.N, EIVE1.R) %>%
+#   gather(Trait, Value, -Analysis.Names) %>%
+#   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+#   group_by(SIVIMID, Trait) %>%
+#   summarise(Value = mean(Value)) %>%
+#   spread(Trait, Value) %>%
+#   merge(read.csv("data/header.csv", fileEncoding = "latin1")) -> plotmeans
+# 
+# save(plotmeans, file = "results/plotmeans/plotmeans.RData")
+
[…]

### 2024-04-10 src/14-publications.R

@@ -1,0 +1,16 @@
+library(tidyverse)
+
+read.csv("../canteunis/data/sivim/invATL.csv", fileEncoding = "latin1") %>%
+  select(codi_xtr2, autors, anny, titol, publicacio) %>%
+  rename(SIVIMID = codi_xtr2) %>%
+  unique %>%
+  filter(SIVIMID %in% header$SIVIMID) %>%
+  select(-SIVIMID) %>%
+  unique %>%
+  filter(autors != "") %>%
+  arrange(autors, anny, titol) %>%
+  rename(Authors = autors,
+         Year = anny,
+         Title = titol,
+         Publication = publicacio) %>%
+  write.csv("results/publications/publications.csv", fileEncoding = "latin1", row.names = FALSE)

### 2024-04-10 src/7-dca.R

@@ -1,186 +1,0 @@
-rm(list = ls())
-
-library(tidyverse); library(vegan)
-
-read.csv("data/header.csv", fileEncoding = "latin1") -> 
-  header
-
-read.csv("data/species.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% header$SIVIMID) -> species
-
-### DCA
-
-species %>%
-  merge(header) %>%
-  filter(Alliance != "Noise") %>%
-  select(SIVIMID, Analysis.Names, Cover.percent) %>%
-  spread(Analysis.Names, Cover.percent, fill = 0) %>%
-  column_to_rownames(var = "SIVIMID") -> df1
-
-decorana(df1) -> dca1
-
-### Paint by classes
-
-vegan::scores(dca1) %>%
-  data.frame() %>%
-  rownames_to_column("SIVIMID") %>%
-  merge(header) %>%
-  mutate(Cluster = as.factor(Class)) %>%
-  mutate(Cluster = fct_recode(Cluster, 
-                              "Artemisietea" = "Artemisietea vulgaris",
-                              "Parietarietea" = "Cymbalario-Parietarietea diffusae",
-                              "Digitario-Eragrostietea" = "Digitario sanguinalis-Eragrostietea minoris",
-                              "Epilobietea" = "Epilobietea angustifolii",
-                              "Papaveretea" = "Papaveretea rhoeadis",
-                              "Polygono-Poetea" = "Polygono-Poetea annuae")) -> df2
-
-aggregate(cbind(DCA1, DCA2) ~ Cluster, data = df2, FUN = mean)%>%
-  mutate(DCA1 = ifelse(Cluster == "Parietarietea", DCA1+.06, DCA1))%>%
-  mutate(DCA1 = ifelse(Cluster == "Chenopodietea", DCA1-.06, DCA1))%>%
-  mutate(DCA2 = ifelse(Cluster == "Papaveretea", DCA2+.08, DCA2))%>%
-  mutate(DCA2 = ifelse(Cluster == "Sisymbrietea", DCA2-.08, DCA2)) -> cent
-merge(df2, setNames(cent, c("Cluster", "oDCA1", "oDCA2")), by = "Cluster", sort = FALSE) -> segs
-
-df2 %>%
-  ggplot(aes(x = DCA1, y = DCA2)) + 
-  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
-  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
[…]

### 2024-04-10 src/8-cwm-pca.R

@@ -1,244 +1,0 @@
-rm(list = ls())
-
-library(tidyverse)
-
-### Reciprocal averaging
-
-# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
-#   select(Analysis.Names,
-#          EIVE1.M:Soil.Disturbance) %>%
-#   gather(Trait, Value, -Analysis.Names) %>%
-#   na.omit %>%
-#   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-#   group_by(SIVIMID, Trait) %>%
-#   summarise(Value = weighted.mean(Value, Cover.percent)) -> plotmeans
-# 
-# read.csv("data/species.csv", fileEncoding = "latin1") %>%
-#   merge(plotmeans) %>%
-#   group_by(Analysis.Names, Trait) %>%
-#   summarise(Value = weighted.mean(Value, Cover.percent)) %>%
-#   rename(ravalue = Value) -> sppmeans
-# 
-# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
-#   select(Analysis.Names,
-#          EIVE1.M:Soil.Disturbance) %>%
-#   gather(Trait, Value, -Analysis.Names) %>%
-#   merge(sppmeans, all.x = TRUE) %>%
-#   mutate(value = ifelse(is.na(Value), ravalue, Value)) %>%
-#   select(-c(Value, ravalue)) %>%
-#   spread(Trait, value) -> raspp
-# 
-# read.csv("data/traits.csv", fileEncoding = "latin1") %>%
-#   select(-c(EIVE1.M:Soil.Disturbance)) %>%
-#   merge(raspp) -> traits
-# 
-# ### CwMs
-# 
-# traits %>%
-#   gather(Trait, Value, -Analysis.Names) %>%
-#   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-#   group_by(SIVIMID, Trait) %>%
-#   summarise(Value = weighted.mean(Value, Cover.percent)) %>%
-#   spread(Trait, Value) %>%
-#   merge(read.csv("data/header.csv", fileEncoding = "latin1")) -> cwms
-# 
-# save(cwms, file = "results/cwms/cwms.RData")
-
-rm(list = ls())
-
-load(file = "results/cwms/cwms.RData")
-  
-### PCA with species traits
-
-cwms %>%
-  filter(Alliance != "Noise") %>%
-  select(SIVIMID,
-         floweringmean,
[…]

### 2024-04-10 src/8-maps.R

@@ -1,0 +1,128 @@
+rm(list = ls())
+
+library(tidyverse); library(vegan)
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>%
+  filter(! Alliance %in% "Noise") %>%
+  mutate(Class = fct_relevel(Class,
+                             "Cymbalario-Parietarietea diffusae",
+                             "Polygono-Poetea annuae",
+                             "Papaveretea rhoeadis",
+                             "Digitario sanguinalis-Eragrostietea minoris",
+                             "Chenopodietea",
+                             "Sisymbrietea",
+                             "Artemisietea vulgaris",
+                             "Epilobietea angustifolii",
+                             "Bidentetea")) %>%
+  mutate(Class = fct_recode(Class, 
+                            "Artemisietea" = "Artemisietea vulgaris",
+                            "Parietarietea" = "Cymbalario-Parietarietea diffusae",
+                            "Digitario-Eragrostietea" = "Digitario sanguinalis-Eragrostietea minoris",
+                            "Epilobietea" = "Epilobietea angustifolii",
+                            "Papaveretea" = "Papaveretea rhoeadis",
+                            "Polygono-Poetea" = "Polygono-Poetea annuae"))%>%
+  mutate(Alliance = fct_relevel(Alliance, 
+                               "Galio valantiae-Parietarion judaicae",
+                               "Cymbalario-Asplenion",
+                               "Polycarpion tetraphylli",
+                               "Polygono-Coronopodion",
+                               "Saginion procumbentis",
+                               "Caucalidion lappulae",
+                               "Scleranthion annui",
+                               "Oxalidion europeae",
+                               "Spergulo arvensis-Erodion cicutariae",
+                               "Allion triquetri",
+                               "Geranio pusilli-Anthriscion caucalidis",
+                               "Chenopodion muralis",
+                               "Echio-Galactition tomentosae",
+                               "Linario polygalifoliae-Vulpion alopecuri",
+                               "Sisymbrion officinalis",
+                               "Convolvulo arvensis-Agropyrion repentis",
+                               "Carduo carpetani-Cirsion odontolepidis",
+                               "Silybo mariani-Urticion piluliferae", 
+                               "Cirsion richterano-chodati",
+                               "Dauco-Melilotion",
+                               "Geo urbani-Alliarion officinalis",
[…]

### 2024-04-10 src/9-cwm-boxplots.R

@@ -1,150 +1,0 @@
-rm(list = ls())
-
-library(tidyverse)
-
-load(file = "results/cwms/cwms.RData")
-
-cwms %>%
-  filter(Alliance != "Noise") %>%
-  select(SIVIMID,
-         Class,
-         Alliance,
-         floweringmean,
-         floweringwidth,
-         heightmax,
-         geophyte, 
-         therophyte,
-         neophyte,
-         Disturbance.Frequency,
-         Disturbance.Severity,
-         Grazing.Pressure,
-         Mowing.Frequency,
-         Soil.Disturbance,
-         EIVE1.M,
-         EIVE1.N,
-         EIVE1.R,
-         EIVE1.T,
-         EIVE1.L) %>%
-  gather(Trait, Value, floweringmean:EIVE1.L) %>%
-  na.omit  %>%
-  mutate(Trait = fct_relevel(Trait, 
-                                "heightmax",
-                                "floweringmean",
-                                "floweringwidth",
-                                "neophyte",
-                                "geophyte",
-                                "therophyte",
-                                "EIVE1.T",
-                                "EIVE1.M",
-                                "EIVE1.L",
-                                "EIVE1.N",
-                                "EIVE1.R",
-                                "Disturbance.Frequency",
-                                "Disturbance.Severity",
-                                "Soil.Disturbance",
-                                "Mowing.Frequency",
-                                "Grazing.Pressure")) %>%
-  mutate(Trait = fct_recode(Trait, 
-                            "Vegetation height (cm)" = "heightmax",
-                            "Flowering peak (month)" = "floweringmean",
-                            "Flowering length (months)" = "floweringwidth",
-                            "Neophytes (% of total cover)" = "neophyte",
-                            "Geophytes (% of total cover)" = "geophyte",
-                            "Therophytes (% of total cover)" = "therophyte",
-                            "Temperature (indicator)" = "EIVE1.T",
-                            "Moisture (indicator)" = "EIVE1.M",
-                            "Light (indicator)" = "EIVE1.L",
-                            "Nutrients (indicator)" = "EIVE1.N",
-                            "Soil reaction (indicator)" = "EIVE1.R",
-                            "Disturbance frequency (indicator)" = "Disturbance.Frequency",
-                            "Disturbance severity (indicator)" = "Disturbance.Severity",
-                            "Soil disturbance (indicator)" = "Soil.Disturbance",
-                            "Mowing frequency (indicator)" = "Mowing.Frequency",
[…]

### 2024-04-10 src/9-ordination.R

@@ -1,0 +1,487 @@
+rm(list = ls())
+
+library(tidyverse); library(vegan)
+
+read.csv("data/header.csv", fileEncoding = "latin1") -> 
+  header
+
+read.csv("data/species.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% header$SIVIMID) -> species
+
+### PCA
+
+species %>%
+  merge(header) %>%
+  filter(Alliance != "Noise") %>%
+  select(SIVIMID, Analysis.Names, Cover.percent) %>%
+  spread(Analysis.Names, Cover.percent, fill = 0) %>%
+  column_to_rownames(var = "SIVIMID") %>% 
+  decostand(method = "hellinger") -> df1
+
+df1 %>%
+  FactoMineR::PCA(graph = TRUE, scale.unit = FALSE) -> pca1
+
+pca1$eig; pca1$var
+
+### Paint by classes
+
+header %>%
+  filter(Alliance != "Noise") %>%
+  select(SIVIMID, Alliance, Class) %>%
+  merge(pca1$ind$coord %>%
+          data.frame %>%
+          rownames_to_column(var = "SIVIMID") %>%
+          select(SIVIMID, Dim.1, Dim.2)) %>%
+  mutate(Class = fct_relevel(Class,
+                             "Cymbalario-Parietarietea diffusae",
+                             "Polygono-Poetea annuae",
+                             "Papaveretea rhoeadis",
+                             "Digitario sanguinalis-Eragrostietea minoris",
+                             "Chenopodietea",
+                             "Sisymbrietea",
+                             "Artemisietea vulgaris",
+                             "Epilobietea angustifolii",
+                             "Bidentetea")) %>%
+  mutate(Class = fct_recode(Class, 
+                            "Artemisietea" = "Artemisietea vulgaris",
+                            "Parietarietea" = "Cymbalario-Parietarietea diffusae",
+                            "Digitario-Eragrostietea" = "Digitario sanguinalis-Eragrostietea minoris",
+                            "Epilobietea" = "Epilobietea angustifolii",
+                            "Papaveretea" = "Papaveretea rhoeadis",
+                            "Polygono-Poetea" = "Polygono-Poetea annuae")) -> df2
+
+aggregate(cbind(Dim.1, Dim.2) ~ Class, data = df2, FUN = mean) -> cent
[…]

### 2024-04-11 doc/manuscript.Rmd
revised methods

@@
 output: 
-  bookdown::word_document2:
+  bookdown::github_document2:
     reference_docx: template.docx
@@
 
+We performed all data management and analysis with R version 4.3.1 [@RN5387], using the R package *tidyverse* [@RN4662] for data processing and visualization.
+We homogenized all taxon names in the analysis and the article using Euro+Med [@RN5044] or Plants of the World Online for taxa not in Euro+Med [@RN5376].
+For the nomenclature of sintaxa we follow Mucina *et al.* [-@RN4698] for alliances and higher ranks, and Rivas-Martínez *et al.* [-@RN4697] for associations.
+The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/manmade>.
+
 ## Study ecoregion
 
-We studied the anthropogenic plant communities of the Iberian Atlantic ecoregion, i.e.
+We studied the anthropogenic plant communities of the Iberian Atlantic ecoregion [@RN5372], i.e.
 the territories with a temperate climate in the north-western Iberian Peninsula.
-Our study ecoregion broadly corresponds with the Cantabrian Mixed Forests ecoregion *sensu* Olson *et al.* [-@RN4667], with the Iberian part of the European Atlantic province *sensu* Rivas-Martínez *et al.* [-@RN4699] and with the Iberian section of the Atlantic biogeographical region of the European Environmental Agency (https://www.eea.europa.eu/data-and-maps/figures/biogeographical-regions-in-europe-2).
-To define in detail the study ecoregion we did a fine-scale revision of its geographic borders at high spatial resolution using the most recent biogeographical map of the Iberian Atlantic territories [@RN5372].
-The revised Iberian Atlantic ecoregion encompasses 79,891 km^2^ of Portugal, Spain and France.
-Its climate is temperate with average monthly temperatures ranging from 6 to 20°C, and with total annual precipitation between 1,100 and 2,600 mm, with some areas having warm-dry summers but without the distinctive 2-month drought of the Mediterranean climate [@RN5373].
-Nonetheless, the Iberian Atlantic ecoregion is located at the transition between the temperate and the Mediterranean climates in southern Europe.
+Our study ecoregion broadly corresponds with the Cantabrian Mixed Forests ecoregion *sensu* Olson *et al.* [-@RN4667], with the Iberian part of the European Atlantic province *sensu* Rivas-Martínez *et al.* [-@RN4699] and with the Iberian section of the Atlantic biogeographical region of the European Environmental Agency (<https://www.eea.europa.eu/data-and-maps/figures/biogeographical-regions-in-europe-2>).
 
@@
 
-To define our target vegetation, we followed the definitions of anthropogenic vegetation in the recent classification of the vegetation of Europe by Mucina *et al.* [-@RN4698].
-In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Bidentetea*.
-For a complete synthesis, we added the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [-@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls in our ecoregion.
-For the nomenclature of sintaxa we follow Mucina *et al.* [-@RN4698] for alliances and higher ranks, and Rivas-Martínez *et al.* [-@RN4697] for associations.
+To circumscribe our study vegetation, we followed the definition of anthropogenic vegetation in the recent classification of the vegetation of Europe by Mucina *et al.* [-@RN4698].
+In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*.
+To these, and for a complete synthesis, we added to our study the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [-@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls.
 
-## Vegetation data
+## Checklist of anthropogenic sintaxa
+
+To assist in our classification, we prepared a checklist of anthropogenic sintaxa that could potentially be present in our study ecoregion (**Supplementary 1**), based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
+The checklist inclused 38 anthropogenic vegetation alliances that could be present in the Iberian Atlantic ecoregion (**Supplementary 2**).
+A goal of our study was to validate the presence of these alliances in our ecoregion, and to assign to them the existing vegetation data.
+
+## Vegetation data selection
 
 We retrieved anthropogenic vegetation plots from SIVIM, the Iberian and Macaronesian Vegetation Information System [@RN5374].
-To identify plots as anthropogenic we followed two criteria: (1) We used the expert system created by Chytrý *et al.* [-@RN5061] to classify vegetation plots into EUNIS pan-European habitat types.
+To identify plots as anthropogenic, we selected those plots that had been classified by the original authors of the plot as belonging to any sintaxa in our sintaxonomical checklist.
+However, since not all plots in SIVIM have an assigned sintaxon, we also retrieved additional plots using the expert system created by Chytrý *et al.* [-@RN5061] to classify vegetation plots into EUNIS pan-European habitat types.
 This expert system uses the floristic composition of plots to assign them to habitat types depending on the presence and abundance of characteristic species.
-We used the habitats assigned by the expert system to retrieve plots that had been assigned to any habitat related to the vegetation classes we had previously defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation) (R55 and R57 include communities that are classified as "grasslands" by EUNIS but as "anthropogenic" by by Mucina *et al.* [-@RN4698]).
-(2) We also retrieved vegetation plots from SIVIM if they had been classified by the original authors of the relevé into any sintaxa belonging to our target anthropogenic vegetation classes (**Supplementary 1**).
-This produced an initial pool of 3,160 vegetation plots from SIVIM, to which we added 89 vegetation plots of urban or peri-urban plant communities sampled by us or extracted from local literature not in SIVIM [@RN5377; @RN5378].
-After initial exploration of this dataset using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379], we identified 652 outlier plots which either did not belong to the target anthropogenic classes or to the geographical area of the ecoregion.
+We retrieved plots that had been assigned to any habitat related to the vegetation classes we had previously defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation).
+It must be noted that habitats R55 and R57 include communities that are classified as *grasslands and lands dominated by forbs, mosses or lichens * by EUNIS but as *anthropogenic vegetation* by by Mucina *et al.* [-@RN4698].
+The extraction from SIVIM produced an initial pool of 3,160 vegetation plots, to which we added 89 vegetation plots of urban or peri-urban plant communities sampled by us or extracted from local literature not in SIVIM [@RN5377; @RN5378].
+
+## Vegetation data cleaning
+
+We performed an exploratory analysis of these 3,249 plots using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379].
+We used the R package *twinspanR* [@RN5380] to do the TWINSPANs in this article, using 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity.
+During this exploratory analysis we identified 741 outlier plots which either did not belong to the target anthropogenic vegetation or to the geographical area of the ecoregion.
 The majority of these plots corresponded to coastal plant communities and had been misclassified by the expert system.
-We removed these, leaving a dataset of 2,508 vegetation  plots for data analyses: 2,419 plots from SIVIM, originally recorded in 89 publications (**Supplementary 2**); plus 89 plots added by us.
+We removed these, leaving a dataset of 2,508 vegetation plots for data analysis: 2,419 plots from SIVIM, originally recorded in 89 publications (**Supplementary 3**); plus 89 plots added by us.
 
-## Statistical analyses
+## Training dataset for semi-supervised classification and validation of alliances
 
-We performed all data analyses with R version 4.3.1 [@RN5387], using the R package 'tidyverse' [@RN4662] for data processing and visualization.
-The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
-We homogenized all taxon names in the analyses and the article using Euro+Med [@RN5044] or Plants of the World Online for taxa not in Euro+Med [@RN5376].
+To create a training dataset for semi-supervised classification (see below), we started by keeping only those plots (n = 2,201) that had been assigned by the original authors of the plot as belonging to any of the 38 alliances in our sintaxonomical checklist (i.e., we removed the plots without sintaxa, that had been retrieved solely by the expert system).
+We further cleaned this subset by subjecting it to a preliminary numerical classification using TWINSPAN, in which we attempted to match the plots into the 38 alliances our sintaxonomical checklist.
+This step allowed us to identify those plots in which there was an agreement between (1) our TWINSPAN-based classification and (2) the classification based on the original-author sintaxa; these plots (n = 1,725) became our training dataset for semi-supervised classification.
+Furthermore, this step allowed us to validate which alliances from our sintaxonomical checklist were present in the study ecoregion (n = 28) and which were absent (n = 10).
 
-[Definition of target alliances and training dataset.]{.underline} As a previous step before classification, we prepared a revised checklist of potential anthropogenic alliances that could be present in our study ecoregion, based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
-Then, we filtered our dataset to extract a subset with the vegetation plots which had been assigned by the authors to any of these alliances.
-We used TWINSPAN to conduct a preliminary numerical classification of this subset, attempting to match the plots into the *a priori* defined checklist of alliances.
-We used 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity, and the R package 'twinspanR' [@RN5380] to do the TWINSPANs.
-To start, we did a TWINSPAN classification into 15 clusters, followed by separate TWINSPANS within each cluster.
-This preliminary TWINSPAN allowed us to (1) validate which alliances were present in the study ecoregion; and (2) prepare a training dataset for semi-supervised classification (see below).
+## Semi-supervised classification
 
-[Semi-supervised classification.]{.underline} Next, we applied a semi-supervised classification to the whole dataset.
+Next, we conducted a semi-supervised classification of the whole dataset (n = 2,508 plots) into the 28 validated anthropogenic alliances.
 Semi-supervised classification uses a training subset of *a priori* classified vegetation plots to classify a secondary subset of unclassified plots [@RN5383].
-Our training subset included all the plots that (1) had been classified by the authors of the relevé into any of the alliances that we had validated during the previous TWINSPANs, and (2) had also been matched by us into these alliances during our TWINSPANs.
-The secondary subset included all the plots which did not meet these two criteria.
-Our goal was to refine and not to simply extend the classification, so we applied the semi-supervised classification to the whole dataset, i.e.
-we allowed plots from the training subset to be re-assigned to other groups during the classification.
-In addition, semi-supervised classification can create new groups to place data points that do not match the already existing *a priori* groups.
-Furthermore, we applied a noise clustering fuzzy algorithm, which allows the plots to be classified either into the *a priori* groups or into a 'noise' group which includes outliers and transitional plots [@RN5382].
+Since our goal was to refine the classification of the whole dataset, we allowed plots from the training subset to be re-assigned to other alliances during the classification.
+In addition, semi-supervised classification can create new groups to place data points that do not match the already existing *a priori* groups, but attempts to do so resulted in new groups with no ecological significance, and thus we kept the  28 alliances as final vegetation groups.
+Furthermore, we applied a noise clustering fuzzy algorithm, which allows the plots to be classified either into the *a priori* alliances or into a *noise* group which includes outliers and transitional plots [@RN5382].
 We set the fuzziness coefficient to a low value (m = 1) to accommodate a high number of transitional plots; and we set the distance to the noise class to d = 1.
-We performed this analysis with the R package 'vegclust' [@RN5383].
-We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package 'RVAideMemoire' [@RN5381] to test the significance of the final vegetation groups; along with Detrended Correspondence Analysis (DCA) as implemented in the R package 'vegan' [@RN5386] to visualize the relationships between the groups.
+We performed this analysis with the R package *vegclust* [@RN5383].
+The semi-supervised classification with noise clustering resulted in the final classification of 2,081 plots into 28 alliances and 427 plots left out in the noise group.
+We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package *RVAideMemoire* [@RN5381] to test the significance of the final vegetation alliances; along with Principal Component Analysis (PCA) as implemented in the R package *FactoMineR* [@RN3166] to visualize the relationships between the alliances.
 
-[Characteristic species and EUNIS habitat regionalization.]{.underline} We calculated sets of characteristic species [@RN5061] for each of the final vegetation groups.
-We defined dominant species as species with more than 25% cover in at least 5% of the vegetation plots of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.05, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations.
-Additionally, we assigned to each group a regionalized level 4 EUNIS habitat code, using pre-existing codes when they existed (https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp), and creating new codes when not.
+## Characteristic species and EUNIS habitat regionalization
 
-[Community traits and ecological requirements.]{.underline} To describe the resulting vegetation groups, we compiled a dataset of species traits and ecological preferences for the species pool of the classified vegetation plots.
-For each species, we extracted from *Flora iberica* [-@RN3243] its maximum height, its median month of flowering, the length of the flowering period, and whether it was a neophyte or not.
-From FloraVeg.EU [-@RN5385] we extracted the species’ life forms and kept 'therophyte' and 'geophyte' for further analysis since these two categories had the largest contribution to variance as per an exploratory Principal Component Analysis (PCA) performed using the R package 'FactoMineR' [@RN3166].
-We also collected the ecological indicator values [@RN5384] and the disturbance indicator values [@RN5101] of each species.
-In the case of the ecological and disturbance indicators, we assigned values to species lacking them by (1) calculating weighted average values of the species with values for each plot (weighting by species cover), and (2) assigning to the missing species the weighted average values of the plots where they were present (weighting by species cover).
-Then, we calculated the plot-level community-weighted mean (weighting by species cover) for each trait and ecological preference, and explored the main axes of variation using PCA, separately for species traits and ecological preferences.
+We calculated sets of characteristic species [@RN5061] for each of the final 28 vegetation alliances.
+We defined dominant species as species with more than 25% cover in at least 5% of the vegetation plots of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species with an *IndVal* value with a p-value lower than 0.05, as calculated with the R package *labdsv* [@RN2311] using 1.000.000 iterations.
+Additionally, we assigned to each alliance a regionalized level 4 EUNIS habitat code, using pre-existing codes when they existed (<https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp>), and creating new codes when not.
+
+## Species origin as native, archaeophyte and neophyte
+
+We classified the species as native, archaophytes and neophytes using the information in *Flora iberica* [-@RN3243] and catalogues of archaeophytes for Britain [@RN5490] and the Czech Republic [@RN5495].
+We must stress that identifying archaeophytes is southern Europe is highly problematic and our classification must be taken as an indication of putative archaeophyte character for the purposes of vegetation description, rather than a definitive classification of the species.
+
+## Species life form
+
+From *FloraVeg.EU* [-@RN5385] we extracted the species’ life forms and kept *therophyte* and *geophyte* for further analysis since these two categories had the largest contribution to variance as per an exploratory Principal Component Analysis (PCA) performed using the R package *FactoMineR* [@RN3166].
+
+## Species and vegetation height and flowering phenology
+
+From *Flora iberica* [-@RN3243] we extracted the maximum height, median month of flowering and length of the flowering period of each species in the dataset.
+We used these to calculate plot-level community-weighted means (weighting by species cover) for each trait, to characterize the vegetation height and flowering phenology of the anthropogenic plant communities.
+
+## Species ecological indicator values
+
+We collected, for each species in the dataset, its ecological indicator values of temperature, moisture, light, nutrients and soil reaction [@RN5384]; and its disturbance frequency and severity indicator values [@RN5101].
+To those species lacking an indicator value in the accessed references, we assigned a value by reciprocal averaging, i.e.
+by (1) calculating weighted average values of the species with values for each plot (weighting by species cover), and (2) assigning to the missing species the weighted average values of the plots where they were present (weighting by species cover).
+Then, for each vegetation plot, we calculated the plot-level mean of each indicator value.
 
@@
 
-Our revision of the literature produced a provisional list of 38 anthropogenic vegetation alliances that could be present in the Iberian Atlantic ecoregion (**Supplementary 3**).
-In our study dataset (n = 2,508 vegetation plots), there were 2,201 plots that had been assigned by the original authors to one of these alliances.
-We used these 2,201 plots for a preliminary re-classification based on TWINSPAN, attempting to re-match the plots into the *a priori* defined checklist of 38 alliances.
-This procedure allowed us to validate 28 alliances as present in the ecoregion.
-A total of 1,725 plots had been classified by both the original authors and our TWINSPAN into these alliances, and these became our training subset for the semi-supervised classification.
-Attempts by the semi-supervised classification to create new groups resulted in new groups with no ecological significance, and thus we kept the original 28 alliances as final vegetation groups (**Table 1**).
-Most of the alliances had a wide distribution in the ecoregion according to our dataset (**Figure 1**), although five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis*, *Paspalo-Agrostion semiverticillati*) had only isolated occurrences in a few points towards the southern limit of the Iberian Atlantic territories.
+The semi-supervised classification resulted in the classification of 2,081 vegetation plots into 28 anthropogenic alliances of the Iberian Atlantic ecoregion (**Table 1**, **Figure 1**).
+Most of the alliances had a wide distribution in the ecoregion according to our dataset, although five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis*, *Paspalo-Agrostion semiverticillati*) had only isolated occurrences in a few points towards the southern limit of the Iberian Atlantic territories.
 The regional literature included in our dataset recognized 72 associations (**Supplementary 4**) within these alliances.
-The semi-supervised classification with noise clustering resulted in the final classification of 2,081 plots into 28 alliances and 427 plots left out in the noise group.
 
@@
 
+***Bidentetea*** Summer-annual pioneer vegetation of temporarily-flooded sites.
+
+16.
+*Bidention tripartitae* Summer-annual pioneer vegetation of temporarily-flooded sites.
+Occupies nutrient-richer sites than the other alliance in the class and has a higher proportion of therophytes (**Fig.
+4**).
+Dispersed throughout the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Persicaria hydropiper*, *Bidens frondosus*, *Lythrum salicaria*.
+
+17.
+*Paspalo-Agrostion semiverticillati* Summer-annual pioneer vegetation of temporarily-flooded and warm sites.
+Occupies warmer and wetter situations than the other alliance in the class, has a higher proportion of neophytes and has a longer flowering season (**Fig.
+4**).
+Isolated occurrences in the south of the ecoregion (**Fig.
+1**).
+Diagnostic species (**Supplementary 5**): *Paspalum distichum*, *Cyperus longus*, *Schoenoplectus lacustris*.
+
 ***Artemisietea vulgaris*** Perennial ruderal vegetation of dry sites.
 
-16.
+18.
 *Convolvulo arvensis-Agropyrion repentis* Semiruderal grasslands.
@@
 
-17.
+19.
 *Carduo carpetani-Cirsion odontolepidis* Thistle ruderal vegetation of warm sites.
@@
 
-18.
+20.
 *Cirsion richterano-chodati* Thistle ruderal vegetation of cold sites.
@@
 
-19.
+21.
 *Dauco-Melilotion* Biennial ruderal vegetation.
@@
 
-20.
+22.
 *Geo urbani-Alliarion officinalis* Low-herb short-lived semiruderal and fringe vegetation.
@@
 
-21.
+23.
 *Arction lappae* Low-herb short-lived ruderal vegetation.
@@
 
-22.
+24.
 *Balloto-Conion maculati* Tall-herb perennial ruderal vegetation.
@@
 
-23.
+25.
 *Aegopodion podagrariae* Tall-herb perennial vegetation of forest margins and clearings.
@@
 
-24.
+26.
 *Epilobion angustifolii* Tall-herb perennial vegetation of forest margins and clearings in acidic soils.
@@
 
-25.
+27.
 *Cynancho-Convolvulion sepium* Tall-herb vegetation of nutrient-rich riparian habitats.
@@
 
-26.
+28.
 *Senecionion fluviatilis* Tall-herb vegetation of nutrient-rich river banks and ditches.
@@
 Diagnostic species (**Supplementary 5**): *Arundo donax*, *Silene latifolia*.
-
-***Bidentetea*** Summer-annual pioneer vegetation of temporarily-flooded sites.
-
-27.
-*Bidention tripartitae* Summer-annual pioneer vegetation of temporarily-flooded sites.
-Occupies nutrient-richer sites than the other alliance in the class and has a higher proportion of therophytes (**Fig.
-4**).
-Dispersed throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Persicaria hydropiper*, *Bidens frondosus*, *Lythrum salicaria*.
-
-28.
-*Paspalo-Agrostion semiverticillati* Summer-annual pioneer vegetation of temporarily-flooded and warm sites.
-Occupies warmer and wetter situations than the other alliance in the class, has a higher proportion of neophytes and has a longer flowering season (**Fig.
-4**).
-Isolated occurrences in the south of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Paspalum distichum*, *Cyperus longus*, *Schoenoplectus lacustris*.
 
@@
 
-Compare proportion of neophytes to other areas [@RN5429; @@RN5435]
+Compare proportion of neophytes to other areas [@RN5429; \@ @RN5435]
 
@@
 
+History of synanthropic veg in czech [@RN5496]
+
 # Data availability
 
-The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
+The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/manmade>.
 Upon publication, a version of record of the repository will be deposited in Zenodo.
@@
 ```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Figure 1.
-Distribution of the anthropogenic alliances in the Iberian Atlantic ecoregion.
-Each dot represents a relevé.
-Dots are colored by vegetation classes."}
+Anthropogenic vegetation alliances of the Iberian Atlantic ecoregion.
+Each dot is a vegetation plot.
+Dot colors indicate the vegetation class."}
 knitr::include_graphics("../results/figures/F1 - maps.png")
@@
 ```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Figure 2.
-Ordination of the floristic composition of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
-The ordination is the results of a Detrended Correspondence Analysis grouped by vegetation classes (A, left) and alliances (B, right)."}
+Floristic composition of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Biplots produced by Principal Component Analysis (PCA) with Hellinger transformation.
+PCAs were conducted for the whole vegetation (A) and for specific subsets (B-F).
+Colors indicate the vegetation class."}
 knitr::include_graphics("../results/figures/F2 - pca.png")
@@
 ```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Figure 3.
-Ordination of the community-level species traits and ecological preferences.
-The ordinations are the results of a Principal Component Analysis calculated for the community-weighted means for species traits (A, left) and ecological and disturbance indicator values (B, right)."}
+Proportion of neophytes, archaeophytes and native species in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Each bar represents the plot-level average proportion of neophytes (dark shade), archaeophytes (medium shade) and natives (light shade) in each vegetation alliance.
+Colors indicate the vegetation class."}
 knitr::include_graphics("../results/figures/F3 - origins.png")
@@
 ```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
-Community-level species traits and ecological preferences by alliance.
-Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
+Proportion of therophytes, geophytes and other life forms in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Each bar represents the plot-level average proportion of therophytes (dark shade), geophytes (medium shade) and other life forms (light shade) in each vegetation alliance.
+Colors indicate the vegetation class."}
 knitr::include_graphics("../results/figures/F4 - lifeforms.png")
@@
 
-```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
-Community-level species traits and ecological preferences by alliance.
-Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
+```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Figure 5.
+Vegetation heigh and flowering phenology of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Boxplots show the community-weighted mean for plant heigh, month of peak flowering and length of the flowering season for each vegetation alliance.
+Colors indicate the vegetation class."}
 knitr::include_graphics("../results/figures/F5 - height-flowering.png")
@@
 
-```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
-Community-level species traits and ecological preferences by alliance.
-Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
+```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Figure 6.
+Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Boxplots show the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity, calculated for each vegetation alliance.
+Colors indicate the vegetation class."}
 knitr::include_graphics("../results/figures/F6 - indicators.png")
@@
 
-```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
-Community-level species traits and ecological preferences by alliance.
-Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
+```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Figure 7.
+Pattern of variation in the ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Biplot produced by Principal Component Analysis (PCA) of the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.
+Labels and arrows indicate the contribution of each indicator to the first and second principal components.
+Colors indicate the vegetation class."}
 knitr::include_graphics("../results/figures/F7 - pca.png")

### 2024-04-11 src/10-origins.R

@@ -26,8 +26,6 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                             "Archaeophytes" = "archaeophyte",
                             "Neophytes" = "neophyte")) %>%
   mutate(Alliance = fct_relevel(Alliance,
-                                "Paspalo-Agrostion semiverticillati",
-                                "Bidention tripartitae",
                                 "Senecionion fluviatilis",
                                 "Cynancho-Convolvulion sepium",
                                 "Epilobion angustifolii",
@@ -40,6 +38,8 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                                 "Silybo mariani-Urticion piluliferae",
                                 "Carduo carpetani-Cirsion odontolepidis",
                                 "Convolvulo arvensis-Agropyrion repentis",
+                                "Paspalo-Agrostion semiverticillati",
+                                "Bidention tripartitae",
                                 "Sisymbrion officinalis",
                                 "Linario polygalifoliae-Vulpion alopecuri",
                                 "Echio-Galactition tomentosae",
@@ -62,9 +62,9 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                              "Digitario sanguinalis-Eragrostietea minoris",
                              "Chenopodietea",
                              "Sisymbrietea",
+                             "Bidentetea",
                              "Artemisietea vulgaris",
-                             "Epilobietea angustifolii",
-                             "Bidentetea")) %>%
+                             "Epilobietea angustifolii")) %>%
   mutate(Class = fct_recode(Class, 
                             "Artemisietea" = "Artemisietea vulgaris",
                             "Parietarietea" = "Cymbalario-Parietarietea diffusae",
@@ -83,9 +83,9 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                                "chocolate4",
                                "firebrick3",
                                "khaki1",
+                               "goldenrod1",
                                "darkmagenta", 
-                               "limegreen",
-                               "goldenrod1")) +
+                               "limegreen")) +
   ggthemes::theme_tufte() +
   xlab("Alliance") + ylab("Proportion of neophytes, archaeophytes and natives at the plot level") +
   coord_flip() +

### 2024-04-11 src/11-lifeforms.R

@@ -32,8 +32,6 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                              "Archaeophytes" = "archaeophyte",
                              "Neophytes" = "neophyte")) %>%
   mutate(Alliance = fct_relevel(Alliance,
-                                "Paspalo-Agrostion semiverticillati",
-                                "Bidention tripartitae",
                                 "Senecionion fluviatilis",
                                 "Cynancho-Convolvulion sepium",
                                 "Epilobion angustifolii",
@@ -46,6 +44,8 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                                 "Silybo mariani-Urticion piluliferae",
                                 "Carduo carpetani-Cirsion odontolepidis",
                                 "Convolvulo arvensis-Agropyrion repentis",
+                                "Paspalo-Agrostion semiverticillati",
+                                "Bidention tripartitae",
                                 "Sisymbrion officinalis",
                                 "Linario polygalifoliae-Vulpion alopecuri",
                                 "Echio-Galactition tomentosae",
@@ -68,9 +68,9 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                              "Digitario sanguinalis-Eragrostietea minoris",
                              "Chenopodietea",
                              "Sisymbrietea",
+                             "Bidentetea",
                              "Artemisietea vulgaris",
-                             "Epilobietea angustifolii",
-                             "Bidentetea")) %>%
+                             "Epilobietea angustifolii")) %>%
   mutate(Class = fct_recode(Class, 
                             "Artemisietea" = "Artemisietea vulgaris",
                             "Parietarietea" = "Cymbalario-Parietarietea diffusae",
@@ -89,9 +89,9 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                                "chocolate4",
                                "firebrick3",
                                "khaki1",
+                               "goldenrod1",
                                "darkmagenta", 
-                               "limegreen",
-                               "goldenrod1")) +
+                               "limegreen")) +
   ggthemes::theme_tufte() +
   xlab("Alliance") + ylab("Proportion of therophytes, geophytes and other life forms at the plot level") +
   coord_flip() +

### 2024-04-11 src/12-height-flowering.R

@@ -121,6 +121,8 @@ cwms %>%
                                "Echio-Galactition tomentosae",
                                "Linario polygalifoliae-Vulpion alopecuri",
                                "Sisymbrion officinalis",
+                               "Bidention tripartitae",
+                               "Paspalo-Agrostion semiverticillati",
                                "Convolvulo arvensis-Agropyrion repentis",
                                "Carduo carpetani-Cirsion odontolepidis",
                                "Silybo mariani-Urticion piluliferae", 
@@ -132,9 +134,7 @@ cwms %>%
                                "Aegopodion podagrariae",
                                "Epilobion angustifolii",
                                "Cynancho-Convolvulion sepium",
-                               "Senecionion fluviatilis",
-                               "Bidention tripartitae",
-                               "Paspalo-Agrostion semiverticillati")) %>%
+                               "Senecionion fluviatilis")) %>%
   mutate(Class = fct_relevel(Class,
                              "Cymbalario-Parietarietea diffusae",
                              "Polygono-Poetea annuae",
@@ -142,9 +142,9 @@ cwms %>%
                              "Digitario sanguinalis-Eragrostietea minoris",
                              "Chenopodietea",
                              "Sisymbrietea",
+                             "Bidentetea",
                              "Artemisietea vulgaris",
-                             "Epilobietea angustifolii",
-                             "Bidentetea")) %>%
+                             "Epilobietea angustifolii")) %>%
   mutate(Class = fct_recode(Class, 
                             "Artemisietea" = "Artemisietea vulgaris",
                             "Parietarietea" = "Cymbalario-Parietarietea diffusae",
@@ -165,9 +165,9 @@ cwms %>%
                                "chocolate4",
                                "firebrick3",
                                "khaki1",
+                               "goldenrod1",
                                "darkmagenta", 
-                               "limegreen",
-                               "goldenrod1")) +
+                               "limegreen")) +
   ggthemes::theme_tufte() +
   xlab("Alliance") + ylab("Value") +
   theme(text = element_text(family = "sans"),
@@ -181,13 +181,14 @@ cwms %>%
         legend.text = element_text(size = 9, face = "italic"), 
         panel.background = element_rect(color = "black", fill = NULL),
         strip.text = element_text(size = 10, hjust = 0, margin = margin(l = 0, b = 4)),
[…]

### 2024-04-11 src/13-indicators.R

@@ -112,6 +112,8 @@ plotmeans %>%
                                 "Echio-Galactition tomentosae",
                                 "Linario polygalifoliae-Vulpion alopecuri",
                                 "Sisymbrion officinalis",
+                                "Bidention tripartitae",
+                                "Paspalo-Agrostion semiverticillati",
                                 "Convolvulo arvensis-Agropyrion repentis",
                                 "Carduo carpetani-Cirsion odontolepidis",
                                 "Silybo mariani-Urticion piluliferae", 
@@ -123,9 +125,7 @@ plotmeans %>%
                                 "Aegopodion podagrariae",
                                 "Epilobion angustifolii",
                                 "Cynancho-Convolvulion sepium",
-                                "Senecionion fluviatilis",
-                                "Bidention tripartitae",
-                                "Paspalo-Agrostion semiverticillati")) %>%
+                                "Senecionion fluviatilis")) %>%
   mutate(Class = fct_relevel(Class,
                              "Cymbalario-Parietarietea diffusae",
                              "Polygono-Poetea annuae",
@@ -133,9 +133,9 @@ plotmeans %>%
                              "Digitario sanguinalis-Eragrostietea minoris",
                              "Chenopodietea",
                              "Sisymbrietea",
+                             "Bidentetea",
                              "Artemisietea vulgaris",
-                             "Epilobietea angustifolii",
-                             "Bidentetea")) %>%
+                             "Epilobietea angustifolii")) %>%
   mutate(Class = fct_recode(Class, 
                             "Artemisietea" = "Artemisietea vulgaris",
                             "Parietarietea" = "Cymbalario-Parietarietea diffusae",
@@ -156,9 +156,9 @@ plotmeans %>%
                                "chocolate4",
                                "firebrick3",
                                "khaki1",
+                               "goldenrod1",
                                "darkmagenta", 
-                               "limegreen",
-                               "goldenrod1")) +
+                               "limegreen")) +
   ggthemes::theme_tufte() +
   xlab("Alliance") + ylab("Value") +
   theme(text = element_text(family = "sans"),
[…]

### 2024-04-11 src/8-maps.R

@@ -11,9 +11,9 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
                              "Digitario sanguinalis-Eragrostietea minoris",
                              "Chenopodietea",
                              "Sisymbrietea",
+                             "Bidentetea",
                              "Artemisietea vulgaris",
-                             "Epilobietea angustifolii",
-                             "Bidentetea")) %>%
+                             "Epilobietea angustifolii")) %>%
   mutate(Class = fct_recode(Class, 
                             "Artemisietea" = "Artemisietea vulgaris",
                             "Parietarietea" = "Cymbalario-Parietarietea diffusae",
@@ -37,6 +37,8 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
                                "Echio-Galactition tomentosae",
                                "Linario polygalifoliae-Vulpion alopecuri",
                                "Sisymbrion officinalis",
+                               "Bidention tripartitae",
+                               "Paspalo-Agrostion semiverticillati",
                                "Convolvulo arvensis-Agropyrion repentis",
                                "Carduo carpetani-Cirsion odontolepidis",
                                "Silybo mariani-Urticion piluliferae", 
@@ -48,9 +50,7 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
                                "Aegopodion podagrariae",
                                "Epilobion angustifolii",
                                "Cynancho-Convolvulion sepium",
-                               "Senecionion fluviatilis",
-                               "Bidention tripartitae",
-                               "Paspalo-Agrostion semiverticillati")) %>%
+                               "Senecionion fluviatilis")) %>%
   mutate(Alliance = fct_recode(Alliance, 
                                "Galio valantiae-\nParietarion judaicae" = "Galio valantiae-Parietarion judaicae",
                                 # "Cymbalario-Asplenion",
@@ -104,13 +104,14 @@ header %>%
         legend.spacing.y = unit(0, "mm"),
         legend.text = element_text(size = 9, face = "italic"), 
         panel.background = element_rect(color = "black", fill = NULL),
+        panel.spacing = unit(0.05, "lines"),
         strip.text = element_text(size = 7, face = "italic"),
         # strip.text = element_blank(),
         plot.title = element_text(size = 12),
         axis.title = element_text(size = 10),
         axis.text.x = element_text(size = 8, color = "black"),
[…]

### 2024-04-11 src/9-ordination.R

@@ -38,9 +38,9 @@ header %>%
                              "Digitario sanguinalis-Eragrostietea minoris",
                              "Chenopodietea",
                              "Sisymbrietea",
+                             "Bidentetea",
                              "Artemisietea vulgaris",
-                             "Epilobietea angustifolii",
-                             "Bidentetea")) %>%
+                             "Epilobietea angustifolii")) %>%
   mutate(Class = fct_recode(Class, 
                             "Artemisietea" = "Artemisietea vulgaris",
                             "Parietarietea" = "Cymbalario-Parietarietea diffusae",
@@ -58,6 +58,7 @@ df2 %>%
   geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
   geom_segment(data = segs, mapping = aes(xend = oDCA1, yend = oDCA2, color = Class), show.legend = FALSE, alpha = 0.5) +
   ggrepel::geom_text_repel(data = cent, aes(label = Class), size = 2, fontface = "italic", min.segment.length = .25) +
+  #geom_text(data = cent, aes(label = Class), size = 2, fontface = "italic") +
   ggthemes::theme_tufte() +
   xlab("PCA1") + ylab("PCA2") +
   scale_y_continuous(labels = function(x) format(x, nsmall = 2)) +
@@ -79,16 +80,16 @@ df2 %>%
         axis.title = element_blank(),
         axis.text.x = element_text(size = 6, color = "black"),
         axis.text.y = element_text(size = 6, color = "black"),
-        plot.margin = unit(c(0.3,0.2,0,0.2), "cm")) +
+        plot.margin = unit(c(.1,.1,0,0), "cm")) +
   scale_color_manual(values = c("grey60",
                                 "cadetblue4",
                                 "chocolate1", 
                                 "chocolate4",
                                 "firebrick3",
                                 "khaki1",
+                                "goldenrod1",
                                 "darkmagenta", 
-                                "limegreen",
-                                "goldenrod1")) -> Fig1A; Fig1A
+                                "limegreen")) -> Fig1A; Fig1A
 
 ### Get legend
 
@@ -106,9 +107,9 @@ header %>%
                              "Digitario sanguinalis-Eragrostietea minoris",
                              "Chenopodietea",
                              "Sisymbrietea",
+                             "Bidentetea",
[…]

### 2024-04-26 doc/manuscript.Rmd
revised ms

@@
 output: 
-  bookdown::github_document2:
+  bookdown::word_document2:
     reference_docx: template.docx
@@
 
+Running title: Iberian Atlantic anthropogenic vegetation
+
+Eduardo Fernández-Pascual^1,a^ (<https://orcid.org/0000-0002-4743-9577>), Víctor González-García^1^ (<https://orcid.org/0000-0002-8949-7943>), Greta Ivesdal^2^, Adrián Lázaro-Lobo^1^ (<https://orcid.org/0000-0002-0509-2085>), Borja Jiménez-Alfaro^1^ (<https://orcid.org/0000-0001-6601-9597>)
+
+^1^ Biodiversity Research Institute (IMIB), University of Oviedo - CSIC - Principality of Asturias, Mieres, Spain; ^2^ Jardín Botánico Atlántico, Gijón/Xixón, Spain
+
+Correspondence: Eduardo Fernández-Pascual, Instituto Mixto de Investigación en Biodiversidad, Campus de Mieres, Edificio de Investigación, 5ª planta, c/ Gonzalo Gutiérrez Quirós s/n, E-33600 Mieres, Spain.
+Email: [fernandezpeduardo\@uniovi.es](mailto:fernandezpeduardo@uniovi.es){.email}.
+Telephone: +34985104781.
+
+# Author contributions
+
+Eduardo Fernández-Pascual: Conceptualization; Methodology; Data Curation; Formal Analysis; Visualization; Writing – Original Draft Preparation; Writing – Review & Editing.
+Víctor González-García: Methodology; Investigation; Formal Analysis; Writing – Review & Editing.
+Greta Ivesdal: Investigation; Formal Analysis; Writing – Review & Editing.
+Adrián Lázaro-Lobo: Investigation; Writing – Review & Editing.
+Borja Jiménez-Alfaro: Methodology; Writing – Review & Editing.
+
+# Funding
+
+Ayuntamiento de Gijón/Xixón, Jardín Botánico Atlántico (SV-23-GIJON-JBA).
+
+# Conflict of interest statement
+
+None declared.
+
+# Data availability statement
+
+Original datasets, R code for analysis and Rmarkdown code for creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/manmade>.
+Upon publication, a version of record of the repository will be deposited in Zenodo.
+
+# Abstract (300 words max.)
+
+* *Questions* Which is the diversity of anthropogenic plant communities at the ecoregional level in a transitional territory between temperate and Mediterranean Europe?
+How are these communities composed by native plants, archaeophytes and neophytes?
+What are the main environmental and disturbance drivers of the diversity in anthropogenic vegetation?
+
+* *Location* Iberian Atlantic ecoregion, Iberian Peninsula, south-western Europe.
+
+* *Methods* We compiled a vegetation database of 2,508 synanthropic plots and revised the regional checklist of anthropogenic alliances.
+We used modified TWINSPAN and semi-supervised classification to classify the plots into the revised alliances.
+We determined the proportion of natives, archaeophytes and neophytes.
+We described the alliances in terms of vegetation height, flowering phenology and ecological requirements for temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.
+
+* *Results* We classified 2,081 vegetation plots into 28 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*).
+The plots included 1,162 vascular plant taxa: 78 % natives, 15 % archaeophytes and 7 % neophytes.
+The main ecological gradients behind this diversity went from dry-sunny to moist-shady habitats, followed by a gradient of low to high disturbance frequency and severity.
+
+* *Conclusions* The human-made habitats of the Iberian Atlantic ecoregion host one third of the ecoregional species pool and one fifth of the Iberian flora.
+Wet anthropogenic habitats are most affected by neophyte invasions.
+Annual vegetation of arable lands is a refuge for archaeophytes.
+Mesic perennial ruderal vegetation is specially rich in native species and can be a biodiversity asset in urban and peri-urban settings.
+
+# Keywords (10 max)
+
+Synanthropic vegetation, human-made habitats, urban biodiversity, neophytes, archaeophytes, alien plants, biological invasions, habitat classification, semi-supervised classification
+
 # Introduction
@@
 Anthropogenic vegetation cover can also impact regional-level extreme climatic events [@RN5440].
-Furthermore, many synanthropic plants can accumulate heavy metals in their biomass [@RN5433], making them valuable candidates for nature-based solutions such as bioremediation and restoration of degraded landscapes [@RN5447].
+Furthermore, many synanthropic plants can accumulate heavy metals in their biomass [@RN5433], making them valuable candidates for nature-based solutions such as bioremediation and restoration of degraded post-industrial landscapes [@RN5447].
 
-The anthropogenic vegetation of temperate Europe is formed by an original pool of native species favored by humans, but this pool has been constantly enriched by the arrival of alien species: first archaeophytes starting in the Neolithic and peaking in the Bronze Age; and then neophytes from the Modern Age (after 1.500 A.D.) to the present day [@RN5439].
-Synanthropic archaeophytes include many Mediterranean taxa that started their expansion into temperate areas during the Mid-Holocene warm period, but it is difficult to disentangle to what extent this expansion was favored by human activities and/or ongoing climatic changes [@RN5432; @RN5443], especially in southern European areas close to the transition between the Mediterranean and the temperate climates [@RN5444].
-Meanwhile, synanthropic neophytes include many American, African and Asian taxa that arrived as a consequence of the global trade networks established by the Western European colonial powers [@RN5451].
-In the last century, anthropogenic vegetation has shown a decrease in species richness and diversity [@RN5423], as rare synanthropic species have become rarer, and neophytes have increased their abundance at the expense of natives and archaeophytes [@RN5441].
+The anthropogenic vegetation of temperate Europe is formed by an original pool of native species favored by humans, but this pool has been constantly enriched by the arrival of alien species [@RN5504].
+First, by archaeophytes, whose arrival started in the Neolithic and peaked in the Bronze Age, with successive increases during periods of accelerated human colonization such as the early Middle Ages [@RN5504].
+Synanthropic archaeophytes include many Mediterranean taxa originated within the segetal flora of the Near East [@RN5494].
+Some of these archaeophytes are 'obligate weeds' that do not have non-anthropogenic habitats and that are assumed to have evolved together with human agriculture [@RN5494].
+Archaeophytes started their expansion into temperate areas during the Mid-Holocene warm period, but it is difficult to disentangle to what extent this expansion was favored by human activities and/or ongoing climatic changes [@RN5432; @RN5443], especially in southern European areas close to the transition between the Mediterranean and the temperate climates [@RN5444].
+The synanthropic flora of Europe has been further enriched by neophytes, alien taxa that arrived from the start of the Modern Age (after 1.500 A.D.) to the present day [@RN5439].
+Synanthropic neophytes include many American, African and Asian taxa that got to Europe as a consequence of the global trade networks established by the Western European colonial powers [@RN5451].
+In the last century, the European anthropogenic vegetation has shown a decrease in species richness and diversity [@RN5423], as rare synanthropic species have become rarer, and neophytes have increased their abundance at the expense of natives and archaeophytes [@RN5441].
 
-Antropogenic vegetation is likely to be similar within ecological regions (ecoregions, hereafter), since they are assumed to encompass areas with a similar biogeographic history and recurrent local ecosystems [@RN5491].
-However, ecoregional-scale studies on the diversity of antropogenic vegetation are scarce in the literature.
+Anthropogenic vegetation is likely to be similar within ecological regions (ecoregions, hereafter), since they are assumed to encompass areas with a similar biogeographic history and recurrent local ecosystems [@RN5491].
+However, ecoregional-scale studies on the diversity of anthropogenic vegetation are scarce in the literature.
 The Iberian Atlantic ecoregion (a.k.a.
@@
 brownfields) whose management and restoration requires an understanding of anthropogenic communities able to colonize or restore them [@RN5448; @RN5449].
-Although there is a long tradition of studying weed and ruderal communities in the ecoregion [@RN5364; @RN5361; @RN5359; @RN5360] there is a current need for an ecoregional-level synthesis that revises and updates the classification and updates in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
-In this article we have performed such a synthesis with two main objectives: (1) provide an updated classification of anthropogenic plant communities in the Iberian Atlantic ecoregion, and (2) describe the main patterns of vegetation diversity in terms of floristic origin (natives, archaeophytes, neophytes); life forms; functional traits (plant height and flowering time); and ecological indicator values (temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity).
+Although there is a long tradition of studying weed and ruderal communities in the ecoregion [@RN5364; @RN5361; @RN5359; @RN5360] there is a current need for an ecoregional-level synthesis that revises and updates the classification in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
+In this article, we have performed such a synthesis with the following objectives: (1) provide an updated classification of anthropogenic plant communities in the Iberian Atlantic ecoregion; (2) quantify the proportion of natives, archaeophytes and neophytes in different anthropogenic plant communities; (3) characterize the plant communities in terms of life forms, plant height and flowering phenology; (4) describe the ecological requirements of the different communities (temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity); and (5) provide a summary description of each of the plant communities.
 
@@
 
+## Data curation and taxonomic criteria
+
 We performed all data management and analysis with R version 4.3.1 [@RN5387], using the R package *tidyverse* [@RN4662] for data processing and visualization.
-We homogenized all taxon names in the analysis and the article using Euro+Med [@RN5044] or Plants of the World Online for taxa not in Euro+Med [@RN5376].
-For the nomenclature of sintaxa we follow Mucina *et al.* [-@RN4698] for alliances and higher ranks, and Rivas-Martínez *et al.* [-@RN4697] for associations.
-The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/manmade>.
+We homogenized all taxon names in the analysis and the article using Euro+Med [@RN5044] or Plants of the World Online [@RN5376] for taxa not in Euro+Med.
+For the nomenclature of sintaxa, we followed Mucina *et al.* [-@RN4698] for alliances and higher ranks, and Rivas-Martínez *et al.* [-@RN4697] for associations.
+The original datasets, as well as R code for analysis and creation of the manuscript, can be accessed at the GitHub repository <https://github.com/efernandezpascual/manmade>.
 
@@
 We studied the anthropogenic plant communities of the Iberian Atlantic ecoregion [@RN5372], i.e.
-the territories with a temperate climate in the north-western Iberian Peninsula.
+the territories with a temperate climate in the north-western part of the Iberian Peninsula.
 Our study ecoregion broadly corresponds with the Cantabrian Mixed Forests ecoregion *sensu* Olson *et al.* [-@RN4667], with the Iberian part of the European Atlantic province *sensu* Rivas-Martínez *et al.* [-@RN4699] and with the Iberian section of the Atlantic biogeographical region of the European Environmental Agency (<https://www.eea.europa.eu/data-and-maps/figures/biogeographical-regions-in-europe-2>).
@@
 In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*.
-To these, and for a complete synthesis, we added to our study the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [-@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls.
+To these, and for a complete synthesis, we added the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [-@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls.
 
@@
 
-To assist in our classification, we prepared a checklist of anthropogenic sintaxa that could potentially be present in our study ecoregion (**Supplementary 1**), based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
-The checklist inclused 38 anthropogenic vegetation alliances that could be present in the Iberian Atlantic ecoregion (**Supplementary 2**).
-A goal of our study was to validate the presence of these alliances in our ecoregion, and to assign to them the existing vegetation data.
+To assist in our classification, we prepared a checklist of anthropogenic sintaxa that could be present in our study ecoregion (**Appendix S1**), based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
+The sintaxonomical checklist included 38 anthropogenic vegetation alliances that could be present in the Iberian Atlantic ecoregion.
+A goal of our study was to validate the presence of these alliances in the ecoregion, and to assign to them the existing vegetation data.
 
@@
 
-We retrieved anthropogenic vegetation plots from SIVIM, the Iberian and Macaronesian Vegetation Information System [@RN5374].
+We retrieved Iberian Atlantic anthropogenic vegetation plots from SIVIM, the Iberian and Macaronesian Vegetation Information System [@RN5374].
 To identify plots as anthropogenic, we selected those plots that had been classified by the original authors of the plot as belonging to any sintaxa in our sintaxonomical checklist.
 However, since not all plots in SIVIM have an assigned sintaxon, we also retrieved additional plots using the expert system created by Chytrý *et al.* [-@RN5061] to classify vegetation plots into EUNIS pan-European habitat types.
-This expert system uses the floristic composition of plots to assign them to habitat types depending on the presence and abundance of characteristic species.
-We retrieved plots that had been assigned to any habitat related to the vegetation classes we had previously defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation).
-It must be noted that habitats R55 and R57 include communities that are classified as *grasslands and lands dominated by forbs, mosses or lichens * by EUNIS but as *anthropogenic vegetation* by by Mucina *et al.* [-@RN4698].
+We retrieved plots that had been assigned to any habitat related to the vegetation classes we had previously defined: all habitats in the level 1 code *V Vegetated man-made habitats*, plus level 3 codes *R55 Lowland moist or wet tall-herb and fern fringe* and *R57 Herbaceous forest clearing vegetation*.
+It must be noted that habitats *R55* and *R57* include communities that are classified as *R Grasslands and lands dominated by forbs, mosses or lichens* by EUNIS but as *Anthropogenic vegetation* by Mucina *et al.* [-@RN4698].
 The extraction from SIVIM produced an initial pool of 3,160 vegetation plots, to which we added 89 vegetation plots of urban or peri-urban plant communities sampled by us or extracted from local literature not in SIVIM [@RN5377; @RN5378].
@@
 
-We performed an exploratory analysis of these 3,249 plots using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379].
-We used the R package *twinspanR* [@RN5380] to do the TWINSPANs in this article, using 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity.
-During this exploratory analysis we identified 741 outlier plots which either did not belong to the target anthropogenic vegetation or to the geographical area of the ecoregion.
+We performed an exploratory analysis of these 3,249 plots using modified Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379].
+We used the R package *twinspanR* [@RN5380] for TWINSPAN, with 3 pseudospecies cut levels (0, 15, 25), a minimum group size of 10 plots and Sørensen's average dissimilarity.
+During this exploratory analysis we identified 741 outlier plots which did not belong to either the target anthropogenic vegetation or to the geographical area of the ecoregion.
 The majority of these plots corresponded to coastal plant communities and had been misclassified by the expert system.
-We removed these, leaving a dataset of 2,508 vegetation plots for data analysis: 2,419 plots from SIVIM, originally recorded in 89 publications (**Supplementary 3**); plus 89 plots added by us.
+We removed these, leaving a dataset of 2,508 vegetation plots for data analysis: 2,419 plots from SIVIM, originally recorded in 89 publications (**Appendix S2**); plus 89 plots added by us.
 
@@
 To create a training dataset for semi-supervised classification (see below), we started by keeping only those plots (n = 2,201) that had been assigned by the original authors of the plot as belonging to any of the 38 alliances in our sintaxonomical checklist (i.e., we removed the plots without sintaxa, that had been retrieved solely by the expert system).
-We further cleaned this subset by subjecting it to a preliminary numerical classification using TWINSPAN, in which we attempted to match the plots into the 38 alliances our sintaxonomical checklist.
-This step allowed us to identify those plots in which there was an agreement between (1) our TWINSPAN-based classification and (2) the classification based on the original-author sintaxa; these plots (n = 1,725) became our training dataset for semi-supervised classification.
+We further cleaned this subset by subjecting it to a preliminary numerical classification using TWINSPAN, in which we attempted to match the plots into the 38 alliances in our sintaxonomical checklist.
+This step allowed us to identify those plots in which there was an agreement between (1) our TWINSPAN-based classification and (2) the classification based on the original author's sintaxon; these plots (n = 1,725) became our training dataset for semi-supervised classification.
 Furthermore, this step allowed us to validate which alliances from our sintaxonomical checklist were present in the study ecoregion (n = 28) and which were absent (n = 10).
@@
 In addition, semi-supervised classification can create new groups to place data points that do not match the already existing *a priori* groups, but attempts to do so resulted in new groups with no ecological significance, and thus we kept the 28 alliances as final vegetation groups.
-Furthermore, we applied a noise clustering fuzzy algorithm, which allows the plots to be classified either into the *a priori* alliances or into a *noise* group which includes outliers and transitional plots [@RN5382].
+Furthermore, we applied a noise clustering fuzzy algorithm, which allows plots to be classified either into the *a priori* alliances or into a *noise* group which includes outliers and transitional plots [@RN5382].
 We set the fuzziness coefficient to a low value (m = 1) to accommodate a high number of transitional plots; and we set the distance to the noise class to d = 1.
@@
 The semi-supervised classification with noise clustering resulted in the final classification of 2,081 plots into 28 alliances and 427 plots left out in the noise group.
-We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package *RVAideMemoire* [@RN5381] to test the significance of the final vegetation alliances; along with Principal Component Analysis (PCA) as implemented in the R package *FactoMineR* [@RN3166] to visualize the relationships between the alliances.
+We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package *RVAideMemoire* [@RN5381] to test the significance of the final vegetation alliances; along with Principal Component Analysis (PCA) as implemented in the R package *FactoMineR* [@RN3166] to visualize the relationships between the classes and alliances.
 
-## Characteristic species and EUNIS habitat regionalization
+## Characteristic species and EUNIS habitat correspondence
 
-We calculated sets of characteristic species [@RN5061] for each of the final 28 vegetation alliances.
+To facilitate the usage of the classification by local managers and environmental consultants, we calculated sets of characteristic species (**Appendix S3**) [@RN5061] for each of the final 28 vegetation alliances.
 We defined dominant species as species with more than 25% cover in at least 5% of the vegetation plots of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species with an *IndVal* value with a p-value lower than 0.05, as calculated with the R package *labdsv* [@RN2311] using 1.000.000 iterations.
-Additionally, we assigned to each alliance a regionalized level 4 EUNIS habitat code, using pre-existing codes when they existed (<https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp>), and creating new codes when not.
+Additionally, we assigned to each alliance its correspondence to level-3 EUNIS habitat codes (<https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp>).
 
@@
 
-We classified the species as native, archaophytes and neophytes using the information in *Flora iberica* [-@RN3243] and catalogues of archaeophytes for Britain [@RN5490] and the Czech Republic [@RN5495].
-We must stress that identifying archaeophytes is southern Europe is highly problematic and our classification must be taken as an indication of putative archaeophyte character for the purposes of vegetation description, rather than a definitive classification of the species.
+We classified the species as native, archaeophytes and neophytes in the Iberian Atlantic ecoregion using the information in *Flora iberica* [-@RN3243] and catalogues of archaeophytes for Britain [@RN5490] and the Czech Republic [@RN5495].
+We must stress that identifying archaeophytes in southern Europe is highly problematic and our classification must be taken as an indication of putative archaeophyte character for the purposes of vegetation description, rather than a definitive classification of the species: as stated by Preston *et al.* [@RN5490], the classification of a species as an archaeophyte should be interpreted as a hypothesis to be tested by further studies.
 
@@
 
-## Species and vegetation height and flowering phenology
+## Species height and flowering phenology
 
 From *Flora iberica* [-@RN3243] we extracted the maximum height, median month of flowering and length of the flowering period of each species in the dataset.
-We used these to calculate plot-level community-weighted means (weighting by species cover) for each trait, to characterize the vegetation height and flowering phenology of the anthropogenic plant communities.
+We used these variables to calculate plot-level community-weighted means (weighting by species cover) for each trait, to characterize the vegetation height and flowering phenology of the anthropogenic plant communities.
 
@@
 
-The semi-supervised classification resulted in the classification of 2,081 vegetation plots into 28 anthropogenic alliances of the Iberian Atlantic ecoregion (**Table 1**, **Figure 1**).
-Most of the alliances had a wide distribution in the ecoregion according to our dataset, although five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis*, *Paspalo-Agrostion semiverticillati*) had only isolated occurrences in a few points towards the southern limit of the Iberian Atlantic territories.
-The regional literature included in our dataset recognized 72 associations (**Supplementary 4**) within these alliances.
+The semi-supervised classification resulted in the classification of 2,081 vegetation plots into 28 anthropogenic alliances (**Fig.
+1**) representing 9 vegetation classes.
+Most of the alliances had a wide distribution in the ecoregion according to our dataset, but five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis* and *Paspalo-Agrostion semiverticillati*) had isolated occurrences in a few points towards the southern limit of the Iberian Atlantic territories.
+The regional literature included in our dataset recognized 72 associations (**Appendix S4**) within these alliances, but exploratory analysis (not shown) indicated that the separation of these associations within the alliances had a generally weak support based on numerical classification methods.
 
-Focusing on the classified plots, the dataset included 1,167 taxa or taxa aggregates.
-Of these, 101 were neophytes (i.e.
-9% of the species pool).
-Regarding life forms, the species pool was dominated by hemicryptophytes (43%) and therophytes (41%).
+The classified plots included 1,162 taxa or taxa aggregates.
 The 10 most frequent species were *Ochlopoa annua* (621 occurrences), *Urtica dioica* (592), *Sonchus oleraceus* (540), *Stellaria media* (516), *Capsella bursa-pastoris* (390), *Polygonum aviculare* (375), *Dactylis glomerata* (363), *Senecio vulgaris* (332) and *Anisantha sterilis* (301).
 Considering only those plots with the most frequent plot size (20 m^2^, n = 256 plots), the average species richness per plot was 16 (minimum = 3, maximum = 37).
+The class with the richest species pool was *Epilobietea angustifolii* (n = 615) and the poorest was *Cymbalario-Parietarietea diffusae* (n = 171).
 
-DCA ordination of the floristic composition of the classified plots (**Fig.
-2**) suggested a high consistency of the vegetation classes *Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Sisymbrietea* and *Bidentetea*.
-The alliances belonging to the two perennial ruderal vegetation classes (*Artemisietea vulgaris* and *Epilobietea angustifolii*) were mixed together.
-The winter-annual class *Chenopodietea* also showed high heterogeneity, especially because of the isolation of the altered dune alliance *Linario polygalifoliae-Vulpion alopecuri*.
+PCA ordination of the floristic composition of the classified plots (**Fig.
+2A**) suggested a high consistency of three major vegetation groups: (1) trampled vegetation of class *Polygono-Poetea annuae*; (2) crop weeds of classes *Papaveretea rhoeadis* and *Digitario sanguinalis-Eragrostietea minoris*; and (3) perennial ruderal vegetation of classes *Artemisietea vulgaris* and *Epilobietea angustifolii*.
+The central position of the floristic space, between these three major groups, was occupied by the annual ruderal vegetation of classes *Chenopodietea*, *Sisymbrietea* and *Bidentetea*; as well as by the wall vegetation of class *Cymbalario-Parietarietea diffusae*.
+Separate PCAs within the major vegetation groups indicated a coherent separation between most of the alliances, but with some cases of relatively high overlap: for example, between (i) *Galio valantiae-Parietarion judaicae* and *Cymbalario-Asplenion*; (ii) *Geo urbani-Alliarion officinalis* and *Aegopodion podagrariae*.
 
-## Community-level traits and ecological preferences
+## Proportion of natives, archaeophytes and neophytes
 
-PCA ordination of the community-weighted means of the species traits (**Fig.
-3A**) indicated that the first two axes explained 65% of the variability.
-The major contributors to axis 1 (41% variance explained) were community height, proportion of geophytes, mean month of flowering and proportion of therophytes; it separated taller communities with more geophytes and later flowering (classes *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Bidentetea*) from shorter communities with more therophytes and earlier flowering (classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis* and *Sisymbrietea*).
-The major contributors to axis 2 (24% variance explained) were proportion of neophytes and width of the flowering season; it separated communities with more neophytes and longer flowering seasons (classes *Bidentetea*, *Digitario sanguinalis-Eragrostietea minoris* and *Polygono-Poetea annuae*) from communities with less neophytes and shorter flowering seasons (classes *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Chenopodietea*).
+The total synanthropic species pool (n = 1,162) was composed of 78 % natives (n = 908), 15 % putative archaeophytes (n = 178) and 7 % neophytes (n = 76).
+The most frequent putative archaeophytes were *Capsella bursa-pastoris* (390 occurrences), *Senecio vulgaris* (332), *Anisantha sterilis* (301), *Anthemis arvensis* (280), *Hordeum murinum* (274), *Raphanus raphanistrum* (233), *Vicia sativa* (228), *Malva sylvestris* (197), *Sisymbrium officinale* (166) and *Echinochloa crus-galli* (165); while the most frequent neophytes were *Veronica persica* (212), *Erigeron canadensis* (138), *Amaranthus hybridus* (73), *Oxalis corniculata* (67), *Lepidium didymum* (57), *Oxalis latifolia* (51), *Paspalum distichum* (48), *Bidens frondosus* (35), *Cyperus eragrostis* (34) and *Symphyotrichum squamatum* (33).
 
-PCA ordination of the community-weighted means of the species ecological and disturbance preferences (**Fig.
-3B**) indicated that the first two axes explained 56% of the variability.
-The major contributors to axis 1 (35% variance explained) were disturbance frequency, soil disturbance, mowing frequency and light requirements; it separated communities thriving with frequent disturbances in open situations (classes *Polygono-Poetea annuae*, *Sisymbrietea*, *Digitario-Eragrostietea* and *Papaveretea rhoeadis*) from communities preferring less frequent disturbances and shadier areas (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii* and *Bidentetea*).
-The major contributors to axis 2 (21% variance explained) were nutrient requirements and disturbance severity; it separated more nitrophilous communities adapted to more severe disturbances (classes *Digitario-Eragrostietea*, *Sisymbrietea* and *Epilobietea angustifolii*) from less nitrophilous communities adapted to lower intensity disturbances (classes *Cymbalario-Parietarietea diffusae*, *Artemisietea vulgaris* and *Chenopodietea*).
+By vegetation class, *Epilobietea angustifolii* had the highest proportion of natives in its species pool (82 %); *Sisymbrietea* had the highest proportion of putative archaeophytes (31 %); and *Bidentetea* had the highest proportion of neophytes (17 %).
+The alliances with the highest proportions of archaeophytes in their species pools were *Caucalidion lappulae* (41 %), *Chenopodion muralis* (35 %) and *Sisymbrion officinalis* (31 %); while the alliances with the highest proportion of neophytes were *Bidention tripartitae* (19 %), *Allion triquetri* (12 %) and *Spergulo arvensis-Erodion cicutariae* (10%).
 
-## Classified anthropogenic plant communities
+Focusing on plot-level average proportions (**Fig.
+3**), the alliances with the most archaeophytes were *Caucalidion lappulae*, *Spergulo arvensis-Erodion cicutariae* and *Sisymbrion officinalis*, in which only about half of the plot-level proportion was made up of native species.
+The alliances with the highest plot-level proportion of neophytes were *Bidention tripartitae*, *Paspalo-Agrostion semiverticillati* and *Senecionion fluviatilis*.
+Some alliances had high proportions of native species, e.g.
+*Epilobion angustifolii*, *Convolvulo arvensis-Agropyrion repentis* and *Linario polygalifoliae-Vulpion alopecuri*.
+
+## Community-level traits
+
+The most frequent life form in the species pool (n = 1,162 species) were hemicryptophytes (44 %), followed by therophytes (42 %), geophytes (9 %), chamaephytes (6 %), phanerophytes (6 %), bryophytes (1 %) and hydrophytes (1 %).
+The proportion of therophytes across classes and alliances (**Fig.
+4**) agreed with the traditional description of the sintaxa.
+Thus, therophytes dominated the annual communities of trampled-soil vegetation (*Polygono-Poetea annuae*), crops weeds (*Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*) and annual ruderals (*Chenopodietea*, *Sisymbrietea*, *Bidentetea*); while perennial life forms (especially hemicryptophytes) dominated walls (*Cymbalario-Parietarietea diffusae*) and perennial ruderal vegetation (*Artemisietea vulgaris*, *Epilobietea angustifolii*).
+Geophytes were a relatively high proportion of the vegetation in three alliances: *Allion triquetri*, *Convolvulo arvensis-Agropyrion repentis* and *Senecionion fluviatilis*.
+
+Community-weighted means for plant height (**Fig.
+5A**) also agreed with the expected description of the sintaxa.
+The shortest communities were the dwarf-herb vegetation of trampled sites (*Polygono-Poetea annuae*, average height = 45.2 cm) and crop weeds (*Papaveretea rhoeadis*, 55.4 cm), while the tallest were the perennial ruderal vegetation of dry (*Artemisietea vulgaris*, 129 cm) and mesic (*Epilobietea angustifolii*, 136 cm) sites.
+The median month of flowering (**Fig.
+5B**) was May (*Sisymbrietea*, *Chenopodietea*, *Papaveretea rhoeadis*), June (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Epilobietea angustifolii*, *Artemisietea vulgaris*) or July (*Digitario sanguinalis-Eragrostietea minoris*, *Bidentetea*).
+The length of the flowering season (**Fig.
+5C**) was generally high, from 5 months (*Artemisietea vulgaris*, *Epilobietea angustifolii*) to 6 (*Chenopodietea*, *Sisymbrietea*), 7 (*Bidentetea*, *Digitario sanguinalis-Eragrostietea minoris*, *Papaveretea rhoeadis*), 8 (*Polygono-Poetea annuae*) or 9 (*Cymbalario-Parietarietea diffusae*).
+
+## Community-level ecological preferences
+
+To visualize the major patterns of variation in the ecological and disturbance preferences of the different communities (**Fig.
+6**) we performed a PCA ordination of plot-level means (**Fig.
+7**).
+The first two PCA axes explained 59 % of the variability.
+The major contributors to axis 1 (36 % variance explained) were light, nutrient and moisture requirements; it separated communities of open, dry and comparatively nutrient-poorer sites (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea*, *Artemisietea vulgaris*, *Chenopodietea* and *Digitario sanguinalis-Eragrostietea minoris*) from communities of shady, moist and nutrient-richer sites (classes *Cymbalario-Parietarietea diffusae*, *Bidentetea* and *Epilobietea angustifolii*).
+The major contributors to axis 2 (23% variance explained) were disturbance severity and frequency; this axis separated communities preferring low severity - low frequency disturbances (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii*, *Chenopodietea*, *Artemisietea vulgaris* and *Bidentetea*) from communities adapted to high severity - high frequency disturbances (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea* and *Digitario sanguinalis-Eragrostietea minoris*).
+
+## Description of the Iberian Atlantic anthropogenic communities
 
@@
 1.
-*Galio valantiae-Parietarion judaicae* Vegetation of dry walls.
-Occupies warmer, drier and more nutrient-rich situations than the other alliance in the class, has a shorter flowering season and is more adapted to disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Parietaria judaica* aggr., *Centranthus ruber*, *Asplenium trichomanes*.
+*Galio valantiae-Parietarion judaicae* Hemicryptophyte-rich vegetation of walls.
+Occupies slightly warmer and sunnier situations than the other alliance in the class and has a shorter flowering season.
+Wide distribution in the ecoregion.
+Diagnostic species in the ecoregion: *Parietaria judaica*, *Centranthus ruber*, *Asplenium trichomanes*.
+Correspondence with EUNIS (2012) habitat *J2.5 - Constructed boundaries*.
 
 2.
-*Cymbalario-Asplenion* Vegetation of mesic walls.
-Occupies colder, wetter and less nutrient-rich situations than the other alliance in the class, has a longer flowering season and is less adapted to disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Cymbalaria muralis*.
+*Cymbalario-Asplenion* Fern-rich vegetation of walls.
+Occupies slightly colder and shadier walls than the other alliance in the class and has a longer flowering season.
+Wide distribution in the ecoregion.
+Diagnostic species in the ecoregion: *Cymbalaria muralis*.
+Correspondence with EUNIS (2012) habitat *J2.5 - Constructed boundaries*.
 
-***Polygono-Poetea annuae*** Dwarf vegetation of heavily-trampled sites.
+***Polygono-Poetea annuae*** Dwarf-herb vegetation of heavily-trampled sites.
 
@@
 *Polycarpion tetraphylli* Dwarf-annual trampled vegetation of warm and sunny sites.
-Occupies warmer, sunnier and less nutrient-rich situations than the other alliances in the class and has an earlier flowering season (**Fig.
-4**).
-Wide distribution in the ecoregion, specially along the coast and in the more Mediterranean valleys of the interior (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Sagina apetala*, *Plantago coronopus*, *Crassula tillaea*, *Spergularia marina*.
+Occupies warmer, sunnier and nutrient-poorer situations than the other alliances in the class and has an earlier flowering season.
+Wide distribution in the ecoregion, specially along the coast and in submediterranean valleys of the interior.
+Diagnostic species in the ecoregion: *Sagina apetala*, *Plantago coronopus*, *Crassula tillaea*, *Spergularia marina*.
+Correspondence with EUNIS habitat *V34 Trampled xeric grassland with annuals*.
 
 4.
-*Polygono-Coronopodion* Dwarf-annual trampled vegetation of dry sites.
-Occupies colder, drier and more nutrient-rich situations than the other alliances in the class, and has a higher proportion of neophytes (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Matricaria suaveolens*, *Polygonum aviculare*.
+*Polygono-Coronopodion* Annual trampled vegetation of dry sites.
+Occupies sites that are colder than those favored by *Polycarpion tetraphylli*, but drier than those of *Saginion procumbentis*.
+Has a specially high proportion of neophytes.
+Wide distribution in the ecoregion.
+Diagnostic species in the ecoregion: *Matricaria suaveolens*, *Polygonum aviculare*.
+Correspondence with EUNIS habitat *V34 Trampled xeric grassland with annuals*.
 
 5.
-*Saginion procumbentis* Dwarf trampled vegetation of mesic sites.
-Occupies colder, wetter and more nutrient-rich situations than the other alliance in the class, and is less dominated by therophytes (**Fig.
-4**).
-Distribution concentrated in the central part of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Sagina procumbens*, *Bryum argenteum*.
+*Saginion procumbentis* Bryophyte-rich trampled vegetation of mesic sites.
+Occupies colder and wetter situations than the other alliances in the class, is less dominated by therophytes and has a relatively high proportion of bryophytes.
+Distribution concentrated in the central part of the ecoregion.
+Diagnostic species in the ecoregion: *Sagina procumbens*, *Bryum argenteum*.
+Correspondence with EUNIS habitat *V35 Trampled mesophilous grassland with annuals*.
 
@@
 *Caucalidion lappulae* Annual weed vegetation of cereal crops on base-rich soils.
-Adapted to more frequent and severe disturbances than the other alliances in the class (**Fig.
-4**).
-Distribution mostly limited to the more Mediterranean valleys south of the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Cyanus segetum*, *Papaver rhoeas*, *Trifolium arvense*, *Anacyclus clavatus*, *Valerianella eriocarpa*.
+Very high proportion of putative archaeophytes.
+Distribution mostly limited to the more submediterranean valleys south of the Cantabrian Mountains.
+Diagnostic species in the ecoregion: *Cyanus segetum*, *Papaver rhoeas*, *Trifolium arvense*, *Anacyclus clavatus*, *Valerianella eriocarpa*.
+Correspondence with EUNIS habitats *V11 Intensive unmixed crops* and *V13 Arable land with unmixed crops grown by low-intensity agricultural methods*.
 
@@
 *Scleranthion annui* Annual weed vegetation of cereal crops on base-poor soils.
-Occupies more acidic soils than the other alliances in the class, and flowers earlier (**Fig.
-4**).
-Wide distribution in the ecoregion, but specially in the west, where acidic bedrocks dominate (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Mibora minima*, *Rumex acetosella*.
+Occupies more acidic soils than the other alliances in the class, and flowers earlier.
+High proportion of putative archaeophytes.
+Wide distribution in the ecoregion, but specially in the west, where acidic bedrocks dominate.
+Diagnostic species in the ecoregion: *Mibora minima*, *Rumex acetosella*.
+Correspondence with EUNIS habitats *V11 Intensive unmixed crops*, *V12 Mixed crops of market gardens and horticulture*, *V13 Arable land with unmixed crops grown by low-intensity agricultural methods* and *V15 Bare tilled, fallow or recently abandoned arable land*.
 
@@
 *Oxalidion europeae* Annual weed vegetation of gardens and root crops.
-Occupies wetter, shadier and more nutrient-rich situations than the other alliances in the class, has a longer flowering season and a higher proportion of neophytes (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Stellaria media*, *Veronica persica*.
+Occupies wetter, shadier and nutrient-richer situations than the other alliances in the class, has a longer flowering season.
+High proportion of putative archaeophytes, and more neophytes than the other alliances in the class.
+Wide distribution in the ecoregion.
+Diagnostic species in the ecoregion: *Stellaria media*, *Veronica persica*.
+Correspondence with EUNIS habitats *V11 Intensive unmixed crops*, *V12 Mixed crops of market gardens and horticulture*, *V13 Arable land with unmixed crops grown by low-intensity agricultural methods* and *V15 Bare tilled, fallow or recently abandoned arable land*.
 
@@
 *Spergulo arvensis-Erodion cicutariae* Summer-annual C4 weed vegetation.
-Late-flowering alliance rich in neophytes and grasses, occurring in highly-disturbed and nutrient-rich soils (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Digitaria sanguinalis*, *Echinochloa crus-galli*, *Chenopodium album*, *Amaranthus hybridus* aggr.
+Late-flowering alliance rich in neophytes and grasses, occurring in highly-disturbed and nutrient-rich sites.
+Wide distribution in the ecoregion.
+Diagnostic species in the ecoregion: *Digitaria sanguinalis*, *Echinochloa crus-galli*, *Chenopodium album*, *Amaranthus hybridus*.
+Correspondence with EUNIS habitats *V11 Intensive unmixed crops*, *V12 Mixed crops of market gardens and horticulture* and *V13 Arable land with unmixed crops grown by low-intensity agricultural methods*.
 
-***Chenopodietea*** Winter-annual ruderal vegetation.
+***Chenopodietea*** Winter-annual fringe and ruderal vegetation.
 
@@
 *Allion triquetri* Geophyte-rich fringe vegetation.
-Occupies warmer and shadier situations than the other alliances in the class, has a higher proportion of geophytes and a lower proportion of therophytes, and is adapted to less frequent disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Urtica membranacea*, *Smyrnium olusatrum*, *Tradescantia fluminensis*.
+This vegetation occupies warmer, shadier and less frequently disturbed situations than other alliances in the class, and is comparatively taller and earlier-flowering.
+Has a higher proportion of geophytes and a lower proportion of therophytes.
+The proportion of neophytes is the highest in the class.
+Wide distribution in the ecoregion.
+Diagnostic species in the ecoregion: *Urtica membranacea*, *Smyrnium olusatrum*, *Tradescantia fluminensis*.
+Correspondence with EUNIS habitat *V37 Annual anthropogenic herbaceous vegetation*.
 
@@
 *Geranio pusilli-Anthriscion caucalidis* Winter-annual fringe vegetation.
-Occupies situations similar to the previous alliance, but in colder sites, and is dominated by therophytes instead of geophytes (**Fig.
-4**).
-Mostly distributed in the more Mediterranean valleys south of the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Geranium lucidum*, *Anthriscus caucalis*.
+Occupies situations similar to the previous alliance, but in colder sites, and is a shorter vegetation dominated by therophytes instead of geophytes.
+Mostly distributed in the submediterranean valleys south of the Cantabrian Mountains.
+Diagnostic species in the ecoregion: *Geranium lucidum*, *Anthriscus caucalis*.
+Correspondence with EUNIS habitat *V37 Annual anthropogenic herbaceous vegetation*.
 
@@
 *Chenopodion muralis* Low-growth winter-annual ruderal vegetation.
-Occupies colder situations than the other alliances in the class and more often disturbed sites, and has a longer and later flowering season (**Fig.
-4**).
-Sparse occurrence throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Malva neglecta*, *Urtica urens*.
+Short vegetation occupying colder situations than the other alliances in the class and more disturbed sites.
+Flowers longer and later.
+Has a high proportion of archaeophytes.
+Sparse occurrence throughout the ecoregion.
+Diagnostic species in the ecoregion: *Malva neglecta*, *Urtica urens*.
+Correspondence with EUNIS habitat *V37 Annual anthropogenic herbaceous vegetation*.
 
@@
 *Echio-Galactition tomentosae* Tall-herb winter-annual ruderal vegetation.
-Occupies drier, sunnier and less nutrient-rich situations than the other alliances in the class (**Fig.
-4**).
-Mostly distributed in the west of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Galactites tomentosus*, *Coleostephus myconis*, *Anisantha rigida*.
+Occupies drier, sunnier and less nutrient-rich situations than other alliances in the class.
+Mostly distributed in the west of the ecoregion.
+Diagnostic species in the ecoregion: *Galactites tomentosus*, *Coleostephus myconis*, *Anisantha rigida*.
+Correspondence with EUNIS habitat *V37 Annual anthropogenic herbaceous vegetation*.
 
@@
 *Linario polygalifoliae-Vulpion alopecuri* Ephemeral annual vegetation of disturbed coastal dunes.
-Occupies drier and more nutrient-poor situations than the other alliances in the class (**Fig.
-4**) and has a high number of coastal dune species.
-Isolated occurrence in the southern coasts of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Vulpia alopecuros*, *Malcolmia littorea*, *Crucianella maritima*, *Artemisia campestris* aggr.
+Occupies warm, dry and sunny coastal sites, and has a high number of coastal dune species.
+Isolated occurrence in the southern coast of the ecoregion.
+Diagnostic species in the ecoregion: *Vulpia alopecuros*, *Malcolmia littorea*, *Crucianella maritima*, *Artemisia campestris*.
+Correspondence with EUNIS habitat *V32 Mediterranean subnitrophilous annual grassland*.
 
@@
 *Sisymbrion officinalis* Summer-annual ruderal vegetation.
-Rich in therophytes and adapted to severe disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Hordeum murinum*, *Sisymbrium officinale*.
+Rich in therophytes and archaeophytes and adapted to severe disturbances.
+Wide distribution in the ecoregion.
+Diagnostic species in the ecoregion: *Hordeum murinum*, *Sisymbrium officinale*.
+Correspondence with EUNIS habitats *V15 Bare tilled, fallow or recently abandoned arable land* and *V37 Annual anthropogenic herbaceous vegetation*.
 
-***Bidentetea*** Summer-annual pioneer vegetation of temporarily-flooded sites.
+***Bidentetea*** Summer-annual pioneer vegetation of temporarily flooded sites.
 
 16.
-*Bidention tripartitae* Summer-annual pioneer vegetation of temporarily-flooded sites.
-Occupies nutrient-richer sites than the other alliance in the class and has a higher proportion of therophytes (**Fig.
-4**).
-Dispersed throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Persicaria hydropiper*, *Bidens frondosus*, *Lythrum salicaria*.
+*Bidention tripartitae* Summer-annual pioneer vegetation of temporarily flooded sites.
+Occupies nutrient-richer sites than the other alliance in the class and has a higher proportion of therophytes and neophytes.
+Dispersed throughout the ecoregion.
+Diagnostic species in the ecoregion: *Persicaria hydropiper*, *Bidens frondosus*, *Lythrum salicaria*.
+Correspondence with EUNIS habitat *Q61 Periodically exposed shore with stable, eutrophic sediments with pioneer or ephemeral vegetation*.
 
 17.
-*Paspalo-Agrostion semiverticillati* Summer-annual pioneer vegetation of temporarily-flooded and warm sites.
-Occupies warmer and wetter situations than the other alliance in the class, has a higher proportion of neophytes and has a longer flowering season (**Fig.
-4**).
-Isolated occurrences in the south of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Paspalum distichum*, *Cyperus longus*, *Schoenoplectus lacustris*.
+*Paspalo-Agrostion semiverticillati* Summer-annual pioneer vegetation of temporarily flooded and warm sites.
+Occupies wetter and more frequently disturbed sites than the other alliance in the class and has a longer flowering season.
+Isolated occurrences in the south of the ecoregion.
+Diagnostic species in the ecoregion: *Paspalum distichum*, *Cyperus longus*, *Schoenoplectus lacustris*.
+Correspondence with EUNIS habitat *R55 Lowland moist or wet tall-herb and fern fringe*.
 
@@
 *Convolvulo arvensis-Agropyrion repentis* Semiruderal grasslands.
-Occupies wetter and shadier situations than the other alliances in the class, and has a higher proportion of geophytes and grasses (**Fig.
-4**).
-Isolated occurrence in the south of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Poa compressa*, *Elytrigia repens*, *Potentilla reptans*, *Valerianella locusta*.
+Occupies wetter, shadier and nutrient-poorer situations than the other alliances in the class, and has a higher proportion of geophytes, grasses and native species.
+Isolated occurrence in the south of the ecoregion.
+Diagnostic species in the ecoregion: *Poa compressa*, *Elytrigia repens*, *Potentilla reptans*, *Valerianella locusta*.
+Correspondence with EUNIS habitat *V38 Dry perennial anthropogenic herbaceous vegetation*.
 
@@
 *Carduo carpetani-Cirsion odontolepidis* Thistle ruderal vegetation of warm sites.
-Occupies warmer situations than the other thistle alliance in the class (**Fig.
-4**).
-Mostly distributed in the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Onopordum acanthium*.
+Occupies warmer situations than the other thistle alliance in the class and has a high proportion of archaeophytes.
+Mostly distributed in the Cantabrian Mountains.
+Diagnostic species in the ecoregion: *Onopordum acanthium*.
+Correspondence with EUNIS habitat *V38 Dry perennial anthropogenic herbaceous vegetation*.
 
@@
 *Cirsion richterano-chodati* Thistle ruderal vegetation of cold sites.
-Occupies colder situations than the other thistle alliance in the class (**Fig.
-4**).
-Mostly distributed in the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Cirsium eriophorum*, *Carduus carpetanus*, *Carduus nutans* aggr.
+Occupies colder situations than the other thistle alliance in the class.
+Mostly distributed in the Cantabrian Mountains.
+Diagnostic species in the ecoregion: *Cirsium eriophorum*, *Carduus carpetanus*, *Carduus nutans*.
+Correspondence with EUNIS habitat *V38 Dry perennial anthropogenic herbaceous vegetation*.
 
@@
 *Dauco-Melilotion* Biennial ruderal vegetation.
-Occupies wetter situations than the two thistle alliances in the class, and has a higher proportion of therophytes (**Fig.
-4**).
-Sparse occurrence throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Helminthotheca echioides*, *Daucus carota*, *Melilotus albus*, *Foeniculum vulgare*.
+Occupies wetter situations than the two thistle alliances in the class and has a higher proportion of neophytes.
+Occurrence throughout the ecoregion.
+Diagnostic species in the ecoregion: *Helminthotheca echioides*, *Daucus carota*, *Melilotus albus*, *Foeniculum vulgare*.
+Correspondence with EUNIS habitat *V38 Dry perennial anthropogenic herbaceous vegetation*.
 
-***Epilobietea angustifolii*** Perennial ruderal vegetation of mesic to wet sites.
+***Epilobietea angustifolii*** Perennial fringe and ruderal vegetation of mesic to wet sites.
 
@@
 *Geo urbani-Alliarion officinalis* Low-herb short-lived semiruderal and fringe vegetation.
-Occupies shadier and less disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Urtica dioica*.
+Occupies shadier and less frequently disturbed situations than the other alliances in the class and has a higher proportion of therophytes and shorter plants.
+Wide distribution in the ecoregion.
+Diagnostic species in the ecoregion: *Urtica dioica*.
+Correspondence with EUNIS habitat *V39 Mesic perennial anthropogenic herbaceous vegetation*.
 
@@
 *Arction lappae* Low-herb short-lived ruderal vegetation.
-Occupies sunnier and more frequently-disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants (**Fig.
-4**).
-Distribution concentrated in the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Oxybasis rubra*, *Senecio duriaei*.
+Occupies sunnier and more frequently disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants.
+Distribution concentrated in the Cantabrian Mountains.
+Diagnostic species in the ecoregion: *Oxybasis rubra*, *Senecio duriaei*.
+Correspondence with EUNIS habitat *V39 Mesic perennial anthropogenic herbaceous vegetation*.
 
@@
 *Balloto-Conion maculati* Tall-herb perennial ruderal vegetation.
-Occupies warmer situations than the other alliances in the class (**Fig.
-4**).
-Distribution concentrated in the Cantabrian Mountains and the east of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Sambucus ebulus*.
+Occupies warmer situations than the other alliances in the class.
+Has a high proportion of archaeophytes.
+Distribution concentrated in the Cantabrian Mountains and the east of the ecoregion.
+Diagnostic species in the ecoregion: *Sambucus ebulus*.
+Correspondence with EUNIS habitat *V39 Mesic perennial anthropogenic herbaceous vegetation*.
 
@@
 *Aegopodion podagrariae* Tall-herb perennial vegetation of forest margins and clearings.
-Occupies shadier and less disturbed situations than the other alliances in the class (**Fig.
-4**).
-Sparse occurrence throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Anthriscus sylvestris*, *Heracleum sphondylium*, *Galium aparine*.
+Occupies shadier and less disturbed situations than the other alliances in the class.
+Sparse occurrence throughout the ecoregion.
+Diagnostic species in the ecoregion: *Anthriscus sylvestris*, *Heracleum sphondylium*, *Galium aparine*.
+Correspondence with EUNIS habitats *R55 Lowland moist or wet tall-herb and fern fringe* and *V39 Mesic perennial anthropogenic herbaceous vegetation*.
 
@@
 *Epilobion angustifolii* Tall-herb perennial vegetation of forest margins and clearings in acidic soils.
-Occupies colder, nutrient-poorer, more acidic and less disturbed situations than the other alliances in the class (**Fig.
-4**).
-Distributed towards the west of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Epilobium angustifolium*, *Digitalis purpurea*, *Luzula lactea*.
+Occupies colder, nutrient-poorer, more acidic and less disturbed situations than the other alliances in the class.
+Distributed towards the west of the ecoregion.
+Diagnostic species in the ecoregion: *Epilobium angustifolium*, *Digitalis purpurea*, *Luzula lactea*.
+Correspondence with EUNIS habitats *R54 Pteridium aquilinum vegetation* and *R57 Herbaceous forest clearing vegetation*.
 
 27.
-*Cynancho-Convolvulion sepium* Tall-herb vegetation of nutrient-rich riparian habitats.
-Occupies wetter situations and less disturbed situations than the other alliances in the class, and has a later flowering season (**Fig.
-4**).
-Dispersed throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Eupatorium cannabinum*, *Angelica sylvestris*, *Picris hieracioides*.
+*Cynancho-Convolvulion sepium* Tall-herb fringe vegetation of nutrient-rich riparian habitats.
+Occupies wetter and less disturbed situations than the other alliances in the class and has a later flowering season.
+Dispersed throughout the ecoregion.
+Diagnostic species in the ecoregion: *Eupatorium cannabinum*, *Angelica sylvestris*, *Picris hieracioides*.
+Correspondence with EUNIS habitat *R55 Lowland moist or wet tall-herb and fern fringe*.
 
 28.
-*Senecionion fluviatilis* Tall-herb vegetation of nutrient-rich river banks and ditches.
-Occupies warmer and wetter situations than the other alliances in the class, flowers later, has taller plants and is highly dominated by neophytes and geophytes (**Fig.
-4**).
-Isolated occurrence in the western coast of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Arundo donax*, *Silene latifolia*.
+*Senecionion fluviatilis* Tall-herb vegetation of nutrient-rich riverbanks and ditches.
+Occupies warmer and wetter situations than the other alliances in the class, flowers later, has taller plants and has a high proportion of neophytes and geophytes.
+Isolated occurrence in the western coast of the ecoregion.
+Diagnostic species in the ecoregion: *Arundo donax*, *Silene latifolia*.
+Correspondence with EUNIS habitat *R55 Lowland moist or wet tall-herb and fern fringe*.
 
@@
 
-The major driver of variation in trampled community composition is soil moisture [@RN5434].
+Our analysis supports the existence of 28 anthropogenic alliances in the vegetation of the Iberian Atlantic ecoregion.
+While the core of this vegetation diversity is made up of temperate European alliances, these are enriched by the occurrence of eight Mediterranean alliances (*Galio valantiae-Parietarion judaicae*, *Polycarpion tetraphylli*, *Allion triquetri*, *Chenopodion muralis*, *Echio-Galactition tomentosae*, *Paspalo-Agrostion semiverticillati*, *Balloto-Conion maculati* and *Cynancho-Convolvulion sepium*); plus three alliances that are endemic of the Iberian coasts (*Linario polygalifoliae-Vulpion alopecuri*) and mountains (*Carduo carpetani-Cirsion odontolepidis*, *Cirsion richterano-chodati*) [@RN4698].
+Seven of these alliances had been reported as absent or uncertain in the ecoregion (*Caucalidion lappulae*, *Chenopodion muralis*), or only present in the Portuguese sector of the ecoregion (*Spergulo arvensis-Erodion cicutariae*, *Allion triquetri*, *Echio-Galactition tomentosae*, *Carduo carpetani-Cirsion odontolepidis*, *Cynancho-Convolvulion sepium*) by the recently published distribution maps of vegetation alliances in Europe [@RN5371].
+Our dataset supports that six of these alliances are relatively well distributed in the ecoregion, while *Caucalidion lappulae* only has isolated occurrences.
+On the other hand, our dataset does not support the occurrence in the ecoregion of nine Mediterranean alliances (*Roemerion hybridae*, *Diplotaxion erucoidis*, *Euphorbion prostratae*, *Alysso granatensis-Brassicion barrelieri*, *Hordeion murini*, *Taeniathero-Aegilopion geniculatae*, *Parietarion lusitanico-mauritanicae*, *Onopordion castellani*, *Silybo mariani-Urticion piluliferae*) and one temperate alliance (*Fragarion vescae*) that had been cited as potentially occurring here [@RN5363; @RN5008; @RN5365; @RN4698; @RN5371].
+For habitat classification purposes, we propose that this vegetation diversity can be summarized into 15 EUNIS habitat types: one constructed habitat (*J2.5*), ten vegetated human-made habitats (*V11*, *V12*, *V13*, *V15*, *V32*, *V35*, *V34*, *V37*, *V38*, *V39*), three habitats of woodland fringes and clearings and tall forb stands (*R54*, *R55*, *R57*) and one habitat of periodically-exposed shores (*Q61*).
+In general, we have followed the correspondence between alliances and EUNIS habitats proposed by the EUNIS expert system [@RN5061], except for *Echio-Galactition tomentosae* which, in the context of the ecoregion, we believe is more adequately placed in the habitat *V37 Annual anthropogenic herbaceous vegetation* rather than in the habitat *V33 Dry Mediterranean land with unpalatable non-vernal herbaceous vegetation*.
 
-In medieval Switzerland, summer and winter weeds communities could not be distinguished [@RN5431].
+Since our goal was to match the classification of the ecoregional classification to the EuroVegChecklist [@RN4698], we refrained from merging recognized alliances or changing their assignation to superior sintaxa.
+However, there are some cases that could deserve further scrutiny.
+For example, the two wall alliances (*Galio valantiae-Parietarion judaicae* and *Cymbalario-Asplenion*) had a large overlap in their geographic distribution, their floristic composition, their proportion of alien species, their community traits and their ecological preferences.
+These two alliances are well recognized in the European literature, and they are generally interpreted as representing the Mediterranean (*Galio valantiae-Parietarion judaicae*) and temperate (*Cymbalario-Asplenion*) versions of wall vegetation [@RN5369; @RN5367].
+It is therefore not surprising that in boundary ecoregions - such as the Iberian Atlantic territories - these two alliances coexist and show a high overlap.
+Also, there was overlap between the alliances *Geo urbani-Alliarion officinalis* and *Aegopodion podagrariae*, two alliances that can occupy ruderal to semi-natural nitrophilous fringes, and which are mostly differentiated by dominant vegetation height, as was supported by our dataset.
 
-The class *Digitario sanguinalis-Eragrostietea minoris* includes thermophilic vegetation dominated by C4 plants that occupies arable land as well as trampled and ruderal sites [@RN5426].
+We found a percentage of 22 % non-native species in the anthropogenic vegetation of the Iberian Atlantic ecoregion.
+This percentage is intermediate between that found in urban areas of temperate Europe and North America (30-50 %) [@RN5498; @RN5499; @RN5500; @RN5079] and Italy (12-26 %) [@RN5501].
+A survey of anthropogenic vegetation in the north-west Balkans found 13 % non-native plants, with more neophytes than archaeophytes [@RN5429].
+A similar study in the Czech Republic found 32 % archaeophytes and 7 % neophytes [@RN5435].
+Taken together, these results suggest that the Iberian Atlantic anthropogenic vegetation shows an invasion status that is intermediate between southern and central-northern Europe, which is in agreement with the biogeographical position of our ecoregion.
+Previous studies also suggested that the invasion status was higher in early successional anthropogenic communities [@RN5423], and our results confirmed this as the proportion of native species tended to be higher in the perennial ruderal classes *Artemisietea vulgaris* and *Epilobietea angustifolii*.
+The classes with more alien species also tended to be those adapted to more frequent and more severe disturbances (*Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Sisymbrietea*), also in agreement with previous research [@RN5435].
+Of the alien species, archaeophytes were a higher proportion in the vegetation associated to arable habitats (*Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Sisymbrietea*), in agreement with the ancient origin of archaeophytes during the development of agriculture [@RN5494; @RN5490].
+On the other hand, neophytes tended to be in higher proportions in vegetation types associated to wet conditions (*Bidentetea*), as has been found in other European regions [@RN5497].
 
-Anthropogenic vegetation composition is sensitive to micro-scale management decisions, such as the ownership of domestic dogs [@RN5438].
+The traits of the Iberian Atlantic anthropogenic communities agreed with those found in other European regions [@RN5436; @RN2297], with more annual species in habitats associated to arable land and more perennials in ruderal habitats, and a general dominance of therophytes and hemicryptophytes in the species pool [@RN5428].
+We found that the main drivers of variation in anthropogenic community composition was a gradient from open-dry to shady-wet sites, supporting the importance of the moisture gradient in shaping anthropogenic vegetation [@RN5434].
+Mesic and moist human-made habitats of the Iberian Atlantic territories harbor perennial communities, some of which are especially rich in native species and have semi-natural characters.
+At the same time, the wetter habitats have some of the most invaded communities.
 
-Neophytes were more represented in early successional anthropogenic communities [@RN5423].
+# Conclusions
 
-Arable land plants were mainly therophytes and alien species; while ruderal habitats were richer in perennials and C strategists [@RN5436].
-
-Compare proportion of neophytes to other areas [@RN5429; \@ @RN5435]
-
-Neophytes have no special preferences for local habitat conditions and their highest proportion was found mainly in disturbed habitats at low elevations. [@RN5435]
-
-Therophytes and hemicryptophytes were most abundant in the life form spectra [@RN5428]
-
-Segetal vegetation of Palestine [@RN5494].
-
-History of synanthropic veg in czech [@RN5496]
-
-# Data availability
-
-The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/manmade>.
-Upon publication, a version of record of the repository will be deposited in Zenodo.
+The anthropogenic vegetation of the Iberian Atlantic ecoregion is home to 1,162 vascular plant taxa.
+This is approximately one third of the whole ecoregional species pool represented in the SIVIM database [@RN5374], or one fifth of the Iberian flora [@RN5503].
+More than half of this synanthropic diversity (615 taxa) occurs in mesic to wet perennial ruderal vegetation, highlighting the potential of some anthropogenic habitats to act as source and supply of biodiversity in human-dominated landscapes [@RN5446; @RN5445].
+These results suggest three main objectives for the sustainable management of human-made habitats: (1) focusing efforts on plant invasion management in anthropogenic plant communities of wet habitats; (2) conserving examples of archaeophyte-rich annual weed communities, especially those linked to traditional agricultural practices [@RN5502]; and (3) conserving and promoting mesic ruderal and fringe vegetation as a biodiversity refuge in urban and peri-urban areas.
 
@@
 
+# Supplementary Information
+
+**Appendix S1** Checklist of anthropogenic sintaxa that could be present in the Iberian Atlantic ecoregion according to the literature.
+Sintaxa names are provided as recorded by the original authors of the vegetation plot.
+The file indicates if the presence of the sintaxon in the Iberian Atlantic ecoregion was validated or not by our analysis.
+
+**Appendix S2** Original publications providing the vegetation plots stored in SIVIM and used in our analysis.
+
+**Appendix S3** Characteristic species combinations of the anthropogenic vegetation alliances in a spreadsheet format, divided into diagnostic, constant and dominant species.
+
+**Appendix S4** Revised checklist of vegetation associations described in the regional literature and belonging to the 28 alliances present in the Iberian Atlantic anthropogenic vegetation.
+
 # Figures
 
-```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Figure 1.
-Anthropogenic vegetation alliances of the Iberian Atlantic ecoregion.
-Each dot is a vegetation plot.
-Dot colors indicate the vegetation class."}
+```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Anthropogenic vegetation alliances of the Iberian Atlantic ecoregion. Each dot is a vegetation plot. Dot colors indicate the vegetation class."}
 knitr::include_graphics("../results/figures/F1 - maps.png")
@@
 
-```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Figure 2.
-Floristic composition of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Floristic diversity of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
 Biplots produced by Principal Component Analysis (PCA) with Hellinger transformation.
@@
 
-```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Figure 3.
-Proportion of neophytes, archaeophytes and native species in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of neophytes, archaeophytes and native species in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
 Each bar represents the plot-level average proportion of neophytes (dark shade), archaeophytes (medium shade) and natives (light shade) in each vegetation alliance.
@@
 
-```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
-Proportion of therophytes, geophytes and other life forms in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of therophytes, geophytes and other life forms in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
 Each bar represents the plot-level average proportion of therophytes (dark shade), geophytes (medium shade) and other life forms (light shade) in each vegetation alliance.
@@
 
-```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Figure 5.
-Vegetation heigh and flowering phenology of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
-Boxplots show the community-weighted mean for plant heigh, month of peak flowering and length of the flowering season for each vegetation alliance.
-Colors indicate the vegetation class."}
+```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Vegetation heigh and flowering phenology of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Boxplots show the community-weighted mean for plant heigh, median month of flowering and length of the flowering season for each vegetation alliance.
+Colors indicate the vegetation class.
+Alliance numbers are the same as listed in the text."}
 knitr::include_graphics("../results/figures/F5 - height-flowering.png")
@@
 
-```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Figure 6.
-Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
 Boxplots show the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity, calculated for each vegetation alliance.
-Colors indicate the vegetation class."}
+Colors indicate the vegetation class.
+Alliance numbers are the same as listed in the text."}
 knitr::include_graphics("../results/figures/F6 - indicators.png")
@@
 
-```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Figure 7.
-Pattern of variation in the ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
 Biplot produced by Principal Component Analysis (PCA) of the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.

### 2024-04-26 src/12-height-flowering.R

@@ -90,7 +90,7 @@ cwms %>%
                                 "Grazing.Pressure")) %>%
   mutate(Trait = fct_recode(Trait, 
                             "(A) Vegetation height (cm)" = "heightmax",
-                            "(B) Flowering peak (month)" = "floweringmean",
+                            "(B) Median flowering month" = "floweringmean",
                             "(C) Flowering length (months)" = "floweringwidth",
                             "Neophytes (% of total cover)" = "neophyte",
                             "Geophytes (% of total cover)" = "geophyte",

### 2024-04-26 src/13-indicators.R

@@ -230,6 +230,9 @@ plotmeans %>%
                             "Polygono-Poetea" = "Polygono-Poetea annuae")) ->
   inds
 
+inds %>% group_by(Class) %>% summarise(PC1 = mean (Dim.1), PC2 = mean(Dim.2)) %>% arrange(PC1)
+inds %>% group_by(Class) %>% summarise(PC1 = mean (Dim.1), PC2 = mean(Dim.2)) %>% arrange(PC2)
+
 pca2$var$coord %>%
         data.frame %>%
         rownames_to_column(var = "Variable") %>%

### 2024-04-26 src/7-spp-numbers.R

@@ -1,10 +1,10 @@
 library(tidyverse)
 
+### General species diversity
+
 read.csv("data/header.csv", fileEncoding = "latin1") %>% 
   filter(Alliance != "Noise") %>%
-  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  group_by(Analysis.Names)
-
+  merge(read.csv("data/species.csv", fileEncoding = "latin1"))
 
 read.csv("data/header.csv", fileEncoding = "latin1") %>% 
   filter(Alliance != "Noise") %>%
@@ -13,76 +13,215 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
 
 read.csv("data/header.csv", fileEncoding = "latin1") %>% 
   filter(Alliance != "Noise") %>%
-  group_by(Area) %>% tally %>% arrange(-n)
-
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>% 
-  filter(Alliance != "Noise" & Area == "20") %>%
   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  select(SIVIMID, Analysis.Names) %>%
-  unique %>%
-  group_by(SIVIMID) %>% tally %>% arrange(-n)
+  select(Class, Analysis.Names) %>%
+  unique() %>%
+  group_by(Class) %>% tally %>% arrange(-n)
 
-read.csv("data/header.csv", fileEncoding = "latin1") %>% 
-  filter(Alliance != "Noise" & Area == "20") %>%
-  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  select(SIVIMID, Analysis.Names) %>%
-  unique %>%
-  group_by(SIVIMID) %>% tally %>% arrange(n)
-
-read.csv("data/header.csv", fileEncoding = "latin1") %>% 
-  filter(Alliance != "Noise" & Area == "20") %>%
-  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  select(SIVIMID, Analysis.Names) %>%
-  unique %>%
-  group_by(SIVIMID) %>% tally %>% pull(n) %>% mean
-
+### Species origin
 
 read.csv("data/header.csv", fileEncoding = "latin1") %>% 
   filter(Alliance != "Noise") %>%
   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
-  select(SIVIMID, Analysis.Names) %>%
[…]

### 2024-04-26 README.Rmd
readme

--- 2024-04-10 doc/manuscript.Rmd
@@
 ---
-title: "Anthropogenic plant communities of the Iberian Atlantic ecoregion"
-output: 
-  bookdown::word_document2:
-    reference_docx: template.docx
-    number_sections: false
-bibliography: refs.bib
-csl: journal-of-vegetation-science.csl
-link-citations: yes
-always_allow_html: true
+title: "README.md"
+output: github_document
 ---
@@
 
-```{r message = FALSE, echo = FALSE, warning = FALSE}
-knitr::knit_hooks$set(inline = function(x) {
-  prettyNum(x, big.mark = ",")
-})
-```
+[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.6077618.svg)](https://doi.org/10.5281/zenodo.6077618)
 
-# Introduction
+![Antigua fábrica de La Cerámica Gijonesa](ceramica.JPG)
 
-Anthropogenic vegetation is the set of plant communities that occur as a direct consequence of human activities, and is chiefly composed by the weed vegetation of arable fields and the ruderal vegetation of human settlements and their surroundings [@RN5441].
-The anthropogenic communities of human-made habitats make up an important part of the supply of natural elements in urban and peri-urban habitats [@RN5445].
-In densely-populated city landscapes, the plant communities that colonize urban vacant lots [@RN2300] are a useful biodiversity resource with high ecological and societal potential [@RN5446].
-Anthropogenic vegetation cover can also impact regional-level extreme climatic events [@RN5440].
-Furthermore, many synanthropic plants can accumulate heavy metals in their biomass [@RN5433], making them valuable candidates for nature-based solutions such as bioremediation and restoration of degraded landscapes [@RN5447].
+# Anthropogenic plant communities of the Iberian Atlantic ecoregion
 
-The anthropogenic vegetation of temperate Europe is formed by an original pool of native species favored by humans, but this pool has been constantly enriched by the arrival of alien species: first archaeophytes starting in the Neolithic and peaking in the Bronze Age; and then neophytes from the Modern Age (after 1.500 A.D.) to the present day [@RN5439].
-Synanthropic archaeophytes include many Mediterranean taxa that started their expansion into temperate areas during the Mid-Holocene warm period, but it is difficult to disentangle to what extent this expansion was favored by human activities and/or ongoing climatic changes [@RN5432; @RN5443], especially in southern European areas close to the transition between the Mediterranean and the temperate climates [@RN5444].
-Meanwhile, synanthropic neophytes include many American, African and Asian taxa that arrived as a consequence of the global trade networks established by the Western European colonial powers [@RN5451].
-In the last century, anthropogenic vegetation has shown a decrease in species richness and diversity [@RN5423], as rare synanthropic species have become rarer, and neophytes have increased their abundance at the expense of natives and archaeophytes [@RN5441].
+This repository stores all the information related to the manuscript [*Anthropogenic plant communities of the Iberian Atlantic ecoregion*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
 
-Antropogenic vegetation is likely to be similar within ecological regions (ecoregions, hereafter), since they are assumed to encompass areas with a similar biogeographic history and recurrent local ecosystems [@RN5491].
-However, ecoregional-scale studies on the diversity of antropogenic vegetation are scarce in the literature.
-The Iberian Atlantic ecoregion (a.k.a.
-Cantabrian Mixed Forests ecoregion; NW Iberian Peninsula) has a long history of human habitation dating back to the Cantabrian Upper Paleolithic [@RN5450].
-This, together with its transitional position at the border between the temperate and Mediterranean climatic zones of Europe [@RN4700], suggests a long regional history for archaeophyte-rich anthropogenic plant communities, likely supporting a current high diversity of anthropogenic plant communities with both temperate and Mediterranean floristic elements [@RN5008].
-Historical trade links with America and Asia, together with the warm and humid temperate climate, have made the region a hotspot for neophyte invasions [@RN5492; @RN5493].
-Recent post-industrial land-use changes have left large areas with abandoned industrial sites (i.e.
-brownfields) whose management and restoration requires an understanding of anthropogenic communities able to colonize or restore them [@RN5448; @RN5449].
-Although there is a long tradition of studying weed and ruderal communities in the ecoregion [@RN5364; @RN5361; @RN5359; @RN5360] there is a current need for an ecoregional-level synthesis that revises and updates the classification and updates in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
-In this article we have performed such a synthesis with two main objectives: (1) provide an updated classification of anthropogenic plant communities in the Iberian Atlantic ecoregion, and (2) describe the main patterns of vegetation diversity in terms of floristic origin (natives, archaeophytes, neophytes); life forms; functional traits (plant height and flowering time); and ecological indicator values (temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity).
+## Contents
 
-# Methods
+This repository is organised following the advice of [Wilson et al. 2017](https://doi.org/10.1371/journal.pcbi.1005510) for recording and storing research projects.
 
-## Study ecoregion
+The following materials are available in the folders of this repository:
 
-We studied the anthropogenic plant communities of the Iberian Atlantic ecoregion, i.e.
-the territories with a temperate climate in the north-western Iberian Peninsula.
-Our study ecoregion broadly corresponds with the Cantabrian Mixed Forests ecoregion *sensu* Olson *et al.* [-@RN4667], with the Iberian part of the European Atlantic province *sensu* Rivas-Martínez *et al.* [-@RN4699] and with the Iberian section of the Atlantic biogeographical region of the European Environmental Agency (https://www.eea.europa.eu/data-and-maps/figures/biogeographical-regions-in-europe-2).
-To define in detail the study ecoregion we did a fine-scale revision of its geographic borders at high spatial resolution using the most recent biogeographical map of the Iberian Atlantic territories [@RN5372].
-The revised Iberian Atlantic ecoregion encompasses 79,891 km^2^ of Portugal, Spain and France.
-Its climate is temperate with average monthly temperatures ranging from 6 to 20°C, and with total annual precipitation between 1,100 and 2,600 mm, with some areas having warm-dry summers but without the distinctive 2-month drought of the Mediterranean climate [@RN5373].
-Nonetheless, the Iberian Atlantic ecoregion is located at the transition between the temperate and the Mediterranean climates in southern Europe.
+* `data` Data files including the vegetation plots.
+* `doc` Files to create the [manuscript](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md) using `Rmarkdown`.
+* `results` Output of the `R` scripts, including figures and model outputs.
+* `src` Scripts in `R` language used to clean the raw data and perform the analyses of the manuscript.
 
-## Definition of anthropogenic vegetation
+## Abstract
 
-To define our target vegetation, we followed the definitions of anthropogenic vegetation in the recent classification of the vegetation of Europe by Mucina *et al.* [-@RN4698].
-In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Bidentetea*.
-For a complete synthesis, we added the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [-@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls in our ecoregion.
-For the nomenclature of sintaxa we follow Mucina *et al.* [-@RN4698] for alliances and higher ranks, and Rivas-Martínez *et al.* [-@RN4697] for associations.
+* *Questions* Which is the diversity of anthropogenic plant communities at the ecoregional level in a transitional territory between temperate and Mediterranean Europe?
+How are these communities composed by native plants, archaeophytes and neophytes?
+What are the main environmental and disturbance drivers of the diversity in anthropogenic vegetation?
 
-## Vegetation data
+* *Location* Iberian Atlantic ecoregion, Iberian Peninsula, south-western Europe.
 
-We retrieved anthropogenic vegetation plots from SIVIM, the Iberian and Macaronesian Vegetation Information System [@RN5374].
-To identify plots as anthropogenic we followed two criteria: (1) We used the expert system created by Chytrý *et al.* [-@RN5061] to classify vegetation plots into EUNIS pan-European habitat types.
-This expert system uses the floristic composition of plots to assign them to habitat types depending on the presence and abundance of characteristic species.
-We used the habitats assigned by the expert system to retrieve plots that had been assigned to any habitat related to the vegetation classes we had previously defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation) (R55 and R57 include communities that are classified as "grasslands" by EUNIS but as "anthropogenic" by by Mucina *et al.* [-@RN4698]).
-(2) We also retrieved vegetation plots from SIVIM if they had been classified by the original authors of the relevé into any sintaxa belonging to our target anthropogenic vegetation classes (**Supplementary 1**).
-This produced an initial pool of 3,160 vegetation plots from SIVIM, to which we added 89 vegetation plots of urban or peri-urban plant communities sampled by us or extracted from local literature not in SIVIM [@RN5377; @RN5378].
-After initial exploration of this dataset using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379], we identified 652 outlier plots which either did not belong to the target anthropogenic classes or to the geographical area of the ecoregion.
-The majority of these plots corresponded to coastal plant communities and had been misclassified by the expert system.
-We removed these, leaving a dataset of 2,508 vegetation  plots for data analyses: 2,419 plots from SIVIM, originally recorded in 89 publications (**Supplementary 2**); plus 89 plots added by us.
+* *Methods* We compiled a vegetation database of 2,508 synanthropic plots and revised the regional checklist of anthropogenic alliances.
+We used modified TWINSPAN and semi-supervised classification to classify the plots into the revised alliances.
+We determined the proportion of natives, archaeophytes and neophytes.
+We described the alliances in terms of vegetation height, flowering phenology and ecological requirements for temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.
 
-## Statistical analyses
+* *Results* We classified 2,081 vegetation plots into 28 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*).
+The plots included 1,162 plant taxa: 78% natives, 15% archaeophytes and 7% neophytes.
+The main ecological gradients behind this diversity went from dry-sunny to moist-shady habitats, followed by a gradient of low to high disturbance frequency and severity.
 
-We performed all data analyses with R version 4.3.1 [@RN5387], using the R package 'tidyverse' [@RN4662] for data processing and visualization.
-The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
-We homogenized all taxon names in the analyses and the article using Euro+Med [@RN5044] or Plants of the World Online for taxa not in Euro+Med [@RN5376].
+* *Conclusions* The human-made habitats of the Iberian Atlantic ecoregion host one third of the ecoregional species pool and one fifth of the Iberian flora.
+Wet anthropogenic habitats are most affected by neophyte invasions.
+Annual vegetation of arable lands is a refuge for archaeophytes.
+Mesic perennial ruderal vegetation is specially rich in native species and can be a biodiversity asset in urban and peri-urban settings.
 
-[Definition of target alliances and training dataset.]{.underline} As a previous step before classification, we prepared a revised checklist of potential anthropogenic alliances that could be present in our study ecoregion, based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
-Then, we filtered our dataset to extract a subset with the vegetation plots which had been assigned by the authors to any of these alliances.
-We used TWINSPAN to conduct a preliminary numerical classification of this subset, attempting to match the plots into the *a priori* defined checklist of alliances.
-We used 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity, and the R package 'twinspanR' [@RN5380] to do the TWINSPANs.
-To start, we did a TWINSPAN classification into 15 clusters, followed by separate TWINSPANS within each cluster.
-This preliminary TWINSPAN allowed us to (1) validate which alliances were present in the study ecoregion; and (2) prepare a training dataset for semi-supervised classification (see below).
+## Citation
 
-[Semi-supervised classification.]{.underline} Next, we applied a semi-supervised classification to the whole dataset.
-Semi-supervised classification uses a training subset of *a priori* classified vegetation plots to classify a secondary subset of unclassified plots [@RN5383].
-Our training subset included all the plots that (1) had been classified by the authors of the relevé into any of the alliances that we had validated during the previous TWINSPANs, and (2) had also been matched by us into these alliances during our TWINSPANs.
-The secondary subset included all the plots which did not meet these two criteria.
-Our goal was to refine and not to simply extend the classification, so we applied the semi-supervised classification to the whole dataset, i.e.
-we allowed plots from the training subset to be re-assigned to other groups during the classification.
-In addition, semi-supervised classification can create new groups to place data points that do not match the already existing *a priori* groups.
-Furthermore, we applied a noise clustering fuzzy algorithm, which allows the plots to be classified either into the *a priori* groups or into a 'noise' group which includes outliers and transitional plots [@RN5382].
-We set the fuzziness coefficient to a low value (m = 1) to accommodate a high number of transitional plots; and we set the distance to the noise class to d = 1.
-We performed this analysis with the R package 'vegclust' [@RN5383].
-We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package 'RVAideMemoire' [@RN5381] to test the significance of the final vegetation groups; along with Detrended Correspondence Analysis (DCA) as implemented in the R package 'vegan' [@RN5386] to visualize the relationships between the groups.
+Please cite the repository, datasets and article as:
 
-[Characteristic species and EUNIS habitat regionalization.]{.underline} We calculated sets of characteristic species [@RN5061] for each of the final vegetation groups.
-We defined dominant species as species with more than 25% cover in at least 5% of the vegetation plots of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species with an 'IndVal' value with a p-value lower than 0.05, as calculated with the R package 'labdsv' [@RN2311] using 1.000.000 iterations.
-Additionally, we assigned to each group a regionalized level 4 EUNIS habitat code, using pre-existing codes when they existed (https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp), and creating new codes when not.
-
-[Community traits and ecological requirements.]{.underline} To describe the resulting vegetation groups, we compiled a dataset of species traits and ecological preferences for the species pool of the classified vegetation plots.
-For each species, we extracted from *Flora iberica* [-@RN3243] its maximum height, its median month of flowering, the length of the flowering period, and whether it was a neophyte or not.
-From FloraVeg.EU [-@RN5385] we extracted the species’ life forms and kept 'therophyte' and 'geophyte' for further analysis since these two categories had the largest contribution to variance as per an exploratory Principal Component Analysis (PCA) performed using the R package 'FactoMineR' [@RN3166].
-We also collected the ecological indicator values [@RN5384] and the disturbance indicator values [@RN5101] of each species.
-In the case of the ecological and disturbance indicators, we assigned values to species lacking them by (1) calculating weighted average values of the species with values for each plot (weighting by species cover), and (2) assigning to the missing species the weighted average values of the plots where they were present (weighting by species cover).
-Then, we calculated the plot-level community-weighted mean (weighting by species cover) for each trait and ecological preference, and explored the main axes of variation using PCA, separately for species traits and ecological preferences.
-
-# Results
-
-## Overview of the classification
-
-Our revision of the literature produced a provisional list of 38 anthropogenic vegetation alliances that could be present in the Iberian Atlantic ecoregion (**Supplementary 3**).
-In our study dataset (n = 2,508 vegetation plots), there were 2,201 plots that had been assigned by the original authors to one of these alliances.
-We used these 2,201 plots for a preliminary re-classification based on TWINSPAN, attempting to re-match the plots into the *a priori* defined checklist of 38 alliances.
-This procedure allowed us to validate 28 alliances as present in the ecoregion.
-A total of 1,725 plots had been classified by both the original authors and our TWINSPAN into these alliances, and these became our training subset for the semi-supervised classification.
-Attempts by the semi-supervised classification to create new groups resulted in new groups with no ecological significance, and thus we kept the original 28 alliances as final vegetation groups (**Table 1**).
-Most of the alliances had a wide distribution in the ecoregion according to our dataset (**Figure 1**), although five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis*, *Paspalo-Agrostion semiverticillati*) had only isolated occurrences in a few points towards the southern limit of the Iberian Atlantic territories.
-The regional literature included in our dataset recognized 72 associations (**Supplementary 4**) within these alliances.
-The semi-supervised classification with noise clustering resulted in the final classification of 2,081 plots into 28 alliances and 427 plots left out in the noise group.
-
-Focusing on the classified plots, the dataset included 1,167 taxa or taxa aggregates.
-Of these, 101 were neophytes (i.e.
-9% of the species pool).
-Regarding life forms, the species pool was dominated by hemicryptophytes (43%) and therophytes (41%).
-The 10 most frequent species were *Ochlopoa annua* (621 occurrences), *Urtica dioica* (592), *Sonchus oleraceus* (540), *Stellaria media* (516), *Capsella bursa-pastoris* (390), *Polygonum aviculare* (375), *Dactylis glomerata* (363), *Senecio vulgaris* (332) and *Anisantha sterilis* (301).
-Considering only those plots with the most frequent plot size (20 m^2^, n = 256 plots), the average species richness per plot was 16 (minimum = 3, maximum = 37).
-
-DCA ordination of the floristic composition of the classified plots (**Fig.
-2**) suggested a high consistency of the vegetation classes *Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Sisymbrietea* and *Bidentetea*.
-The alliances belonging to the two perennial ruderal vegetation classes (*Artemisietea vulgaris* and *Epilobietea angustifolii*) were mixed together.
-The winter-annual class *Chenopodietea* also showed high heterogeneity, especially because of the isolation of the altered dune alliance *Linario polygalifoliae-Vulpion alopecuri*.
-
-## Community-level traits and ecological preferences
-
-PCA ordination of the community-weighted means of the species traits (**Fig.
-3A**) indicated that the first two axes explained 65% of the variability.
-The major contributors to axis 1 (41% variance explained) were community height, proportion of geophytes, mean month of flowering and proportion of therophytes; it separated taller communities with more geophytes and later flowering (classes *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Bidentetea*) from shorter communities with more therophytes and earlier flowering (classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis* and *Sisymbrietea*).
-The major contributors to axis 2 (24% variance explained) were proportion of neophytes and width of the flowering season; it separated communities with more neophytes and longer flowering seasons (classes *Bidentetea*, *Digitario sanguinalis-Eragrostietea minoris* and *Polygono-Poetea annuae*) from communities with less neophytes and shorter flowering seasons (classes *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Chenopodietea*).
-
-PCA ordination of the community-weighted means of the species ecological and disturbance preferences (**Fig.
-3B**) indicated that the first two axes explained 56% of the variability.
-The major contributors to axis 1 (35% variance explained) were disturbance frequency, soil disturbance, mowing frequency and light requirements; it separated communities thriving with frequent disturbances in open situations (classes *Polygono-Poetea annuae*, *Sisymbrietea*, *Digitario-Eragrostietea* and *Papaveretea rhoeadis*) from communities preferring less frequent disturbances and shadier areas (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii* and *Bidentetea*).
-The major contributors to axis 2 (21% variance explained) were nutrient requirements and disturbance severity; it separated more nitrophilous communities adapted to more severe disturbances (classes *Digitario-Eragrostietea*, *Sisymbrietea* and *Epilobietea angustifolii*) from less nitrophilous communities adapted to lower intensity disturbances (classes *Cymbalario-Parietarietea diffusae*, *Artemisietea vulgaris* and *Chenopodietea*).
-
-## Classified anthropogenic plant communities
-
-***Cymbalario-Parietarietea diffusae*** Vegetation of human-made walls.
-
-1.
-*Galio valantiae-Parietarion judaicae* Vegetation of dry walls.
-Occupies warmer, drier and more nutrient-rich situations than the other alliance in the class, has a shorter flowering season and is more adapted to disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Parietaria judaica* aggr., *Centranthus ruber*, *Asplenium trichomanes*.
-
-2.
-*Cymbalario-Asplenion* Vegetation of mesic walls.
-Occupies colder, wetter and less nutrient-rich situations than the other alliance in the class, has a longer flowering season and is less adapted to disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Cymbalaria muralis*.
-
-***Polygono-Poetea annuae*** Dwarf vegetation of heavily-trampled sites.
-
-3.
-*Polycarpion tetraphylli* Dwarf-annual trampled vegetation of warm and sunny sites.
-Occupies warmer, sunnier and less nutrient-rich situations than the other alliances in the class and has an earlier flowering season (**Fig.
-4**).
-Wide distribution in the ecoregion, specially along the coast and in the more Mediterranean valleys of the interior (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Sagina apetala*, *Plantago coronopus*, *Crassula tillaea*, *Spergularia marina*.
-
-4.
-*Polygono-Coronopodion* Dwarf-annual trampled vegetation of dry sites.
-Occupies colder, drier and more nutrient-rich situations than the other alliances in the class, and has a higher proportion of neophytes (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Matricaria suaveolens*, *Polygonum aviculare*.
-
-5.
-*Saginion procumbentis* Dwarf trampled vegetation of mesic sites.
-Occupies colder, wetter and more nutrient-rich situations than the other alliance in the class, and is less dominated by therophytes (**Fig.
-4**).
-Distribution concentrated in the central part of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Sagina procumbens*, *Bryum argenteum*.
-
-***Papaveretea rhoeadis*** Annual weed vegetation of crops and gardens.
-
-6.
-*Caucalidion lappulae* Annual weed vegetation of cereal crops on base-rich soils.
-Adapted to more frequent and severe disturbances than the other alliances in the class  (**Fig.
-4**).
-Distribution mostly limited to the more Mediterranean valleys south of the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Cyanus segetum*, *Papaver rhoeas*, *Trifolium arvense*, *Anacyclus clavatus*, *Valerianella eriocarpa*.
-
-7.
-*Scleranthion annui* Annual weed vegetation of cereal crops on base-poor soils.
-Occupies more acidic soils than the other alliances in the class, and flowers earlier (**Fig.
-4**).
-Wide distribution in the ecoregion, but specially in the west, where acidic bedrocks dominate (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Mibora minima*, *Rumex acetosella*.
-
-8.
-*Oxalidion europeae* Annual weed vegetation of gardens and root crops.
-Occupies wetter, shadier and more nutrient-rich situations than the other alliances in the class, has a longer flowering season and a higher proportion of neophytes (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Stellaria media*, *Veronica persica*.
-
-***Digitario sanguinalis-Eragrostietea minoris*** Summer-annual C4 weed vegetation.
-
-9.
-*Spergulo arvensis-Erodion cicutariae* Summer-annual C4 weed vegetation.
-Late-flowering alliance rich in neophytes and grasses, occurring in highly-disturbed and nutrient-rich soils (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Digitaria sanguinalis*, *Echinochloa crus-galli*, *Chenopodium album*, *Amaranthus hybridus* aggr.
-
-***Chenopodietea*** Winter-annual ruderal vegetation.
-
-10.
-*Allion triquetri* Geophyte-rich fringe vegetation.
-Occupies warmer and shadier situations than the other alliances in the class, has a higher proportion of geophytes and a lower proportion of therophytes, and is adapted to less frequent disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Urtica membranacea*, *Smyrnium olusatrum*, *Tradescantia fluminensis*.
-
-11.
-*Geranio pusilli-Anthriscion caucalidis* Winter-annual fringe vegetation.
-Occupies situations similar to the previous alliance, but in colder sites, and is dominated by therophytes instead of geophytes (**Fig.
-4**).
-Mostly distributed in the more Mediterranean valleys south of the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Geranium lucidum*, *Anthriscus caucalis*.
-
-12.
-*Chenopodion muralis* Low-growth winter-annual ruderal vegetation.
-Occupies colder situations than the other alliances in the class and more often disturbed sites, and has a longer and later flowering season (**Fig.
-4**).
-Sparse occurrence throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Malva neglecta*, *Urtica urens*.
-
-13.
-*Echio-Galactition tomentosae* Tall-herb winter-annual ruderal vegetation.
-Occupies drier, sunnier and less nutrient-rich situations than the other alliances in the class (**Fig.
-4**).
-Mostly distributed in the west of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Galactites tomentosus*, *Coleostephus myconis*, *Anisantha rigida*.
-
-14.
-*Linario polygalifoliae-Vulpion alopecuri* Ephemeral annual vegetation of disturbed coastal dunes.
-Occupies drier and more nutrient-poor situations than the other alliances in the class (**Fig.
-4**) and has a high number of coastal dune species.
-Isolated occurrence in the southern coasts of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Vulpia alopecuros*, *Malcolmia littorea*, *Crucianella maritima*, *Artemisia campestris* aggr.
-
-***Sisymbrietea*** Summer-annual ruderal vegetation.
-
-15.
-*Sisymbrion officinalis* Summer-annual ruderal vegetation.
-Rich in therophytes and adapted to severe disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Hordeum murinum*, *Sisymbrium officinale*.
-
-***Artemisietea vulgaris*** Perennial ruderal vegetation of dry sites.
-
-16.
-*Convolvulo arvensis-Agropyrion repentis* Semiruderal grasslands.
-Occupies wetter and shadier situations than the other alliances in the class, and has a higher proportion of geophytes and grasses (**Fig.
-4**).
-Isolated occurrence in the south of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Poa compressa*, *Elytrigia repens*, *Potentilla reptans*, *Valerianella locusta*.
-
-17.
-*Carduo carpetani-Cirsion odontolepidis* Thistle ruderal vegetation of warm sites.
-Occupies warmer situations than the other thistle alliance in the class (**Fig.
-4**).
-Mostly distributed in the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Onopordum acanthium*.
-
-18.
-*Cirsion richterano-chodati* Thistle ruderal vegetation of cold sites.
-Occupies colder situations than the other thistle alliance in the class (**Fig.
-4**).
-Mostly distributed in the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Cirsium eriophorum*, *Carduus carpetanus*, *Carduus nutans* aggr.
-
-19.
-*Dauco-Melilotion* Biennial ruderal vegetation.
-Occupies wetter situations than the two thistle alliances in the class, and has a higher proportion of therophytes (**Fig.
-4**).
-Sparse occurrence throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Helminthotheca echioides*, *Daucus carota*, *Melilotus albus*, *Foeniculum vulgare*.
-
-***Epilobietea angustifolii*** Perennial ruderal vegetation of mesic to wet sites.
-
-20.
-*Geo urbani-Alliarion officinalis* Low-herb short-lived semiruderal and fringe vegetation.
-Occupies shadier and less disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Urtica dioica*.
-
-21.
-*Arction lappae* Low-herb short-lived ruderal vegetation.
-Occupies sunnier and more frequently-disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants (**Fig.
-4**).
-Distribution concentrated in the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Oxybasis rubra*, *Senecio duriaei*.
-
-22.
-*Balloto-Conion maculati* Tall-herb perennial ruderal vegetation.
-Occupies warmer situations than the other alliances in the class (**Fig.
-4**).
-Distribution concentrated in the Cantabrian Mountains and the east of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Sambucus ebulus*.
-
-23.
-*Aegopodion podagrariae* Tall-herb perennial vegetation of forest margins and clearings.
-Occupies shadier and less disturbed situations than the other alliances in the class (**Fig.
-4**).
-Sparse occurrence throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Anthriscus sylvestris*, *Heracleum sphondylium*, *Galium aparine*.
-
-24.
-*Epilobion angustifolii* Tall-herb perennial vegetation of forest margins and clearings in acidic soils.
-Occupies colder, nutrient-poorer, more acidic and less disturbed situations than the other alliances in the class (**Fig.
-4**).
-Distributed towards the west of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Epilobium angustifolium*, *Digitalis purpurea*, *Luzula lactea*.
-
-25.
-*Cynancho-Convolvulion sepium* Tall-herb vegetation of nutrient-rich riparian habitats.
-Occupies wetter situations and less disturbed situations than the other alliances in the class, and has a later flowering season (**Fig.
-4**).
-Dispersed throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Eupatorium cannabinum*, *Angelica sylvestris*, *Picris hieracioides*.
-
-26.
-*Senecionion fluviatilis* Tall-herb vegetation of nutrient-rich river banks and ditches.
-Occupies warmer and wetter situations than the other alliances in the class, flowers later, has taller plants and is highly dominated by neophytes and geophytes (**Fig.
-4**).
-Isolated occurrence in the western coast of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Arundo donax*, *Silene latifolia*.
-
-***Bidentetea*** Summer-annual pioneer vegetation of temporarily-flooded sites.
-
-27.
-*Bidention tripartitae* Summer-annual pioneer vegetation of temporarily-flooded sites.
-Occupies nutrient-richer sites than the other alliance in the class and has a higher proportion of therophytes (**Fig.
-4**).
-Dispersed throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Persicaria hydropiper*, *Bidens frondosus*, *Lythrum salicaria*.
-
-28.
-*Paspalo-Agrostion semiverticillati* Summer-annual pioneer vegetation of temporarily-flooded and warm sites.
-Occupies warmer and wetter situations than the other alliance in the class, has a higher proportion of neophytes and has a longer flowering season (**Fig.
-4**).
-Isolated occurrences in the south of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Paspalum distichum*, *Cyperus longus*, *Schoenoplectus lacustris*.
-
-# Discussion
-
-The major driver of variation in trampled community composition is soil moisture [@RN5434].
-
-In medieval Switzerland, summer and winter weeds communities could not be distinguished [@RN5431].
-
-The class *Digitario sanguinalis-Eragrostietea minoris* includes thermophilic vegetation dominated by C4 plants that occupies arable land as well as trampled and ruderal sites [@RN5426].
-
-Anthropogenic vegetation composition is sensitive to micro-scale management decisions, such as the ownership of domestic dogs [@RN5438].
-
-Neophytes were more represented in early successional anthropogenic communities [@RN5423].
-
-Arable land plants were mainly therophytes and alien species; while ruderal habitats were richer in perennials and C strategists [@RN5436].
-
-Compare proportion of neophytes to other areas [@RN5429; @@RN5435]
-
-Neophytes have no special preferences for local habitat conditions and their highest proportion was found mainly in disturbed habitats at low elevations. [@RN5435]
-
-Therophytes and hemicryptophytes were most abundant in the life form spectra [@RN5428]
-
-Segetal vegetation of Palestine [@RN5494].
-
-# Data availability
-
-The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository https://github.com/efernandezpascual/manmade.
-Upon publication, a version of record of the repository will be deposited in Zenodo.
-
-# References
-
-::: {#refs}
-:::
-
-# Figures
-
-```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Figure 1.
-Distribution of the anthropogenic alliances in the Iberian Atlantic ecoregion.
-Each dot represents a relevé.
-Dots are colored by vegetation classes."}
-knitr::include_graphics("../results/figures/F1 - maps.png")
-```
-
-```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Figure 2.
-Ordination of the floristic composition of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
-The ordination is the results of a Detrended Correspondence Analysis grouped by vegetation classes (A, left) and alliances (B, right)."}
-knitr::include_graphics("../results/figures/F2 - pca.png")
-```
-
-```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Figure 3.
-Ordination of the community-level species traits and ecological preferences.
-The ordinations are the results of a Principal Component Analysis calculated for the community-weighted means for species traits (A, left) and ecological and disturbance indicator values (B, right)."}
-knitr::include_graphics("../results/figures/F3 - origins.png")
-```
-
-```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
-Community-level species traits and ecological preferences by alliance.
-Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
-knitr::include_graphics("../results/figures/F4 - lifeforms.png")
-```
-
-```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
-Community-level species traits and ecological preferences by alliance.
-Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
-knitr::include_graphics("../results/figures/F5 - height-flowering.png")
-```
-
-```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
-Community-level species traits and ecological preferences by alliance.
-Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
-knitr::include_graphics("../results/figures/F6 - indicators.png")
-```
-
-```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
-Community-level species traits and ecological preferences by alliance.
-Boxplots calculated for the community-weighted means for species traits and ecological and disturbance indicator values."}
-knitr::include_graphics("../results/figures/F7 - pca.png")
-```
+> Fernández-Pascual E, González-García, Ivesdal G, Lázaro-Lobo A, Jiménez-Alfaro B.
+[*Anthropogenic plant communities of the Iberian Atlantic ecoregion*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), *in preparation*.

### 2024-04-26 doc/manuscript.Rmd
readme

@@
 * *Results* We classified 2,081 vegetation plots into 28 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*).
-The plots included 1,162 vascular plant taxa: 78 % natives, 15 % archaeophytes and 7 % neophytes.
+The plots included 1,162 plant taxa: 78% natives, 15% archaeophytes and 7% neophytes.
 The main ecological gradients behind this diversity went from dry-sunny to moist-shady habitats, followed by a gradient of low to high disturbance frequency and severity.
@@
 
-The total synanthropic species pool (n = 1,162) was composed of 78 % natives (n = 908), 15 % putative archaeophytes (n = 178) and 7 % neophytes (n = 76).
+The total synanthropic species pool (n = 1,162) was composed of 78% natives (n = 908), 15% putative archaeophytes (n = 178) and 7% neophytes (n = 76).
 The most frequent putative archaeophytes were *Capsella bursa-pastoris* (390 occurrences), *Senecio vulgaris* (332), *Anisantha sterilis* (301), *Anthemis arvensis* (280), *Hordeum murinum* (274), *Raphanus raphanistrum* (233), *Vicia sativa* (228), *Malva sylvestris* (197), *Sisymbrium officinale* (166) and *Echinochloa crus-galli* (165); while the most frequent neophytes were *Veronica persica* (212), *Erigeron canadensis* (138), *Amaranthus hybridus* (73), *Oxalis corniculata* (67), *Lepidium didymum* (57), *Oxalis latifolia* (51), *Paspalum distichum* (48), *Bidens frondosus* (35), *Cyperus eragrostis* (34) and *Symphyotrichum squamatum* (33).
 
-By vegetation class, *Epilobietea angustifolii* had the highest proportion of natives in its species pool (82 %); *Sisymbrietea* had the highest proportion of putative archaeophytes (31 %); and *Bidentetea* had the highest proportion of neophytes (17 %).
-The alliances with the highest proportions of archaeophytes in their species pools were *Caucalidion lappulae* (41 %), *Chenopodion muralis* (35 %) and *Sisymbrion officinalis* (31 %); while the alliances with the highest proportion of neophytes were *Bidention tripartitae* (19 %), *Allion triquetri* (12 %) and *Spergulo arvensis-Erodion cicutariae* (10%).
+By vegetation class, *Epilobietea angustifolii* had the highest proportion of natives in its species pool (82%); *Sisymbrietea* had the highest proportion of putative archaeophytes (31%); and *Bidentetea* had the highest proportion of neophytes (17%).
+The alliances with the highest proportions of archaeophytes in their species pools were *Caucalidion lappulae* (41%), *Chenopodion muralis* (35%%) and *Sisymbrion officinalis* (31%); while the alliances with the highest proportion of neophytes were *Bidention tripartitae* (19%), *Allion triquetri* (12%) and *Spergulo arvensis-Erodion cicutariae* (10%).
 
@@
 
-The most frequent life form in the species pool (n = 1,162 species) were hemicryptophytes (44 %), followed by therophytes (42 %), geophytes (9 %), chamaephytes (6 %), phanerophytes (6 %), bryophytes (1 %) and hydrophytes (1 %).
+The most frequent life form in the species pool (n = 1,162 species) were hemicryptophytes (44%), followed by therophytes (42%), geophytes (9%), chamaephytes (6%), phanerophytes (6%), bryophytes (1%) and hydrophytes (1%).
 The proportion of therophytes across classes and alliances (**Fig.
@@
 7**).
-The first two PCA axes explained 59 % of the variability.
-The major contributors to axis 1 (36 % variance explained) were light, nutrient and moisture requirements; it separated communities of open, dry and comparatively nutrient-poorer sites (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea*, *Artemisietea vulgaris*, *Chenopodietea* and *Digitario sanguinalis-Eragrostietea minoris*) from communities of shady, moist and nutrient-richer sites (classes *Cymbalario-Parietarietea diffusae*, *Bidentetea* and *Epilobietea angustifolii*).
+The first two PCA axes explained 59% of the variability.
+The major contributors to axis 1 (36% variance explained) were light, nutrient and moisture requirements; it separated communities of open, dry and comparatively nutrient-poorer sites (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea*, *Artemisietea vulgaris*, *Chenopodietea* and *Digitario sanguinalis-Eragrostietea minoris*) from communities of shady, moist and nutrient-richer sites (classes *Cymbalario-Parietarietea diffusae*, *Bidentetea* and *Epilobietea angustifolii*).
 The major contributors to axis 2 (23% variance explained) were disturbance severity and frequency; this axis separated communities preferring low severity - low frequency disturbances (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii*, *Chenopodietea*, *Artemisietea vulgaris* and *Bidentetea*) from communities adapted to high severity - high frequency disturbances (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea* and *Digitario sanguinalis-Eragrostietea minoris*).
@@
 
-We found a percentage of 22 % non-native species in the anthropogenic vegetation of the Iberian Atlantic ecoregion.
-This percentage is intermediate between that found in urban areas of temperate Europe and North America (30-50 %) [@RN5498; @RN5499; @RN5500; @RN5079] and Italy (12-26 %) [@RN5501].
-A survey of anthropogenic vegetation in the north-west Balkans found 13 % non-native plants, with more neophytes than archaeophytes [@RN5429].
-A similar study in the Czech Republic found 32 % archaeophytes and 7 % neophytes [@RN5435].
+We found a percentage of 22% non-native species in the anthropogenic vegetation of the Iberian Atlantic ecoregion.
+This percentage is intermediate between that found in urban areas of temperate Europe and North America (30-50%) [@RN5498; @RN5499; @RN5500; @RN5079] and Italy (12-26%) [@RN5501].
+A survey of anthropogenic vegetation in the north-west Balkans found 13% non-native plants, with more neophytes than archaeophytes [@RN5429].
+A similar study in the Czech Republic found 32% archaeophytes and 7% neophytes [@RN5435].
 Taken together, these results suggest that the Iberian Atlantic anthropogenic vegetation shows an invasion status that is intermediate between southern and central-northern Europe, which is in agreement with the biogeographical position of our ecoregion.
@@
 
-The anthropogenic vegetation of the Iberian Atlantic ecoregion is home to 1,162 vascular plant taxa.
+The anthropogenic vegetation of the Iberian Atlantic ecoregion is home to 1,162 plant taxa.
 This is approximately one third of the whole ecoregional species pool represented in the SIVIM database [@RN5374], or one fifth of the Iberian flora [@RN5503].

### 2024-06-12 README.Rmd
readme 📚
submission-ready 🏙️

@@
 
-![Antigua fábrica de La Cerámica Gijonesa](ceramica.JPG)
+![](ceramica.JPG)
 
-# Anthropogenic plant communities of the Iberian Atlantic ecoregion
+# Classification and characterization of anthropogenic plant communities at the ecoregion level
 
-This repository stores all the information related to the manuscript [*Anthropogenic plant communities of the Iberian Atlantic ecoregion*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
+This repository stores all the information related to the manuscript [*Classification and characterization of anthropogenic plant communities at the ecoregion level*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), including the datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
 
@@
 
-* *Questions* Which is the diversity of anthropogenic plant communities at the ecoregional level in a transitional territory between temperate and Mediterranean Europe?
-How are these communities composed by native plants, archaeophytes and neophytes?
-What are the main environmental and disturbance drivers of the diversity in anthropogenic vegetation?
+* *Questions* How can we classify the diversity of anthropogenic plant communities at the ecoregion level?
+How are these communities characterized by species origins (native plants, archaeophytes, and neophytes), species traits (lifeforms, plant height, and flowering phenology) and ecological preferences (temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity)?
 
-* *Location* Iberian Atlantic ecoregion, Iberian Peninsula, south-western Europe.
+* *Location* Iberian Atlantic ecoregion (a.k.a. Cantabrian Mixed Forests); Portugal, Spain, and France; south-western Europe.
 
-* *Methods* We compiled a vegetation database of 2,508 synanthropic plots and revised the regional checklist of anthropogenic alliances.
+* *Methods* We compiled a vegetation database of 2,508 synanthropic plots and revised the regional checklist of anthropogenic vegetation alliances.
 We used modified TWINSPAN and semi-supervised classification to classify the plots into the revised alliances.
-We determined the proportion of natives, archaeophytes and neophytes.
-We described the alliances in terms of vegetation height, flowering phenology and ecological requirements for temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.
+We determined the proportion of natives, archaeophytes, and neophytes.
+We also described the alliances in terms of species traits and ecological requirements.
 
 * *Results* We classified 2,081 vegetation plots into 28 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*).
-The plots included 1,162 plant taxa: 78% natives, 15% archaeophytes and 7% neophytes.
-The main ecological gradients behind this diversity went from dry-sunny to moist-shady habitats, followed by a gradient of low to high disturbance frequency and severity.
+The plots included 1,162 plant taxa: 78% natives, 15% archaeophytes, and 7% neophytes.
+Vegetation groups were organized along a principal axis of variation related to abiotic stress (dry-sunny to moist-shady habitats), and a second axis related to disturbance (low to high disturbance frequency and severity).
 
-* *Conclusions* The human-made habitats of the Iberian Atlantic ecoregion host one third of the ecoregional species pool and one fifth of the Iberian flora.
-Wet anthropogenic habitats are most affected by neophyte invasions.
-Annual vegetation of arable lands is a refuge for archaeophytes.
-Mesic perennial ruderal vegetation is specially rich in native species and can be a biodiversity asset in urban and peri-urban settings.
+* *Conclusions* Ecoregion-level synthesis detected a discrepancy between the number of anthropogenic vegetation units described in the literature (38 alliances) and the number supported by numerical classification (28).
+The diversity of anthropogenic vegetation can be organized into three groups: trampled, weeds, and ruderals.
+In the Iberian Atlantic ecoregion, anthropogenic habitats host one third of the ecoregion plant species pool and one fifth of the Iberian flora.
+Mesic perennial ruderal vegetation is especially rich in native species and can be a biodiversity asset in urban ecosystems.
+Our ecoregion-level framework can improve the assessment and management of anthropogenic plant communities within biogeographically meaningful regions.
 
@@
 > Fernández-Pascual E, González-García, Ivesdal G, Lázaro-Lobo A, Jiménez-Alfaro B.
-[*Anthropogenic plant communities of the Iberian Atlantic ecoregion*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), *in preparation*.
+[*Classification and characterization of anthropogenic plant communities at the ecoregion level*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), *in preparation*.

### 2024-06-12 doc/manuscript.Rmd
submission-ready 🏙️

--- 2024-04-11 doc/manuscript.Rmd
@@
 ---
-title: "Anthropogenic plant communities of the Iberian Atlantic ecoregion"
+title: "Classification and characterization of anthropogenic plant communities at the ecoregion level"
 output: 
-  bookdown::github_document2:
+  bookdown::word_document2:
     reference_docx: template.docx
@@
 
+# Abstract (300 words max.)
+
+* *Questions* How can we classify the diversity of anthropogenic plant communities at the ecoregion level?
+How are these communities characterized by species origins (native plants, archaeophytes, and neophytes), species traits (lifeforms, plant height, and flowering phenology) and ecological preferences (temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity)?
+
+* *Location* Iberian Atlantic ecoregion (a.k.a. Cantabrian Mixed Forests); Portugal, Spain, and France; south-western Europe.
+
+* *Methods* We compiled a vegetation database of 2,508 synanthropic plots and revised the regional checklist of anthropogenic vegetation alliances.
+We used modified TWINSPAN and semi-supervised classification to classify the plots into the revised alliances.
+We determined the proportion of natives, archaeophytes, and neophytes.
+We also described the alliances in terms of species traits and ecological requirements.
+
+* *Results* We classified 2,081 vegetation plots into 28 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*).
+The plots included 1,162 plant taxa: 78% natives, 15% archaeophytes, and 7% neophytes.
+Vegetation groups were organized along a principal axis of variation related to abiotic stress (dry-sunny to moist-shady habitats), and a second axis related to disturbance (low to high disturbance frequency and severity).
+
+* *Conclusions* Ecoregion-level synthesis detected a discrepancy between the number of anthropogenic vegetation units described in the literature (38 alliances) and the number supported by numerical classification (28).
+The diversity of anthropogenic vegetation can be organized into three groups: trampled, weeds, and ruderals.
+In the Iberian Atlantic ecoregion, anthropogenic habitats host one third of the ecoregion plant species pool and one fifth of the Iberian flora.
+Mesic perennial ruderal vegetation is especially rich in native species and can be a biodiversity asset in urban ecosystems.
+Our ecoregion-level framework can improve the assessment and management of anthropogenic plant communities within biogeographically meaningful regions.
+
+# Keywords (10 max)
+
+Synanthropic vegetation, manmade habitats, human-made habitats, urban biodiversity, neophytes, archaeophytes, alien plants, habitat classification, semi-supervised classification, Iberian Atlantic ecoregion
+
 # Introduction
 
-Anthropogenic vegetation is the set of plant communities that occur as a direct consequence of human activities, and is chiefly composed by the weed vegetation of arable fields and the ruderal vegetation of human settlements and their surroundings [@RN5441].
-The anthropogenic communities of human-made habitats make up an important part of the supply of natural elements in urban and peri-urban habitats [@RN5445].
-In densely-populated city landscapes, the plant communities that colonize urban vacant lots [@RN2300] are a useful biodiversity resource with high ecological and societal potential [@RN5446].
-Anthropogenic vegetation cover can also impact regional-level extreme climatic events [@RN5440].
-Furthermore, many synanthropic plants can accumulate heavy metals in their biomass [@RN5433], making them valuable candidates for nature-based solutions such as bioremediation and restoration of degraded landscapes [@RN5447].
+Anthropogenic vegetation consists of plant communities assembled as a direct consequence of human activities.
+In general, this vegetation is composed of so-called weeds of arable fields and ruderal plants of human settlements and nearby habitats [@RN5441].
+Anthropogenic communities make up an important part of the supply of biodiversity in urban and peri-urban areas [@RN5445].
+In densely populated landscapes, the plant communities that colonize urban vacant lots [@RN2300] are a useful natural resource with high ecological and societal potential [@RN5446].
+Anthropogenic vegetation can also mitigate regional-level extreme climatic events [@RN5440].
+Furthermore, many synanthropic plants can accumulate heavy metals in their biomass [@RN5433], making them valuable candidates for nature-based solutions such as bioremediation and restoration of degraded post-industrial landscapes [@RN5447].
 
-The anthropogenic vegetation of temperate Europe is formed by an original pool of native species favored by humans, but this pool has been constantly enriched by the arrival of alien species: first archaeophytes starting in the Neolithic and peaking in the Bronze Age; and then neophytes from the Modern Age (after 1.500 A.D.) to the present day [@RN5439].
-Synanthropic archaeophytes include many Mediterranean taxa that started their expansion into temperate areas during the Mid-Holocene warm period, but it is difficult to disentangle to what extent this expansion was favored by human activities and/or ongoing climatic changes [@RN5432; @RN5443], especially in southern European areas close to the transition between the Mediterranean and the temperate climates [@RN5444].
-Meanwhile, synanthropic neophytes include many American, African and Asian taxa that arrived as a consequence of the global trade networks established by the Western European colonial powers [@RN5451].
-In the last century, anthropogenic vegetation has shown a decrease in species richness and diversity [@RN5423], as rare synanthropic species have become rarer, and neophytes have increased their abundance at the expense of natives and archaeophytes [@RN5441].
+In temperate Europe, anthropogenic vegetation is a melting pot of floras from different origins.
+Starting with a pool of native species favored by humans, and constantly enriched by human-assisted arrival of alien (i.e.
+non-native) species [@RN5504].
+The introduction of archaeophytes (i.e.
+alien species arriving before 1.500 CE) likely started in the early Neolithic and peaked in the Bronze Age, with successive waves during periods of accelerated human colonization such as the early Middle Ages [@RN5504].
+In Europe, synanthropic archaeophytes include many Mediterranean taxa originating from the segetal flora of the Near East [@RN5494].
+Some of these archaeophytes are 'obligate weeds' that do not occur in non-anthropogenic habitats and are assumed to have evolved together with human agriculture [@RN5494].
+Since many archaeophytes started their expansion into temperate areas during the Mid-Holocene warm period, it is difficult to disentangle to what extent this expansion was favored by human activities and/or ongoing climatic changes [@RN5432; @RN5443].
+The synanthropic flora of Europe has been further enriched by neophytes, alien taxa that arrived after the expansion of intercontinental trade in the Modern Age (after 1.500 CE) [@RN5439].
+Synanthropic neophytes therefore include many American, African, and Asian taxa that arrived in Europe because of the global trade networks established by western European colonialism [@RN5451].
+In the last century, the European anthropogenic vegetation has shown a decrease in species richness and diversity [@RN5423], as rare synanthropic species have become rarer, and neophytes have increased their abundance at the expense of natives and archaeophytes [@RN5441].
 
-Antropogenic vegetation is likely to be similar within ecological regions (ecoregions, hereafter), since they are assumed to encompass areas with a similar biogeographic history and recurrent local ecosystems [@RN5491].
-However, ecoregional-scale studies on the diversity of antropogenic vegetation are scarce in the literature.
+Despite the potential of anthropogenic vegetation as a biodiversity asset in human-made landscapes, and its ecological interest as an assemblage of species pools from different biogeographical origins, human-made vegetation is largely absent from ecosystem management.
+A comprehensive management will require a solid definition, classification, and description of anthropogenic plant communities.
+However, the classification of anthropogenic vegetation is subject to particularities making it a special case compared to the classification of non-anthropogenic communities.
+In human-made habitats, the traditional understanding of community assembly based on sequential filtering (dispersal, abiotic, and biotic) must incorporate an additional filtering based on human preferences and actions [@RN5505].
+Human landscapes are a mosaic of land uses [@RN5527], having specific effects on plant community assembly, and driven by sociological rather than traditional ecological factors [@RN2300].
+In addition to these potentially confounding factors, human-assisted dispersal means that anthropogenic floras are relatively homogeneous across large biogeographical scales.
+However, traditional phytosociology has tended to classify them with a level of detail similar to the one dedicated to non-anthropogenic vegetation, leading to an inflation of anthropogenic syntaxa [@RN5528].
+Another issue is defining anthropogenic vegetation, especially in a continent such as Europe where human impact on the landscape has been widespread since ancient times.
+The two major vegetation/habitat classifications coexisting in Europe today, phytosociological classification [@RN4698] and the EUNIS habitat classification [@RN5061], have non-overlapping definitions of which plant communities fall within the ‘anthropogenic’ or ‘manmade’ categories.
+These issues are further muddled by the fact that most classification studies of anthropogenic vegetation have focused either on a restricted political unit [@RN5424; @RN5426] or in a subset of the vegetation, such as segetal [@RN5494], ruderal [@RN5422] or trampled [@RN5434] vegetation.
+There is a need for a numerical classification and synthesis of the full spectrum of anthropogenic vegetation, and an appropriate scale to conduct such study is the ecological region (ecoregion, hereafter), since these biogeographical units are assumed to encompass areas with a similar biogeographic history and recurrent local ecosystems [@RN5491].
+
 The Iberian Atlantic ecoregion (a.k.a.
 Cantabrian Mixed Forests ecoregion; NW Iberian Peninsula) has a long history of human habitation dating back to the Cantabrian Upper Paleolithic [@RN5450].
-This, together with its transitional position at the border between the temperate and Mediterranean climatic zones of Europe [@RN4700], suggests a long regional history for archaeophyte-rich anthropogenic plant communities, likely supporting a current high diversity of anthropogenic plant communities with both temperate and Mediterranean floristic elements [@RN5008].
-Historical trade links with America and Asia, together with the warm and humid temperate climate, have made the region a hotspot for neophyte invasions [@RN5492; @RN5493].
+This, together with its transitional position at the border between the temperate and Mediterranean climatic zones of Europe [@RN4700], suggests a long regional history for archaeophyte-rich anthropogenic plant communities, supporting the current diverse anthropogenic vegetation with both temperate and Mediterranean floristic elements [@RN5008].
+Moreover, historical trade links with America and Asia, together with the warm and humid temperate climate, have made the region a hotspot for biological invasions [@RN5492; @RN5493].
 Recent post-industrial land-use changes have left large areas with abandoned industrial sites (i.e.
-brownfields) whose management and restoration requires an understanding of anthropogenic communities able to colonize or restore them [@RN5448; @RN5449].
-Although there is a long tradition of studying weed and ruderal communities in the ecoregion [@RN5364; @RN5361; @RN5359; @RN5360] there is a current need for an ecoregional-level synthesis that revises and updates the classification and updates in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
-In this article we have performed such a synthesis with two main objectives: (1) provide an updated classification of anthropogenic plant communities in the Iberian Atlantic ecoregion, and (2) describe the main patterns of vegetation diversity in terms of floristic origin (natives, archaeophytes, neophytes); life forms; functional traits (plant height and flowering time); and ecological indicator values (temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity).
+brownfields) whose management and restoration requires an understanding of the anthropogenic communities able to colonize them [@RN5448; @RN5449].
+Although there is a long tradition of studying weed and ruderal communities in the ecoregion [@RN5364; @RN5361; @RN5359; @RN5360] there is a current need for an ecoregion-level synthesis that revises and updates the classification in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
+In this article, we have performed such a synthesis with the following two objectives: (1) to provide an updated classification of anthropogenic plant communities in the Iberian Atlantic ecoregion; and (2) to characterize the diversity of anthropogenic vegetation in terms of species pools (natives, archaeophytes, neophytes); species traits (life form, plant height, flowering phenology); and ecological requirements (temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity).
 
@@
 
+## Data curation and taxonomic criteria
+
 We performed all data management and analysis with R version 4.3.1 [@RN5387], using the R package *tidyverse* [@RN4662] for data processing and visualization.
-We homogenized all taxon names in the analysis and the article using Euro+Med [@RN5044] or Plants of the World Online for taxa not in Euro+Med [@RN5376].
-For the nomenclature of sintaxa we follow Mucina *et al.* [-@RN4698] for alliances and higher ranks, and Rivas-Martínez *et al.* [-@RN4697] for associations.
-The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/manmade>.
+We homogenized all taxon names using Euro+Med [@RN5044] or Plants of the World Online [@RN5376] for taxa not in Euro+Med.
+For the nomenclature of syntaxa, we followed Mucina *et al.* [-@RN4698] for alliances and higher ranks, and Rivas-Martínez *et al.* [-@RN4697] for associations.
+The original datasets, as well as R code for analysis and creation of the manuscript, are stored in a GitHub repository available at Zenodo (see 'Data availability statement').
 
@@
 We studied the anthropogenic plant communities of the Iberian Atlantic ecoregion [@RN5372], i.e.
-the territories with a temperate climate in the north-western Iberian Peninsula.
-Our study ecoregion broadly corresponds with the Cantabrian Mixed Forests ecoregion *sensu* Olson *et al.* [-@RN4667], with the Iberian part of the European Atlantic province *sensu* Rivas-Martínez *et al.* [-@RN4699] and with the Iberian section of the Atlantic biogeographical region of the European Environmental Agency (<https://www.eea.europa.eu/data-and-maps/figures/biogeographical-regions-in-europe-2>).
+the territories with a temperate climate in the north-western part of the Iberian Peninsula (NW Portugal, N Spain, SW France).
+Our study ecoregion broadly corresponds to the Cantabrian Mixed Forests ecoregion *sensu* Olson *et al.* [-@RN4667], to the Iberian part of the European Atlantic province *sensu* Rivas-Martínez *et al.* [-@RN4699] and to the Iberian section of the Atlantic biogeographical region of the European Environmental Agency (<https://www.eea.europa.eu/data-and-maps/figures/biogeographical-regions-in-europe-2>).
 
@@
 
-To circumscribe our study vegetation, we followed the definition of anthropogenic vegetation in the recent classification of the vegetation of Europe by Mucina *et al.* [-@RN4698].
-In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*.
-To these, and for a complete synthesis, we added to our study the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [-@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls.
+To circumscribe our study vegetation, we followed the definition of anthropogenic vegetation in the revised classification of the vegetation of Europe by Mucina *et al.* [-@RN4698].
+In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris*, and *Epilobietea angustifolii*.
+For the sake of completeness, we also considered the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [-@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls.
 
-## Checklist of anthropogenic sintaxa
+## Checklist of anthropogenic syntaxa
 
-To assist in our classification, we prepared a checklist of anthropogenic sintaxa that could potentially be present in our study ecoregion (**Supplementary 1**), based on regional sintaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
-The checklist inclused 38 anthropogenic vegetation alliances that could be present in the Iberian Atlantic ecoregion (**Supplementary 2**).
-A goal of our study was to validate the presence of these alliances in our ecoregion, and to assign to them the existing vegetation data.
+To assist in our classification, we prepared a checklist of anthropogenic syntaxa that could be present in our study ecoregion (**Appendix S1**), based on regional syntaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
+This syntaxonomical checklist included 38 anthropogenic vegetation alliances that could be present in the Iberian Atlantic ecoregion according to the literature.
 
@@
 
-We retrieved anthropogenic vegetation plots from SIVIM, the Iberian and Macaronesian Vegetation Information System [@RN5374].
-To identify plots as anthropogenic, we selected those plots that had been classified by the original authors of the plot as belonging to any sintaxa in our sintaxonomical checklist.
-However, since not all plots in SIVIM have an assigned sintaxon, we also retrieved additional plots using the expert system created by Chytrý *et al.* [-@RN5061] to classify vegetation plots into EUNIS pan-European habitat types.
-This expert system uses the floristic composition of plots to assign them to habitat types depending on the presence and abundance of characteristic species.
-We retrieved plots that had been assigned to any habitat related to the vegetation classes we had previously defined: all habitats in the level 1 code V (vegetated man-made habitats), plus level 3 codes R55 (lowland moist or wet tall-herb and fern fringe) and R57 (herbaceous forest clearing vegetation).
-It must be noted that habitats R55 and R57 include communities that are classified as *grasslands and lands dominated by forbs, mosses or lichens * by EUNIS but as *anthropogenic vegetation* by by Mucina *et al.* [-@RN4698].
-The extraction from SIVIM produced an initial pool of 3,160 vegetation plots, to which we added 89 vegetation plots of urban or peri-urban plant communities sampled by us or extracted from local literature not in SIVIM [@RN5377; @RN5378].
+We retrieved Iberian Atlantic anthropogenic vegetation plots from the Iberian and Macaronesian Vegetation Information System (SIVIM) [@RN5374].
+To identify plots as anthropogenic, we selected those plots that had been assigned by the original authors of the plot into any anthropogenic syntaxa in our syntaxonomical checklist.
+However, since not all plots in SIVIM have an assigned syntaxon, we also retrieved additional plots using the expert system created by Chytrý *et al.* [-@RN5061] to classify vegetation plots into EUNIS pan-European habitat types.
+We retrieved plots that had been assigned to any habitat related to the anthropogenic vegetation classes we had previously defined: all habitats in the level 1 code *V Vegetated man-made habitats*, plus level 3 codes *R55 Lowland moist or wet tall-herb and fern fringe* and *R57 Herbaceous forest clearing vegetation*.
+It must be noted that habitats *R55* and *R57* include communities that are classified as *R Grasslands and lands dominated by forbs, mosses or lichens* by EUNIS but as *Anthropogenic vegetation* by Mucina *et al.* [-@RN4698].
+The extraction of data from SIVIM produced an initial pool of 3,160 vegetation plots, to which we added 89 vegetation plots of urban or peri-urban plant communities sampled by us or extracted from local literature not included in SIVIM [@RN5377; @RN5378].
 
@@
 
-We performed an exploratory analysis of these 3,249 plots using Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379].
-We used the R package *twinspanR* [@RN5380] to do the TWINSPANs in this article, using 3 pseudospecies cut levels (0, 15, 25) with a minimum group size of 10 plots and Sørensen's average dissimilarity.
-During this exploratory analysis we identified 741 outlier plots which either did not belong to the target anthropogenic vegetation or to the geographical area of the ecoregion.
+We performed an exploratory data analysis of these 3,249 plots using modified Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379] with the R package *twinspanR* [@RN5380], 3 pseudospecies cut levels (0, 15, 25), a minimum group size of 10 plots, and Sørensen's average dissimilarity.
+During this analysis we identified 741 outlier plots which did not belong to either the target anthropogenic vegetation or to the geographical area of the ecoregion.
 The majority of these plots corresponded to coastal plant communities and had been misclassified by the expert system.
-We removed these, leaving a dataset of 2,508 vegetation plots for data analysis: 2,419 plots from SIVIM, originally recorded in 89 publications (**Supplementary 3**); plus 89 plots added by us.
+We removed these plots, leaving a dataset of 2,508 vegetation plots for data analysis: 2,419 plots from SIVIM, originally recorded in 89 publications (**Appendix S2**); plus 89 plots added by us.
 
@@
 
-To create a training dataset for semi-supervised classification (see below), we started by keeping only those plots (n = 2,201) that had been assigned by the original authors of the plot as belonging to any of the 38 alliances in our sintaxonomical checklist (i.e., we removed the plots without sintaxa, that had been retrieved solely by the expert system).
-We further cleaned this subset by subjecting it to a preliminary numerical classification using TWINSPAN, in which we attempted to match the plots into the 38 alliances our sintaxonomical checklist.
-This step allowed us to identify those plots in which there was an agreement between (1) our TWINSPAN-based classification and (2) the classification based on the original-author sintaxa; these plots (n = 1,725) became our training dataset for semi-supervised classification.
-Furthermore, this step allowed us to validate which alliances from our sintaxonomical checklist were present in the study ecoregion (n = 28) and which were absent (n = 10).
+To create a training dataset for the semi-supervised classification (see below), we started by keeping only those plots (n = 2,201) that had been assigned by the original authors of the plot to any of the 38 anthropogenic alliances defined in our syntaxonomical checklist (i.e., we removed the plots without syntaxa, that had been retrieved solely by the expert system).
+We further cleaned this subset using a preliminary TWINSPAN classification, in which we attempted to match the plots into the 38 alliances.
+This step allowed us to identify those plots in which there was an agreement between (1) our TWINSPAN-based classification and (2) the classification based on the original author's syntaxon.
+These plots (n = 1,725) became our training dataset for semi-supervised classification.
+Furthermore, this step allowed us to validate which alliances from our syntaxonomical checklist were present in the study ecoregion (n = 28) and which were absent (n = 10).
 
@@
 Since our goal was to refine the classification of the whole dataset, we allowed plots from the training subset to be re-assigned to other alliances during the classification.
-In addition, semi-supervised classification can create new groups to place data points that do not match the already existing *a priori* groups, but attempts to do so resulted in new groups with no ecological significance, and thus we kept the  28 alliances as final vegetation groups.
-Furthermore, we applied a noise clustering fuzzy algorithm, which allows the plots to be classified either into the *a priori* alliances or into a *noise* group which includes outliers and transitional plots [@RN5382].
+In addition, semi-supervised classification can create new groups to place data points that do not match the already existing *a priori* groups, but attempts to do so resulted in new groups with no ecological significance, and thus we kept the 28 alliances as final anthropogenic vegetation groups.
+Furthermore, we applied a noise clustering fuzzy algorithm, which allows plots to be classified either into the *a priori* alliances or into a *noise* group which includes outliers and transitional plots [@RN5382].
 We set the fuzziness coefficient to a low value (m = 1) to accommodate a high number of transitional plots; and we set the distance to the noise class to d = 1.
@@
 The semi-supervised classification with noise clustering resulted in the final classification of 2,081 plots into 28 alliances and 427 plots left out in the noise group.
-We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package *RVAideMemoire* [@RN5381] to test the significance of the final vegetation alliances; along with Principal Component Analysis (PCA) as implemented in the R package *FactoMineR* [@RN3166] to visualize the relationships between the alliances.
+We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances, and Holm's p-value correction) fitted with the R package *RVAideMemoire* [@RN5381] to test the significance of the final vegetation alliances; along with Principal Component Analysis (PCA) as implemented in the R package *FactoMineR* [@RN3166] to visualize the relationships between the classes and alliances.
 
-## Characteristic species and EUNIS habitat regionalization
+## Characteristic species and EUNIS habitat correspondence
 
-We calculated sets of characteristic species [@RN5061] for each of the final 28 vegetation alliances.
-We defined dominant species as species with more than 25% cover in at least 5% of the vegetation plots of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species with an *IndVal* value with a p-value lower than 0.05, as calculated with the R package *labdsv* [@RN2311] using 1.000.000 iterations.
-Additionally, we assigned to each alliance a regionalized level 4 EUNIS habitat code, using pre-existing codes when they existed (<https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp>), and creating new codes when not.
+To facilitate the use of the classification by local managers and environmental consultants, we calculated sets of characteristic species (**Appendix S3**) [@RN5061] for each one of the final 28 vegetation alliances.
+We defined dominant species as species with more than 25% cover in at least 5% of the vegetation plots of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species whose *IndVal* had a p-value lower than 0.05, as calculated with the R package *labdsv* [@RN2311] using 1.000.000 iterations.
+Additionally, we assigned to each alliance its corresponding level-3 EUNIS habitat code or codes (<https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp>).
 
-## Species origin as native, archaeophyte and neophyte
+## Species origin as native, archaeophyte, or neophyte
 
-We classified the species as native, archaophytes and neophytes using the information in *Flora iberica* [-@RN3243] and catalogues of archaeophytes for Britain [@RN5490] and the Czech Republic [@RN5495].
-We must stress that identifying archaeophytes is southern Europe is highly problematic and our classification must be taken as an indication of putative archaeophyte character for the purposes of vegetation description, rather than a definitive classification of the species.
+We classified the species as native, archaeophytes, or neophytes in the Iberian Atlantic ecoregion using the information in *Flora iberica* [-@RN3243] and catalogues of archaeophytes for Britain [@RN5490] and the Czech Republic [@RN5495].
+We must stress that identifying archaeophytes in southern Europe is highly problematic [@RN5444] and our classification must be taken as an indication of putative archaeophyte character for the purposes of vegetation description and comparison with other European regions, rather than a definitive classification of the species; as stated by Preston *et al.* [@RN5490], the classification of a species as an archaeophyte should be interpreted as a hypothesis to be tested by further studies.
 
@@
 
-From *FloraVeg.EU* [-@RN5385] we extracted the species’ life forms and kept *therophyte* and *geophyte* for further analysis since these two categories had the largest contribution to variance as per an exploratory Principal Component Analysis (PCA) performed using the R package *FactoMineR* [@RN3166].
+Using *FloraVeg.EU* [-@RN5385] we extracted the species’ life forms and kept *therophyte* and *geophyte* for further analysis since these two categories had the largest contribution to variance among alliances, as per an exploratory Principal Component Analysis (PCA) performed using the R package *FactoMineR* [@RN3166].
 
-## Species and vegetation height and flowering phenology
+## Species height and flowering phenology
 
-From *Flora iberica* [-@RN3243] we extracted the maximum height, median month of flowering and length of the flowering period of each species in the dataset.
-We used these to calculate plot-level community-weighted means (weighting by species cover) for each trait, to characterize the vegetation height and flowering phenology of the anthropogenic plant communities.
+From *Flora iberica* [-@RN3243] we extracted the maximum height, median month of flowering, and length of the flowering period of each species in the dataset.
+We used these variables to calculate plot-level community-weighted means (weighting by species cover) for each trait, to characterize the vegetation height and flowering phenology of the anthropogenic plant communities.
 
@@
 
-We collected, for each species in the dataset, its ecological indicator values of temperature, moisture, light, nutrients and soil reaction [@RN5384]; and its disturbance frequency and severity indicator values [@RN5101].
+For each species in the dataset, we collected its ecological indicator values of temperature, moisture, light, nutrients, and soil reaction [@RN5488]; and its disturbance frequency and severity indicator values [@RN5101].
 To those species lacking an indicator value in the accessed references, we assigned a value by reciprocal averaging, i.e.
@@
 
-The semi-supervised classification resulted in the classification of 2,081 vegetation plots into 28 anthropogenic alliances of the Iberian Atlantic ecoregion (**Table 1**, **Figure 1**).
-Most of the alliances had a wide distribution in the ecoregion according to our dataset, although five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis*, *Paspalo-Agrostion semiverticillati*) had only isolated occurrences in a few points towards the southern limit of the Iberian Atlantic territories.
-The regional literature included in our dataset recognized 72 associations (**Supplementary 4**) within these alliances.
+The semi-supervised classification resulted in the classification of 2,081 vegetation plots into 28 anthropogenic alliances (**Fig.
+1**, **Appendix S4**) representing 9 vegetation classes.
+Most of the alliances had a wide distribution in the ecoregion according to our dataset, but five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis* and *Paspalo-Agrostion semiverticillati*) had isolated occurrences towards the southern limit of the Iberian Atlantic territories.
+The regional literature included in our dataset recognized 72 associations (**Appendix S5**) within these alliances, but exploratory analysis (not shown) indicated that the separation of these associations within the alliances had a generally weak support based on numerical classification methods.
 
-Focusing on the classified plots, the dataset included 1,167 taxa or taxa aggregates.
-Of these, 101 were neophytes (i.e.
-9% of the species pool).
-Regarding life forms, the species pool was dominated by hemicryptophytes (43%) and therophytes (41%).
-The 10 most frequent species were *Ochlopoa annua* (621 occurrences), *Urtica dioica* (592), *Sonchus oleraceus* (540), *Stellaria media* (516), *Capsella bursa-pastoris* (390), *Polygonum aviculare* (375), *Dactylis glomerata* (363), *Senecio vulgaris* (332) and *Anisantha sterilis* (301).
-Considering only those plots with the most frequent plot size (20 m^2^, n = 256 plots), the average species richness per plot was 16 (minimum = 3, maximum = 37).
+The classified plots included 1,162 taxa or taxa aggregates.
+The 10 most frequent species were *Ochlopoa annua* (621 occurrences), *Urtica dioica* (592), *Sonchus oleraceus* (540), *Stellaria media* (516), *Capsella bursa-pastoris* (390), *Polygonum aviculare* (375), *Dactylis glomerata* (363), *Senecio vulgaris* (332), and *Anisantha sterilis* (301).
+Considering only those plots with the most frequent plot size range (10-30 m^2^, n = 867 plots), the average species richness per plot was 16 (minimum = 3, maximum = 48).
+The class with the richest species pool was *Epilobietea angustifolii* (n = 615) and the poorest was *Cymbalario-Parietarietea diffusae* (n = 171).
 
-DCA ordination of the floristic composition of the classified plots (**Fig.
-2**) suggested a high consistency of the vegetation classes *Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Sisymbrietea* and *Bidentetea*.
-The alliances belonging to the two perennial ruderal vegetation classes (*Artemisietea vulgaris* and *Epilobietea angustifolii*) were mixed together.
-The winter-annual class *Chenopodietea* also showed high heterogeneity, especially because of the isolation of the altered dune alliance *Linario polygalifoliae-Vulpion alopecuri*.
+PCA ordination of the floristic composition (**Fig.
+2A**) clearly differentiated three vegetation groups: (1) trampled vegetation of class *Polygono-Poetea annuae*; (2) crop weeds of classes *Papaveretea rhoeadis* and *Digitario sanguinalis-Eragrostietea minoris*; and (3) perennial ruderal vegetation of classes *Artemisietea vulgaris* and *Epilobietea angustifolii*.
+The central position of the floristic space, between these three major groups, was occupied by the annual ruderal vegetation of classes *Chenopodietea*, *Sisymbrietea*, and *Bidentetea*; as well as by the wall vegetation of class *Cymbalario-Parietarietea diffusae*.
+Separate PCAs within the major vegetation groups indicated a coherent separation between most of the alliances, but with some cases of relatively high overlap: for example, between (i) *Galio valantiae-Parietarion judaicae* and *Cymbalario-Asplenion*; or (ii) *Geo urbani-Alliarion officinalis* and *Aegopodion podagrariae*.
 
-## Community-level traits and ecological preferences
+## Proportion of natives, archaeophytes and neophytes
 
-PCA ordination of the community-weighted means of the species traits (**Fig.
-3A**) indicated that the first two axes explained 65% of the variability.
-The major contributors to axis 1 (41% variance explained) were community height, proportion of geophytes, mean month of flowering and proportion of therophytes; it separated taller communities with more geophytes and later flowering (classes *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Bidentetea*) from shorter communities with more therophytes and earlier flowering (classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis* and *Sisymbrietea*).
-The major contributors to axis 2 (24% variance explained) were proportion of neophytes and width of the flowering season; it separated communities with more neophytes and longer flowering seasons (classes *Bidentetea*, *Digitario sanguinalis-Eragrostietea minoris* and *Polygono-Poetea annuae*) from communities with less neophytes and shorter flowering seasons (classes *Artemisietea vulgaris*, *Epilobietea angustifolii* and *Chenopodietea*).
+The total synanthropic species pool (n = 1,162) was composed of 78% natives (n = 908), 15% putative archaeophytes (n = 178), and 7% neophytes (n = 76).
+The most frequent putative archaeophytes were *Capsella bursa-pastoris* (390 occurrences), *Senecio vulgaris* (332), *Anisantha sterilis* (301), *Anthemis arvensis* (280), *Hordeum murinum* (274), *Raphanus raphanistrum* (233), *Vicia sativa* (228), *Malva sylvestris* (197), *Sisymbrium officinale* (166), and *Echinochloa crus-galli* (165); while the most frequent neophytes were *Veronica persica* (212), *Erigeron canadensis* (138), *Amaranthus hybridus* (73), *Oxalis corniculata* (67), *Lepidium didymum* (57), *Oxalis latifolia* (51), *Paspalum distichum* (48), *Bidens frondosus* (35), *Cyperus eragrostis* (34), and *Symphyotrichum squamatum* (33).
 
-PCA ordination of the community-weighted means of the species ecological and disturbance preferences (**Fig.
-3B**) indicated that the first two axes explained 56% of the variability.
-The major contributors to axis 1 (35% variance explained) were disturbance frequency, soil disturbance, mowing frequency and light requirements; it separated communities thriving with frequent disturbances in open situations (classes *Polygono-Poetea annuae*, *Sisymbrietea*, *Digitario-Eragrostietea* and *Papaveretea rhoeadis*) from communities preferring less frequent disturbances and shadier areas (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii* and *Bidentetea*).
-The major contributors to axis 2 (21% variance explained) were nutrient requirements and disturbance severity; it separated more nitrophilous communities adapted to more severe disturbances (classes *Digitario-Eragrostietea*, *Sisymbrietea* and *Epilobietea angustifolii*) from less nitrophilous communities adapted to lower intensity disturbances (classes *Cymbalario-Parietarietea diffusae*, *Artemisietea vulgaris* and *Chenopodietea*).
+By vegetation class, *Epilobietea angustifolii* had the highest proportion of natives in its species pool (82%); *Sisymbrietea* had the highest proportion of putative archaeophytes (31%); and *Bidentetea* had the highest proportion of neophytes (17%).
+The alliances with the highest proportions of archaeophytes in their species pools were *Caucalidion lappulae* (41%), *Chenopodion muralis* (35%%), and *Sisymbrion officinalis* (31%); while the alliances with the highest proportion of neophytes were *Bidention tripartitae* (19%), *Allion triquetri* (12%), and *Spergulo arvensis-Erodion cicutariae* (10%).
 
-## Classified anthropogenic plant communities
+At the community level, focusing on plot average proportions (**Fig.
+3**), the alliances with the most archaeophytes were *Caucalidion lappulae*, *Spergulo arvensis-Erodion cicutariae*, and *Sisymbrion officinalis*, in which only about half of the plot-level proportion was made up of native species.
+The alliances with the highest plot-level proportion of neophytes were *Bidention tripartitae*, *Paspalo-Agrostion semiverticillati*, and *Senecionion fluviatilis*.
+Some alliances had high proportions of native species, e.g.
+*Epilobion angustifolii*, *Convolvulo arvensis-Agropyrion repentis*, and *Linario polygalifoliae-Vulpion alopecuri*.
 
-***Cymbalario-Parietarietea diffusae*** Vegetation of human-made walls.
+## Community-level traits
 
-1.
-*Galio valantiae-Parietarion judaicae* Vegetation of dry walls.
-Occupies warmer, drier and more nutrient-rich situations than the other alliance in the class, has a shorter flowering season and is more adapted to disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Parietaria judaica* aggr., *Centranthus ruber*, *Asplenium trichomanes*.
+The most frequent life form in the species pool (n = 1,162 species) were hemicryptophytes (44%), followed by therophytes (42%), geophytes (9%), chamaephytes (6%), phanerophytes (6%), bryophytes (1%), and hydrophytes (1%).
+The proportion of therophytes across classes and alliances (**Fig.
+4**) agreed with the traditional description of the syntaxa: therophytes dominated the annual communities of trampled-soil vegetation (*Polygono-Poetea annuae*), crops weeds (*Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*), and annual ruderals (*Chenopodietea*, *Sisymbrietea*, *Bidentetea*); while perennial life forms (especially hemicryptophytes) dominated walls (*Cymbalario-Parietarietea diffusae*) and perennial ruderal vegetation (*Artemisietea vulgaris*, *Epilobietea angustifolii*).
+Geophytes represented a relatively high proportion of the vegetation in three alliances: *Allion triquetri*, *Convolvulo arvensis-Agropyrion repentis*, and *Senecionion fluviatilis*.
 
-2.
-*Cymbalario-Asplenion* Vegetation of mesic walls.
-Occupies colder, wetter and less nutrient-rich situations than the other alliance in the class, has a longer flowering season and is less adapted to disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Cymbalaria muralis*.
+Community-weighted means for plant height (**Fig.
+5A**) also agreed with the expected description of the syntaxa: the shortest communities were the dwarf-herb vegetation of trampled sites (*Polygono-Poetea annuae*, average height = 45.2 cm) and crop weeds (*Papaveretea rhoeadis*, 55.4 cm), while the tallest were the perennial ruderal vegetation of dry (*Artemisietea vulgaris*, 129 cm) and mesic (*Epilobietea angustifolii*, 136 cm) sites.
+The median month of flowering (**Fig.
+5B**) corresponded to May (*Sisymbrietea*, *Chenopodietea*, *Papaveretea rhoeadis*), June (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Epilobietea angustifolii*, *Artemisietea vulgaris*) or July (*Digitario sanguinalis-Eragrostietea minoris*, *Bidentetea*).
+The length of the flowering season (**Fig.
+5C**) was generally high, from 5 months (*Artemisietea vulgaris*, *Epilobietea angustifolii*) to 6 (*Chenopodietea*, *Sisymbrietea*), 7 (*Bidentetea*, *Digitario sanguinalis-Eragrostietea minoris*, *Papaveretea rhoeadis*), 8 (*Polygono-Poetea annuae*), or 9 (*Cymbalario-Parietarietea diffusae*).
 
-***Polygono-Poetea annuae*** Dwarf vegetation of heavily-trampled sites.
+## Community-level ecological preferences
 
-3.
-*Polycarpion tetraphylli* Dwarf-annual trampled vegetation of warm and sunny sites.
-Occupies warmer, sunnier and less nutrient-rich situations than the other alliances in the class and has an earlier flowering season (**Fig.
-4**).
-Wide distribution in the ecoregion, specially along the coast and in the more Mediterranean valleys of the interior (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Sagina apetala*, *Plantago coronopus*, *Crassula tillaea*, *Spergularia marina*.
-
-4.
-*Polygono-Coronopodion* Dwarf-annual trampled vegetation of dry sites.
-Occupies colder, drier and more nutrient-rich situations than the other alliances in the class, and has a higher proportion of neophytes (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Matricaria suaveolens*, *Polygonum aviculare*.
-
-5.
-*Saginion procumbentis* Dwarf trampled vegetation of mesic sites.
-Occupies colder, wetter and more nutrient-rich situations than the other alliance in the class, and is less dominated by therophytes (**Fig.
-4**).
-Distribution concentrated in the central part of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Sagina procumbens*, *Bryum argenteum*.
-
-***Papaveretea rhoeadis*** Annual weed vegetation of crops and gardens.
-
-6.
-*Caucalidion lappulae* Annual weed vegetation of cereal crops on base-rich soils.
-Adapted to more frequent and severe disturbances than the other alliances in the class (**Fig.
-4**).
-Distribution mostly limited to the more Mediterranean valleys south of the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Cyanus segetum*, *Papaver rhoeas*, *Trifolium arvense*, *Anacyclus clavatus*, *Valerianella eriocarpa*.
-
-7.
-*Scleranthion annui* Annual weed vegetation of cereal crops on base-poor soils.
-Occupies more acidic soils than the other alliances in the class, and flowers earlier (**Fig.
-4**).
-Wide distribution in the ecoregion, but specially in the west, where acidic bedrocks dominate (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Mibora minima*, *Rumex acetosella*.
-
-8.
-*Oxalidion europeae* Annual weed vegetation of gardens and root crops.
-Occupies wetter, shadier and more nutrient-rich situations than the other alliances in the class, has a longer flowering season and a higher proportion of neophytes (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Stellaria media*, *Veronica persica*.
-
-***Digitario sanguinalis-Eragrostietea minoris*** Summer-annual C4 weed vegetation.
-
-9.
-*Spergulo arvensis-Erodion cicutariae* Summer-annual C4 weed vegetation.
-Late-flowering alliance rich in neophytes and grasses, occurring in highly-disturbed and nutrient-rich soils (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Digitaria sanguinalis*, *Echinochloa crus-galli*, *Chenopodium album*, *Amaranthus hybridus* aggr.
-
-***Chenopodietea*** Winter-annual ruderal vegetation.
-
-10.
-*Allion triquetri* Geophyte-rich fringe vegetation.
-Occupies warmer and shadier situations than the other alliances in the class, has a higher proportion of geophytes and a lower proportion of therophytes, and is adapted to less frequent disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Urtica membranacea*, *Smyrnium olusatrum*, *Tradescantia fluminensis*.
-
-11.
-*Geranio pusilli-Anthriscion caucalidis* Winter-annual fringe vegetation.
-Occupies situations similar to the previous alliance, but in colder sites, and is dominated by therophytes instead of geophytes (**Fig.
-4**).
-Mostly distributed in the more Mediterranean valleys south of the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Geranium lucidum*, *Anthriscus caucalis*.
-
-12.
-*Chenopodion muralis* Low-growth winter-annual ruderal vegetation.
-Occupies colder situations than the other alliances in the class and more often disturbed sites, and has a longer and later flowering season (**Fig.
-4**).
-Sparse occurrence throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Malva neglecta*, *Urtica urens*.
-
-13.
-*Echio-Galactition tomentosae* Tall-herb winter-annual ruderal vegetation.
-Occupies drier, sunnier and less nutrient-rich situations than the other alliances in the class (**Fig.
-4**).
-Mostly distributed in the west of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Galactites tomentosus*, *Coleostephus myconis*, *Anisantha rigida*.
-
-14.
-*Linario polygalifoliae-Vulpion alopecuri* Ephemeral annual vegetation of disturbed coastal dunes.
-Occupies drier and more nutrient-poor situations than the other alliances in the class (**Fig.
-4**) and has a high number of coastal dune species.
-Isolated occurrence in the southern coasts of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Vulpia alopecuros*, *Malcolmia littorea*, *Crucianella maritima*, *Artemisia campestris* aggr.
-
-***Sisymbrietea*** Summer-annual ruderal vegetation.
-
-15.
-*Sisymbrion officinalis* Summer-annual ruderal vegetation.
-Rich in therophytes and adapted to severe disturbances (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Hordeum murinum*, *Sisymbrium officinale*.
-
-***Bidentetea*** Summer-annual pioneer vegetation of temporarily-flooded sites.
-
-16.
-*Bidention tripartitae* Summer-annual pioneer vegetation of temporarily-flooded sites.
-Occupies nutrient-richer sites than the other alliance in the class and has a higher proportion of therophytes (**Fig.
-4**).
-Dispersed throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Persicaria hydropiper*, *Bidens frondosus*, *Lythrum salicaria*.
-
-17.
-*Paspalo-Agrostion semiverticillati* Summer-annual pioneer vegetation of temporarily-flooded and warm sites.
-Occupies warmer and wetter situations than the other alliance in the class, has a higher proportion of neophytes and has a longer flowering season (**Fig.
-4**).
-Isolated occurrences in the south of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Paspalum distichum*, *Cyperus longus*, *Schoenoplectus lacustris*.
-
-***Artemisietea vulgaris*** Perennial ruderal vegetation of dry sites.
-
-18.
-*Convolvulo arvensis-Agropyrion repentis* Semiruderal grasslands.
-Occupies wetter and shadier situations than the other alliances in the class, and has a higher proportion of geophytes and grasses (**Fig.
-4**).
-Isolated occurrence in the south of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Poa compressa*, *Elytrigia repens*, *Potentilla reptans*, *Valerianella locusta*.
-
-19.
-*Carduo carpetani-Cirsion odontolepidis* Thistle ruderal vegetation of warm sites.
-Occupies warmer situations than the other thistle alliance in the class (**Fig.
-4**).
-Mostly distributed in the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Onopordum acanthium*.
-
-20.
-*Cirsion richterano-chodati* Thistle ruderal vegetation of cold sites.
-Occupies colder situations than the other thistle alliance in the class (**Fig.
-4**).
-Mostly distributed in the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Cirsium eriophorum*, *Carduus carpetanus*, *Carduus nutans* aggr.
-
-21.
-*Dauco-Melilotion* Biennial ruderal vegetation.
-Occupies wetter situations than the two thistle alliances in the class, and has a higher proportion of therophytes (**Fig.
-4**).
-Sparse occurrence throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Helminthotheca echioides*, *Daucus carota*, *Melilotus albus*, *Foeniculum vulgare*.
-
-***Epilobietea angustifolii*** Perennial ruderal vegetation of mesic to wet sites.
-
-22.
-*Geo urbani-Alliarion officinalis* Low-herb short-lived semiruderal and fringe vegetation.
-Occupies shadier and less disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants (**Fig.
-4**).
-Wide distribution in the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Urtica dioica*.
-
-23.
-*Arction lappae* Low-herb short-lived ruderal vegetation.
-Occupies sunnier and more frequently-disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants (**Fig.
-4**).
-Distribution concentrated in the Cantabrian Mountains (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Oxybasis rubra*, *Senecio duriaei*.
-
-24.
-*Balloto-Conion maculati* Tall-herb perennial ruderal vegetation.
-Occupies warmer situations than the other alliances in the class (**Fig.
-4**).
-Distribution concentrated in the Cantabrian Mountains and the east of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Sambucus ebulus*.
-
-25.
-*Aegopodion podagrariae* Tall-herb perennial vegetation of forest margins and clearings.
-Occupies shadier and less disturbed situations than the other alliances in the class (**Fig.
-4**).
-Sparse occurrence throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Anthriscus sylvestris*, *Heracleum sphondylium*, *Galium aparine*.
-
-26.
-*Epilobion angustifolii* Tall-herb perennial vegetation of forest margins and clearings in acidic soils.
-Occupies colder, nutrient-poorer, more acidic and less disturbed situations than the other alliances in the class (**Fig.
-4**).
-Distributed towards the west of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Epilobium angustifolium*, *Digitalis purpurea*, *Luzula lactea*.
-
-27.
-*Cynancho-Convolvulion sepium* Tall-herb vegetation of nutrient-rich riparian habitats.
-Occupies wetter situations and less disturbed situations than the other alliances in the class, and has a later flowering season (**Fig.
-4**).
-Dispersed throughout the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Eupatorium cannabinum*, *Angelica sylvestris*, *Picris hieracioides*.
-
-28.
-*Senecionion fluviatilis* Tall-herb vegetation of nutrient-rich river banks and ditches.
-Occupies warmer and wetter situations than the other alliances in the class, flowers later, has taller plants and is highly dominated by neophytes and geophytes (**Fig.
-4**).
-Isolated occurrence in the western coast of the ecoregion (**Fig.
-1**).
-Diagnostic species (**Supplementary 5**): *Arundo donax*, *Silene latifolia*.
+To visualize the major patterns of variation in the ecological and disturbance preferences of the different anthropogenic communities (**Fig.
+6**) we performed a PCA ordination of plot-level means (**Fig.
+7**).
+The first two PCA axes explained 59% of the variability.
+The major contributors to axis 1 (36% variance explained) were light, nutrient, and moisture requirements; it separated communities of open, dry and comparatively nutrient-poorer sites (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea*, *Artemisietea vulgaris*, *Chenopodietea*, and *Digitario sanguinalis-Eragrostietea minoris*) from communities of shady, moist and nutrient-richer sites (classes *Cymbalario-Parietarietea diffusae*, *Bidentetea*, and *Epilobietea angustifolii*).
+The major contributors to axis 2 (23% variance explained) were disturbance severity and frequency; this axis separated communities preferring less severe and less frequent disturbances (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii*, *Chenopodietea*, *Artemisietea vulgaris* and *Bidentetea*) from communities adapted to more severe and more frequent disturbances (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea* and *Digitario sanguinalis-Eragrostietea minoris*).
 
@@
 
-The major driver of variation in trampled community composition is soil moisture [@RN5434].
+Our synthesis of the anthropogenic vegetation at the ecoregion level has detected a discrepancy between the number of vegetation units described in the literature (i.e.
+38 alliances) and the number supported by numerical classification (28 alliances).
+Our classification supported that anthropogenic vegetation can be broadly divided into three groups belonging to trampled, weed, and ruderal communities.
+These groups are organized along a principal axis of variation related to abiotic stress, and a second axis related to disturbance.
+These two axes determine the characteristics of the communities in terms of species origins and species traits.
+This synthesis allows for a better understanding of anthropogenic vegetation, as a first step towards a better integration of this biodiversity asset into ecosystem management and nature-based solutions.
 
-In medieval Switzerland, summer and winter weeds communities could not be distinguished [@RN5431].
+## Vegetation and habitat classification of anthropogenic communities
 
-The class *Digitario sanguinalis-Eragrostietea minoris* includes thermophilic vegetation dominated by C4 plants that occupies arable land as well as trampled and ruderal sites [@RN5426].
+Our analysis supports the existence of 28 anthropogenic alliances in the vegetation of the Iberian Atlantic ecoregion.
+While the core of this vegetation diversity is made up of temperate European alliances, these are enriched by the occurrence of eight Mediterranean alliances (*Galio valantiae-Parietarion judaicae*, *Polycarpion tetraphylli*, *Allion triquetri*, *Chenopodion muralis*, *Echio-Galactition tomentosae*, *Paspalo-Agrostion semiverticillati*, *Balloto-Conion maculati* and *Cynancho-Convolvulion sepium*); plus three alliances that are endemic of the Iberian coasts (*Linario polygalifoliae-Vulpion alopecuri*) and mountains (*Carduo carpetani-Cirsion odontolepidis*, *Cirsion richterano-chodati*) [@RN4698].
+Seven of these alliances had been reported as absent or uncertain in the ecoregion (*Caucalidion lappulae*, *Chenopodion muralis*), or only present in the Portuguese sector of the ecoregion (*Spergulo arvensis-Erodion cicutariae*, *Allion triquetri*, *Echio-Galactition tomentosae*, *Carduo carpetani-Cirsion odontolepidis*, *Cynancho-Convolvulion sepium*) by the recently published distribution maps of vegetation alliances in Europe [@RN5371].
+Our dataset supports that six of these alliances are relatively well distributed in the ecoregion, while *Caucalidion lappulae* only has isolated occurrences.
+On the other hand, our dataset does not support the occurrence in the ecoregion of nine Mediterranean alliances (*Roemerion hybridae*, *Diplotaxion erucoidis*, *Euphorbion prostratae*, *Alysso granatensis-Brassicion barrelieri*, *Hordeion murini*, *Taeniathero-Aegilopion geniculatae*, *Parietarion lusitanico-mauritanicae*, *Onopordion castellani*, *Silybo mariani-Urticion piluliferae*) and one temperate alliance (*Fragarion vescae*) that had been cited as potentially occurring here [@RN5363; @RN5008; @RN5365; @RN4698; @RN5371].
+For habitat classification purposes, we propose that this vegetation diversity can be summarized into 15 EUNIS habitat types: one constructed habitat (*J2.5*), ten vegetated human-made habitats (*V11*, *V12*, *V13*, *V15*, *V32*, *V35*, *V34*, *V37*, *V38*, *V39*), three habitats of woodland fringes and clearings and tall forb stands (*R54*, *R55*, *R57*) and one habitat of periodically-exposed shores (*Q61*).
+In general, we have followed the correspondence between alliances and EUNIS habitats proposed by the EUNIS expert system [@RN5061], except for *Echio-Galactition tomentosae* which, in the context of the ecoregion, seems to be more adequately placed in the habitat *V37 Annual anthropogenic herbaceous vegetation* rather than in the habitat *V33 Dry Mediterranean land with unpalatable non-vernal herbaceous vegetation*.
 
-Anthropogenic vegetation composition is sensitive to micro-scale management decisions, such as the ownership of domestic dogs [@RN5438].
+Since our goal was to match the classification of the ecoregion classification to the EuroVegChecklist [@RN4698], we refrained from merging recognized alliances or changing their assignation to superior syntaxa.
+However, there are some cases that could deserve further scrutiny.
+For example, the two wall alliances (*Galio valantiae-Parietarion judaicae* and *Cymbalario-Asplenion*) had a large overlap in their geographic distribution, their floristic composition, their proportion of alien species, their community traits, and their ecological preferences.
+These two alliances are well recognized in the European literature, and they are generally interpreted as representing the Mediterranean (*Galio valantiae-Parietarion judaicae*) and temperate (*Cymbalario-Asplenion*) versions of wall vegetation [@RN5369; @RN5367].
+It is therefore not surprising that in biogeographic transitional regions - such as the Iberian Atlantic territories - these two alliances coexist and show a high overlap.
+Also, there was overlap between the alliances *Geo urbani-Alliarion officinalis* and *Aegopodion podagrariae*, two alliances that can occupy ruderal to semi-natural nitrophilous fringes, and which are mostly differentiated by dominant vegetation height, as was supported by our dataset.
 
-Neophytes were more represented in early successional anthropogenic communities [@RN5423].
+## Characterization of anthropogenic communities
 
-Arable land plants were mainly therophytes and alien species; while ruderal habitats were richer in perennials and C strategists [@RN5436].
+We found 22% non-native species (including neophytes and archaeophytes) in the anthropogenic vegetation of the Iberian Atlantic ecoregion.
+This percentage is intermediate between that found in urban areas of temperate Europe and North America (30-50%) [@RN5498; @RN5499; @RN5500; @RN5079] and Italy (12-26%) [@RN5501].
+A survey of anthropogenic vegetation in the north-west Balkans found 13% non-native plants, with more neophytes than archaeophytes [@RN5429].
+A similar study in the Czech Republic found 32% archaeophytes and 7% neophytes [@RN5435].
+Taken together, these results suggest that the Iberian Atlantic anthropogenic vegetation shows an incidence of non-native species that is intermediate between southern and central-northern Europe, which is in agreement with the biogeographical position of our ecoregion.
+Previous studies also suggested that the non-natives were more frequent in early successional anthropogenic communities [@RN5423], and our results confirmed this as the proportion of native species tended to be higher in the perennial ruderal classes *Artemisietea vulgaris* and *Epilobietea angustifolii*.
+The classes with more non-native species also tended to be those adapted to more frequent and more severe disturbances (*Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Sisymbrietea*), also in agreement with previous research [@RN5435].
+Archaeophytes had a higher proportion in the vegetation associated to arable habitats (*Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Sisymbrietea*), in agreement with the ancient origin of archaeophytes during the development of agriculture [@RN5494; @RN5490].
+On the other hand, neophytes tended to be in higher proportions in vegetation types associated to wet conditions (*Bidentetea*), as has been found in other European regions [@RN5497].
 
-Compare proportion of neophytes to other areas [@RN5429; \@ @RN5435]
+The traits of the Iberian Atlantic anthropogenic communities agreed with those found in other European regions [@RN5436; @RN2297], with more annual species in habitats associated to arable land and more perennials in ruderal habitats, and a general dominance of therophytes and hemicryptophytes in the species pool [@RN5428].
+We found that the main drivers of variation in anthropogenic community composition was a gradient from open-dry to shady-wet sites, supporting the importance of the moisture gradient in shaping anthropogenic vegetation [@RN5434].
+Mesic and moist human-made habitats of the Iberian Atlantic territories harbor perennial communities, some of which are especially rich in native species and have semi-natural characters.
+At the same time, the wetter habitats have some of the higher prevalences of non-native species.
 
-Neophytes have no special preferences for local habitat conditions and their highest proportion was found mainly in disturbed habitats at low elevations. [@RN5435]
+# Conclusions
 
-Therophytes and hemicryptophytes were most abundant in the life form spectra [@RN5428]
-
-Segetal vegetation of Palestine [@RN5494].
-
-History of synanthropic veg in czech [@RN5496]
-
-# Data availability
-
-The original datasets, as well as R code for analysis and creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/manmade>.
-Upon publication, a version of record of the repository will be deposited in Zenodo.
+Ecoregion-level synthesis detected a discrepancy between the number of anthropogenic vegetation units described in the literature and the number supported by numerical classification.
+The diversity of anthropogenic vegetation can be organized into trampled, weed and ruderal communities varying primarily along an axis of abiotic stress.
+Importantly, the anthropogenic vegetation of the Iberian Atlantic ecoregion is home to 1,162 plant taxa.
+This is approximately one third of the whole ecoregion species pool represented in the SIVIM database [@RN5374], and one fifth of the Iberian flora [@RN5503].
+More than half of this synanthropic diversity (615 taxa) occurs in mesic to wet perennial ruderal vegetation, highlighting the potential of some anthropogenic habitats as a source of biodiversity in human-dominated landscapes [@RN5446; @RN5445].
+These results suggest three main objectives for the sustainable management of human-made habitats: (1) focusing efforts on alien plant management in anthropogenic plant communities of wet habitats; (2) conserving examples of archaeophyte-rich annual weed communities, especially those linked to traditional agricultural practices [@RN5502]; and (3) conserving and promoting mesic ruderal and fringe vegetation as a biodiversity refuge in urban and peri-urban areas.
 
@@
 
+# Supplementary Information
+
+**Appendix S1** Checklist of anthropogenic syntaxa that could be present in the Iberian Atlantic ecoregion according to the literature.
+Syntaxa names are provided as recorded by the original authors of the vegetation plot.
+The file indicates if the presence of the syntaxon in the Iberian Atlantic ecoregion was validated or not by our analysis.
+
+**Appendix S2** Original publications providing the vegetation plots stored in SIVIM and used in our analysis.
+
+**Appendix S3** Characteristic species combinations of the anthropogenic vegetation alliances in spreadsheet format, divided into diagnostic, constant, and dominant species.
+
+**Appendix S4** Synthetic description of the 28 alliances present in the Iberian Atlantic anthropogenic vegetation.
+
+**Appendix S5** Revised checklist of vegetation associations described in the regional literature and belonging to the 28 alliances present in the Iberian Atlantic anthropogenic vegetation.
+
 # Figures
 
-```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Figure 1.
-Anthropogenic vegetation alliances of the Iberian Atlantic ecoregion.
-Each dot is a vegetation plot.
-Dot colors indicate the vegetation class."}
+```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Anthropogenic vegetation alliances of the Iberian Atlantic ecoregion. Each dot is a vegetation plot. Dot colors indicate the vegetation class."}
 knitr::include_graphics("../results/figures/F1 - maps.png")
@@
 
-```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Figure 3.
-Proportion of neophytes, archaeophytes and native species in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
-Each bar represents the plot-level average proportion of neophytes (dark shade), archaeophytes (medium shade) and natives (light shade) in each vegetation alliance.
+```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of neophytes, archaeophytes, and native species in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Each bar represents the plot-level average proportion of neophytes (dark shade), archaeophytes (medium shade), and natives (light shade) in each vegetation alliance.
 Colors indicate the vegetation class."}
@@
 
-```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Figure 4.
-Proportion of therophytes, geophytes and other life forms in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
-Each bar represents the plot-level average proportion of therophytes (dark shade), geophytes (medium shade) and other life forms (light shade) in each vegetation alliance.
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of therophytes, geophytes, and other life forms (hemicryptophytes, chamaephytes, phanerophytes, bryophytes, hydrophytes) in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Each bar represents the plot-level average proportion of therophytes (dark shade), geophytes (medium shade), and other life forms (light shade) in each vegetation alliance.
 Colors indicate the vegetation class."}
@@
 
-```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Figure 5.
-Vegetation heigh and flowering phenology of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
-Boxplots show the community-weighted mean for plant heigh, month of peak flowering and length of the flowering season for each vegetation alliance.
-Colors indicate the vegetation class."}
+```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Vegetation height and flowering phenology of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Boxplots show the community-weighted mean for plant height, median month of flowering, and length of the flowering season for each vegetation alliance.
+Colors indicate the vegetation class.
+Alliances as ordered as in Figures 1 and 3-4, and alliance numbers are the same as listed in Appendix S4."}
 knitr::include_graphics("../results/figures/F5 - height-flowering.png")
@@
 
-```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Figure 6.
-Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
-Boxplots show the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity, calculated for each vegetation alliance.
-Colors indicate the vegetation class."}
+```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Boxplots show the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity, calculated for each vegetation alliance.
+Colors indicate the vegetation class.
+Alliances as ordered as in Figures 1 and 3-4, and alliance numbers are the same as listed in Appendix S4."}
 knitr::include_graphics("../results/figures/F6 - indicators.png")
@@
 
-```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Figure 7.
-Pattern of variation in the ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
-Biplot produced by Principal Component Analysis (PCA) of the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.
+```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+Biplot produced by Principal Component Analysis (PCA) of the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity.
 Labels and arrows indicate the contribution of each indicator to the first and second principal components.

### 2024-06-12 doc/title-page.Rmd
submission-ready 🏙️

---
title: "Classification and characterization of anthropogenic plant communities at the ecoregion level"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: journal-of-vegetation-science.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

Running title: Iberian Atlantic anthropogenic vegetation

Eduardo Fernández-Pascual^1,a^ (<https://orcid.org/0000-0002-4743-9577>), Víctor González-García^1^ (<https://orcid.org/0000-0002-8949-7943>), Greta Ivesdal^2^, Adrián Lázaro-Lobo^1^ (<https://orcid.org/0000-0002-0509-2085>), Borja Jiménez-Alfaro^1^ (<https://orcid.org/0000-0001-6601-9597>)

^1^ Biodiversity Research Institute (IMIB), University of Oviedo - CSIC - Principality of Asturias, Mieres, Spain; ^2^ Jardín Botánico Atlántico, Gijón/Xixón, Spain

Correspondence: Eduardo Fernández-Pascual, Instituto Mixto de Investigación en Biodiversidad, Campus de Mieres, Edificio de Investigación, 5ª planta, c/ Gonzalo Gutiérrez Quirós s/n, E-33600 Mieres, Spain. Email: [fernandezpeduardo\@uniovi.es](mailto:fernandezpeduardo@uniovi.es){.email}. Telephone: +34985104781.

# Author contributions

Eduardo Fernández-Pascual: Conceptualization; Methodology; Data Curation; Formal Analysis; Visualization; Writing – Original Draft Preparation; Writing – Review & Editing. Víctor González-García: Methodology; Investigation; Formal Analysis; Writing – Review & Editing. Greta Ivesdal: Investigation; Formal Analysis; Writing – Review & Editing. Adrián Lázaro-Lobo: Investigation; Writing – Review & Editing. Borja Jiménez-Alfaro: Methodology; Writing – Review & Editing.

# Funding

Ayuntamiento de Gijón/Xixón, Jardín Botánico Atlántico (SV-23-GIJON-JBA).

# Conflict of interest statement

None declared.

# Data availability statement

Original datasets, R code for analysis and Rmarkdown code for creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/anthropogenic>. Upon publication, a version of record of the repository will be deposited in Zenodo.

### 2024-06-12 results/supporting/AppendixS4.Rmd
submission-ready 🏙️

---
title: "Appendix S4 Synthetic description of the 28 alliances present in the Iberian Atlantic anthropogenic vegetation."
output: 
  bookdown::pdf_document2:
    number_sections: false
    toc: false
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

***Cymbalario-Parietarietea diffusae*** Vegetation of human-made walls.

1.  *Galio valantiae-Parietarion judaicae* Hemicryptophyte-rich vegetation of walls. Occupies slightly warmer and sunnier situations than the other alliance in the class and has a shorter flowering season. Wide distribution in the ecoregion. Diagnostic species in the ecoregion: *Parietaria judaica*, *Centranthus ruber*, *Asplenium trichomanes*. Correspondence with EUNIS (2012) habitat *J2.5 - Constructed boundaries*.

2.  *Cymbalario-Asplenion* Fern-rich vegetation of walls. Occupies slightly colder and shadier walls than the other alliance in the class and has a longer flowering season. Wide distribution in the ecoregion. Diagnostic species in the ecoregion: *Cymbalaria muralis*. Correspondence with EUNIS (2012) habitat *J2.5 - Constructed boundaries*.

***Polygono-Poetea annuae*** Dwarf-herb vegetation of heavily-trampled sites.

3.  *Polycarpion tetraphylli* Dwarf-annual trampled vegetation of warm and sunny sites. Occupies warmer, sunnier and nutrient-poorer situations than the other alliances in the class and has an earlier flowering season. Wide distribution in the ecoregion, specially along the coast and in submediterranean valleys of the interior. Diagnostic species in the ecoregion: *Sagina apetala*, *Plantago coronopus*, *Crassula tillaea*, *Spergularia marina*. Correspondence with EUNIS habitat *V34 Trampled xeric grassland with annuals*.

4.  *Polygono-Coronopodion* Annual trampled vegetation of dry sites. Occupies sites that are colder than those preferred by *Polycarpion tetraphylli*, but drier than those of *Saginion procumbentis*. Has a specially high proportion of neophytes. Wide distribution in the ecoregion. Diagnostic species in the ecoregion: *Matricaria suaveolens*, *Polygonum aviculare*. Correspondence with EUNIS habitat *V34 Trampled xeric grassland with annuals*.

5.  *Saginion procumbentis* Bryophyte-rich trampled vegetation of mesic sites. Occupies colder and wetter situations than the other alliances in the class, is less dominated by therophytes and has a relatively high proportion of bryophytes. Distribution concentrated in the central part of the ecoregion. Diagnostic species in the ecoregion: *Sagina procumbens*, *Bryum argenteum*. Correspondence with EUNIS habitat *V35 Trampled mesophilous grassland with annuals*.

***Papaveretea rhoeadis*** Annual weed vegetation of crops and gardens.

6.  *Caucalidion lappulae* Annual weed vegetation of cereal crops on base-rich soils. Very high proportion of putative archaeophytes. Distribution mostly limited to the more submediterranean valleys south of the Cantabrian Mountains. Diagnostic species in the ecoregion: *Cyanus segetum*, *Papaver rhoeas*, *Trifolium arvense*, *Anacyclus clavatus*, *Valerianella eriocarpa*. Correspondence with EUNIS habitats *V11 Intensive unmixed crops* and *V13 Arable land with unmixed crops grown by low-intensity agricultural methods*.

7.  *Scleranthion annui* Annual weed vegetation of cereal crops on base-poor soils. Occupies more acidic soils than the other alliances in the class, and flowers earlier. High proportion of putative archaeophytes. Wide distribution in the ecoregion, but especially in the west, where acidic bedrocks dominate. Diagnostic species in the ecoregion: *Mibora minima*, *Rumex acetosella*. Correspondence with EUNIS habitats *V11 Intensive unmixed crops*, *V12 Mixed crops of market gardens and horticulture*, *V13 Arable land with unmixed crops grown by low-intensity agricultural methods* and *V15 Bare tilled, fallow or recently abandoned arable land*.

8.  *Oxalidion europeae* Annual weed vegetation of gardens and root crops. Occupies wetter, shadier and nutrient-richer situations, and has a longer flowering season, than the other alliances in the class. High proportion of putative archaeophytes, and more neophytes than the other alliances in the class. Wide distribution in the ecoregion. Diagnostic species in the ecoregion: *Stellaria media*, *Veronica persica*. Correspondence with EUNIS habitats *V11 Intensive unmixed crops*, *V12 Mixed crops of market gardens and horticulture*, *V13 Arable land with unmixed crops grown by low-intensity agricultural methods* and *V15 Bare tilled, fallow or recently abandoned arable land*.

***Digitario sanguinalis-Eragrostietea minoris*** Summer-annual C4 weed vegetation.

9.  *Spergulo arvensis-Erodion cicutariae* Summer-annual C4 weed vegetation. Late-flowering alliance rich in neophytes and grasses, occurring in highly-disturbed and nutrient-rich sites. Wide distribution in the ecoregion. Diagnostic species in the ecoregion: *Digitaria sanguinalis*, *Echinochloa crus-galli*, *Chenopodium album*, *Amaranthus hybridus*. Correspondence with EUNIS habitats *V11 Intensive unmixed crops*, *V12 Mixed crops of market gardens and horticulture* and *V13 Arable land with unmixed crops grown by low-intensity agricultural methods*.

***Chenopodietea*** Winter-annual fringe and ruderal vegetation.

10. *Allion triquetri* Geophyte-rich fringe vegetation. This vegetation occupies warmer, shadier and less frequently disturbed situations than other alliances in the class, and is comparatively taller and earlier-flowering. Has a higher proportion of geophytes and a lower proportion of therophytes. The proportion of neophytes is the highest in the class. Wide distribution in the ecoregion. Diagnostic species in the ecoregion: *Urtica membranacea*, *Smyrnium olusatrum*, *Tradescantia fluminensis*. Correspondence with EUNIS habitat *V37 Annual anthropogenic herbaceous vegetation*.

11. *Geranio pusilli-Anthriscion caucalidis* Winter-annual fringe vegetation. Occupies situations similar to the previous alliance, but in colder sites, and is a shorter vegetation dominated by therophytes instead of geophytes. Mostly distributed in the submediterranean valleys south of the Cantabrian Mountains. Diagnostic species in the ecoregion: *Geranium lucidum*, *Anthriscus caucalis*. Correspondence with EUNIS habitat *V37 Annual anthropogenic herbaceous vegetation*.

12. *Chenopodion muralis* Low-growth winter-annual ruderal vegetation. Short vegetation occupying colder situations than the other alliances in the class and more disturbed sites. Flowering season longer and later. Has a high proportion of archaeophytes. Sparse occurrence throughout the ecoregion. Diagnostic species in the ecoregion: *Malva neglecta*, *Urtica urens*. Correspondence with EUNIS habitat *V37 Annual anthropogenic herbaceous vegetation*.

13. *Echio-Galactition tomentosae* Tall-herb winter-annual ruderal vegetation. Occupies drier, sunnier and less nutrient-rich situations than other alliances in the class. Mostly distributed in the west of the ecoregion. Diagnostic species in the ecoregion: *Galactites tomentosus*, *Coleostephus myconis*, *Anisantha rigida*. Correspondence with EUNIS habitat *V37 Annual anthropogenic herbaceous vegetation*.

14. *Linario polygalifoliae-Vulpion alopecuri* Ephemeral annual vegetation of disturbed coastal dunes. Occupies warm, dry and sunny coastal sites, and has a high number of coastal dune species. Isolated occurrence in the southern coast of the ecoregion. Diagnostic species in the ecoregion: *Vulpia alopecuros*, *Malcolmia littorea*, *Crucianella maritima*, *Artemisia campestris*. Correspondence with EUNIS habitat *V32 Mediterranean subnitrophilous annual grassland*.

***Sisymbrietea*** Summer-annual ruderal vegetation.

15. *Sisymbrion officinalis* Summer-annual ruderal vegetation. Rich in therophytes and archaeophytes and adapted to severe disturbances. Wide distribution in the ecoregion. Diagnostic species in the ecoregion: *Hordeum murinum*, *Sisymbrium officinale*. Correspondence with EUNIS habitats *V15 Bare tilled, fallow or recently abandoned arable land* and *V37 Annual anthropogenic herbaceous vegetation*.

***Bidentetea*** Summer-annual pioneer vegetation of temporarily flooded sites.

16. *Bidention tripartitae* Summer-annual pioneer vegetation of temporarily flooded sites. Occupies nutrient-richer sites than the other alliance in the class and has a higher proportion of therophytes and neophytes. Dispersed throughout the ecoregion. Diagnostic species in the ecoregion: *Persicaria hydropiper*, *Bidens frondosus*, *Lythrum salicaria*. Correspondence with EUNIS habitat *Q61 Periodically exposed shore with stable, eutrophic sediments with pioneer or ephemeral vegetation*.

17. *Paspalo-Agrostion semiverticillati* Summer-annual pioneer vegetation of temporarily flooded and warm sites. Occupies wetter and more frequently disturbed sites than the other alliance in the class and has a longer flowering season. Isolated occurrences in the south of the ecoregion. Diagnostic species in the ecoregion: *Paspalum distichum*, *Cyperus longus*, *Schoenoplectus lacustris*. Correspondence with EUNIS habitat *R55 Lowland moist or wet tall-herb and fern fringe*.

***Artemisietea vulgaris*** Perennial ruderal vegetation of dry sites.

18. *Convolvulo arvensis-Agropyrion repentis* Semiruderal grasslands. Occupies wetter, shadier and nutrient-poorer situations than the other alliances in the class, and has a higher proportion of geophytes, grasses and native species. Isolated occurrence in the south of the ecoregion. Diagnostic species in the ecoregion: *Poa compressa*, *Elytrigia repens*, *Potentilla reptans*, *Valerianella locusta*. Correspondence with EUNIS habitat *V38 Dry perennial anthropogenic herbaceous vegetation*.

19. *Carduo carpetani-Cirsion odontolepidis* Thistle ruderal vegetation of warm sites. Occupies warmer situations than the other thistle alliance in the class and has a high proportion of archaeophytes. Mostly distributed in the Cantabrian Mountains. Diagnostic species in the ecoregion: *Onopordum acanthium*. Correspondence with EUNIS habitat *V38 Dry perennial anthropogenic herbaceous vegetation*.

20. *Cirsion richterano-chodati* Thistle ruderal vegetation of cold sites. Occupies colder situations than the other thistle alliance in the class. Mostly distributed in the Cantabrian Mountains. Diagnostic species in the ecoregion: *Cirsium eriophorum*, *Carduus carpetanus*, *Carduus nutans*. Correspondence with EUNIS habitat *V38 Dry perennial anthropogenic herbaceous vegetation*.

21. *Dauco-Melilotion* Biennial ruderal vegetation. Occupies wetter situations than the two thistle alliances in the class and has a higher proportion of neophytes. Occurrence throughout the ecoregion. Diagnostic species in the ecoregion: *Helminthotheca echioides*, *Daucus carota*, *Melilotus albus*, *Foeniculum vulgare*. Correspondence with EUNIS habitat *V38 Dry perennial anthropogenic herbaceous vegetation*.

***Epilobietea angustifolii*** Perennial fringe and ruderal vegetation of mesic to wet sites.

22. *Geo urbani-Alliarion officinalis* Low-herb short-lived semiruderal and fringe vegetation. Occupies shadier and less frequently disturbed situations than the other alliances in the class and has a higher proportion of therophytes and shorter plants. Wide distribution in the ecoregion. Diagnostic species in the ecoregion: *Urtica dioica*. Correspondence with EUNIS habitat *V39 Mesic perennial anthropogenic herbaceous vegetation*.

23. *Arction lappae* Low-herb short-lived ruderal vegetation. Occupies sunnier and more frequently disturbed situations than the other alliances in the class, and has a higher proportion of therophytes and shorter plants. Distribution concentrated in the Cantabrian Mountains. Diagnostic species in the ecoregion: *Oxybasis rubra*, *Senecio duriaei*. Correspondence with EUNIS habitat *V39 Mesic perennial anthropogenic herbaceous vegetation*.

24. *Balloto-Conion maculati* Tall-herb perennial ruderal vegetation. Occupies warmer situations than the other alliances in the class. Has a high proportion of archaeophytes. Distribution concentrated in the Cantabrian Mountains and the east of the ecoregion. Diagnostic species in the ecoregion: *Sambucus ebulus*. Correspondence with EUNIS habitat *V39 Mesic perennial anthropogenic herbaceous vegetation*.

25. *Aegopodion podagrariae* Tall-herb perennial vegetation of forest margins and clearings. Occupies shadier and less disturbed situations than the other alliances in the class. Sparse occurrence throughout the ecoregion. Diagnostic species in the ecoregion: *Anthriscus sylvestris*, *Heracleum sphondylium*, *Galium aparine*. Correspondence with EUNIS habitats *R55 Lowland moist or wet tall-herb and fern fringe* and *V39 Mesic perennial anthropogenic herbaceous vegetation*.

26. *Epilobion angustifolii* Tall-herb perennial vegetation of forest margins and clearings in acidic soils. Occupies colder, nutrient-poorer, more acidic and less disturbed situations than the other alliances in the class. Distributed towards the west of the ecoregion. Diagnostic species in the ecoregion: *Epilobium angustifolium*, *Digitalis purpurea*, *Luzula lactea*. Correspondence with EUNIS habitats *R54 Pteridium aquilinum vegetation* and *R57 Herbaceous forest clearing vegetation*.

27. *Cynancho-Convolvulion sepium* Tall-herb fringe vegetation of nutrient-rich riparian habitats. Occupies wetter and less disturbed situations than the other alliances in the class and has a later flowering season. Dispersed throughout the ecoregion. Diagnostic species in the ecoregion: *Eupatorium cannabinum*, *Angelica sylvestris*, *Picris hieracioides*. Correspondence with EUNIS habitat *R55 Lowland moist or wet tall-herb and fern fringe*.

28. *Senecionion fluviatilis* Tall-herb vegetation of nutrient-rich riverbanks and ditches. Occupies warmer and wetter situations than the other alliances in the class, flowers later, has taller plants and has a high proportion of neophytes and geophytes. Isolated occurrence in the western coast of the ecoregion. Diagnostic species in the ecoregion: *Arundo donax*, *Silene latifolia*. Correspondence with EUNIS habitat *R55 Lowland moist or wet tall-herb and fern fringe*.

### 2024-06-12 src/7-spp-numbers.R

@@ -18,6 +18,15 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
   unique() %>%
   group_by(Class) %>% tally %>% arrange(-n)
 
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise") %>% 
+  filter(Area >= 10) %>%
+  filter(Area <= 30) %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+  group_by(SIVIMID) %>%
+  summarise(l = length(Analysis.Names)) %>%
+  pull(l) %>% mean
+
 ### Species origin
 
 read.csv("data/header.csv", fileEncoding = "latin1") %>% 

### 2024-06-12 doc/title-page.Rmd
Zenodo link ⛓️

@@
 
-Original datasets, R code for analysis and Rmarkdown code for creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/anthropogenic>.
-Upon publication, a version of record of the repository will be deposited in Zenodo.
+Original datasets, R code for analysis and Rmarkdown code for creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/anthropogenic> stored at https://zenodo.org/doi/10.5281/zenodo.11613571.
+
+. Upon publication, a version of record of the repository will be deposited in Zenodo.

### 2024-10-03 doc/response.Rmd
revised classification, keeping Aegopodion

---
title: "Response letter on the manuscript Classification and characterization of anthropogenic plant communities at the ecoregion level"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
    number_sections: false
bibliography: refs.bib
csl: journal-of-vegetation-science.csl
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

# Co-ordinating Editor Comments to Author

> Dear authors, I found this study interesting, even if limited to a relatively short geographical area. I believe this study can also be relevant at the European scale; therefore, I think it is suitable to be published in AVS. However, I agree with the referees that the manuscript can be significantly improved before publication.

We would like to thank the editor and the reviewers for the time invested in reviewing our manuscript and providing constructive comments. We have revised the manuscript to incorporate all the suggestions. Please find below a point-by-point response to each comment.

# Referee 1

> In general, the survey is interesting and has been conducted correctly in the analysis procedure and the results yielded are consistent and respond to the initial questions. For that reason I consider it can be published in AVS. However, in the comments below are suggested some changes, mostly in the introduction, and other changes are highlighted in the notes over the manuscript. I recommend to accept the paper with major changes.

Thank you for your positive evaluation and your useful comments, which we have integrated into a revised version of the manuscript.

> The title does not have any indication of the geographical frame of the survey and it is relevant to indicate it to potential readers. It is in the keywords but it should be in the title.

We agree and we have changed the title to *Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*.

> The entire article is built up using the frame of the concept of ecoregion (Bailey 2004, 1989; Olson et al. 2001). About this there are two related questions: 1 What is the benefit of using such biogeographic frame in terms of how the results or the interpretations could be altered or deteriorated if not used? Is it a merely geographical description of the study area? If so, then, 2 What is the difference of considering the ecoregions instead the classical biogeographic units such as regions, in the sense of Good or Takhtatajan? It is important to quote that biogeographic regions also include vegetation units living within them as was established early in the 20th century by Flauhaut and Braun-Blanquet (Loidi 2021)

We would argue that the manuscript is built up based on the idea that this type of studies should be conducted using biogeographical units, regardless of the rank or classification system of the unit. In our specific example, focusing on the Iberian Atlantic territories (which rank would be "province", but see below for a discussion of the definition of the Iberian Atlantic territories and its correspondence with different biogeographical classifications) allows to study a territory that shares (1) a similar macroclimate and (2) a shared biogeographical history, including a history of biological invasions. This is especially important when studying anthropogenic vegetation, to define lists of regional neophytes and archaeophytes, which would be not possible if considering a larger area (for example, most "archaeophytes" and some "neophytes" would be considered "native" at the European level) or a smaller area or political unit (for example, species native to some parts of the Iberian Atlantic territories are some times considered "alien" in some other parts, e.g. *Pinus pinaster*). This can allow to make comparisons between invasion levels in different vegetation types in different regions, see for example Chytrý et al. [-@RN5497].

We have revised the usage of the term "ecoregion" throughout the manuscript, to avoid giving the false impression that we are specially interested in the ecoregion concept and classification by Olson et al. [-@RN4667], rather than in studying the Iberian Atlantic territories as a biogeographically-coherent unit.

> It seems that the plots used are original of NW Iberia, as shown in Figure 1. The boundaries of the grey area, supposedly taken from the map of Olson et al. 2001, are quite coincident with what has been profusely named as Cantabro-Atlantic subprovince by Rivas-Martínez (Rivas-Martínez et al. 2017) and related authors such as Fernández Prieto et al. 2020, 2023. I would suggest to use these units and names to prevent from nomenclatural inflation and cite the corresponding papers (see below). If we comment the names, the name “Iberian Atlantic” is not so precise as “Cantabro-Atlantic”. The former could include southern Portugal (Algarve or Vicentine Coast), which is not at all the case. To use the term Cantabrian or Cantabrico is particularly precise to indicate the geographic position of this territory. Additionally, it would be necessary to make a more formal description of the “Iberian Atlantic ecoregion” and discuss the differences from this ecoregion with the “Cantabro-Atlantic subprovince”.

We agree that the definition and name of the studied biogeographical unit requires further clarification First, we must make it clear how we define our study region: "the territories with a temperate climate in the north-western part of the Iberian Peninsula". In other words, we are following precisely the definition by Fernández Prieto et al. [-@RN5372] in *Naturalia Cantabricae* 8(2): 30-37. More specifically, we are following the geographical limits presented in Figure 1 of Fernández Prieto et al. [-@RN5372].

The concept of this biogeographical unit as presented by Fernández Prieto et al. [-@RN5372] is essentially equivalent to the *Cantabrian Mixed Forests ecoregion* by Olson et al. [-@RN4667]. The major difference between the two concepts is in the southern borders of the region, which are much more precisely drawn in Fernández Prieto et al. [-@RN5372].

However, the biogeographical unit does not correspond with Rivas-Martínez's *Cantabro-Atlantic subprovince*, as there are two deviations from that concept. First, our study region also includes Rivas-Martínez's *Orocantabrian subprovince*. Second, Rivas-Martínez's *Cantabro-Atlantic subprovince* includes a larger portion of French territory. In the third (and, to our knowledge, last) edition of the Biogeographic map of Europe [@RN5592], the *Cantabro-Atlantic subprovince* extends in France from the French Basque country in the south to the Cotentin Peninsula in Normandy. In sum, the more precise correspondence between our study region and Rivas-Martínez's biogegraphical classification would be "the Iberian part of the *European Atlantic province*". We include this equivalence in our methods section, in the subsection "Biogeographical territory under study".

Finally, we agree that, from a geographical point of view, "Iberian Atlantic" can refer to the Iberian coast down to Gibraltar, but the term "Atlantic" is used here in a bioclimatic sense, in accordance with the definition by Fernández Prieto et al. [-@RN5372] and the biogeographical unit names by Rivas-Martínez (cf. *European Atlantic province* in Rivas-Martínez et al. [-@RN5592]) and the European Union's Natura 2000 framework (cf. *Atlantic biogeographical region*, https://www.eea.europa.eu/en/analysis/maps-and-charts/biogeographical-regions-in-europe-2).

In any case, to avoid confusion, we have revised the manuscript's section "Biogeographical territory under study" to better explain the study area and its equivalence in different classifications. We have also revised the whole manuscript to avoid using the term *Iberian Atlantic ecoregion*, which mixes the concepts *Cantabrian Mixed Forests ecoregion* by Olson et al. [-@RN4667] and *Iberian Atlantic territories* by Fernández Prieto et al. [-@RN5372]. Now, we use *Iberian Atlantic territories* in accordance with Fernández Prieto et al. [-@RN5372]. We have also added the reference to Fernández Prieto et al. [-@RN5588] which further supports the definition of the studied biogeographical unit.

> One of the main results of this survey is the selection of 28 alliances (appendix S4) from an initial set of 38 that could be present in the Iberian Atlantic ecoregion. So, I understand that there are 10 alliances have not been detected. It would be interesting to list them somewhere (in appendix S2, for instance) and make clear that no data from them have been found in the datasets.

The missing 10 alliances are listed in the discussion, section *Vegetation and habitat classification of anthropogenic communities*. Some of these alliances were present in the dataset in the sense that the relevés had been assigned to syntaxa belonging to these alliances by the original authors of the relevé. However, the alliances were not supported by our numerical classification, i.e. their relevés were asigned to another alliance, or to the noise group.

> Concerning the weakly represented alliances. This is inevitable if we start with a definition of the study area prior to surveying its plant communities. There will be always some types which are found in the extreme or border of its geographical area or ecological amplitude and are, thus, weakly represented. It is an inherent problem of such a territorial planning and has to be conveniently considered. Other possibility is to choose the vegetation alliances and study them in all their geographic amplitude. This would entail that their geographical territory would be diverse but the survey would bring more accurate results about their character and differences.

We agree that the risk exist, although we believe that this risk is lessened by focusing on a biogeographical unit rather than in e.g. a political division. Furthermore, there were alliances that are only found in the extremes (geographical or ecological) of the study area, but still they were supported by the numerical classification. We believe that the fact that the missing alliances were not supported by the numerical classification indicates that the relevés that had been assigned to these alliances are poor representatives of the vegetation types, and more likely transitional communities at the extreme of the alliance distributions.

> In S2 author’s names should be corrected and homogenized e.g. Lence Paz is C. or Carmen, all the patronimic names such as Pérez, Fernández, etc,, have an accent but have instead a strange signal in the involved letter. It is Rivas-Martínez, not Rivas Martínez; Is it Romero or Romero Buján?, etc.

In Appendix S2, author names are provided exactly as they are stored in the SIVIM database. This is necessary to ensure cross-referencing between our dataset and the SIVIM database.

> In the figure 1 the area does not include any French terrotory [...] In the maps of figure 1 there is no French territory included

The figure includes a portion of French territory in the Pyrénées-Atlantiques department, following the biogeographical borders shown by Fernández Prieto et al. [-@RN5372] in their Figure 1, although our figure may be to small to fully appreciate this. We have revised the mention in the abstract to make clear that we are speaking about SW France.

> Anthropogenic syntaxa have a larger biogeographical scope than other vegetation types, particularly grasslands and scrub, but they have  narrower distribution than others such as aquatic vegetation. On the other hand, they have been described by traditional phytsociology after fine ecological variations. If there is inflation it should be somhow documented.

The inflation is reflected in the 78 associations which had been described in the original relevés as stored in SIVIM (see "Results - Overview of the classification"). It was not our goal to classify to the level of these associations, but our exploratory analysis indicated that they were not supported by numerical classification (i.e. TWINSPAN within the alliances did not result in splits along association lines, and many locally-described associations were only represented by a low number of relevés).

> There are several papers on this topic published in the area by Loidi and other authors, they are quoted in the adjointed sheet

Thank you for these references, we have incorporated them into our manuscript.

> classification at the alliance level ..

We have updated the text as suggested.

> In this paper only forest and mantle communities are dealt, no nitrophilous vegetation.

We cite this reference as the basis for defining the borders of our study biogeographical territory, we have revised the text to make this clear.

> In these disturbed communities, geophytes, with some exceptions, use to be  scarce. Do you think neccesary to highlight their proportion more than other life-forms? It seems to me that the main point is to describe the relative importance of hemicriptophytes in relation to therophytes as they become dominant in their respective alliances.

We agree that, in most of the alliances, only the variation of hemicriptophytes and therophytes is relevant. However, the presence of geophytes is an important part of the description of some of the alliances (e.g. see the description of *Allion triquetri* by Mucina et al. [-@RN4698]: *Mesic nitrophilous geophyte-rich fringe vegetation of the Western Mediterranean*). Furthermore, the variation in geophytes explained a significant part of the overall variance in a exploratory PCA. For these two reasons, we decided to highlight the propotion of geophytes in our manuscript.

> This figure is difficult to read. Point clouds overlap and the information is partially not visible. I suggest to use two figures for the same PCA, each one representing a half of the alliances selected in form that overlpings are avoided. This could also allow to use larger dots which would be more easily identified by the color.

Thank you for your suggestions to improve Figure 7. We have provided a new version of the figure, where each vegetation class is represented separately.

> It should be cited as: Castroviejo, S. (coord. gen.). 1986-2020. Flora iberica. Real Jardín Botánico, CSIC, Madrid [and other comments regarding the references]

Thank you for checking our references, we have corrected them as needed.

# Referee 2

> This paper presents a classification and characterisation of the anthropogenic vegetation alliances of the Iberian Atlantic ecoregion in Europe. The study area is not very large, but I think that the paper is a valuable contribution to the knowledge of this vegetation type in Europe, and it can be considered suitable for AVS after some revision.

Thank you for your positive evaluation and suggestions.

> The authors claim that (1) there is a need for a numerical classification and synthesis of the full spectrum of anthropogenic vegetation, and (2) an appropriate scale to conduct such study is the ecological region. While I strongly agree with point 1, I'm less convinced about point 2. Why should a revision at the ecoregion level be superior to a continental pan-European synthesis? I think, both is necessary, and the authors should acknowledge this fact.

We fully agree that a pan-European synthesis would be worthwhile goal, and never intended to suggest the contrary. In fact, we would be very happy of any research efforts on continental studies of anthropogenic vegetation. We have added a new final sentence to our conclusions: *We hope that our synthesis of the anthropogenic communities of the Iberian Atlantic territories serves as a contribution towards a pan-European synthesis of human-made vegetation, a synthesis that could shed light on the coexistence between humans and plants.* We also added a similar closing statement to the abstract. We have also modified our text to avoid giving the impression that we are suggesting that the ecoregion (or a similarly ranked biogeographical unit) is the *only* appropriate scale at which one can produce a synthesis of anthropogenic vegetation.

> The title is a bit too general. Please consider to add a subtitle such as "a case-study from SW Europe", or something similar.

We agree and have changed the title to *Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*.

> The merits of the applied two-step approach (first a Twinspan classification, then a "semi-supervised" noise clustering) remain unclear to me. More details are needed to fully understand what the authors have done. How many Twinspan clusters did you choose, and how did you determine the optimal number of clusters? How did you match the individual Twinspan clusters with EVC alliances? A crosstab showing the Twinspan clusters and the original alliance assignment should be presented. In the noise clustering, no additional clusters were detected, but plots were allowed to be re-assigned, and a considerable part of the data (17%) remained unclassified in a "noise group". Still, the "semi-supervised" classification is a numerical cluster algorithm. In my opinion, a pure unsupervised classification with subsequent interpretation of the result would have been more straightforward. Re-assignment of plots should be done on the basis of well-defined diagnostic species or using formal definitions, not by applying two numerical algorithms one after another. Or did I miss something important?**

The main aim of our classification was for it to be consistent with regional phytosociological expertise, and to match that expertise with current EuroVegChecklist alliances [@RN4698], not to create a new classification. To achieve this, one advantage of using the SIVIM database is that a majority of the plots had been assigned to associations or alliances by the original authors [@RN5374]. However, there could be misclassification due to unclear concepts or human error. Therefore, we used TWINSPAN to revise the classification by the original authors, to define the number of clusters that better aligned with our syntaxonomical checklist, and to remove outliers or misclassified plots. The obtained clusters therefore provided an agreement between phytosociological and numerical classification, and contained only those plots that better represented each alliance. These clusters were then used to assign other plots to the most similar cluster. We think semi-supervised classification is an optimal procedure to reach an agreement between existing and revised classifications (following Tichý et al. [-@RN5593], the aim is "preserving the good old units and searching for new ones"). We found that this important point was not clearly explained in the text or the abstract, so we revised the text accordingly. We have also provided more details on the TWINSPAN procedure as requested by the reviewer.

> The authors conclude that ten alliances previously reported from the study area do not occur there. This seems a bit hasty. Maybe, these alliances were only represented by a few plots, so they were not detected by the numerical methods.

We note that some alliances were supported by the methods, despite being extreme (ecologically or geographically) for the study region and/or being represented by a few plots. Therefore, we believe that the fact that the missing alliances were not supported by the numerical classification supports that the relevés that had been assigned to these alliances are poor representatives of the vegetation types, and more likely transitional communities at the extreme of the alliance distributions. Nonetheless, we have modified the text to be less conclusive about their absence, by adding the following sentence to the discussion: *Of course, the fact that these ten alliances are not supported by our numerical methods only indicates that they are poorly represented in our dataset, and should not be taken as absolute proof that they are absent from the Iberian Atlantic territories*.

> Most importantly, a synoptic table of the alliances is missing. I consider a table showing the constancy of all species in the 28 alliance, with the diagnostic species highlighted, an absolute prerequisite for the paper to be acceptable. A shorted table in the main text would be nice, but there should be a least the full table as supplementary information.

**Pendiente de hacer (Víctor?)**

> p5, l11: "non-overlapping definitions" ... different, but certainly overlapping definitions!

You are right, of course. We have change it to *not fully equivalent*.

> p9, l15: Why did you use 15% as cutlevel, and not 5%? The latter would correspond to the border between Braun-Blanquet class 1 and 2, while 15 cuts the Braun-Blanquet class 2 in the middle. Which cutlevel then corresponds to Braun-Blanquet class 2?**

Our transformation of Braun-Blanquet values follows the standard of assigning intermediate cover values. Thus, the BrBl value 2 (representing 5-25% cover) is transformed to 15% and the value 3 (25-50%) is transformed to 37.5%. Although we could use more approximate pseudospecies levels, the results of TWINSPAN will be the same when the transformation values are selected within the original cover ranges.

> p10, l5: Please clarify if the same Twinspan settings were used as before.

Yes, and we have added this information to the manuscript.

> p20, l10: "which are mostly differentiated by dominant vegetation height" ... Is this a new result of your study, or has this been mentioned in the literature (if so, please provide a reference)?

Yes, this is a traditional part of the alliance descriptions, and we have added this to our text with a reference.

> p21, l12: "neophytes tended to be in higher proportions in vegetation types associated to wet conditions (Bidentetea)" ... I would have expected an even higher portion of neophytes in the Senecionion fluviatilis. Maybe, plots of alluvial forb communities dominated by neophytes are underrepresented in your dataset?

Our *Senecionion fluviatilis* plots are certainly dominated by alien species, especifically by *Arundo donax*, which has a mean cover of 74%. However, *Arundo donax* is accompanied by many native species (mean = 12 native species per plot) and so the plot-level proportion of neophytes seems low (as it is just the propotion, not weighted by species cover). That being said, it might be possible that plots with neophytes are generally underrepresented across the dataset, because of a sampling bias towards "natural" communities. But this underrepresentation is not necessarily higher in *Senecionion fluviatilis* vs. other alliances.

> References: In several references the page numbers are missing!

We have revised the references.

> Appendix S3: What is the value for dominant species? Constancy? Mean cover value? Something else? Please clarify.

We have added information on how we calculated the values for diagnostic, constant and dominant species.

# References

::: {#refs}
:::

### 2024-10-03 doc/manuscript.Rmd
revised classification, keeping Aegopodion

@@
 ---
-title: "Classification and characterization of anthropogenic plant communities at the ecoregion level"
+title: "Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula"
+
 output: 
@@
 
-* *Questions* How can we classify the diversity of anthropogenic plant communities at the ecoregion level?
-How are these communities characterized by species origins (native plants, archaeophytes, and neophytes), species traits (lifeforms, plant height, and flowering phenology) and ecological preferences (temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity)?
+* *Questions* How can we reconcile regional and European classifications of anthropogenic plant communities at the biogeographical scale?
+How are these communities characterized by species origins, species traits and ecological preferences?
 
-* *Location* Iberian Atlantic ecoregion (a.k.a. Cantabrian Mixed Forests); Portugal, Spain, and France; south-western Europe.
+* *Location* Atlantic territories in the NW Iberian Peninsula (a.k.a. Cantabrian Mixed Forests ecoregion); NW Portugal, NW Spain, and SW France; south-western Europe.
 
-* *Methods* We compiled a vegetation database of 2,508 synanthropic plots and revised the regional checklist of anthropogenic vegetation alliances.
-We used modified TWINSPAN and semi-supervised classification to classify the plots into the revised alliances.
+* *Methods* We classified 2,508 synanthropic plots with the aim of being consistent with regional phytosociological expertise, while matching that expertise with current EuroVegChecklist alliances.
+We used modified TWINSPAN to revise the original phytosociological classification of the plots, followed by semi-supervised re-classification of the whole dataset.
 We determined the proportion of natives, archaeophytes, and neophytes.
-We also described the alliances in terms of species traits and ecological requirements.
+We also described the alliances in terms of species traits (lifeforms, plant height, and flowering phenology) and ecological requirements (temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity).
 
@@
 
-* *Conclusions* Ecoregion-level synthesis detected a discrepancy between the number of anthropogenic vegetation units described in the literature (38 alliances) and the number supported by numerical classification (28).
+* *Conclusions* Biogeographical-level synthesis detected a discrepancy between the number of anthropogenic vegetation units described in the literature (39 alliances) and the number supported by numerical classification (28).
 The diversity of anthropogenic vegetation can be organized into three groups: trampled, weeds, and ruderals.
-In the Iberian Atlantic ecoregion, anthropogenic habitats host one third of the ecoregion plant species pool and one fifth of the Iberian flora.
+In the Iberian Atlantic territories, anthropogenic habitats host one third of the regional plant species pool and one fifth of the Iberian flora.
 Mesic perennial ruderal vegetation is especially rich in native species and can be a biodiversity asset in urban ecosystems.
-Our ecoregion-level framework can improve the assessment and management of anthropogenic plant communities within biogeographically meaningful regions.
+Our biogeographical-level framework can improve the assessment and management of anthropogenic plant communities within biogeographically meaningful regions, and contribute towards a European-level synthesis of human-made vegetation.
 
@@
 
-Synanthropic vegetation, manmade habitats, human-made habitats, urban biodiversity, neophytes, archaeophytes, alien plants, habitat classification, semi-supervised classification, Iberian Atlantic ecoregion
+Synanthropic vegetation, manmade habitats, human-made habitats, urban biodiversity, neophytes, archaeophytes, alien plants, habitat classification, semi-supervised classification, Cantabrian Mixed Forests
 
@@
 Another issue is defining anthropogenic vegetation, especially in a continent such as Europe where human impact on the landscape has been widespread since ancient times.
-The two major vegetation/habitat classifications coexisting in Europe today, phytosociological classification [@RN4698] and the EUNIS habitat classification [@RN5061], have non-overlapping definitions of which plant communities fall within the ‘anthropogenic’ or ‘manmade’ categories.
+The two major vegetation/habitat classifications coexisting in Europe today, phytosociological classification [@RN4698] and the EUNIS habitat classification [@RN5061], have not fully equivalent definitions of which plant communities fall within the ‘anthropogenic’ or ‘manmade’ categories.
 These issues are further muddled by the fact that most classification studies of anthropogenic vegetation have focused either on a restricted political unit [@RN5424; @RN5426] or in a subset of the vegetation, such as segetal [@RN5494], ruderal [@RN5422] or trampled [@RN5434] vegetation.
-There is a need for a numerical classification and synthesis of the full spectrum of anthropogenic vegetation, and an appropriate scale to conduct such study is the ecological region (ecoregion, hereafter), since these biogeographical units are assumed to encompass areas with a similar biogeographic history and recurrent local ecosystems [@RN5491].
+There is a need for a numerical classification and synthesis of the full spectrum of anthropogenic vegetation, focusing on coherent biogeographical units, i.e.
+areas with a similar bioclimate, a shared biogeographic history and recurrent local ecosystems [@RN5491].
 
-The Iberian Atlantic ecoregion (a.k.a.
-Cantabrian Mixed Forests ecoregion; NW Iberian Peninsula) has a long history of human habitation dating back to the Cantabrian Upper Paleolithic [@RN5450].
-This, together with its transitional position at the border between the temperate and Mediterranean climatic zones of Europe [@RN4700], suggests a long regional history for archaeophyte-rich anthropogenic plant communities, supporting the current diverse anthropogenic vegetation with both temperate and Mediterranean floristic elements [@RN5008].
+The Atlantic territories in the NW Iberian Peninsula (a.k.a.
+Cantabrian Mixed Forests ecoregion) have a long history of human habitation dating back to the Cantabrian Upper Paleolithic [@RN5450].
+This, together with their transitional position at the border between the temperate and Mediterranean climatic zones of Europe [@RN4700], suggests a long regional history for archaeophyte-rich anthropogenic plant communities, supporting the current diverse anthropogenic vegetation with both temperate and Mediterranean floristic elements [@RN5008].
 Moreover, historical trade links with America and Asia, together with the warm and humid temperate climate, have made the region a hotspot for biological invasions [@RN5492; @RN5493].
@@
 brownfields) whose management and restoration requires an understanding of the anthropogenic communities able to colonize them [@RN5448; @RN5449].
-Although there is a long tradition of studying weed and ruderal communities in the ecoregion [@RN5364; @RN5361; @RN5359; @RN5360] there is a current need for an ecoregion-level synthesis that revises and updates the classification in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
-In this article, we have performed such a synthesis with the following two objectives: (1) to provide an updated classification of anthropogenic plant communities in the Iberian Atlantic ecoregion; and (2) to characterize the diversity of anthropogenic vegetation in terms of species pools (natives, archaeophytes, neophytes); species traits (life form, plant height, flowering phenology); and ecological requirements (temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity).
+Although there is a long tradition of studying weed and ruderal communities within the area [@RN5364; @RN5361; @RN5359; @RN5360; @RN5589; @RN5590; @RN5591] there is a current need for an biogeographical-level synthesis that revises and updates the classification in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
+In this article, we have performed such a synthesis with the following two objectives: (1) to provide an updated classification at the alliance level of anthropogenic plant communities in the Iberian Atlantic territories; and (2) to characterize the diversity of anthropogenic vegetation in terms of species pools (natives, archaeophytes, neophytes); species traits (life form, plant height, flowering phenology); and ecological requirements (temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity).
 
@@
 
-## Study ecoregion
+## Biogeographical territory under study
 
-We studied the anthropogenic plant communities of the Iberian Atlantic ecoregion [@RN5372], i.e.
-the territories with a temperate climate in the north-western part of the Iberian Peninsula (NW Portugal, N Spain, SW France).
-Our study ecoregion broadly corresponds to the Cantabrian Mixed Forests ecoregion *sensu* Olson *et al.* [-@RN4667], to the Iberian part of the European Atlantic province *sensu* Rivas-Martínez *et al.* [-@RN4699] and to the Iberian section of the Atlantic biogeographical region of the European Environmental Agency (<https://www.eea.europa.eu/data-and-maps/figures/biogeographical-regions-in-europe-2>).
+We studied the anthropogenic plant communities of the Iberian Atlantic territories, a biogeographical unit defined by Fernández Prieto *et al.* [-@RN5372; -@RN5588] as the territories with a temperate climate in the north-western part of the Iberian Peninsula (NW Portugal, N Spain, SW France).
+Our study area follows the detailed borders provided by Fernández Prieto *et al.* [-@RN5372; -@RN5588] and is broadly equivalent to the Cantabrian Mixed Forests ecoregion *sensu* Olson *et al.* [-@RN4667], to the Iberian part of the European Atlantic province *sensu* Rivas-Martínez *et al.* [-@RN4699; -@RN5592] (i.e.
+the Orocantabrian subprovince and parts of the Cantabro Atlantic subprovince) and to the Iberian section of the Atlantic biogeographical region of the European Environmental Agency (<https://www.eea.europa.eu/data-and-maps/figures/biogeographical-regions-in-europe-2>).
 
@@
 
-To assist in our classification, we prepared a checklist of anthropogenic syntaxa that could be present in our study ecoregion (**Appendix S1**), based on regional syntaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
-This syntaxonomical checklist included 38 anthropogenic vegetation alliances that could be present in the Iberian Atlantic ecoregion according to the literature.
+To assist in our classification, we prepared a checklist of anthropogenic syntaxa that could be present in our study area (**Appendix S1**), based on regional syntaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
+This syntaxonomical checklist included 39 anthropogenic vegetation alliances that could be present in the Iberian Atlantic territories according to the literature.
 
@@
 We performed an exploratory data analysis of these 3,249 plots using modified Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379] with the R package *twinspanR* [@RN5380], 3 pseudospecies cut levels (0, 15, 25), a minimum group size of 10 plots, and Sørensen's average dissimilarity.
-During this analysis we identified 741 outlier plots which did not belong to either the target anthropogenic vegetation or to the geographical area of the ecoregion.
+During this analysis we identified 741 outlier plots which did not belong to either the target anthropogenic vegetation or to the geographical area of the Iberian Atlantic territories.
 The majority of these plots corresponded to coastal plant communities and had been misclassified by the expert system.
@@
 
-To create a training dataset for the semi-supervised classification (see below), we started by keeping only those plots (n = 2,201) that had been assigned by the original authors of the plot to any of the 38 anthropogenic alliances defined in our syntaxonomical checklist (i.e., we removed the plots without syntaxa, that had been retrieved solely by the expert system).
-We further cleaned this subset using a preliminary TWINSPAN classification, in which we attempted to match the plots into the 38 alliances.
-This step allowed us to identify those plots in which there was an agreement between (1) our TWINSPAN-based classification and (2) the classification based on the original author's syntaxon.
+The main aim of our classification was for it to be consistent with regional phytosociological expertise, and to match that expertise with current EuroVegChecklist alliances [@RN4698].
+To achieve this, one advantage of using the SIVIM database is that a majority of the plots had been assigned to associations or alliances by the original authors [@RN5374].
+However, there could be misclassification due to unclear concepts or human error.
+Therefore, we used TWINSPAN to revise the classification by the original authors, to define the number of clusters that better aligned with our syntaxonomical checklist, and to remove outliers or misclassified plots.
+This allowed us to create a training dataset for a posterior semi-supervised classification (see below), which contained only those plots that better represented each alliance.
+
+We started this stage of the analyses by keeping only those plots (n = 2,218) that had been assigned by the original authors of the plot to any of the 39 anthropogenic alliances defined in our syntaxonomical checklist (i.e., we removed the plots without syntaxa, that had been retrieved solely by the expert system).
+We further cleaned this subset using TWINSPAN classification (with the same settings as above), in which we attempted to match the plots into the 39 alliances.
+In a first round of TWINSPAN, we divided the dataset into an increasing number of clusters, until reaching 15 clusters, beyond which the new created clusters were not phytosociologically meaningful.
+Then, we performed a second round of TWINSPAN within each of the 15 clusters.
+In this second round, we assigned each of the clusters to an alliance in our checklist, based on the composition of diagnostic species as described in the regional literature.
+Within each of the 15 primary clusters, we created an increasing number of secondary clusters until the newly created clusters splited one of the alliances that had already been assigned.
+
+This stage of the analyses allowed us to identify those plots in which there was an agreement between (1) our TWINSPAN-based classification (i.e.
+numerical classification) and (2) the classification based on the original author's syntaxon (i.e.
+regional phytosociological expertise).
 These plots (n = 1,725) became our training dataset for semi-supervised classification.
-Furthermore, this step allowed us to validate which alliances from our syntaxonomical checklist were present in the study ecoregion (n = 28) and which were absent (n = 10).
+Furthermore, this step allowed us to validate which alliances from our syntaxonomical checklist were present in the dataset (n = 28) and which were absent (n = 10).
 
@@
 
-Next, we conducted a semi-supervised classification of the whole dataset (n = 2,508 plots) into the 28 validated anthropogenic alliances.
+Next, we conducted a semi-supervised classification [@RN5593] of the whole dataset (n = 2,508 plots) into the 28 validated anthropogenic alliances.
 Semi-supervised classification uses a training subset of *a priori* classified vegetation plots to classify a secondary subset of unclassified plots [@RN5383].
@@
 
-We classified the species as native, archaeophytes, or neophytes in the Iberian Atlantic ecoregion using the information in *Flora iberica* [-@RN3243] and catalogues of archaeophytes for Britain [@RN5490] and the Czech Republic [@RN5495].
+We classified the species as native, archaeophytes, or neophytes in the Iberian Atlantic territories using the information in *Flora iberica* [-@RN3243] and catalogues of archaeophytes for Britain [@RN5490] and the Czech Republic [@RN5495].
 We must stress that identifying archaeophytes in southern Europe is highly problematic [@RN5444] and our classification must be taken as an indication of putative archaeophyte character for the purposes of vegetation description and comparison with other European regions, rather than a definitive classification of the species; as stated by Preston *et al.* [@RN5490], the classification of a species as an archaeophyte should be interpreted as a hypothesis to be tested by further studies.
@@
 1**, **Appendix S4**) representing 9 vegetation classes.
-Most of the alliances had a wide distribution in the ecoregion according to our dataset, but five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis* and *Paspalo-Agrostion semiverticillati*) had isolated occurrences towards the southern limit of the Iberian Atlantic territories.
+Most of the alliances had a wide distribution in the Iberian Atlantic territories according to our dataset, but five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis* and *Paspalo-Agrostion semiverticillati*) had isolated occurrences towards the southern limit of the Iberian Atlantic territories.
 The regional literature included in our dataset recognized 72 associations (**Appendix S5**) within these alliances, but exploratory analysis (not shown) indicated that the separation of these associations within the alliances had a generally weak support based on numerical classification methods.
@@
 
-Our synthesis of the anthropogenic vegetation at the ecoregion level has detected a discrepancy between the number of vegetation units described in the literature (i.e.
-38 alliances) and the number supported by numerical classification (28 alliances).
+Our synthesis of the anthropogenic vegetation has detected a discrepancy between the number of vegetation units described in the literature (i.e.
+39 alliances) and the number supported by numerical classification (28 alliances).
 Our classification supported that anthropogenic vegetation can be broadly divided into three groups belonging to trampled, weed, and ruderal communities.
@@
 
-Our analysis supports the existence of 28 anthropogenic alliances in the vegetation of the Iberian Atlantic ecoregion.
-While the core of this vegetation diversity is made up of temperate European alliances, these are enriched by the occurrence of eight Mediterranean alliances (*Galio valantiae-Parietarion judaicae*, *Polycarpion tetraphylli*, *Allion triquetri*, *Chenopodion muralis*, *Echio-Galactition tomentosae*, *Paspalo-Agrostion semiverticillati*, *Balloto-Conion maculati* and *Cynancho-Convolvulion sepium*); plus three alliances that are endemic of the Iberian coasts (*Linario polygalifoliae-Vulpion alopecuri*) and mountains (*Carduo carpetani-Cirsion odontolepidis*, *Cirsion richterano-chodati*) [@RN4698].
-Seven of these alliances had been reported as absent or uncertain in the ecoregion (*Caucalidion lappulae*, *Chenopodion muralis*), or only present in the Portuguese sector of the ecoregion (*Spergulo arvensis-Erodion cicutariae*, *Allion triquetri*, *Echio-Galactition tomentosae*, *Carduo carpetani-Cirsion odontolepidis*, *Cynancho-Convolvulion sepium*) by the recently published distribution maps of vegetation alliances in Europe [@RN5371].
-Our dataset supports that six of these alliances are relatively well distributed in the ecoregion, while *Caucalidion lappulae* only has isolated occurrences.
-On the other hand, our dataset does not support the occurrence in the ecoregion of nine Mediterranean alliances (*Roemerion hybridae*, *Diplotaxion erucoidis*, *Euphorbion prostratae*, *Alysso granatensis-Brassicion barrelieri*, *Hordeion murini*, *Taeniathero-Aegilopion geniculatae*, *Parietarion lusitanico-mauritanicae*, *Onopordion castellani*, *Silybo mariani-Urticion piluliferae*) and one temperate alliance (*Fragarion vescae*) that had been cited as potentially occurring here [@RN5363; @RN5008; @RN5365; @RN4698; @RN5371].
+Our analysis supports the existence of 28 anthropogenic alliances in the vegetation of the Iberian Atlantic territories.
+While the core of this vegetation diversity is made up of temperate European alliances, these are enriched by the occurrence of eight Mediterranean alliances (*Galio valantiae-Parietarion judaicae*, *Polycarpion tetraphylli*, *Allion triquetri*, *Chenopodion muralis*, *Echio-Galactition tomentosae*, *Paspalo-Agrostion semiverticillati*, *Balloto-Conion maculati* and *Cynancho-Convolvulion sepium*); plus two alliances that are endemic of the Iberian coasts (*Linario polygalifoliae-Vulpion alopecuri*) and mountains (*Cirsion richterano-chodati*) [@RN4698].
+Seven of these alliances had been reported as absent or uncertain in the Iberian Atlantic territories (*Caucalidion lappulae*, *Chenopodion muralis*, *Onopordion acanthii*), or only present in the Portuguese sector of the area (*Spergulo arvensis-Erodion cicutariae*, *Allion triquetri*, *Echio-Galactition tomentosae*, *Cynancho-Convolvulion sepium*) by the recently published distribution maps of vegetation alliances in Europe [@RN5371].
+Our dataset supports that six of these alliances are relatively well distributed in the study area, while *Caucalidion lappulae* only has isolated occurrences.
+
+On the other hand, our dataset does not support the occurrence in the Iberian Atlantic territories of ten Mediterranean alliances (*Roemerion hybridae*, *Diplotaxion erucoidis*, *Euphorbion prostratae*, *Alysso granatensis-Brassicion barrelieri*, *Hordeion murini*, *Taeniathero-Aegilopion geniculatae*, *Parietarion lusitanico-mauritanicae*, *Onopordion castellani*, *Carduo carpetani-Cirsion odontolepidis*, *Silybo mariani-Urticion piluliferae*) and one temperate alliance (*Fragarion vescae*) that had been cited as potentially occurring here [@RN5363; @RN5008; @RN5365; @RN4698; @RN5371].
+Of course, the fact that these eleven alliances are not supported by our numerical methods only indicates that they are poorly represented in our dataset, and should not be taken as absolute proof that they are absent from the Iberian Atlantic territories.
+
 For habitat classification purposes, we propose that this vegetation diversity can be summarized into 15 EUNIS habitat types: one constructed habitat (*J2.5*), ten vegetated human-made habitats (*V11*, *V12*, *V13*, *V15*, *V32*, *V35*, *V34*, *V37*, *V38*, *V39*), three habitats of woodland fringes and clearings and tall forb stands (*R54*, *R55*, *R57*) and one habitat of periodically-exposed shores (*Q61*).
-In general, we have followed the correspondence between alliances and EUNIS habitats proposed by the EUNIS expert system [@RN5061], except for *Echio-Galactition tomentosae* which, in the context of the ecoregion, seems to be more adequately placed in the habitat *V37 Annual anthropogenic herbaceous vegetation* rather than in the habitat *V33 Dry Mediterranean land with unpalatable non-vernal herbaceous vegetation*.
+In general, we have followed the correspondence between alliances and EUNIS habitats proposed by the EUNIS expert system [@RN5061], except for *Echio-Galactition tomentosae* which, in the context of the study area, seems to be more adequately placed in the habitat *V37 Annual anthropogenic herbaceous vegetation* rather than in the habitat *V33 Dry Mediterranean land with unpalatable non-vernal herbaceous vegetation*.
 
-Since our goal was to match the classification of the ecoregion classification to the EuroVegChecklist [@RN4698], we refrained from merging recognized alliances or changing their assignation to superior syntaxa.
+Since our goal was to match the classification to the EuroVegChecklist [@RN4698], we refrained from merging recognized alliances or changing their assignation to superior syntaxa.
 However, there are some cases that could deserve further scrutiny.
@@
 It is therefore not surprising that in biogeographic transitional regions - such as the Iberian Atlantic territories - these two alliances coexist and show a high overlap.
-Also, there was overlap between the alliances *Geo urbani-Alliarion officinalis* and *Aegopodion podagrariae*, two alliances that can occupy ruderal to semi-natural nitrophilous fringes, and which are mostly differentiated by dominant vegetation height, as was supported by our dataset.
+Also, there was overlap between the alliances *Geo urbani-Alliarion officinalis* and *Aegopodion podagrariae*, two alliances that can occupy ruderal to semi-natural nitrophilous fringes, and which are traditionally differentiated in their descriptions [@RN4698] by vegetation height, as was supported by our dataset.
 
@@
 
-We found 22% non-native species (including neophytes and archaeophytes) in the anthropogenic vegetation of the Iberian Atlantic ecoregion.
+We found 22% non-native species (including neophytes and archaeophytes) in the anthropogenic vegetation of the Iberian Atlantic territories.
 This percentage is intermediate between that found in urban areas of temperate Europe and North America (30-50%) [@RN5498; @RN5499; @RN5500; @RN5079] and Italy (12-26%) [@RN5501].
@@
 A similar study in the Czech Republic found 32% archaeophytes and 7% neophytes [@RN5435].
-Taken together, these results suggest that the Iberian Atlantic anthropogenic vegetation shows an incidence of non-native species that is intermediate between southern and central-northern Europe, which is in agreement with the biogeographical position of our ecoregion.
+Taken together, these results suggest that the Iberian Atlantic anthropogenic vegetation shows an incidence of non-native species that is intermediate between southern and central-northern Europe, which is in agreement with the biogeographical position of our study area.
 Previous studies also suggested that the non-natives were more frequent in early successional anthropogenic communities [@RN5423], and our results confirmed this as the proportion of native species tended to be higher in the perennial ruderal classes *Artemisietea vulgaris* and *Epilobietea angustifolii*.
@@
 
-Ecoregion-level synthesis detected a discrepancy between the number of anthropogenic vegetation units described in the literature and the number supported by numerical classification.
+Biogeographical-level synthesis detected a discrepancy between the number of anthropogenic vegetation units described in the literature and the number supported by numerical classification.
 The diversity of anthropogenic vegetation can be organized into trampled, weed and ruderal communities varying primarily along an axis of abiotic stress.
-Importantly, the anthropogenic vegetation of the Iberian Atlantic ecoregion is home to 1,162 plant taxa.
-This is approximately one third of the whole ecoregion species pool represented in the SIVIM database [@RN5374], and one fifth of the Iberian flora [@RN5503].
+Importantly, the anthropogenic vegetation of the Iberian Atlantic territories is home to 1,162 plant taxa.
+This is approximately one third of the whole regional species pool represented in the SIVIM database [@RN5374], and one fifth of the Iberian flora [@RN5503].
 More than half of this synanthropic diversity (615 taxa) occurs in mesic to wet perennial ruderal vegetation, highlighting the potential of some anthropogenic habitats as a source of biodiversity in human-dominated landscapes [@RN5446; @RN5445].
 These results suggest three main objectives for the sustainable management of human-made habitats: (1) focusing efforts on alien plant management in anthropogenic plant communities of wet habitats; (2) conserving examples of archaeophyte-rich annual weed communities, especially those linked to traditional agricultural practices [@RN5502]; and (3) conserving and promoting mesic ruderal and fringe vegetation as a biodiversity refuge in urban and peri-urban areas.
+We hope that our synthesis of the anthropogenic communities of the Iberian Atlantic territories serves as a contribution towards a pan-European synthesis of human-made vegetation, a synthesis that could shed light on the coexistence between humans and plants.
 
@@
 
-**Appendix S1** Checklist of anthropogenic syntaxa that could be present in the Iberian Atlantic ecoregion according to the literature.
+**Appendix S1** Checklist of anthropogenic syntaxa that could be present in the Iberian Atlantic territories according to the literature.
 Syntaxa names are provided as recorded by the original authors of the vegetation plot.
-The file indicates if the presence of the syntaxon in the Iberian Atlantic ecoregion was validated or not by our analysis.
+The file indicates if the presence of the syntaxon in the Iberian Atlantic territories was validated or not by our analysis.
 
@@
 **Appendix S3** Characteristic species combinations of the anthropogenic vegetation alliances in spreadsheet format, divided into diagnostic, constant, and dominant species.
+The values are as follows: Diagnostic species = *IndVal* multiplied by 100; Constant species = percentage occurrence frequency (constancy); Dominant species = percentage frequency of plots in which the species occurs with a cover larger than 25%.
 
@@
 
-```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Anthropogenic vegetation alliances of the Iberian Atlantic ecoregion. Each dot is a vegetation plot. Dot colors indicate the vegetation class."}
+```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Anthropogenic vegetation alliances of the Iberian Atlantic territories. Each dot is a vegetation plot. Dot colors indicate the vegetation class."}
 knitr::include_graphics("../results/figures/F1 - maps.png")
@@
 
-```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Floristic diversity of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Floristic diversity of the anthropogenic plant communities of the Iberian Atlantic territories.
 Biplots produced by Principal Component Analysis (PCA) with Hellinger transformation.
@@
 
-```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of neophytes, archaeophytes, and native species in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of neophytes, archaeophytes, and native species in the anthropogenic plant communities of the Iberian Atlantic territories.
 Each bar represents the plot-level average proportion of neophytes (dark shade), archaeophytes (medium shade), and natives (light shade) in each vegetation alliance.
@@
 
-```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of therophytes, geophytes, and other life forms (hemicryptophytes, chamaephytes, phanerophytes, bryophytes, hydrophytes) in the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of therophytes, geophytes, and other life forms (hemicryptophytes, chamaephytes, phanerophytes, bryophytes, hydrophytes) in the anthropogenic plant communities of the Iberian Atlantic territories.
 Each bar represents the plot-level average proportion of therophytes (dark shade), geophytes (medium shade), and other life forms (light shade) in each vegetation alliance.
@@
 
-```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Vegetation height and flowering phenology of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Vegetation height and flowering phenology of the anthropogenic plant communities of the Iberian Atlantic territories.
 Boxplots show the community-weighted mean for plant height, median month of flowering, and length of the flowering season for each vegetation alliance.
@@
 
-```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic territories.
 Boxplots show the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity, calculated for each vegetation alliance.
@@
 
-```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic ecoregion.
+```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic territories.
 Biplot produced by Principal Component Analysis (PCA) of the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity.
-Labels and arrows indicate the contribution of each indicator to the first and second principal components.
-Colors indicate the vegetation class."}
+Each dot is a plot.
+In the left panel, labels and arrows indicate the contribution of each indicator to the first and second principal components.
+The right panels show the space occupied by each vegetation class (colored dots) with respect to the whole anthropogenic vegetation (grey dots)."}
 knitr::include_graphics("../results/figures/F7 - pca.png")

### 2024-10-03 src/10-origins.R

@@ -36,7 +36,7 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                                 "Dauco-Melilotion", 
                                 "Cirsion richterano-chodati",
                                 "Silybo mariani-Urticion piluliferae",
-                                "Carduo carpetani-Cirsion odontolepidis",
+                                "Onopordion acanthii",
                                 "Convolvulo arvensis-Agropyrion repentis",
                                 "Paspalo-Agrostion semiverticillati",
                                 "Bidention tripartitae",

### 2024-10-03 src/11-lifeforms.R

@@ -42,7 +42,7 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                                 "Dauco-Melilotion", 
                                 "Cirsion richterano-chodati",
                                 "Silybo mariani-Urticion piluliferae",
-                                "Carduo carpetani-Cirsion odontolepidis",
+                                "Onopordion acanthii",
                                 "Convolvulo arvensis-Agropyrion repentis",
                                 "Paspalo-Agrostion semiverticillati",
                                 "Bidention tripartitae",

### 2024-10-03 src/12-height-flowering.R

@@ -124,7 +124,7 @@ cwms %>%
                                "Bidention tripartitae",
                                "Paspalo-Agrostion semiverticillati",
                                "Convolvulo arvensis-Agropyrion repentis",
-                               "Carduo carpetani-Cirsion odontolepidis",
+                               "Onopordion acanthii",
                                "Silybo mariani-Urticion piluliferae", 
                                "Cirsion richterano-chodati",
                                "Dauco-Melilotion",

### 2024-10-03 src/13-indicators.R

@@ -115,7 +115,7 @@ plotmeans %>%
                                 "Bidention tripartitae",
                                 "Paspalo-Agrostion semiverticillati",
                                 "Convolvulo arvensis-Agropyrion repentis",
-                                "Carduo carpetani-Cirsion odontolepidis",
+                                "Onopordion acanthii",
                                 "Silybo mariani-Urticion piluliferae", 
                                 "Cirsion richterano-chodati",
                                 "Dauco-Melilotion",
@@ -233,6 +233,204 @@ plotmeans %>%
 inds %>% group_by(Class) %>% summarise(PC1 = mean (Dim.1), PC2 = mean(Dim.2)) %>% arrange(PC1)
 inds %>% group_by(Class) %>% summarise(PC1 = mean (Dim.1), PC2 = mean(Dim.2)) %>% arrange(PC2)
 
+### Plot each class
+
+inds %>% 
+  filter(Class == "Parietarietea") %>%
+  ggplot(aes(Dim.1, Dim.2)) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "grey60") +
+  geom_vline(xintercept = 0, linetype = "dashed", color = "grey60") +
+  geom_point(data = inds, color = "grey",
+             show.legend = TRUE, size = 1) +
+  geom_point(color = "grey60",
+             show.legend = TRUE, size = 2.2) +
+  ggthemes::theme_tufte() +
+  annotate("text", x = 0, y = -4, size = 3, family = "sans", fontface = "italic", label = "Parietarietea") +
+  theme(text = element_text(family = "sans"),
+        strip.background = element_blank(),
+        panel.background = element_rect(color = "black", fill = NULL),
+        legend.title = element_text(size = 10, face = "italic"),
+        axis.text = element_blank(),
+        axis.title = element_blank(),
+        axis.ticks = element_blank(),
+        plot.margin = unit(c(0,0,0,0), "cm")) -> 
+  Fig3B; Fig3B
+
+inds %>% 
+  filter(Class == "Polygono-Poetea") %>%
+  ggplot(aes(Dim.1, Dim.2)) +
+  geom_hline(yintercept = 0, linetype = "dashed", color = "grey60") +
+  geom_vline(xintercept = 0, linetype = "dashed", color = "grey60") +
+  geom_point(data = inds, color = "grey",
+             show.legend = TRUE, size = 1) +
+  geom_point(color = "cadetblue4",
[…]

### 2024-10-03 src/2-twinspan NEW.R

@@ -1,0 +1,1153 @@
+library(tidyverse); library(vegan); library(labdsv)
+
+rm(list = ls())
+
+read.csv("results/sintaxonomy/original-sintaxonomy.csv", fileEncoding = "latin1") %>%
+  filter(Anthropogenic == "Yes") %>% 
+  select(Original, Alliance, Order, Class) %>%
+  unique -> syntaxonomy
+
+syntaxonomy %>% group_by(Alliance)
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>%
+  select(-c(Alliance, Order, Class)) %>%
+  merge(syntaxonomy)-> header
+
+header %>% group_by(Alliance)
+header %>% group_by(SIVIMID)
+
+header %>% pull(Class) %>% unique
+
+#### Cluster 1
+
+header %>% filter(Class == "Cymbalario-Parietarietea diffusae") -> header3
+header3 %>% group_by(Original) %>% tally
+header3 %>% group_by(Alliance) %>% tally
+
+read.csv("data/species.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% header3$SIVIMID) -> species
+
+species %>%
+  select(SIVIMID, Analysis.Names, Cover.percent) %>%
+  spread(Analysis.Names, Cover.percent, fill = 0) %>%
+  column_to_rownames(var = "SIVIMID") -> df1
+
+### Twinspan 
+
+twinspanR::twinspan(
+  df1,
+  modif = TRUE,
+  cut.levels = c(0, 15, 25),
+  min.group.size = 10,
+  clusters = 2,
+  diss = "multi.sorensen",
+  mean.median = "mean",
+  show.output.on.console = FALSE,
+  quiet = TRUE) %>%
+  cut() %>%
+  cbind(df1) %>%
+  select(".") %>%
+  rename("Cluster" = ".") %>%
+  rownames_to_column(var = "SIVIMID") -> kclusters
+
+header %>% merge(kclusters, by = "SIVIMID") -> header4
+
+### DCA
+
+decorana(df1) -> dca1
+
+vegan::scores(dca1) %>%
+  data.frame() %>%
+  rownames_to_column("SIVIMID") %>%
+  merge(header4) %>%
+  mutate(Cluster = as.factor(Cluster)) -> df2
+
+cent <- aggregate(cbind(DCA1, DCA2) ~ Cluster, data = df2, FUN = mean)
[…]

### 2024-10-03 src/3-vegclust.R

@@ -16,7 +16,7 @@ read.csv("data/urban-header-5.1.csv", fileEncoding = "latin1") %>%
 
 ### Prepare community composition matrices
 
-read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") %>% 
+read.csv("data/species.csv", fileEncoding = "latin1") %>% 
   filter(SIVIMID %in% headerClassified$SIVIMID) %>%
   select(SIVIMID, Analysis.Names, Cover.percent) %>%
   spread(Analysis.Names, Cover.percent, fill = 0) %>%
@@ -24,7 +24,7 @@ read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") %>%
   decostand("normalize") -> # Chord transformation of community data
   dfClassified
 
-read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") %>% 
+read.csv("data/species.csv", fileEncoding = "latin1") %>% 
   filter(SIVIMID %in% headerUnclassified$SIVIMID) %>%
   select(SIVIMID, Analysis.Names, Cover.percent) %>%
   spread(Analysis.Names, Cover.percent, fill = 0) %>%
@@ -110,7 +110,7 @@ header2 %>%
 
 ### DCA of new groups
 
-read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") -> species
+read.csv("data/species.csv", fileEncoding = "latin1") -> species
 
 species %>%
   merge(header2) %>%
@@ -190,6 +190,126 @@ data.frame(Community = indicators$maxcls, Indicator = indicators$indcls,
   slice_max(Indicator, n = 4) %>%
   data.frame
 
+### M
+
+header2 %>% filter(Cluster == "M1") -> headerM1
+
+headerM1 %>%
+  group_by(Alliance) %>% tally
+
+read.csv("data/species.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% headerM1$SIVIMID) -> speciesM1
+
+speciesM1 %>%
+  select(SIVIMID, Analysis.Names, Cover.percent) %>%
+  spread(Analysis.Names, Cover.percent, fill = 0) %>%
+  column_to_rownames(var = "SIVIMID") -> df1
+
+### Twinspan 
+
+twinspanR::twinspan(
+  df1,
+  modif = TRUE,
+  cut.levels = c(0, 15, 25),
+  min.group.size = 10,
+  clusters = 3,
[…]

### 2024-10-03 src/8-maps.R

@@ -40,7 +40,7 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
                                "Bidention tripartitae",
                                "Paspalo-Agrostion semiverticillati",
                                "Convolvulo arvensis-Agropyrion repentis",
-                               "Carduo carpetani-Cirsion odontolepidis",
+                               "Onopordion acanthii",
                                "Silybo mariani-Urticion piluliferae", 
                                "Cirsion richterano-chodati",
                                "Dauco-Melilotion",
@@ -67,7 +67,7 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
                                "Echio-\nGalactition tomentosae" = "Echio-Galactition tomentosae",
                                "Linario polygalifoliae-\nVulpion alopecuri" = "Linario polygalifoliae-Vulpion alopecuri",
                                 # "Sisymbrion officinalis",
-                               "Carduo carpetani-\nCirsion odontolepidis" = "Carduo carpetani-Cirsion odontolepidis",
+                               "Carduo carpetani-\nCirsion odontolepidis" = "Onopordion acanthii",
                                "Cirsion\nrichterano-chodati" = "Cirsion richterano-chodati",
                                 # "Dauco-Melilotion",
                                "Geo urbani-\nAlliarion officinalis" = "Geo urbani-Alliarion officinalis",

### 2024-10-03 src/2-twinspan NEW.R
recovered header species

@@ -857,6 +857,7 @@ header4 %>%
 
 header %>% filter(Class == "Epilobietea angustifolii" ) -> header3
 # header %>% filter(Order %in% c( "Convolvuletalia sepium") ) -> header3
+# header %>% filter(Alliance %in% c( "Aegopodion podagrariae", "Balloto-Conion maculati", "Geo urbani-Alliarion officinalis") ) -> header3
 header3 %>% group_by(Original) %>% tally %>% print(n =222)
 header3 %>% group_by(Alliance) %>% tally
 
@@ -874,7 +875,7 @@ twinspanR::twinspan(
   modif = TRUE,
   cut.levels = c(0, 15, 25),
   min.group.size = 10,
-  clusters = 11,
+  clusters = 2,
   diss = "multi.sorensen",
   mean.median = "mean",
   show.output.on.console = FALSE,

### 2024-10-04 doc/manuscript.Rmd
Revised classification, no Aegopodion

@@
 To assist in our classification, we prepared a checklist of anthropogenic syntaxa that could be present in our study area (**Appendix S1**), based on regional syntaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
-This syntaxonomical checklist included 39 anthropogenic vegetation alliances that could be present in the Iberian Atlantic territories according to the literature.
+This syntaxonomical checklist included 36 anthropogenic vegetation alliances that could be present in the Iberian Atlantic territories according to the literature.
 
@@
 
-We started this stage of the analyses by keeping only those plots (n = 2,218) that had been assigned by the original authors of the plot to any of the 39 anthropogenic alliances defined in our syntaxonomical checklist (i.e., we removed the plots without syntaxa, that had been retrieved solely by the expert system).
+We started this stage of the analyses by keeping only those plots (n = 2,100) that had been assigned by the original authors of the plot to any of the 39 anthropogenic alliances defined in our syntaxonomical checklist (i.e., we removed the plots without syntaxa, that had been retrieved solely by the expert system).
 We further cleaned this subset using TWINSPAN classification (with the same settings as above), in which we attempted to match the plots into the 39 alliances.
@@
 regional phytosociological expertise).
-These plots (n = 1,725) became our training dataset for semi-supervised classification.
-Furthermore, this step allowed us to validate which alliances from our syntaxonomical checklist were present in the dataset (n = 28) and which were absent (n = 10).
+These plots (n = 1,767) became our training dataset for semi-supervised classification.
+Furthermore, this step allowed us to validate which alliances from our syntaxonomical checklist were present in the dataset (n = 27) and which were absent (n = 9).
+
+## Not supported by TWINSPAN
+
+*Caucalidion lappulae* (4)
+*Roemerion hybridae* (2)
+*Hordeion murini* (2)
+*Taeniathero-Aegilopion geniculatae* (1)
+*Onopordion castellani* (6)
+*Euphorbion prostratae* (4)
+
+*Onopordion acanthii* (5)
+*Fragarion vescae* (1)
+*Aegopodion podagrariae* (39)
 
@@
 We performed this analysis with the R package *vegclust* [@RN5383].
-The semi-supervised classification with noise clustering resulted in the final classification of 2,081 plots into 28 alliances and 427 plots left out in the noise group.
+The semi-supervised classification with noise clustering resulted in the final classification of 2,050 plots into 27 alliances and 458 plots left out in the noise group.
 We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances, and Holm's p-value correction) fitted with the R package *RVAideMemoire* [@RN5381] to test the significance of the final vegetation alliances; along with Principal Component Analysis (PCA) as implemented in the R package *FactoMineR* [@RN3166] to visualize the relationships between the classes and alliances.

### 2024-10-04 src/2-twinspan NEW.R

@@ -3,19 +3,17 @@ library(tidyverse); library(vegan); library(labdsv)
 rm(list = ls())
 
 read.csv("results/sintaxonomy/original-sintaxonomy.csv", fileEncoding = "latin1") %>%
-  filter(Anthropogenic == "Yes") %>% 
   select(Original, Alliance, Order, Class) %>%
   unique -> syntaxonomy
 
 syntaxonomy %>% group_by(Alliance)
 
 read.csv("data/header.csv", fileEncoding = "latin1") %>%
-  select(-c(Alliance, Order, Class)) %>%
+  select(-c(Alliance, Order, Class, CANTEUNIS)) %>%
   merge(syntaxonomy)-> header
 
 header %>% group_by(Alliance)
 header %>% group_by(SIVIMID)
-
 header %>% pull(Class) %>% unique
 
 #### Cluster 1
@@ -241,7 +239,6 @@ header4 %>%
   select(-Cluster) %>%
   write.csv("data/urban-header-5.1B.csv", fileEncoding = "latin1")
 
-
 #### Cluster 3
 
 header %>% filter(Class == "Papaveretea rhoeadis") -> header3
@@ -307,7 +304,6 @@ header4 %>%
   arrange(Cluster, -n) %>%
   data.frame
 
-
 ### Dominant
 
 species %>%
@@ -360,8 +356,6 @@ header4 %>%
                          "Scleranthion annui" = "3")) %>%
   select(-Cluster) %>%
   write.csv("data/urban-header-5.1C.csv", fileEncoding = "latin1")
-
-
 
 #### Cluster 4
 
@@ -428,7 +422,6 @@ header4 %>%
   arrange(Cluster, -n) %>%
   data.frame
 
-
 ### Dominant
 
 species %>%
@@ -483,9 +476,6 @@ header4 %>%
                          "Linario polygalifoliae-Vulpion alopecuri" = "5")) %>%
   select(-Cluster) %>%
   write.csv("data/urban-header-5.1D.csv", fileEncoding = "latin1")
-
-
-
 
 #### Cluster 5
 
@@ -605,9 +595,6 @@ header4 %>%
   select(-Cluster) %>%
   write.csv("data/urban-header-5.1E.csv", fileEncoding = "latin1")
 
-
-
-
 #### Cluster 6
 
 header %>% filter(Class == "Bidentetea") -> header3
@@ -673,7 +660,6 @@ header4 %>%
   arrange(Cluster, -n) %>%
   data.frame
 
-
[…]

### 2024-10-04 src/3-vegclust.R

@@ -108,6 +108,11 @@ header2 %>%
   arrange(Cluster, -n) %>%
   print(n = 500)
 
+header2 %>%
+  group_by(Cluster) %>%
+  tally %>% 
+  print(n = 100)
+
 ### DCA of new groups
 
 read.csv("data/species.csv", fileEncoding = "latin1") -> species
@@ -250,65 +255,56 @@ header4 %>%
   arrange(Cluster, -n) %>%
   data.frame
 
-### N
-
-header2 %>% filter(Cluster == "N") -> headerM1
-
-headerM1 %>%
-  group_by(Alliance) %>% tally
-
-read.csv("data/species.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% headerM1$SIVIMID) -> speciesM1
-
-speciesM1 %>%
-  select(SIVIMID, Analysis.Names, Cover.percent) %>%
-  spread(Analysis.Names, Cover.percent, fill = 0) %>%
-  column_to_rownames(var = "SIVIMID") -> df1
-
-### Twinspan 
-
-twinspanR::twinspan(
-  df1,
-  modif = TRUE,
-  cut.levels = c(0, 15, 25),
-  min.group.size = 10,
-  clusters = 8,
-  diss = "multi.sorensen",
-  mean.median = "mean",
-  show.output.on.console = FALSE,
-  quiet = TRUE) %>%
-  cut() %>%
-  cbind(df1) %>%
-  select(".") %>%
-  rename("Cluster" = ".") %>%
-  rownames_to_column(var = "SIVIMID") -> kclusters
-
-headerM1 %>% select(-Cluster) %>% merge(kclusters, by = "SIVIMID") -> header4
-
-### DCA
-
-decorana(df1) -> dca1
-
-vegan::scores(dca1) %>%
-  data.frame() %>%
-  rownames_to_column("SIVIMID") %>%
-  merge(header4) %>%
-  mutate(Cluster = as.factor(Cluster)) -> df2
-
-cent <- aggregate(cbind(DCA1, DCA2) ~ Cluster, data = df2, FUN = mean)
-segs <- merge(df2, setNames(cent, c("Cluster", "oDCA1", "oDCA2")), by = "Cluster", sort = FALSE)
-
-df2 %>%
-  ggplot(aes(x = DCA1, y = DCA2)) + 
[…]

### 2024-10-04 results/supporting/AppendixS4.Rmd
revising alliance descriptions

@@
 6.
-*Caucalidion lappulae* Annual weed vegetation of cereal crops on base-rich soils.
-Very high proportion of putative archaeophytes.
-Distribution mostly limited to the more submediterranean valleys south of the Cantabrian Mountains.
-Diagnostic species in the ecoregion: *Cyanus segetum*, *Papaver rhoeas*, *Trifolium arvense*, *Anacyclus clavatus*, *Valerianella eriocarpa*.
-Correspondence with EUNIS habitats *V11 Intensive unmixed crops* and *V13 Arable land with unmixed crops grown by low-intensity agricultural methods*.
-
-7.
 *Scleranthion annui* Annual weed vegetation of cereal crops on base-poor soils.
@@
 
-8.
+7.
 *Oxalidion europeae* Annual weed vegetation of gardens and root crops.
@@
 
-9.
+8.
 *Spergulo arvensis-Erodion cicutariae* Summer-annual C4 weed vegetation.
@@
 
-10.
+9.
 *Allion triquetri* Geophyte-rich fringe vegetation.
@@
 
-11.
+10.
 *Geranio pusilli-Anthriscion caucalidis* Winter-annual fringe vegetation.
@@
 
-12.
+11.
 *Chenopodion muralis* Low-growth winter-annual ruderal vegetation.
@@
 
-13.
+12.
 *Echio-Galactition tomentosae* Tall-herb winter-annual ruderal vegetation.
@@
 
-14.
+13.
 *Linario polygalifoliae-Vulpion alopecuri* Ephemeral annual vegetation of disturbed coastal dunes.
@@
 
-15.
+14.
 *Sisymbrion officinalis* Summer-annual ruderal vegetation.
@@
 
-16.
+15.
 *Bidention tripartitae* Summer-annual pioneer vegetation of temporarily flooded sites.
@@
 
-17.
+16.
 *Paspalo-Agrostion semiverticillati* Summer-annual pioneer vegetation of temporarily flooded and warm sites.
@@
 
-18.
+17.
 *Convolvulo arvensis-Agropyrion repentis* Semiruderal grasslands.
@@
 Correspondence with EUNIS habitat *V38 Dry perennial anthropogenic herbaceous vegetation*.
+
+17. *Silybo mariani-Urticion piluliferae*
 
@@
 25.
-*Aegopodion podagrariae* Tall-herb perennial vegetation of forest margins and clearings.
-Occupies shadier and less disturbed situations than the other alliances in the class.
-Sparse occurrence throughout the ecoregion.
-Diagnostic species in the ecoregion: *Anthriscus sylvestris*, *Heracleum sphondylium*, *Galium aparine*.
-Correspondence with EUNIS habitats *R55 Lowland moist or wet tall-herb and fern fringe* and *V39 Mesic perennial anthropogenic herbaceous vegetation*.
-
-26.
 *Epilobion angustifolii* Tall-herb perennial vegetation of forest margins and clearings in acidic soils.
@@
 
-27.
+26.
 *Cynancho-Convolvulion sepium* Tall-herb fringe vegetation of nutrient-rich riparian habitats.
@@
 
-28.
+27.
 *Senecionion fluviatilis* Tall-herb vegetation of nutrient-rich riverbanks and ditches.

### 2024-10-11 doc/response.Rmd
revision for AVS

@@
 output: 
-  bookdown::word_document2:
-    reference_docx: template.docx
+  bookdown::pdf_document2:
     number_sections: false
+    toc: false
 bibliography: refs.bib
@@
 
-> Dear authors, I found this study interesting, even if limited to a relatively short geographical area.
+## Dear authors, I found this study interesting, even if limited to a relatively short geographical area.
 I believe this study can also be relevant at the European scale; therefore, I think it is suitable to be published in AVS.
@@
 
-> In general, the survey is interesting and has been conducted correctly in the analysis procedure and the results yielded are consistent and respond to the initial questions.
+## In general, the survey is interesting and has been conducted correctly in the analysis procedure and the results yielded are consistent and respond to the initial questions.
 For that reason I consider it can be published in AVS.
@@
 
-> The title does not have any indication of the geographical frame of the survey and it is relevant to indicate it to potential readers. It is in the keywords but it should be in the title.
+## The title does not have any indication of the geographical frame of the survey and it is relevant to indicate it to potential readers. It is in the keywords but it should be in the title.
 
@@
 
-> The entire article is built up using the frame of the concept of ecoregion (Bailey 2004, 1989; Olson et al.
+## The entire article is built up using the frame of the concept of ecoregion (Bailey 2004, 1989; Olson et al.
 2001).
@@
 
-> It seems that the plots used are original of NW Iberia, as shown in Figure 1.
+## It seems that the plots used are original of NW Iberia, as shown in Figure 1.
 The boundaries of the grey area, supposedly taken from the map of Olson et al.
@@
 
-We agree that the definition and name of the studied biogeographical unit requires further clarification First, we must make it clear how we define our study region: "the territories with a temperate climate in the north-western part of the Iberian Peninsula".
+We agree that the definition and name of the studied biogeographical unit requires further clarification.
+First, we must make it clear how we define our study region: "the territories with a temperate climate in the north-western part of the Iberian Peninsula".
 In other words, we are following precisely the definition by Fernández Prieto et al.
@@
 First, our study region also includes Rivas-Martínez's *Orocantabrian subprovince*.
-Second, Rivas-Martínez's *Cantabro-Atlantic subprovince* includes a larger portion of French territory.
-In the third (and, to our knowledge, last) edition of the Biogeographic map of Europe [@RN5592], the *Cantabro-Atlantic subprovince* extends in France from the French Basque country in the south to the Cotentin Peninsula in Normandy.
-In sum, the more precise correspondence between our study region and Rivas-Martínez's biogegraphical classification would be "the Iberian part of the *European Atlantic province*".
+Second, Rivas-Martínez's *Cantabro-Atlantic subprovince* includes a larger portion of French territory than our study region.
+Moreover, the *Cantabro-Atlantic subprovince* was not a stable entity in the third versions of Rivas-Martínez's map.
+For example, in the first edition, it extended in France from the French Basque country in the south to Calais in the north.
+In the second edition, the extent in the north was reduced to the Cotentin Peninsula in Normandy.
+Regarding the third edition, we have been unable to check it, since it is not available in digital format.
+In sum, the more precise correspondence between our study region and Rivas-Martínez's biogeographical classification would be "the Iberian part of the *European Atlantic province*".
 We include this equivalence in our methods section, in the subsection "Biogeographical territory under study".
@@
 
-In any case, to avoid confusion, we have revised the manuscript's section "Biogeographical territory under study" to better explain the study area and its equivalence in different classifications.
+In any case, to avoid confusion, we have revised the manuscript's section "Biogeographical territory under study" [page 7, lines 8-17] to better explain the study area and its equivalence in different classifications.
 We have also revised the whole manuscript to avoid using the term *Iberian Atlantic ecoregion*, which mixes the concepts *Cantabrian Mixed Forests ecoregion* by Olson et al.
@@
 
-> One of the main results of this survey is the selection of 28 alliances (appendix S4) from an initial set of 38 that could be present in the Iberian Atlantic ecoregion.
+## One of the main results of this survey is the selection of 28 alliances (appendix S4) from an initial set of 38 that could be present in the Iberian Atlantic ecoregion.
 So, I understand that there are 10 alliances have not been detected.
@@
 
-The missing 10 alliances are listed in the discussion, section *Vegetation and habitat classification of anthropogenic communities*.
-Some of these alliances were present in the dataset in the sense that the relevés had been assigned to syntaxa belonging to these alliances by the original authors of the relevé.
+We list them in the methods [from page 10, line 21 to page 11, line 4] and the discussion [page 19, lines 14-22].
+These alliances were present in the dataset in the sense that the relevés had been assigned to syntaxa belonging to these alliances by the original authors of the relevé.
 However, the alliances were not supported by our numerical classification, i.e.
-their relevés were asigned to another alliance, or to the noise group.
+their relevés were assigned to another alliance, or to the noise group.
 
-> Concerning the weakly represented alliances.
+## Concerning the weakly represented alliances.
 This is inevitable if we start with a definition of the study area prior to surveying its plant communities.
@@
 a political division.
-Furthermore, there were alliances that are only found in the extremes (geographical or ecological) of the study area, but still they were supported by the numerical classification.
+Furthermore, there were alliances that are only found in the extremes (i.e.
+the contact with the Mediterranean region) of the study area, but still they were supported by the numerical classification.
 We believe that the fact that the missing alliances were not supported by the numerical classification indicates that the relevés that had been assigned to these alliances are poor representatives of the vegetation types, and more likely transitional communities at the extreme of the alliance distributions.
 
-> In S2 author’s names should be corrected and homogenized e.g.
+## In S2 author’s names should be corrected and homogenized e.g.
 Lence Paz is C.
@@
 
-In Appendix S2, author names are provided exactly as they are stored in the SIVIM database. This is necessary to ensure cross-referencing between our dataset and the SIVIM database.
+In Appendix S2, author names are provided exactly as they are stored in the SIVIM database.
+This is necessary to ensure cross-referencing between our dataset and the SIVIM database.
+We have specified this in the heading of the Appendix S2.
 
-> In the figure 1 the area does not include any French terrotory [...] In the maps of figure 1 there is no French territory included
+## In the figure 1 the area does not include any French terrotory [...] In the maps of figure 1 there is no French territory included
 
@@
 
-> Anthropogenic syntaxa have a larger biogeographical scope than other vegetation types, particularly grasslands and scrub, but they have  narrower distribution than others such as aquatic vegetation.
+## Anthropogenic syntaxa have a larger biogeographical scope than other vegetation types, particularly grasslands and scrub, but they have  narrower distribution than others such as aquatic vegetation.
 On the other hand, they have been described by traditional phytsociology after fine ecological variations.
@@
 
-The inflation is reflected in the 78 associations which had been described in the original relevés as stored in SIVIM (see "Results - Overview of the classification").
-It was not our goal to classify to the level of these associations, but our exploratory analysis indicated that they were not supported by numerical classification (i.e.
-TWINSPAN within the alliances did not result in splits along association lines, and many locally-described associations were only represented by a low number of relevés).
+We have removed this sentence from the revised version.
 
-> There are several papers on this topic published in the area by Loidi and other authors, they are quoted in the adjointed sheet
+## There are several papers on this topic published in the area by Loidi and other authors, they are quoted in the adjointed sheet
 
@@
 
-> classification at the alliance level ..
+## classification at the alliance level ..
 
@@
 
-> In this paper only forest and mantle communities are dealt, no nitrophilous vegetation.
+## In this paper only forest and mantle communities are dealt, no nitrophilous vegetation.
 
@@
 
-> In these disturbed communities, geophytes, with some exceptions, use to be  scarce.
+## In these disturbed communities, geophytes, with some exceptions, use to be  scarce.
 Do you think neccesary to highlight their proportion more than other life-forms?
@@
 
-We agree that, in most of the alliances, only the variation of hemicriptophytes and therophytes is relevant.
+We agree that, in most of the alliances, only the variation of hemicryptophytes vs.
+therophytes is relevant.
 However, the presence of geophytes is an important part of the description of some of the alliances (e.g.
@@
 [-@RN4698]: *Mesic nitrophilous geophyte-rich fringe vegetation of the Western Mediterranean*).
-Furthermore, the variation in geophytes explained a significant part of the overall variance in a exploratory PCA.
-For these two reasons, we decided to highlight the propotion of geophytes in our manuscript.
+Furthermore, the variation in geophytes explained a significant part of the overall variance in an exploratory PCA.
+For these two reasons, we decided to highlight the proportion of geophytes in our manuscript.
 
-> This figure is difficult to read.
+## This figure is difficult to read.
 Point clouds overlap and the information is partially not visible.
@@
 
-> It should be cited as: Castroviejo, S. (coord. gen.). 1986-2020. Flora iberica. Real Jardín Botánico, CSIC, Madrid [and other comments regarding the references]
+## It should be cited as: Castroviejo, S. (coord. gen.). 1986-2020. Flora iberica. Real Jardín Botánico, CSIC, Madrid [and other comments regarding the references]
 
@@
 
-> This paper presents a classification and characterisation of the anthropogenic vegetation alliances of the Iberian Atlantic ecoregion in Europe.
+## This paper presents a classification and characterisation of the anthropogenic vegetation alliances of the Iberian Atlantic ecoregion in Europe.
 The study area is not very large, but I think that the paper is a valuable contribution to the knowledge of this vegetation type in Europe, and it can be considered suitable for AVS after some revision.
@@
 
-> The authors claim that (1) there is a need for a numerical classification and synthesis of the full spectrum of anthropogenic vegetation, and (2) an appropriate scale to conduct such study is the ecological region.
+## The authors claim that (1) there is a need for a numerical classification and synthesis of the full spectrum of anthropogenic vegetation, and (2) an appropriate scale to conduct such study is the ecological region.
 While I strongly agree with point 1, I'm less convinced about point 2.
@@
 
-We fully agree that a pan-European synthesis would be worthwhile goal, and never intended to suggest the contrary.
+We fully agree that a pan-European synthesis would be a worthwhile goal, and we never intended to suggest the contrary.
 In fact, we would be very happy of any research efforts on continental studies of anthropogenic vegetation.
@@
 
-> The title is a bit too general. Please consider to add a subtitle such as "a case-study from SW Europe", or something similar.
+## The title is a bit too general. Please consider to add a subtitle such as "a case-study from SW Europe", or something similar.
 
@@
 
-> The merits of the applied two-step approach (first a Twinspan classification, then a "semi-supervised" noise clustering) remain unclear to me.
+## The merits of the applied two-step approach (first a Twinspan classification, then a "semi-supervised" noise clustering) remain unclear to me.
 More details are needed to fully understand what the authors have done.
@@
 
+We agree that we may have done a poor job of explaining our classification approach.
 The main aim of our classification was for it to be consistent with regional phytosociological expertise, and to match that expertise with current EuroVegChecklist alliances [@RN4698], not to create a new classification.
@@
 [-@RN5593], the aim is "preserving the good old units and searching for new ones").
-We found that this important point was not clearly explained in the text or the abstract, so we revised the text accordingly.
-We have also provided more details on the TWINSPAN procedure as requested by the reviewer.
 
-> The authors conclude that ten alliances previously reported from the study area do not occur there.
+We found that this important point was not clearly explained in the text or the abstract, so we revised the text accordingly [from page 9, line 19 to page 11, line 9].
+In the same lines, we have also provided more details on the TWINSPAN procedure as requested by the reviewer.
+
+While revising these aspects, we have repeated all the classification analysis.
+During the new classification, we found and fixed some errors in the syntaxonomical checklist, as well as in the assignation of TWINSPAN clusters.
+This has resulted in an improved classification into 25 alliances.
+
+## The authors conclude that ten alliances previously reported from the study area do not occur there.
 This seems a bit hasty.
@@
 Therefore, we believe that the fact that the missing alliances were not supported by the numerical classification supports that the relevés that had been assigned to these alliances are poor representatives of the vegetation types, and more likely transitional communities at the extreme of the alliance distributions.
-Nonetheless, we have modified the text to be less conclusive about their absence, by adding the following sentence to the discussion: *Of course, the fact that these ten alliances are not supported by our numerical methods only indicates that they are poorly represented in our dataset, and should not be taken as absolute proof that they are absent from the Iberian Atlantic territories*.
+Nonetheless, we have modified the text to be less conclusive about their absence, by adding the following sentence to the discussion: *Of course, the fact that these nine alliances are not supported by our numerical methods only indicates that they are poorly represented in our dataset, and should not be taken as absolute proof that they are absent from the Iberian Atlantic territories*.
 
-> Most importantly, a synoptic table of the alliances is missing.
+## Most importantly, a synoptic table of the alliances is missing.
 I consider a table showing the constancy of all species in the 28 alliance, with the diagnostic species highlighted, an absolute prerequisite for the paper to be acceptable.
@@
 
-**Pendiente de hacer (Víctor?)**
+We have included the full synoptic table as Appendix S5.
 
-> p5, l11: "non-overlapping definitions" ... different, but certainly overlapping definitions!
+## p5, l11: "non-overlapping definitions" ... different, but certainly overlapping definitions!
 
-You are right, of course. We have change it to *not fully equivalent*.
+We agree, and we have changed it to *"not fully equivalent"*.
 
-> p9, l15: Why did you use 15% as cutlevel, and not 5%?
+## p9, l15: Why did you use 15% as cut level, and not 5%?
 The latter would correspond to the border between Braun-Blanquet class 1 and 2, while 15 cuts the Braun-Blanquet class 2 in the middle.
-Which cutlevel then corresponds to Braun-Blanquet class 2?**
+Which cut level then corresponds to Braun-Blanquet class 2?**
 
 Our transformation of Braun-Blanquet values follows the standard of assigning intermediate cover values.
-Thus, the BrBl value 2 (representing 5-25% cover) is transformed to 15% and the value 3 (25-50%) is transformed to 37.5%.
+Thus, the Burble value 2 (representing 5-25% cover) is transformed to 15% and the value 3 (25-50%) is transformed to 37.5%.
 Although we could use more approximate pseudospecies levels, the results of TWINSPAN will be the same when the transformation values are selected within the original cover ranges.
 
-> p10, l5: Please clarify if the same Twinspan settings were used as before.
+## p10, l5: Please clarify if the same TwinspanR settings were used as before.
 
@@
 
-> p20, l10: "which are mostly differentiated by dominant vegetation height" ... Is this a new result of your study, or has this been mentioned in the literature (if so, please provide a reference)?
+## p20, l10: "which are mostly differentiated by dominant vegetation height" ... Is this a new result of your study, or has this been mentioned in the literature (if so, please provide a reference)?
 
-Yes, this is a traditional part of the alliance descriptions, and we have added this to our text with a reference.
+Yes, this is a traditional part of the alliance descriptions, but in any case we have removed the sentence from the manuscript.
 
-> p21, l12: "neophytes tended to be in higher proportions in vegetation types associated to wet conditions (Bidentetea)" ...
+## p21, l12: "neophytes tended to be in higher proportions in vegetation types associated to wet conditions (Bidentetea)" ...
 I would have expected an even higher portion of neophytes in the Senecionion fluviatilis.
@@
 
-Our *Senecionion fluviatilis* plots are certainly dominated by alien species, especifically by *Arundo donax*, which has a mean cover of 74%.
-However, *Arundo donax* is accompanied by many native species (mean = 12 native species per plot) and so the plot-level proportion of neophytes seems low (as it is just the propotion, not weighted by species cover).
+Our *Senecionion fluviatilis* plots are certainly rich in alien species, specially some communities invaded by *Arundo donax*, in which this species has a mean cover of 74%.
+However, in these communities, *Arundo donax* is accompanied by many native species (mean = 12 native species per plot) and so the plot-level proportion of neophytes seems low (as it is just the proportion, not weighted by species cover).
 That being said, it might be possible that plots with neophytes are generally underrepresented across the dataset, because of a sampling bias towards "natural" communities.
@@
 
-> References: In several references the page numbers are missing!
+## References: In several references the page numbers are missing!
 
@@
 
-> Appendix S3: What is the value for dominant species? Constancy? Mean cover value? Something else? Please clarify.
+## Appendix S3: What is the value for dominant species? Constancy? Mean cover value? Something else? Please clarify.
 
-We have added information on how we calculated the values for diagnostic, constant and dominant species.
+We have added to the Appendix header the information on how we calculated the values for diagnostic, constant and dominant species.
 

### 2024-10-11 doc/manuscript.Rmd
revision for AVS

--- 2024-10-03 doc/manuscript.Rmd
@@
 
-* *Questions* How can we reconcile regional and European classifications of anthropogenic plant communities at the biogeographical scale?
-How are these communities characterized by species origins, species traits and ecological preferences?
+* *Questions* Can we reconcile regional and European classifications of anthropogenic plant communities at the biogeographical scale?
+How are these communities characterized by species origins, traits and ecological preferences?
 
-* *Location* Atlantic territories in the NW Iberian Peninsula (a.k.a. Cantabrian Mixed Forests ecoregion); NW Portugal, NW Spain, and SW France; south-western Europe.
+* *Location* Atlantic territories in the NW Iberian Peninsula (a.k.a. Cantabrian Mixed Forests ecoregion); south-western Europe.
 
-* *Methods* We classified 2,508 synanthropic plots with the aim of being consistent with regional phytosociological expertise, while matching that expertise with current EuroVegChecklist alliances.
-We used modified TWINSPAN to revise the original phytosociological classification of the plots, followed by semi-supervised re-classification of the whole dataset.
-We determined the proportion of natives, archaeophytes, and neophytes.
-We also described the alliances in terms of species traits (lifeforms, plant height, and flowering phenology) and ecological requirements (temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity).
+* *Methods* We classified 2,508 plots with the aim of being consistent with regional phytosociological expertise, while matching that expertise with current EuroVegChecklist alliances.
+We used modified TWINSPAN to revise the original phytosociological classification, followed by semi-supervised re-classification of the whole dataset.
+We determined the proportion of natives, archaeophytes and neophytes.
+We also described the alliances in terms of species traits (lifeforms, height and flowering phenology) and ecological requirements (temperature, moisture, light, nutrients, soil reaction, disturbance frequency and severity).
 
-* *Results* We classified 2,081 vegetation plots into 28 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*).
-The plots included 1,162 plant taxa: 78% natives, 15% archaeophytes, and 7% neophytes.
-Vegetation groups were organized along a principal axis of variation related to abiotic stress (dry-sunny to moist-shady habitats), and a second axis related to disturbance (low to high disturbance frequency and severity).
+* *Results* We assigned 2,086 vegetation plots to 25 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*).
+The plots included 1,149 taxa: 78% natives, 15% archaeophytes and 7% neophytes.
+Vegetation groups were organized along a principal axis of abiotic stress (dry-sunny to moist-shady habitats) and a secondary axis of disturbance.
 
-* *Conclusions* Biogeographical-level synthesis detected a discrepancy between the number of anthropogenic vegetation units described in the literature (39 alliances) and the number supported by numerical classification (28).
-The diversity of anthropogenic vegetation can be organized into three groups: trampled, weeds, and ruderals.
-In the Iberian Atlantic territories, anthropogenic habitats host one third of the regional plant species pool and one fifth of the Iberian flora.
-Mesic perennial ruderal vegetation is especially rich in native species and can be a biodiversity asset in urban ecosystems.
-Our biogeographical-level framework can improve the assessment and management of anthropogenic plant communities within biogeographically meaningful regions, and contribute towards a European-level synthesis of human-made vegetation.
+* *Conclusions* In the Iberian Atlantic territories, anthropogenic habitats host one third of the regional plant species pool and one fifth of the Iberian flora.
+Mesic perennial ruderal vegetation is especially rich in native species and can be a biodiversity asset in urban landscapes.
+Our biogeographical-level synthesis can improve the management of anthropogenic plant communities and contribute towards a European-level synthesis of human-made vegetation.
 
@@
 
-Anthropogenic vegetation consists of plant communities assembled as a direct consequence of human activities.
-In general, this vegetation is composed of so-called weeds of arable fields and ruderal plants of human settlements and nearby habitats [@RN5441].
-Anthropogenic communities make up an important part of the supply of biodiversity in urban and peri-urban areas [@RN5445].
-In densely populated landscapes, the plant communities that colonize urban vacant lots [@RN2300] are a useful natural resource with high ecological and societal potential [@RN5446].
+Anthropogenic vegetation is the sum of plant communities assembled as a direct consequence of human activities.
+In general, this vegetation is composed of the so-called weeds of arable fields and the ruderal plants growing on human settlements and nearby habitats [@RN5441].
+In densely populated landscapes, the anthropogenic communities that colonize vacant lots [@RN2300] are a useful natural resource with high ecological and societal potential [@RN5446; @RN5445].
 Anthropogenic vegetation can also mitigate regional-level extreme climatic events [@RN5440].
-Furthermore, many synanthropic plants can accumulate heavy metals in their biomass [@RN5433], making them valuable candidates for nature-based solutions such as bioremediation and restoration of degraded post-industrial landscapes [@RN5447].
+Furthermore, many synanthropic plants are adapted to disturbance and can accumulate heavy metals in their biomass [@RN5433], making them valuable candidates for nature-based solutions such as bioremediation and restoration of degraded post-industrial landscapes [@RN5447].
 
-In temperate Europe, anthropogenic vegetation is a melting pot of floras from different origins.
-Starting with a pool of native species favored by humans, and constantly enriched by human-assisted arrival of alien (i.e.
+In temperate Europe, anthropogenic vegetation is a melting pot of floras from different biogeographical origins.
+The starting native pool of species favored by humans has been constantly enriched by human-assisted arrival of alien (i.e.
 non-native) species [@RN5504].
@@
 Some of these archaeophytes are 'obligate weeds' that do not occur in non-anthropogenic habitats and are assumed to have evolved together with human agriculture [@RN5494].
-Since many archaeophytes started their expansion into temperate areas during the Mid-Holocene warm period, it is difficult to disentangle to what extent this expansion was favored by human activities and/or ongoing climatic changes [@RN5432; @RN5443].
 The synanthropic flora of Europe has been further enriched by neophytes, alien taxa that arrived after the expansion of intercontinental trade in the Modern Age (after 1.500 CE) [@RN5439].
-Synanthropic neophytes therefore include many American, African, and Asian taxa that arrived in Europe because of the global trade networks established by western European colonialism [@RN5451].
-In the last century, the European anthropogenic vegetation has shown a decrease in species richness and diversity [@RN5423], as rare synanthropic species have become rarer, and neophytes have increased their abundance at the expense of natives and archaeophytes [@RN5441].
+Synanthropic neophytes therefore include many American, African and Asian taxa that arrived in Europe because of the global trade networks established by western European colonialism [@RN5451].
+In the last century, the European anthropogenic vegetation has shown a decrease in species richness and diversity [@RN5423], as rare synanthropic species have become rarer and neophytes have increased their abundance at the expense of natives and archaeophytes [@RN5441].
 
-Despite the potential of anthropogenic vegetation as a biodiversity asset in human-made landscapes, and its ecological interest as an assemblage of species pools from different biogeographical origins, human-made vegetation is largely absent from ecosystem management.
-A comprehensive management will require a solid definition, classification, and description of anthropogenic plant communities.
+Despite the potential of anthropogenic vegetation as a biodiversity asset in human-made landscapes and its ecological interest as an assemblage of species pools from different biogeographical origins, human-made vegetation is largely absent from ecosystem management.
+A comprehensive management will require a solid definition, classification and description of anthropogenic plant communities.
 However, the classification of anthropogenic vegetation is subject to particularities making it a special case compared to the classification of non-anthropogenic communities.
-In human-made habitats, the traditional understanding of community assembly based on sequential filtering (dispersal, abiotic, and biotic) must incorporate an additional filtering based on human preferences and actions [@RN5505].
-Human landscapes are a mosaic of land uses [@RN5527], having specific effects on plant community assembly, and driven by sociological rather than traditional ecological factors [@RN2300].
-In addition to these potentially confounding factors, human-assisted dispersal means that anthropogenic floras are relatively homogeneous across large biogeographical scales.
-However, traditional phytosociology has tended to classify them with a level of detail similar to the one dedicated to non-anthropogenic vegetation, leading to an inflation of anthropogenic syntaxa [@RN5528].
+In human-made habitats, the traditional understanding of community assembly based on sequential filtering (dispersal, abiotic and biotic) must incorporate an additional filtering based on human preferences and actions [@RN5505].
+Human landscapes are a mosaic of land uses [@RN5527], having specific effects on plant community assembly and driven by sociological rather than traditional ecological factors [@RN2300].
 Another issue is defining anthropogenic vegetation, especially in a continent such as Europe where human impact on the landscape has been widespread since ancient times.
-The two major vegetation/habitat classifications coexisting in Europe today, phytosociological classification [@RN4698] and the EUNIS habitat classification [@RN5061], have not fully equivalent definitions of which plant communities fall within the ‘anthropogenic’ or ‘manmade’ categories.
+The two major vegetation/habitat classifications coexisting in Europe today, phytosociological classification [@RN4698] and the EUNIS habitat classification [@RN5061], have not fully equivalent definitions of which plant communities fall within their ‘anthropogenic’ or ‘manmade’ categories.
 These issues are further muddled by the fact that most classification studies of anthropogenic vegetation have focused either on a restricted political unit [@RN5424; @RN5426] or in a subset of the vegetation, such as segetal [@RN5494], ruderal [@RN5422] or trampled [@RN5434] vegetation.
@@
 Cantabrian Mixed Forests ecoregion) have a long history of human habitation dating back to the Cantabrian Upper Paleolithic [@RN5450].
-This, together with their transitional position at the border between the temperate and Mediterranean climatic zones of Europe [@RN4700], suggests a long regional history for archaeophyte-rich anthropogenic plant communities, supporting the current diverse anthropogenic vegetation with both temperate and Mediterranean floristic elements [@RN5008].
+This, together with their transitional position at the border between the temperate and Mediterranean climatic zones of Europe, suggests a long regional history for archaeophyte-rich anthropogenic plant communities, supporting the current diverse anthropogenic vegetation with both temperate and Mediterranean floristic elements [@RN5008].
 Moreover, historical trade links with America and Asia, together with the warm and humid temperate climate, have made the region a hotspot for biological invasions [@RN5492; @RN5493].
@@
 brownfields) whose management and restoration requires an understanding of the anthropogenic communities able to colonize them [@RN5448; @RN5449].
-Although there is a long tradition of studying weed and ruderal communities within the area [@RN5364; @RN5361; @RN5359; @RN5360; @RN5589; @RN5590; @RN5591] there is a current need for an biogeographical-level synthesis that revises and updates the classification in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
-In this article, we have performed such a synthesis with the following two objectives: (1) to provide an updated classification at the alliance level of anthropogenic plant communities in the Iberian Atlantic territories; and (2) to characterize the diversity of anthropogenic vegetation in terms of species pools (natives, archaeophytes, neophytes); species traits (life form, plant height, flowering phenology); and ecological requirements (temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity).
+Although there is a long tradition of studying weed and ruderal communities within the area [@RN5364; @RN5361; @RN5359; @RN5360; @RN5589; @RN5590; @RN5591] there is a current need for a biogeographical-level synthesis that revises and updates the classification in accord with recent developments in European vegetation science [@RN4698] and habitat classification [@RN5061].
+In this article, we have performed such a synthesis with the following two objectives: (1) to provide an updated classification at the alliance level of anthropogenic plant communities in the Iberian Atlantic territories; and (2) to characterize the diversity of anthropogenic vegetation in terms of species pools (natives, archaeophytes, neophytes); species traits (life form, plant height, flowering phenology); and ecological requirements (temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity).
 
@@
 We homogenized all taxon names using Euro+Med [@RN5044] or Plants of the World Online [@RN5376] for taxa not in Euro+Med.
-For the nomenclature of syntaxa, we followed Mucina *et al.* [-@RN4698] for alliances and higher ranks, and Rivas-Martínez *et al.* [-@RN4697] for associations.
+For the nomenclature of syntaxa, we followed Mucina *et al.* [-@RN4698] for alliances and higher ranks and Rivas-Martínez *et al.* [-@RN4697] for associations.
 The original datasets, as well as R code for analysis and creation of the manuscript, are stored in a GitHub repository available at Zenodo (see 'Data availability statement').
@@
 To circumscribe our study vegetation, we followed the definition of anthropogenic vegetation in the revised classification of the vegetation of Europe by Mucina *et al.* [-@RN4698].
-In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris*, and *Epilobietea angustifolii*.
+In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*.
 For the sake of completeness, we also considered the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [-@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls.
@@
 To assist in our classification, we prepared a checklist of anthropogenic syntaxa that could be present in our study area (**Appendix S1**), based on regional syntaxonomical checklists [@RN5363; @RN5008; @RN5365] and recent revisions at the European level [@RN4698; @RN5371].
-This syntaxonomical checklist included 39 anthropogenic vegetation alliances that could be present in the Iberian Atlantic territories according to the literature.
+This syntaxonomical checklist included 34 anthropogenic vegetation alliances that could be present in the Iberian Atlantic territories according to the literature.
 
@@
 
-We performed an exploratory data analysis of these 3,249 plots using modified Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379] with the R package *twinspanR* [@RN5380], 3 pseudospecies cut levels (0, 15, 25), a minimum group size of 10 plots, and Sørensen's average dissimilarity.
+We performed an exploratory data analysis of these 3,249 plots using modified Two-Way Indicator Species Analysis (TWINSPAN) [@RN5379] with the R package *twinspanR* [@RN5380].
+We used 3 pseudospecies cut levels (0, 15, 25), a minimum group size of 10 plots and Sørensen's average dissimilarity.
 During this analysis we identified 741 outlier plots which did not belong to either the target anthropogenic vegetation or to the geographical area of the Iberian Atlantic territories.
 The majority of these plots corresponded to coastal plant communities and had been misclassified by the expert system.
-We removed these plots, leaving a dataset of 2,508 vegetation plots for data analysis: 2,419 plots from SIVIM, originally recorded in 89 publications (**Appendix S2**); plus 89 plots added by us.
+We removed these plots, resulting in a dataset of 2,508 vegetation plots for data analysis: 2,419 plots from SIVIM, originally recorded in 89 publications (**Appendix S2**); plus 89 plots added by us.
 
@@
 
-The main aim of our classification was for it to be consistent with regional phytosociological expertise, and to match that expertise with current EuroVegChecklist alliances [@RN4698].
+The main aim of our classification was for it to be consistent with regional phytosociological expertise and to match that expertise with current EuroVegChecklist alliances [@RN4698].
 To achieve this, one advantage of using the SIVIM database is that a majority of the plots had been assigned to associations or alliances by the original authors [@RN5374].
 However, there could be misclassification due to unclear concepts or human error.
-Therefore, we used TWINSPAN to revise the classification by the original authors, to define the number of clusters that better aligned with our syntaxonomical checklist, and to remove outliers or misclassified plots.
-This allowed us to create a training dataset for a posterior semi-supervised classification (see below), which contained only those plots that better represented each alliance.
+Therefore, we used TWINSPAN to revise the classification by the original authors, to define the number of clusters that better aligned with our syntaxonomical checklist and to remove outliers or misclassified plots.
 
-We started this stage of the analyses by keeping only those plots (n = 2,218) that had been assigned by the original authors of the plot to any of the 39 anthropogenic alliances defined in our syntaxonomical checklist (i.e., we removed the plots without syntaxa, that had been retrieved solely by the expert system).
-We further cleaned this subset using TWINSPAN classification (with the same settings as above), in which we attempted to match the plots into the 39 alliances.
-In a first round of TWINSPAN, we divided the dataset into an increasing number of clusters, until reaching 15 clusters, beyond which the new created clusters were not phytosociologically meaningful.
-Then, we performed a second round of TWINSPAN within each of the 15 clusters.
-In this second round, we assigned each of the clusters to an alliance in our checklist, based on the composition of diagnostic species as described in the regional literature.
-Within each of the 15 primary clusters, we created an increasing number of secondary clusters until the newly created clusters splited one of the alliances that had already been assigned.
+We started this stage of the analyses by keeping only those plots (n = 2,095) that had been assigned by the original authors of the plot to any of the 9 anthropogenic vegetation classes defined in our syntaxonomical checklist (i.e., we removed the plots without syntaxa, that had been retrieved solely by the expert system).
+Then, we performed a TWINSPAN classification (with the same settings as above) within each vegetation class, attempting to classify its plots into a number of clusters equal to the number of alliances of that class in our checklist.
+We assigned the resulting clusters to an alliance in our checklist based on their composition of diagnostic species as described in the regional literature.
+In some cases, this resulted in a classification into a number of clusters in which (1) more than one cluster was assigned to the same alliance in the checklist; and (2) some alliances in the checklist could not be assigned to any of the clusters (i.e.
+the plots that had been assigned by the original authors to that alliance were separated in different clusters).
+By the end of this process, of the 34 alliances in our checklist, 25 had been assigned to clusters.
+Nine alliances were not assigned to any cluster.
+Of these, eight had been assigned to a few plots by the original authors: *Onopordion castellani* (n = 6 plots), *Onopordion acanthii* (5), *Euphorbion prostratae* (4), *Caucalidion lappulae* (4), *Roemerion hybridae* (2), *Hordeion murini* (2), *Taeniathero-Aegilopion geniculatae* (1) and *Fragarion vescae* (1).
+The alliance *Aegopodion podagrariae* had been assigned by the original authors to 39 plots, but our TWINSPAN analysis could not separate this alliance from the alliances *Geo urbani-Alliarion officinalis* and *Balloto-Conion maculati*.
 
@@
 regional phytosociological expertise).
-These plots (n = 1,725) became our training dataset for semi-supervised classification.
-Furthermore, this step allowed us to validate which alliances from our syntaxonomical checklist were present in the dataset (n = 28) and which were absent (n = 10).
+These plots (n = 1,850) were the best representatives of each alliance and became our training dataset for semi-supervised classification.
 
@@
 
-Next, we conducted a semi-supervised classification [@RN5593] of the whole dataset (n = 2,508 plots) into the 28 validated anthropogenic alliances.
+Next, we conducted a semi-supervised classification [@RN5593] of the whole dataset (n = 2,508 plots) into the 25 TWINSPAN-validated anthropogenic alliances.
 Semi-supervised classification uses a training subset of *a priori* classified vegetation plots to classify a secondary subset of unclassified plots [@RN5383].
-Since our goal was to refine the classification of the whole dataset, we allowed plots from the training subset to be re-assigned to other alliances during the classification.
-In addition, semi-supervised classification can create new groups to place data points that do not match the already existing *a priori* groups, but attempts to do so resulted in new groups with no ecological significance, and thus we kept the 28 alliances as final anthropogenic vegetation groups.
+Since our goal was to refine the classification of the whole dataset, we allowed plots from the training subset to be re-assigned to other alliances during the re-classification.
+In addition, semi-supervised classification can create new groups to place data points that do not match the already existing *a priori* groups, but attempts to do so resulted in new groups with no ecological significance and thus we kept the 25 alliances as final anthropogenic vegetation groups.
 Furthermore, we applied a noise clustering fuzzy algorithm, which allows plots to be classified either into the *a priori* alliances or into a *noise* group which includes outliers and transitional plots [@RN5382].
@@
 We performed this analysis with the R package *vegclust* [@RN5383].
-The semi-supervised classification with noise clustering resulted in the final classification of 2,081 plots into 28 alliances and 427 plots left out in the noise group.
-We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances, and Holm's p-value correction) fitted with the R package *RVAideMemoire* [@RN5381] to test the significance of the final vegetation alliances; along with Principal Component Analysis (PCA) as implemented in the R package *FactoMineR* [@RN3166] to visualize the relationships between the classes and alliances.
+The semi-supervised classification with noise clustering resulted in the final classification of 2,086 plots into 26 alliances and 422 plots left out in the noise group.
+We used pairwise PERMANOVA (with 100,000 iterations, Euclidean distances and Holm's p-value correction) fitted with the R package *RVAideMemoire* [@RN5381] to test the significance of the final vegetation alliances; along with Principal Component Analysis (PCA) as implemented in the R package *FactoMineR* [@RN3166] to visualize the relationships between the classes and alliances.
 
@@
 
-To facilitate the use of the classification by local managers and environmental consultants, we calculated sets of characteristic species (**Appendix S3**) [@RN5061] for each one of the final 28 vegetation alliances.
+To facilitate the use of the classification by local managers and environmental consultants, we calculated sets of characteristic species [@RN5061] for each one of the final 25 vegetation alliances.
 We defined dominant species as species with more than 25% cover in at least 5% of the vegetation plots of the group; constant species as species with a frequency higher than 50% in the group; and diagnostic species as species whose *IndVal* had a p-value lower than 0.05, as calculated with the R package *labdsv* [@RN2311] using 1.000.000 iterations.
-Additionally, we assigned to each alliance its corresponding level-3 EUNIS habitat code or codes (<https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp>).
+Additionally, we assigned to each alliance a regionalized level 4 EUNIS habitat code (<https://eunis.eea.europa.eu/habitats-code-browser-revised.jsp>).
 
-## Species origin as native, archaeophyte, or neophyte
+## Species origin as native, archaeophyte or neophyte
 
-We classified the species as native, archaeophytes, or neophytes in the Iberian Atlantic territories using the information in *Flora iberica* [-@RN3243] and catalogues of archaeophytes for Britain [@RN5490] and the Czech Republic [@RN5495].
+We classified the species as native, archaeophytes or neophytes in the Iberian Atlantic territories using the information in *Flora iberica* [-@RN3243] and catalogues of archaeophytes for Britain [@RN5490] and the Czech Republic [@RN5495].
 We must stress that identifying archaeophytes in southern Europe is highly problematic [@RN5444] and our classification must be taken as an indication of putative archaeophyte character for the purposes of vegetation description and comparison with other European regions, rather than a definitive classification of the species; as stated by Preston *et al.* [@RN5490], the classification of a species as an archaeophyte should be interpreted as a hypothesis to be tested by further studies.
@@
 
-From *Flora iberica* [-@RN3243] we extracted the maximum height, median month of flowering, and length of the flowering period of each species in the dataset.
+From *Flora iberica* [-@RN3243] we extracted the maximum height, median month of flowering and length of the flowering period of each species in the dataset.
 We used these variables to calculate plot-level community-weighted means (weighting by species cover) for each trait, to characterize the vegetation height and flowering phenology of the anthropogenic plant communities.
@@
 
-For each species in the dataset, we collected its ecological indicator values of temperature, moisture, light, nutrients, and soil reaction [@RN5488]; and its disturbance frequency and severity indicator values [@RN5101].
+For each species in the dataset, we collected its ecological indicator values of temperature, moisture, light, nutrients and soil reaction [@RN5488]; and its disturbance frequency and severity indicator values [@RN5101].
 To those species lacking an indicator value in the accessed references, we assigned a value by reciprocal averaging, i.e.
-by (1) calculating weighted average values of the species with values for each plot (weighting by species cover), and (2) assigning to the missing species the weighted average values of the plots where they were present (weighting by species cover).
+by (1) calculating weighted average values of the species with values for each plot (weighting by species cover) and (2) assigning to the missing species the weighted average values of the plots where they were present (weighting by species cover).
 Then, for each vegetation plot, we calculated the plot-level mean of each indicator value.
@@
 
-The semi-supervised classification resulted in the classification of 2,081 vegetation plots into 28 anthropogenic alliances (**Fig.
-1**, **Appendix S4**) representing 9 vegetation classes.
-Most of the alliances had a wide distribution in the Iberian Atlantic territories according to our dataset, but five of them (*Caucalidion lappulae*, *Linario polygalifoliae-Vulpion alopecuri*, *Convolvulo arvensis-Agropyrion repentis*, *Senecionion fluviatilis* and *Paspalo-Agrostion semiverticillati*) had isolated occurrences towards the southern limit of the Iberian Atlantic territories.
-The regional literature included in our dataset recognized 72 associations (**Appendix S5**) within these alliances, but exploratory analysis (not shown) indicated that the separation of these associations within the alliances had a generally weak support based on numerical classification methods.
+The semi-supervised classification resulted in the classification of 2,086 vegetation plots into 25 anthropogenic alliances (**Fig.
+1**, description of the alliances and correspondence with EUNIS habitat types in **Appendix S3**, groups of characteristic species in **Appendix S4**, synoptic table in **Appendix S5**) representing 9 vegetation classes.
+The regional literature included in our dataset recognized 69 associations (**Appendix S6**) within these alliances, but exploratory analysis (not shown) indicated that the separation of these associations within the alliances had a generally weak support based on numerical classification methods.
 
-The classified plots included 1,162 taxa or taxa aggregates.
-The 10 most frequent species were *Ochlopoa annua* (621 occurrences), *Urtica dioica* (592), *Sonchus oleraceus* (540), *Stellaria media* (516), *Capsella bursa-pastoris* (390), *Polygonum aviculare* (375), *Dactylis glomerata* (363), *Senecio vulgaris* (332), and *Anisantha sterilis* (301).
+The classified plots included 1,149 taxa or taxa aggregates.
+The 10 most frequent species were *Ochlopoa annua* (625 occurrences), *Urtica dioica* (588), *Sonchus oleraceus* (548), *Stellaria media* (520), *Capsella bursa-pastoris* (396), *Polygonum aviculare* (373), *Dactylis glomerata* (369), *Senecio vulgaris* (336) and *Anisantha sterilis* (310).
 Considering only those plots with the most frequent plot size range (10-30 m^2^, n = 867 plots), the average species richness per plot was 16 (minimum = 3, maximum = 48).
-The class with the richest species pool was *Epilobietea angustifolii* (n = 615) and the poorest was *Cymbalario-Parietarietea diffusae* (n = 171).
+The class with the richest species pool was *Epilobietea angustifolii* (n = 605) and the poorest was *Bidentetea* (n = 132).
 
 PCA ordination of the floristic composition (**Fig.
-2A**) clearly differentiated three vegetation groups: (1) trampled vegetation of class *Polygono-Poetea annuae*; (2) crop weeds of classes *Papaveretea rhoeadis* and *Digitario sanguinalis-Eragrostietea minoris*; and (3) perennial ruderal vegetation of classes *Artemisietea vulgaris* and *Epilobietea angustifolii*.
-The central position of the floristic space, between these three major groups, was occupied by the annual ruderal vegetation of classes *Chenopodietea*, *Sisymbrietea*, and *Bidentetea*; as well as by the wall vegetation of class *Cymbalario-Parietarietea diffusae*.
-Separate PCAs within the major vegetation groups indicated a coherent separation between most of the alliances, but with some cases of relatively high overlap: for example, between (i) *Galio valantiae-Parietarion judaicae* and *Cymbalario-Asplenion*; or (ii) *Geo urbani-Alliarion officinalis* and *Aegopodion podagrariae*.
+2A**) differentiated three vegetation groups: (1) trampled vegetation of class *Polygono-Poetea annuae*; (2) crop weeds of classes *Papaveretea rhoeadis* and *Digitario sanguinalis-Eragrostietea minoris*; and (3) perennial ruderal vegetation of classes *Artemisietea vulgaris* and *Epilobietea angustifolii*.
+The central position of the floristic space, between these three major groups, was occupied by the annual ruderal vegetation of classes *Chenopodietea*, *Sisymbrietea* and *Bidentetea*; as well as by the wall vegetation of class *Cymbalario-Parietarietea diffusae*.
+Separate PCAs within the major vegetation groups indicated a coherent separation between most of the alliances, but with some cases of relatively high overlap, especially between *Galio valantiae-Parietarion judaicae* and *Cymbalario-Asplenion*.
 
@@
 
-The total synanthropic species pool (n = 1,162) was composed of 78% natives (n = 908), 15% putative archaeophytes (n = 178), and 7% neophytes (n = 76).
-The most frequent putative archaeophytes were *Capsella bursa-pastoris* (390 occurrences), *Senecio vulgaris* (332), *Anisantha sterilis* (301), *Anthemis arvensis* (280), *Hordeum murinum* (274), *Raphanus raphanistrum* (233), *Vicia sativa* (228), *Malva sylvestris* (197), *Sisymbrium officinale* (166), and *Echinochloa crus-galli* (165); while the most frequent neophytes were *Veronica persica* (212), *Erigeron canadensis* (138), *Amaranthus hybridus* (73), *Oxalis corniculata* (67), *Lepidium didymum* (57), *Oxalis latifolia* (51), *Paspalum distichum* (48), *Bidens frondosus* (35), *Cyperus eragrostis* (34), and *Symphyotrichum squamatum* (33).
+The total synanthropic species pool (n = 1,149) was composed of 78% natives (n = 899), 15% putative archaeophytes (n = 174) and 7% neophytes (n = 76).
+The most frequent putative archaeophytes were *Capsella bursa-pastoris* (396 occurrences), *Senecio vulgaris* (336), *Anisantha sterilis* (310), *Hordeum murinum* (297), *Anthemis arvensis* (290), *Raphanus raphanistrum* (236), *Vicia sativa* (229), *Malva sylvestris* (210), *Sisymbrium officinale* (185) and *Anisantha diandra* (175); while the most frequent neophytes were *Veronica persica* (215), *Erigeron canadensis* (136), *Amaranthus hybridus* (75), *Oxalis corniculata* (66), *Lepidium didymum* (59), *Oxalis latifolia* (56), *Paspalum distichum* (42), *Bidens frondosus* (34), *Cyperus eragrostis* (34) and *Symphyotrichum squamatum* (34).
 
-By vegetation class, *Epilobietea angustifolii* had the highest proportion of natives in its species pool (82%); *Sisymbrietea* had the highest proportion of putative archaeophytes (31%); and *Bidentetea* had the highest proportion of neophytes (17%).
-The alliances with the highest proportions of archaeophytes in their species pools were *Caucalidion lappulae* (41%), *Chenopodion muralis* (35%%), and *Sisymbrion officinalis* (31%); while the alliances with the highest proportion of neophytes were *Bidention tripartitae* (19%), *Allion triquetri* (12%), and *Spergulo arvensis-Erodion cicutariae* (10%).
+By vegetation class, *Epilobietea angustifolii* had the highest proportion of natives in its species pool (82%); *Sisymbrietea* had the highest proportion of putative archaeophytes (30%); and *Bidentetea* had the highest proportion of neophytes (16%).
+The alliances with the highest proportions of archaeophytes in their species pools were *Chenopodion muralis* (35%%), *Silybo mariani-Urticion piluliferae* (31%) and *Sisymbrion officinalis* (30%); while the alliances with the highest proportion of neophytes were *Bidention tripartitae* (20%), *Allion triquetri* (12%) and *Spergulo arvensis-Erodion cicutariae* (11%).
 
 At the community level, focusing on plot average proportions (**Fig.
-3**), the alliances with the most archaeophytes were *Caucalidion lappulae*, *Spergulo arvensis-Erodion cicutariae*, and *Sisymbrion officinalis*, in which only about half of the plot-level proportion was made up of native species.
-The alliances with the highest plot-level proportion of neophytes were *Bidention tripartitae*, *Paspalo-Agrostion semiverticillati*, and *Senecionion fluviatilis*.
+3**), the alliances with the most archaeophytes were *Sisymbrion officinalis*, *Silybo mariani-Urticion piluliferae* and *Spergulo arvensis-Erodion cicutariae*, in which only about half of the plot-level proportion was made up of native species.
+The alliances with the highest plot-level proportion of neophytes were *Bidention tripartitae*, *Paspalo-Agrostion semiverticillati* and *Spergulo arvensis-Erodion cicutariae*.
 Some alliances had high proportions of native species, e.g.
-*Epilobion angustifolii*, *Convolvulo arvensis-Agropyrion repentis*, and *Linario polygalifoliae-Vulpion alopecuri*.
+*Epilobion angustifolii*, *Linario polygalifoliae-Vulpion alopecuri* and *Arction lappae*.
 
@@
 
-The most frequent life form in the species pool (n = 1,162 species) were hemicryptophytes (44%), followed by therophytes (42%), geophytes (9%), chamaephytes (6%), phanerophytes (6%), bryophytes (1%), and hydrophytes (1%).
+The most frequent life form in the species pool (n = 1,149 species) were hemicryptophytes (44%), followed by therophytes (41%), geophytes (8%), chamaephytes (6%), phanerophytes (5%), bryophytes (1%) and hydrophytes (1%).
 The proportion of therophytes across classes and alliances (**Fig.
-4**) agreed with the traditional description of the syntaxa: therophytes dominated the annual communities of trampled-soil vegetation (*Polygono-Poetea annuae*), crops weeds (*Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*), and annual ruderals (*Chenopodietea*, *Sisymbrietea*, *Bidentetea*); while perennial life forms (especially hemicryptophytes) dominated walls (*Cymbalario-Parietarietea diffusae*) and perennial ruderal vegetation (*Artemisietea vulgaris*, *Epilobietea angustifolii*).
-Geophytes represented a relatively high proportion of the vegetation in three alliances: *Allion triquetri*, *Convolvulo arvensis-Agropyrion repentis*, and *Senecionion fluviatilis*.
+4**) agreed with the traditional description of the syntaxa: therophytes dominated the annual communities of trampled-soil vegetation (*Polygono-Poetea annuae*), crops weeds (*Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*) and annual ruderals (*Chenopodietea*, *Sisymbrietea*, *Bidentetea*); while perennial life forms (especially hemicryptophytes) dominated walls (*Cymbalario-Parietarietea diffusae*) and perennial ruderal vegetation (*Artemisietea vulgaris*, *Epilobietea angustifolii*).
+Geophytes represented a relatively high proportion of the vegetation in two alliances: *Allion triquetri* and *Senecionion fluviatilis*.
 
 Community-weighted means for plant height (**Fig.
-5A**) also agreed with the expected description of the syntaxa: the shortest communities were the dwarf-herb vegetation of trampled sites (*Polygono-Poetea annuae*, average height = 45.2 cm) and crop weeds (*Papaveretea rhoeadis*, 55.4 cm), while the tallest were the perennial ruderal vegetation of dry (*Artemisietea vulgaris*, 129 cm) and mesic (*Epilobietea angustifolii*, 136 cm) sites.
+5A**) also agreed with the expected description of the syntaxa: the shortest communities were the dwarf-herb vegetation of trampled sites (*Polygono-Poetea annuae*, average height = 45.2 cm) and crop weeds (*Papaveretea rhoeadis*, 55.6 cm), while the tallest were the perennial ruderal vegetation of dry (*Artemisietea vulgaris*, 131 cm) and mesic (*Epilobietea angustifolii*, 138 cm) sites.
 The median month of flowering (**Fig.
@@
 The length of the flowering season (**Fig.
-5C**) was generally high, from 5 months (*Artemisietea vulgaris*, *Epilobietea angustifolii*) to 6 (*Chenopodietea*, *Sisymbrietea*), 7 (*Bidentetea*, *Digitario sanguinalis-Eragrostietea minoris*, *Papaveretea rhoeadis*), 8 (*Polygono-Poetea annuae*), or 9 (*Cymbalario-Parietarietea diffusae*).
+5C**) was generally high, from 5 months (*Artemisietea vulgaris*, *Epilobietea angustifolii*) to 6 (*Chenopodietea*, *Sisymbrietea*), 7 (*Bidentetea*, *Digitario sanguinalis-Eragrostietea minoris*, *Papaveretea rhoeadis*), 8 (*Polygono-Poetea annuae*) or 9 (*Cymbalario-Parietarietea diffusae*).
 
@@
 7**).
-The first two PCA axes explained 59% of the variability.
-The major contributors to axis 1 (36% variance explained) were light, nutrient, and moisture requirements; it separated communities of open, dry and comparatively nutrient-poorer sites (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea*, *Artemisietea vulgaris*, *Chenopodietea*, and *Digitario sanguinalis-Eragrostietea minoris*) from communities of shady, moist and nutrient-richer sites (classes *Cymbalario-Parietarietea diffusae*, *Bidentetea*, and *Epilobietea angustifolii*).
-The major contributors to axis 2 (23% variance explained) were disturbance severity and frequency; this axis separated communities preferring less severe and less frequent disturbances (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii*, *Chenopodietea*, *Artemisietea vulgaris* and *Bidentetea*) from communities adapted to more severe and more frequent disturbances (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea* and *Digitario sanguinalis-Eragrostietea minoris*).
+The first two PCA axes explained 59% of the variance.
+The major contributors to axis 1 (36% variance explained) were light, nutrient and moisture requirements; it separated communities of open, dry and comparatively nutrient-poorer sites (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea*, *Artemisietea vulgaris*, *Chenopodietea* and *Digitario sanguinalis-Eragrostietea minoris*) from communities of shady, moist and nutrient-richer sites (classes *Bidentetea* and *Epilobietea angustifolii*).
+The major contributors to axis 2 (23% variance explained) were disturbance severity and frequency; this axis separated communities preferring less severe and less frequent disturbances (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii*) from communities adapted to more severe and more frequent disturbances (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea* and *Digitario sanguinalis-Eragrostietea minoris*).
 
@@
 
-Our synthesis of the anthropogenic vegetation has detected a discrepancy between the number of vegetation units described in the literature (i.e.
-39 alliances) and the number supported by numerical classification (28 alliances).
-Our classification supported that anthropogenic vegetation can be broadly divided into three groups belonging to trampled, weed, and ruderal communities.
-These groups are organized along a principal axis of variation related to abiotic stress, and a second axis related to disturbance.
+Our synthesis of the anthropogenic vegetation of the northwestern Iberian Peninsula has detected a discrepancy between the number of vegetation units described in the literature (i.e.
+34 alliances) and the number supported by numerical classification (25 alliances).
+Our classification supported that anthropogenic vegetation can be broadly divided into three groups belonging to trampled, weed and ruderal communities.
+These groups are organized along a principal axis of variation related to abiotic stress and a second axis related to disturbance.
 These two axes determine the characteristics of the communities in terms of species origins and species traits.
@@
 
-Our analysis supports the existence of 28 anthropogenic alliances in the vegetation of the Iberian Atlantic territories.
-While the core of this vegetation diversity is made up of temperate European alliances, these are enriched by the occurrence of eight Mediterranean alliances (*Galio valantiae-Parietarion judaicae*, *Polycarpion tetraphylli*, *Allion triquetri*, *Chenopodion muralis*, *Echio-Galactition tomentosae*, *Paspalo-Agrostion semiverticillati*, *Balloto-Conion maculati* and *Cynancho-Convolvulion sepium*); plus two alliances that are endemic of the Iberian coasts (*Linario polygalifoliae-Vulpion alopecuri*) and mountains (*Cirsion richterano-chodati*) [@RN4698].
-Seven of these alliances had been reported as absent or uncertain in the Iberian Atlantic territories (*Caucalidion lappulae*, *Chenopodion muralis*, *Onopordion acanthii*), or only present in the Portuguese sector of the area (*Spergulo arvensis-Erodion cicutariae*, *Allion triquetri*, *Echio-Galactition tomentosae*, *Cynancho-Convolvulion sepium*) by the recently published distribution maps of vegetation alliances in Europe [@RN5371].
-Our dataset supports that six of these alliances are relatively well distributed in the study area, while *Caucalidion lappulae* only has isolated occurrences.
+Our analysis supports the existence of 25 anthropogenic alliances in the vegetation of the Iberian Atlantic territories.
+While the core of this vegetation diversity is made up of temperate European alliances, these are enriched by the occurrence of eight Mediterranean alliances (*Galio valantiae-Parietarion judaicae*, *Polycarpion tetraphylli*, *Allion triquetri*, *Chenopodion muralis*, *Echio-Galactition tomentosae*, *Paspalo-Agrostion semiverticillati*, *Silybo mariani-Urticion piluliferae* and *Balloto-Conion maculati*); plus three alliances that are endemic of the Iberian coasts (*Linario polygalifoliae-Vulpion alopecuri*) and mountains (*Carduo carpetani-Cirsion odontolepidis*, *Cirsion richterano-chodati*) [@RN4698].
+Five of these alliances had been reported as absent or uncertain in the Iberian Atlantic territories (*Chenopodion muralis*) or only present in the Portuguese sector of the area (*Spergulo arvensis-Erodion cicutariae*, *Allion triquetri*, *Echio-Galactition tomentosae*, *Carduo carpetani-Cirsion odontolepidis*) by the recently published distribution maps of vegetation alliances in Europe [@RN5371].
 
-On the other hand, our dataset does not support the occurrence in the Iberian Atlantic territories of ten Mediterranean alliances (*Roemerion hybridae*, *Diplotaxion erucoidis*, *Euphorbion prostratae*, *Alysso granatensis-Brassicion barrelieri*, *Hordeion murini*, *Taeniathero-Aegilopion geniculatae*, *Parietarion lusitanico-mauritanicae*, *Onopordion castellani*, *Carduo carpetani-Cirsion odontolepidis*, *Silybo mariani-Urticion piluliferae*) and one temperate alliance (*Fragarion vescae*) that had been cited as potentially occurring here [@RN5363; @RN5008; @RN5365; @RN4698; @RN5371].
-Of course, the fact that these eleven alliances are not supported by our numerical methods only indicates that they are poorly represented in our dataset, and should not be taken as absolute proof that they are absent from the Iberian Atlantic territories.
-
-For habitat classification purposes, we propose that this vegetation diversity can be summarized into 15 EUNIS habitat types: one constructed habitat (*J2.5*), ten vegetated human-made habitats (*V11*, *V12*, *V13*, *V15*, *V32*, *V35*, *V34*, *V37*, *V38*, *V39*), three habitats of woodland fringes and clearings and tall forb stands (*R54*, *R55*, *R57*) and one habitat of periodically-exposed shores (*Q61*).
-In general, we have followed the correspondence between alliances and EUNIS habitats proposed by the EUNIS expert system [@RN5061], except for *Echio-Galactition tomentosae* which, in the context of the study area, seems to be more adequately placed in the habitat *V37 Annual anthropogenic herbaceous vegetation* rather than in the habitat *V33 Dry Mediterranean land with unpalatable non-vernal herbaceous vegetation*.
+On the other hand, our dataset does not support the occurrence in the Iberian Atlantic territories of nine alliances (*Caucalidion lappulae*, *Roemerion hybridae*, *Euphorbion prostratae*, *Hordeion murini*, *Taeniathero-Aegilopion geniculatae*, *Onopordion castellani*, *Onopordion acanthii*, *Fragarion vescae* and *Aegopodion podagrariae*) that had been cited as potentially occurring here [@RN5363; @RN5008; @RN5365; @RN4698; @RN5371].
+Of course, the fact that these nine alliances are not supported by our numerical methods only indicates that they are poorly represented in our dataset and should not be taken as absolute proof that they are absent from the Iberian Atlantic territories.
 
@@
 However, there are some cases that could deserve further scrutiny.
-For example, the two wall alliances (*Galio valantiae-Parietarion judaicae* and *Cymbalario-Asplenion*) had a large overlap in their geographic distribution, their floristic composition, their proportion of alien species, their community traits, and their ecological preferences.
-These two alliances are well recognized in the European literature, and they are generally interpreted as representing the Mediterranean (*Galio valantiae-Parietarion judaicae*) and temperate (*Cymbalario-Asplenion*) versions of wall vegetation [@RN5369; @RN5367].
+For example, the two wall alliances (*Galio valantiae-Parietarion judaicae* and *Cymbalario-Asplenion*) had a large overlap in their geographic distribution, their floristic composition, their proportion of alien species, their community traits and their ecological preferences.
+These two alliances are well recognized in the European literature and they are generally interpreted as representing the Mediterranean (*Galio valantiae-Parietarion judaicae*) and temperate (*Cymbalario-Asplenion*) versions of wall vegetation [@RN5369; @RN5367].
 It is therefore not surprising that in biogeographic transitional regions - such as the Iberian Atlantic territories - these two alliances coexist and show a high overlap.
-Also, there was overlap between the alliances *Geo urbani-Alliarion officinalis* and *Aegopodion podagrariae*, two alliances that can occupy ruderal to semi-natural nitrophilous fringes, and which are traditionally differentiated in their descriptions [@RN4698] by vegetation height, as was supported by our dataset.
+
+For habitat classification purposes, we propose that this vegetation diversity can be summarized into 25 EUNIS habitat types: two constructed habitats (*J2.5*), three habitats of arable crops and gardens (*V1*), 16 habitats of anthropogenic grasslands (*V3*), two habitats of woodland clearings and fringes (*R5*) and two habitats of periodically-exposed shores (*Q61*).
+In general, we have followed the correspondence between alliances and EUNIS habitats proposed by the EUNIS expert system [@RN5061], except for *Echio-Galactition tomentosae* which, in the context of the study area, seems to be more adequately placed in the habitat *V37 Annual anthropogenic herbaceous vegetation* rather than in the habitat *V33 Dry Mediterranean land with unpalatable non-vernal herbaceous vegetation*.
 
@@
 Taken together, these results suggest that the Iberian Atlantic anthropogenic vegetation shows an incidence of non-native species that is intermediate between southern and central-northern Europe, which is in agreement with the biogeographical position of our study area.
-Previous studies also suggested that the non-natives were more frequent in early successional anthropogenic communities [@RN5423], and our results confirmed this as the proportion of native species tended to be higher in the perennial ruderal classes *Artemisietea vulgaris* and *Epilobietea angustifolii*.
+Previous studies also suggested that the non-natives were more frequent in early successional anthropogenic communities [@RN5423] and our results confirmed this as the proportion of native species tended to be higher in the perennial ruderal classes *Artemisietea vulgaris* and *Epilobietea angustifolii*.
 The classes with more non-native species also tended to be those adapted to more frequent and more severe disturbances (*Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Sisymbrietea*), also in agreement with previous research [@RN5435].
@@
 
-The traits of the Iberian Atlantic anthropogenic communities agreed with those found in other European regions [@RN5436; @RN2297], with more annual species in habitats associated to arable land and more perennials in ruderal habitats, and a general dominance of therophytes and hemicryptophytes in the species pool [@RN5428].
+The traits of the Iberian Atlantic anthropogenic communities agreed with those found in other European regions [@RN5436; @RN2297], with more annual species in habitats associated to arable land and more perennials in ruderal habitats and a general dominance of therophytes and hemicryptophytes in the species pool [@RN5428].
 We found that the main drivers of variation in anthropogenic community composition was a gradient from open-dry to shady-wet sites, supporting the importance of the moisture gradient in shaping anthropogenic vegetation [@RN5434].
@@
 
-Biogeographical-level synthesis detected a discrepancy between the number of anthropogenic vegetation units described in the literature and the number supported by numerical classification.
-The diversity of anthropogenic vegetation can be organized into trampled, weed and ruderal communities varying primarily along an axis of abiotic stress.
-Importantly, the anthropogenic vegetation of the Iberian Atlantic territories is home to 1,162 plant taxa.
-This is approximately one third of the whole regional species pool represented in the SIVIM database [@RN5374], and one fifth of the Iberian flora [@RN5503].
-More than half of this synanthropic diversity (615 taxa) occurs in mesic to wet perennial ruderal vegetation, highlighting the potential of some anthropogenic habitats as a source of biodiversity in human-dominated landscapes [@RN5446; @RN5445].
+The anthropogenic vegetation of the Iberian Atlantic territories is home to more than a thousand plant taxa.
+This is approximately one third of the whole regional species pool represented in the SIVIM database [@RN5374] and one fifth of the Iberian flora [@RN5503].
+More than half of this synanthropic diversity (c.
+600 taxa) occurs in mesic to wet perennial ruderal vegetation, highlighting the potential of some anthropogenic habitats as a source of biodiversity in human-dominated landscapes [@RN5446; @RN5445].
 These results suggest three main objectives for the sustainable management of human-made habitats: (1) focusing efforts on alien plant management in anthropogenic plant communities of wet habitats; (2) conserving examples of archaeophyte-rich annual weed communities, especially those linked to traditional agricultural practices [@RN5502]; and (3) conserving and promoting mesic ruderal and fringe vegetation as a biodiversity refuge in urban and peri-urban areas.
@@
 
-**Appendix S2** Original publications providing the vegetation plots stored in SIVIM and used in our analysis.
+**Appendix S2** Original publications providing the vegetation plots stored in SIVIM and used in our analysis. Reference details are provided as they are in the SIVIM database.
 
-**Appendix S3** Characteristic species combinations of the anthropogenic vegetation alliances in spreadsheet format, divided into diagnostic, constant, and dominant species.
+**Appendix S3** Description of the 25 alliances present in the anthropogenic vegetation of the northwestern Iberian Peninsula, and correspondence with regionalized level 4 EUNIS habitats.
+
+**Appendix S4** Characteristic species combinations of the anthropogenic vegetation alliances in spreadsheet format, divided into diagnostic, constant and dominant species.
 The values are as follows: Diagnostic species = *IndVal* multiplied by 100; Constant species = percentage occurrence frequency (constancy); Dominant species = percentage frequency of plots in which the species occurs with a cover larger than 25%.
 
-**Appendix S4** Synthetic description of the 28 alliances present in the Iberian Atlantic anthropogenic vegetation.
+**Appendix S5** Synoptic table of the fidelity (phi) of the diagnostic species of the 25 anthropogenic alliances.
 
-**Appendix S5** Revised checklist of vegetation associations described in the regional literature and belonging to the 28 alliances present in the Iberian Atlantic anthropogenic vegetation.
+**Appendix S6** Revised checklist of vegetation associations described in the regional literature and belonging to the 25 alliances present in the Iberian Atlantic anthropogenic vegetation.
 
@@
 
-```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of neophytes, archaeophytes, and native species in the anthropogenic plant communities of the Iberian Atlantic territories.
-Each bar represents the plot-level average proportion of neophytes (dark shade), archaeophytes (medium shade), and natives (light shade) in each vegetation alliance.
+```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of neophytes, archaeophytes and native species in the anthropogenic plant communities of the Iberian Atlantic territories.
+Each bar represents the plot-level average proportion of neophytes (dark shade), archaeophytes (medium shade) and natives (light shade) in each vegetation alliance.
 Colors indicate the vegetation class."}
@@
 
-```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of therophytes, geophytes, and other life forms (hemicryptophytes, chamaephytes, phanerophytes, bryophytes, hydrophytes) in the anthropogenic plant communities of the Iberian Atlantic territories.
-Each bar represents the plot-level average proportion of therophytes (dark shade), geophytes (medium shade), and other life forms (light shade) in each vegetation alliance.
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of therophytes, geophytes and other life forms (hemicryptophytes, chamaephytes, phanerophytes, bryophytes, hydrophytes) in the anthropogenic plant communities of the Iberian Atlantic territories.
+Each bar represents the plot-level average proportion of therophytes (dark shade), geophytes (medium shade) and other life forms (light shade) in each vegetation alliance.
 Colors indicate the vegetation class."}
@@
 ```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Vegetation height and flowering phenology of the anthropogenic plant communities of the Iberian Atlantic territories.
-Boxplots show the community-weighted mean for plant height, median month of flowering, and length of the flowering season for each vegetation alliance.
+Boxplots show the community-weighted mean for plant height, median month of flowering and length of the flowering season for each vegetation alliance.
 Colors indicate the vegetation class.
-Alliances as ordered as in Figures 1 and 3-4, and alliance numbers are the same as listed in Appendix S4."}
+Alliances as ordered as in Figures 1 and 3-4 and alliance numbers are the same as listed in Appendix S3."}
 knitr::include_graphics("../results/figures/F5 - height-flowering.png")
@@
 ```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic territories.
-Boxplots show the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity, calculated for each vegetation alliance.
+Boxplots show the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity, calculated for each vegetation alliance.
 Colors indicate the vegetation class.
-Alliances as ordered as in Figures 1 and 3-4, and alliance numbers are the same as listed in Appendix S4."}
+Alliances as ordered as in Figures 1 and 3-4 and alliance numbers are the same as listed in Appendix S3."}
 knitr::include_graphics("../results/figures/F6 - indicators.png")
@@
 ```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic territories.
-Biplot produced by Principal Component Analysis (PCA) of the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency, and disturbance severity.
+Biplot produced by Principal Component Analysis (PCA) of the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.
 Each dot is a plot.

### 2024-10-11 doc/title-page.Rmd
zenodo link ⛓️
revision for AVS

@@
 ---
-title: "Classification and characterization of anthropogenic plant communities at the ecoregion level"
+title: "Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula"
 output: 
@@
 Original datasets, R code for analysis and Rmarkdown code for creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/anthropogenic> stored at https://zenodo.org/doi/10.5281/zenodo.11613571.
-
-. Upon publication, a version of record of the repository will be deposited in Zenodo.

### 2024-10-11 results/supporting/AppendixS3.Rmd
revision for AVS

---
title: "Appendix S3 Description of the 25 alliances present in the anthropogenic vegetation of the northwestern Iberian Peninsula, and correspondence with regionalized level 4 EUNIS habitats."
output: 
  bookdown::pdf_document2:
    number_sections: false
    toc: false
link-citations: yes
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message = FALSE, echo = FALSE, warning = FALSE}
knitr::knit_hooks$set(inline = function(x) {
  prettyNum(x, big.mark = ",")
})
```

***Cymbalario-Parietarietea diffusae*** Chasmophytic vegetation of human-made walls.

1.  *Galio valantiae-Parietarion judaicae* Hemicryptophyte-rich vegetation of sunny human-made walls. A predominantly Mediterranean alliance that occupies sunnier situations than the other alliance in the class and has a shorter flowering season. Diagnostic species: *Parietaria judaica*, *Centranthus ruber*, *Polypodium cambricum*, *Asplenium ceterach*, *Hypericum hircinum*. Regionalized level 4 EUNIS habitat: *J25Z Sunny constructed walls with hemicryptophytes - Muros construidos soleados, con hemicriptófitos*.

2.  *Cymbalario-Asplenion* Fern-rich vegetation of shady human-made walls. A predominantly temperate alliance that occupies shadier situations than the other alliance in the class and has a longer flowering season. Diagnostic species: *Cymbalaria muralis*, *Asplenium trichomanes*, *Umbilicus rupestris*, *Trachelium caeruleum*, *Asplenium scolopendrium*. Regionalized level 4 EUNIS habitat: *J25Y Shady constructed walls with ferns - Muros contruidos sombríos, con helechos*.

***Polygono-Poetea annuae*** Dwarf-annual vegetation of trampled sites.

3.  *Polycarpion tetraphylli* Dwarf-annual vegetation of trampled sunny sites. Occupies warmer, sunnier and nutrient-poorer situations than the other alliances in the class and has an earlier flowering season. Diagnostic species: *Sagina apetala*, *Plantago coronopus*, *Crassula tillaea*, *Spergularia marina*, *Sagina maritima*. Regionalized level 4 EUNIS habitat: *V34Z Trampled sunny soils with annuals - Suelos pisoteados soleados, con céspedes anuales*.

4.  *Polygono-Coronopodion* Dwarf-annual vegetation of trampled dry sites. Occupies sites that are shadier than those preferred by *Polycarpion tetraphylli*, but drier than those of *Saginion procumbentis*. Diagnostic species: *Polygonum aviculare*, *Matricaria suaveolens*, *Plantago major*. Regionalized level 4 EUNIS habitat: *V34Y Trampled dry soils with annuals - Suelos pisoteados secos, con céspedes anuales*.

5.  *Saginion procumbentis* Bryophyte-rich vegetation of trampled shady sites. Occupies shadier, colder and moister sites than the other alliances in the class, is less dominated by therophytes and has a relatively high proportion of bryophytes. Diagnostic species: *Sagina procumbens*, *Bryum argenteum*, *Ochlopoa annua*. Regionalized level 4 EUNIS habitat: *V35Z Trampled shady soils with bryophytes - Suelos pisoteados sombríos, con briófitos*.

***Papaveretea rhoeadis*** Annual weed vegetation of arable crops and vegetable gardens.

6.  *Scleranthion annui* Annual weeds of cereal crops. Occupies more acidic soils than the other alliance in the class, and flowers earlier. Rich in putative archaeophytes. Diagnostic species: *Mibora minima*, *Anthemis arvensis*, *Rumex acetosella*, *Spergula arvensis*, *Raphanus raphanistrum*, *Arnoseris minima*, *Scleranthus annuus*, *Anthoxanthum ovatum*, *Ornithopus compressus*. Regionalized level 4 EUNIS habitat: *V10Z Cereal crops with winter-spring weeds - Cultivos de cereales, con malas hierbas de invierno y primavera*.

7.  *Oxalidion europeae* Annual weeds of vegetable gardens and root crops. Occupies wetter, shadier and nutrient-richer sites, and has a longer flowering season than the other alliance in the class. Rich in putative archaeophytes, and with more neophytes than the other alliance in the class. Diagnostic species: *Stellaria media*, *Veronica persica*, *Senecio vulgaris*, *Cardamine hirsuta*, *Veronica hederifolia*, *Euphorbia helioscopia*, *Lamium amplexicaule*, *Fumaria capreolata*. Regionalized level 4 EUNIS habitat:s *V10Y Vegetable gardens and root crops with winter-spring weeds - Huertos y cultivos de raíces, con malas hierbas de invierno y primavera*.

***Digitario sanguinalis-Eragrostietea minoris*** Grass-rich weed vegetation with summer-annual C4 species.

8.  *Spergulo arvensis-Erodion cicutariae* Late-summer C4 weeds. Late-flowering alliance rich in neophytes and grasses, occurring in highly-disturbed and nutrient-rich sites. Diagnostic species: *Chenopodium album*, *Echinochloa crus-galli*, *Digitaria sanguinalis*,  *Solanum nigrum*, *Amaranthus hybridus*, *Fallopia convolvulus*, *Persicaria maculosa*, *Oxalis latifolia*, *Portulaca oleracea*. Regionalized level 4 EUNIS habitat: *V10X Crops with C4 summer weeds - Cultivos, con malas hierbas estivales C4*.

***Chenopodietea*** Winter-annual ruderal vegetation.

9. *Allion triquetri* Geophyte-rich vegetation of shady and nutrient-rich forest fringes. Occupies warmer, shadier and less frequently disturbed situations than the other alliances in the class, is comparatively taller and earlier-flowering. Has a higher proportion of geophytes and a lower proportion of therophytes. The proportion of neophytes is the highest in the class. Diagnostic species: *Urtica membranacea*, *Smyrnium olusatrum*, *Tradescantia fluminensis*, *Allium triquetrum*, *Arum italicum*, *Chelidonium majus*. Regionalized level 4 EUNIS habitat: *V37Z Shady and nutrient-rich forest fringes with geophytes - Orlas forestales escionitrófilas, con geófitos*.

10. *Geranio pusilli-Anthriscion caucalidis* Winter-annual vegetation of shady and nutrient-rich forest fringes. Occupies situations similar to the previous alliance, but in colder sites, and is a shorter vegetation dominated by therophytes instead of geophytes. Diagnostic species: *Geranium lucidum*, *Anthriscus caucalis*. Regionalized level 4 EUNIS habitat: *V37Y Shady and nutrient-rich forest fringes with winter annuals - Orlas forestales escionitrófilas, con vegetación anual de invierno*.

11. *Chenopodion muralis* Winter-annual vegetation of disturbed sites. Short vegetation occupying frequently disturbed sites. Flowers later and for longer than other alliances in the class. Rich in putative archaeophytes. This is a predominantly Mediterranean alliance, and has a sparse occurrence throughout the study area. Diagnostic species: *Malva neglecta*, *Urtica urens*, *Amaranthus deflexus*, *Chenopodiastrum murale*. Regionalized level 4 EUNIS habitat: *V37X Disturbed sites with winter annuals - Lugares alterados, con vegetación anual de invierno*.

12. *Echio-Galactition tomentosae* Sub-nitrophilous and winter-annual pioneer vegetation of abandoned sites. Occupies drier, sunnier and less nutrient-rich situations than other alliances in the class. Diagnostic species: *Galactites tomentosus*, *Coleostephus myconis*, *Anisantha rigida*, *Lolium multiflorum*, *Briza maxima*, *Vicia sativa*, *Silene gallica*, *Crepis capillaris*, *Briza minor*. Regionalized level 4 EUNIS habitat: *V37W Subnitrophilous disturbed sites with annual pioneer vegetation - Lugares alterados subnitrófilos, con vegetación pionera anual*.

13. *Linario polygalifoliae-Vulpion alopecuri* Ephemeral annual vegetation of disturbed coastal dunes. Occupies warm, dry and sunny coastal sites, and has a high number of coastal dune species. Isolated occurrence in the southern coast of the region. Diagnostic species: *Vulpia alopecuros*, *Crucianella maritima*, *Malcolmia littorea*, *Artemisia campestris* (full list in **Appendix S4**). Regionalized level 4 EUNIS habitat: *V32Z Disturbed dunes with ephemeral vegetation - Dunas alteradas, con vegetación anual efímera*.

***Sisymbrietea*** Summer-annual ruderal vegetation.

14. *Sisymbrion officinalis* Summer-annual ruderal vegetation. Rich in therophytes and archaeophytes and adapted to severe disturbances. Diagnostic species: *Hordeum murinum*, *Sisymbrium officinale*, *Lolium perenne*. Regionalized level 4 EUNIS habitat: *V37V Disturbed sites with summer annuals - Lugares alterados, con vegetación anual de verano*.

***Bidentetea*** Summer-annual pioneer vegetation of temporarily flooded sites.

15. *Bidention tripartitae* Summer-annual pioneer vegetation of temporarily flooded sites. Occupies nutrient-richer sites than the other alliance in the class and has a higher proportion of therophytes and neophytes. Diagnostic species: *Persicaria hydropiper*, *Bidens frondosus*, *Lythrum salicaria*, *Phalaroides arundinacea*, *Cyperus eragrostis* (full list in **Appendix S4**). Regionalized level 4 EUNIS habitat: *Q61Z Temporarily flooded soils, with summer-annual pioneer vegetation - Suelos temporalmente encharcados, con vegetación pionera anual*.

16. *Paspalo-Agrostion semiverticillati* Mediterranean summer-annual pioneer vegetation of temporarily flooded subsaline sites. Occupies wetter and more frequently disturbed sites than the other alliance in the class and has a longer flowering season. This is a Mediterranean alliance, which in the study area is distributed mostly in the sub-Mediterranean valleys south of the Cantabrian Mountains. Diagnostic species: *Paspalum distichum*, *Cyperus longus*, *Schoenoplectus lacustris* (full list in **Appendix S4**). Regionalized level 4 EUNIS habitat: *Q61Y Submediterranean temporarily flooded soils, with summer-annual pioneer grasslands - Suelos temporalmente encharcados submediterráneos, con céspedes pioneros anuales*.

***Artemisietea vulgaris*** Perennial ruderal vegetation of dry sites.

17. *Silybo mariani-Urticion piluliferae* Mediterranean thistle-dominated vegetation. Occupies sites that are warmer and more severely disturbed than the other alliances in the class. Is taller and flowers earlier than the other alliances. Rich in archaeophytes and therophytes. This is a Mediterranean alliance, which in the study area is distributed mostly in the sub-Mediterranean valleys south of the Cantabrian Mountains. Diagnostic species: *Silybum marianum*, *Carduus tenuiflorus*, *Anisantha diandra*. Regionalized level 4 EUNIS habitat: *V38Z Mediterranean thistle fields - Cardales mediterráneos*.

18. *Carduo carpetani-Cirsion odontolepidis* Supramediterranean thistle-dominated vegetation of the Cantabrian Mountains. Occupies warmer situations than the other alliances in the class and has a high proportion of archaeophytes. Mostly distributed in the Cantabrian Mountains. Diagnostic species: *Onopordum acanthium*, *Reseda luteola*, *Verbascum pulverulentum*, *Lactuca virosa*, *Echium vulgare*. Regionalized level 4 EUNIS habitat: *V38Y Supramediterranean thistle fields of the Cantabrian Mountains - Cardales supramediterráneos orocantábricos*.

19. *Cirsion richterano-chodati* Montane thistle fields of the Cantabrian Mountains. Occupies colder situations than the other alliances in the class. Mostly distributed in the Cantabrian Mountains. Diagnostic species: *Cirsium eriophorum*, *Carduus carpetanus*, *Carduus nutans*, *Jacobaea vulgaris*, *Cirsium richterianum*, *Cirsium arvense*, *Achillea millefolium*. Regionalized level 4 EUNIS habitat: *V38X Montane thistle fields of the Cantabrian Mountains - Cardales montanos orocantábricos*.

20. *Dauco-Melilotion* Biennial pioneer vegetation of abandoned sites. Occupies wetter situations than the thistle alliances in the class and has a higher proportion of neophytes. Diagnostic species: *Helminthotheca echioides*, *Daucus carota*, *Melilotus albus*, *Foeniculum vulgare*, *Erigeron canadensis* (full list in **Appendix S4**). Regionalized level 4 EUNIS habitat: *V38W Dry disturbed sites with biennal pioneer vegetation - Lugares alterados secos, con vegetación pionera bienal*.

***Epilobietea angustifolii*** Perennial ruderal vegetation of mesic to wet sites.

21. *Geo urbani-Alliarion officinalis* Perennial vegetation of shady and nutrient-rich forest fringes. Occupies shadier and less frequently disturbed situations than the other alliances in the class. Includes tall-herb vegetation of forest fringes, which some authors separate as *Aegopodion podagrariae*. Diagnostic species: *Urtica dioica*, *Pentaglottis sempervirens*, *Galium aparine*, *Lamium maculatum*, *Anthriscus sylvestris*. Regionalized level 4 EUNIS habitat: *V39Z Shady and nutrient-rich forest fringes with perennial vegetation - Orlas forestales escionitrófilas, con vegetación perenne*.

22. *Arction lappae* Perennial vegetation of disturbed mesic sites of the Cantabrian Mountains. Occupies sunnier and more frequently disturbed situations than the other alliances in the class, and has shorter plants. Distribution concentrated in the Cantabrian Mountains. Diagnostic species: *Oxybasis rubra*, *Senecio duriaei*. Regionalized level 4 EUNIS habitat: *V39Y Mesic disturbed sites of the Cantabrian Mountains, with perennial vegetation - Lugares alterados mésicos orocantábricos, con vegetación perenne*.

23. *Balloto-Conion maculati* Tall-herb perennial vegetation of mesic to moist disturbed sites. Occupies warmer situations than the other alliances in the class. Rich in archaeophytes. Diagnostic species: *Sambucus ebulus*, *Conium maculatum*. Regionalized level 4 EUNIS habitat: *V39X Mesic to moist disturbed sites, with tall-herb perennial vegetation - Lugares alterados de mésicos a húmedos, con megaforbios perennes*.

24. *Epilobion angustifolii* Tall-herb perennial vegetation of forest fringes in acidic soils. Occupies colder, nutrient-poorer, more acidic and less disturbed situations than the other alliances in the class. Diagnostic species: *Epilobium angustifolium*, *Digitalis purpurea*, *Avenella flexuosa*, *Luzula lactea* (full list in **Appendix S4**). Regionalized level 4 EUNIS habitat: *R57Z Forest fringes and clearings on acid soils, with fire-loving tall-herb vegetation - Orlas y claros forestales de suelos ácidos, con megaforbios pirófitos*.

25. *Senecionion fluviatilis* Tall-herb perennial vegetation of nutrient-rich riverbanks and ditches. Occupies wetter situations than the other alliances in the class, flowers later, has taller plants and has a high proportion of neophytes and geophytes. Diagnostic species: *Eupatorium cannabinum*, *Arundo donax*, *Angelica sylvestris*, *Picris hieracioides*, *Mentha suaveolens*, *Pteridium aquilinum*, *Centaurea debeauxii*, *Silene latifolia*. Regionalized level 4 EUNIS habitat: *R55Z Fringes of nutrient-rich ditches and riverbanks, with tall-herb perennial vegetation - Orlas nitrófilas de cunetas y arroyos, con megaforbios*.

### 2024-10-11 src/10-origins.R

@@ -35,8 +35,8 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                                 "Geo urbani-Alliarion officinalis",
                                 "Dauco-Melilotion", 
                                 "Cirsion richterano-chodati",
+                                "Carduo carpetani-Cirsion odontolepidis",
                                 "Silybo mariani-Urticion piluliferae",
-                                "Onopordion acanthii",
                                 "Convolvulo arvensis-Agropyrion repentis",
                                 "Paspalo-Agrostion semiverticillati",
                                 "Bidention tripartitae",
@@ -110,4 +110,4 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
   geom_hline(yintercept = 0.5, linetype = "dashed") -> Fig4; Fig4
 
 ggsave(Fig4, file = "results/figures/F3 - origins.png", bg = "white", 
-       path = NULL, scale = 1, width = 180, height = 100, units = "mm", dpi = 600)
+       path = NULL, scale = 1, width = 180, height = 90, units = "mm", dpi = 600)

### 2024-10-11 src/11-lifeforms.R

@@ -41,8 +41,8 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                                 "Geo urbani-Alliarion officinalis",
                                 "Dauco-Melilotion", 
                                 "Cirsion richterano-chodati",
+                                "Carduo carpetani-Cirsion odontolepidis",
                                 "Silybo mariani-Urticion piluliferae",
-                                "Onopordion acanthii",
                                 "Convolvulo arvensis-Agropyrion repentis",
                                 "Paspalo-Agrostion semiverticillati",
                                 "Bidention tripartitae",
@@ -116,4 +116,4 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
   geom_hline(yintercept = 0.5, linetype = "dashed") -> Fig4; Fig4
 
 ggsave(Fig4, file = "results/figures/F4 - lifeforms.png", bg = "white", 
-       path = NULL, scale = 1, width = 180, height = 100, units = "mm", dpi = 600)
+       path = NULL, scale = 1, width = 180, height = 90, units = "mm", dpi = 600)

### 2024-10-11 src/12-height-flowering.R

@@ -35,6 +35,7 @@ library(tidyverse)
 # ### CwMs
 # 
 # traits %>%
+#   select(-origin) %>%
 #   gather(Trait, Value, -Analysis.Names) %>%
 #   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
 #   group_by(SIVIMID, Trait) %>%
@@ -124,8 +125,8 @@ cwms %>%
                                "Bidention tripartitae",
                                "Paspalo-Agrostion semiverticillati",
                                "Convolvulo arvensis-Agropyrion repentis",
-                               "Onopordion acanthii",
                                "Silybo mariani-Urticion piluliferae", 
+                               "Carduo carpetani-Cirsion odontolepidis",
                                "Cirsion richterano-chodati",
                                "Dauco-Melilotion",
                                "Geo urbani-Alliarion officinalis",

### 2024-10-11 src/13-indicators.R

@@ -35,7 +35,7 @@ library(tidyverse)
 # ### Plot means
 # 
 # traits %>%
-#   select(Analysis.Names, Disturbance.Frequency, Disturbance.Severity, 
+#   select(Analysis.Names, Disturbance.Frequency, Disturbance.Severity,
 #          EIVE1.T, EIVE1.M, EIVE1.L, EIVE1.N, EIVE1.R) %>%
 #   gather(Trait, Value, -Analysis.Names) %>%
 #   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
@@ -115,8 +115,8 @@ plotmeans %>%
                                 "Bidention tripartitae",
                                 "Paspalo-Agrostion semiverticillati",
                                 "Convolvulo arvensis-Agropyrion repentis",
-                                "Onopordion acanthii",
                                 "Silybo mariani-Urticion piluliferae", 
+                                "Carduo carpetani-Cirsion odontolepidis",
                                 "Cirsion richterano-chodati",
                                 "Dauco-Melilotion",
                                 "Geo urbani-Alliarion officinalis",

### 2024-10-11 src/2-twinspan NEW.R

@@ -732,7 +732,7 @@ twinspanR::twinspan(
   modif = TRUE,
   cut.levels = c(0, 15, 25),
   min.group.size = 10,
-  clusters = 6,
+  clusters = 5,
   diss = "multi.sorensen",
   mean.median = "mean",
   show.output.on.console = FALSE,
@@ -826,10 +826,9 @@ header4 %>%
   mutate(L1 = fct_recode(as.factor(Cluster),
                          "Dauco-Melilotion" = "1",
                          "Silybo mariani-Urticion piluliferae" = "2",
-                         "Convolvulo arvensis-Agropyrion repentis" = "3",
-                         "Cirsion richterano-chodati" = "4",
-                         "Carduo carpetani-Cirsion odontolepidis" = "5",
-                         "Cirsion richterano-chodati" = "6")) %>%
+                         "Cirsion richterano-chodati" = "3",
+                         "Carduo carpetani-Cirsion odontolepidis" = "4",
+                         "Cirsion richterano-chodati" = "5")) %>%
   select(-Cluster) %>%
   write.csv("data/urban-header-5.1G.csv", fileEncoding = "latin1")
 
@@ -855,7 +854,7 @@ twinspanR::twinspan(
   modif = TRUE,
   cut.levels = c(0, 15, 25),
   min.group.size = 10,
-  clusters = 11,
+  clusters = 6,
   diss = "multi.sorensen",
   mean.median = "mean",
   show.output.on.console = FALSE,
@@ -951,17 +950,11 @@ data.frame(Community = indicators$maxcls, Indicator = indicators$indcls,
 header4 %>%
   mutate(L1 = fct_recode(as.factor(Cluster),
                          "Balloto-Conion maculati" = "1",
-                         "NA" = "2",
-                         "NA" = "3",
-                         "Balloto-Conion maculati" = "4",
-                         "NA" = "5",
-                         #"Aegopodion podagrariae" = "5",
-                         "Geo urbani-Alliarion officinalis" = "6",
-                         "NA" = "7",
-                         "Arction lappae" = "8",
-                         "Senecionion fluviatilis" = "9",
-                         "Cynancho-Convolvulion sepium" = "10",
-                         "Epilobion angustifolii" = "11")) %>%
+                         "Balloto-Conion maculati" = "2",
+                         "Geo urbani-Alliarion officinalis" = "3",
[…]

### 2024-10-11 src/2-twinspan.R

@@ -1,1874 +1,0 @@
-library(tidyverse); library(vegan); library(labdsv)
-
-rm(list = ls())
-
-read.csv("results/sintaxonomy/original-sintaxonomy.csv", fileEncoding = "latin1") %>%
-  filter(Anthropogenic == "Yes") %>%
-  #filter(Assigned != "Not Cantabrian") %>%
-  pull(Alliance) %>%
-  unique -> alliances
-
-read.csv("data/urban-header-5.0.csv", fileEncoding = "latin1") %>%
-  filter(Alliance %in% alliances) -> header
-
-read.csv("data/urban-species-5.0.csv", fileEncoding = "latin1") %>%
-  filter(SIVIMID %in% header$SIVIMID) -> species
-
-species %>%
-  select(SIVIMID, Analysis.Names, Cover.percent) %>%
-  spread(Analysis.Names, Cover.percent, fill = 0) %>%
-  column_to_rownames(var = "SIVIMID") -> df1
-
-### Twinspan 
-
-twinspanR::twinspan(
-  df1,
-  modif = TRUE,
-  cut.levels = c(0, 15, 25),
-  min.group.size = 10,
-  clusters = 15,
-  diss = "multi.sorensen",
-  mean.median = "mean",
-  show.output.on.console = FALSE,
-  quiet = TRUE) %>% 
-  cut() %>%
-  cbind(df1) %>%
-  select(".") %>%
-  rename("Cluster" = ".") %>%
-  rownames_to_column(var = "SIVIMID") -> kclusters
-
-header %>% merge(kclusters, by = "SIVIMID") -> header2
-
-### Permanova, find max number of significant clusters = 15
-
-# df1 %>% rownames_to_column(var = "SIVIMID") %>% merge(header2, by = "SIVIMID") -> df1b
-# RVAideMemoire::pairwise.perm.manova(dist(df1, "euclidian"), df1b$Cluster, nperm = 9999, p.method = "holm")
-# adonis2(df1 ~ Cluster, data = df1b, permutations = 9999)
-# anova(betadisper(vegdist(df1, method = "euclidean"), df1b$Cluster))
-
-### DCA
-
-decorana(df1) -> dca1
-
-vegan::scores(dca1) %>%
-  data.frame() %>%
-  rownames_to_column("SIVIMID") %>%
-  merge(header2) %>%
-  mutate(Cluster = as.factor(Cluster)) -> df2
-
[…]

### 2024-10-11 src/3-vegclust.R

@@ -250,6 +250,10 @@ df2 %>%
 ### Sintaxa
 
 header4 %>%
+  group_by(Cluster) %>%
+  tally
+
+header4 %>%
   group_by(Cluster, Alliance) %>%
   tally %>%
   arrange(Cluster, -n) %>%
@@ -313,8 +317,7 @@ header2 %>%
   headerClassifiedM
 
 header2 %>%
-  filter(Cluster %in% c("M1")) %>%
-  filter(! SIVIMID %in% potentillion) -> # Cluster 1 of M1 belong to Potentillion anserinae
+  filter(Cluster %in% c("M1")) -> 
   headerUnclassifiedM
 
 read.csv("data/species.csv", fileEncoding = "latin1") %>% 
@@ -354,10 +357,71 @@ newclassesM %>%
   rownames_to_column(var = "SIVIMID") %>%
   merge(read.csv("data/urban-header-5.1.csv", fileEncoding = "latin1"), by = "SIVIMID") -> header2b
 
+header2b
+
+header2b %>%
+  group_by(Cluster) %>%
+  tally
+
+header2b %>%
+  group_by(Cluster, Alliance) %>%
+  tally %>%
+  arrange(Cluster, -n) %>%
+  data.frame
+
+header2b %>%
+  group_by(Cluster, Original) %>%
+  tally %>%
+  arrange(Cluster, -n) %>%
+  data.frame
+
+### Dominant
+
+read.csv("data/species.csv", fileEncoding = "latin1")  %>%
+  merge(header2b, by = "SIVIMID") %>%
+  group_by(Cluster, Analysis.Names) %>%
+  summarise(D = sum(Cover.percent)) %>%
+  group_by(Cluster) %>%
+  slice_max(D, n = 5) %>%
+  data.frame
+
+### Constant
+
+read.csv("data/species.csv", fileEncoding = "latin1") %>%
+  merge(header2b, by = "SIVIMID") %>%
+  group_by(Cluster, Analysis.Names) %>%
+  summarise(F = length(Analysis.Names)) %>%
+  group_by(Cluster) %>%
+  slice_max(F, n = 5) %>%
+  data.frame
+
+### Indicator spp
+
+read.csv("data/species.csv", fileEncoding = "latin1") %>%
+  merge(header2b, by = "SIVIMID") %>%
+  select(Cluster, SIVIMID, Analysis.Names, Cover.percent) %>%
+  spread(Analysis.Names, Cover.percent, fill = 0) -> plots
+
[…]

### 2024-10-11 src/5-permanova.R

@@ -4,7 +4,7 @@ rm(list = ls())
 
 library(tidyverse); library(vegan)
 
-# read.csv("data/header.csv", fileEncoding = "latin1") %>% filter(! Alliance %in% "Noise") -> 
+# read.csv("data/header.csv", fileEncoding = "latin1") %>% filter(! Alliance %in% "Noise") ->
 #   header
 # 
 # read.csv("data/species.csv", fileEncoding = "latin1") %>% filter(SIVIMID %in% header$SIVIMID) -> species

### 2024-10-11 src/6-characteristic-spp.R

@@ -7,6 +7,8 @@ library(tidyverse); library(indicspecies); library(labdsv)
 read.csv("data/header.csv", fileEncoding = "latin1") %>% filter(! Alliance %in% "Noise")-> 
   header
 
+header %>% group_by(Alliance)
+
 read.csv("data/species.csv", fileEncoding = "latin1") %>% 
   filter(SIVIMID %in% header$SIVIMID) -> species
 
@@ -65,6 +67,7 @@ load(file = "results/characteristic-spp/indval.RData")
 
 read.csv("data/header.csv", fileEncoding = "latin1") %>%
   select(Alliance) %>%
+  filter(Alliance != "Noise") %>%
   unique %>%
   mutate(Community = as.numeric(as.factor(Alliance))) -> communities
 
@@ -105,7 +108,8 @@ species %>%
   mutate(Species = Analysis.Names,
          Type = "Constant",
          Value = Frequency*100) %>%
-  select(Species, Alliance, Type, Value) -> constspp
+  select(Species, Alliance, Type, Value) %>%
+  arrange(Alliance, -Value) -> constspp
 
 ### Dominant
 
@@ -135,11 +139,12 @@ species %>%
   mutate(Species = Analysis.Names,
          Type = "Dominant",
          Value = Frequency*100) %>%
-  select(Species, Alliance, Type, Value) -> domispp
+  select(Species, Alliance, Type, Value) %>%
+  arrange(Alliance, -Value) -> domispp
 
 ### Merge
 
-read.csv("results/sintaxonomy/original-sintaxonomy.csv", fileEncoding = "latin1") %>%
+read.csv("results/sintaxonomy/canteunis.csv", fileEncoding = "latin1") %>%
   select(Alliance, CANTEUNIS) %>%
   na.omit %>% unique -> habitats
 

### 2024-10-11 src/7-spp-numbers.R

@@ -1,5 +1,10 @@
 library(tidyverse)
 
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise") %>%
+  group_by(SIVIMID)
+
 ### General species diversity
 
 read.csv("data/header.csv", fileEncoding = "latin1") %>% 
@@ -26,6 +31,24 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
   group_by(SIVIMID) %>%
   summarise(l = length(Analysis.Names)) %>%
   pull(l) %>% mean
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise") %>% 
+  filter(Area >= 10) %>%
+  filter(Area <= 30) %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+  group_by(SIVIMID) %>%
+  summarise(l = length(Analysis.Names)) %>%
+  pull(l) %>% min
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise") %>% 
+  filter(Area >= 10) %>%
+  filter(Area <= 30) %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+  group_by(SIVIMID) %>%
+  summarise(l = length(Analysis.Names)) %>%
+  pull(l) %>% max
 
 ### Species origin
 
@@ -63,6 +86,20 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
   select(Alliance, origin, per) %>%
   spread(origin, per) %>%
   arrange(-archaeophyte )
+
+read.csv("data/header.csv", fileEncoding = "latin1") %>% 
+  filter(Alliance != "Noise") %>%
+  merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
+  merge(read.csv("data/traits.csv", fileEncoding = "latin1"))  %>%
+  select(Alliance, Analysis.Names, origin) %>%
+  unique %>%
+  group_by(Alliance, origin) %>% tally %>%
+  group_by(Alliance) %>%
+  mutate(T = sum(n)) %>%
+  mutate(per = n/T) %>%
+  select(Alliance, origin, per) %>%
+  spread(origin, per) %>%
+  arrange(-neophyte )
 
 read.csv("data/header.csv", fileEncoding = "latin1") %>% 
   filter(Alliance != "Noise") %>%

### 2024-10-11 src/8-maps.R

@@ -40,8 +40,8 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
                                "Bidention tripartitae",
                                "Paspalo-Agrostion semiverticillati",
                                "Convolvulo arvensis-Agropyrion repentis",
-                               "Onopordion acanthii",
                                "Silybo mariani-Urticion piluliferae", 
+                               "Carduo carpetani-Cirsion odontolepidis",
                                "Cirsion richterano-chodati",
                                "Dauco-Melilotion",
                                "Geo urbani-Alliarion officinalis",
@@ -55,7 +55,7 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
                                "Galio valantiae-\nParietarion judaicae" = "Galio valantiae-Parietarion judaicae",
                                 # "Cymbalario-Asplenion",
                                 # "Polycarpion tetraphylli",
-                                # "Polygono-Coronopodion",
+                               "Polygono-\nCoronopodion" = "Polygono-Coronopodion",
                                 # "Saginion procumbentis",
                                 # "Scleranthion annui",
                                 # "Oxalidion europeae",
@@ -67,18 +67,18 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
                                "Echio-\nGalactition tomentosae" = "Echio-Galactition tomentosae",
                                "Linario polygalifoliae-\nVulpion alopecuri" = "Linario polygalifoliae-Vulpion alopecuri",
                                 # "Sisymbrion officinalis",
-                               "Carduo carpetani-\nCirsion odontolepidis" = "Onopordion acanthii",
+                               "Carduo carpetani-\nCirsion odontolepidis" = "Carduo carpetani-Cirsion odontolepidis",
                                "Cirsion\nrichterano-chodati" = "Cirsion richterano-chodati",
                                 # "Dauco-Melilotion",
                                "Geo urbani-\nAlliarion officinalis" = "Geo urbani-Alliarion officinalis",
                                 # "Arction lappae",
-                                # "Balloto-Conion maculati",
+                                "Balloto-\nConion maculati" = "Balloto-Conion maculati",
                                 # "Aegopodion podagrariae",
[…]

### 2024-10-11 src/9-ordination.R

@@ -275,7 +275,6 @@ df2 %>%
         plot.margin = unit(c(.1,.1,0,0), "cm")) +
   scale_color_manual(values = c("chocolate1",
                                 "chocolate1",
-                                "chocolate1",
                                 "chocolate4")) -> Fig1C; Fig1C
 
 ### Chenopodietea and Sisymbrietea
@@ -486,3 +485,4 @@ cowplot::plot_grid(Fig1, legend, rel_widths = c(3, 1)) -> Figure1; Figure1
 
 ggsave(Figure1, file = "results/figures/F2 - pca.png", bg = "white", 
        path = NULL, scale = 1, width = 180, height = 165, units = "mm", dpi = 600)
+

### 2024-11-15 doc/response.Rmd
AVS R2

--- 2024-10-03 doc/response.Rmd
@@
 ---
-title: "Response letter on the manuscript Classification and characterization of anthropogenic plant communities at the ecoregion level"
+title: "Response letter on the manuscript Classification and characterization of anthropogenic plant communities at the ecoregion level (2nd revision)"
 output: 
-  bookdown::word_document2:
-    reference_docx: template.docx
+  bookdown::pdf_document2:
     number_sections: false
+    toc: false
 bibliography: refs.bib
@@
 
-# Co-ordinating Editor Comments to Author
+# Co-ordinating Editor Comments to Author:
 
-> Dear authors, I found this study interesting, even if limited to a relatively short geographical area.
-I believe this study can also be relevant at the European scale; therefore, I think it is suitable to be published in AVS.
-However, I agree with the referees that the manuscript can be significantly improved before publication.
+*Dear authors, your work has improved consistently.
+However, in this round, I involved one additional referee, Zdenka Lososová, who pointed attention to some critical points of discussion.
+Please revise the manuscript according to her requests.*
 
-We would like to thank the editor and the reviewers for the time invested in reviewing our manuscript and providing constructive comments.
-We have revised the manuscript to incorporate all the suggestions.
-Please find below a point-by-point response to each comment.
+\textcolor{teal}{Thank you for a new chance to improve our manuscript.
+We have implemented all the changes required by the new reviewer, as well as the change requested by reviewer 2 regarding the synoptic table.}
 
-# Referee 1
+# Referee: 1
 
-> In general, the survey is interesting and has been conducted correctly in the analysis procedure and the results yielded are consistent and respond to the initial questions.
-For that reason I consider it can be published in AVS.
-However, in the comments below are suggested some changes, mostly in the introduction, and other changes are highlighted in the notes over the manuscript.
-I recommend to accept the paper with major changes.
+*I consider the paper has been improved and the flaws were seen in the first review have been corrected*
 
-Thank you for your positive evaluation and your useful comments, which we have integrated into a revised version of the manuscript.
+\textcolor{teal}{Thank you again for your contribution to improving our manuscript.}
 
-> The title does not have any indication of the geographical frame of the survey and it is relevant to indicate it to potential readers. It is in the keywords but it should be in the title.
+# Referee: 3
 
-We agree and we have changed the title to *Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*.
+*The manuscript provides a detailed overview of anthropogenic vegetation types in the northwestern Iberian Peninsula.
+The authors assembled a robust dataset of plots.
+These plots were classified using semi-supervised classification methods and assigned to 25 alliances.
+The primary characteristics of these alliances are given using proportions of species with different origins, traits, and indicator values.
+I appreciate the way the classification procedure is described.
+I also appreciate the nice form of all of the figures.
+I read the manuscript with all the supplements and the response letter, and I see that the authors highly improved the manuscript.
+I did not review the manuscript in the first round.
+However, I have some serious comments which should be discussed.*
 
-> The entire article is built up using the frame of the concept of ecoregion (Bailey 2004, 1989; Olson et al.
-2001).
-About this there are two related questions: 1 What is the benefit of using such biogeographic frame in terms of how the results or the interpretations could be altered or deteriorated if not used?
-Is it a merely geographical description of the study area?
-If so, then, 2 What is the difference of considering the ecoregions instead the classical biogeographic units such as regions, in the sense of Good or Takhtatajan?
-It is important to quote that biogeographic regions also include vegetation units living within them as was established early in the 20th century by Flauhaut and Braun-Blanquet (Loidi 2021)
+\textcolor{teal}{Thank you for your kind words and your advice to improve the manuscript.}
 
-We would argue that the manuscript is built up based on the idea that this type of studies should be conducted using biogeographical units, regardless of the rank or classification system of the unit.
-In our specific example, focusing on the Iberian Atlantic territories (which rank would be "province", but see below for a discussion of the definition of the Iberian Atlantic territories and its correspondence with different biogeographical classifications) allows to study a territory that shares (1) a similar macroclimate and (2) a shared biogeographical history, including a history of biological invasions.
-This is especially important when studying anthropogenic vegetation, to define lists of regional neophytes and archaeophytes, which would be not possible if considering a larger area (for example, most "archaeophytes" and some "neophytes" would be considered "native" at the European level) or a smaller area or political unit (for example, species native to some parts of the Iberian Atlantic territories are some times considered "alien" in some other parts, e.g.
-*Pinus pinaster*).
-This can allow to make comparisons between invasion levels in different vegetation types in different regions, see for example Chytrý et al.
-[-@RN5497].
+*The class Galio-Urticetea is not included as it also includes native communities. However, the same is true for Bidentetea, and this class is included. Explain!*
 
-We have revised the usage of the term "ecoregion" throughout the manuscript, to avoid giving the false impression that we are specially interested in the ecoregion concept and classification by Olson et al.
-[-@RN4667], rather than in studying the Iberian Atlantic territories as a biogeographically-coherent unit.
+\textcolor{teal}{The class Galio-Urticetea is included in the manuscript.
+However, we are following the vegetation classification by Mucina et al.
+(https://doi.org/10.1111/avsc.12257) which includes Galio-Urticetea within the class Epilobietea angustifolii.
+We have modified the text (page 7, line 11-12) to make this clear.}
 
-> It seems that the plots used are original of NW Iberia, as shown in Figure 1.
-The boundaries of the grey area, supposedly taken from the map of Olson et al.
-2001, are quite coincident with what has been profusely named as Cantabro-Atlantic subprovince by Rivas-Martínez (Rivas-Martínez et al.
-2017) and related authors such as Fernández Prieto et al.
-2020, 2023.
-I would suggest to use these units and names to prevent from nomenclatural inflation and cite the corresponding papers (see below).
-If we comment the names, the name “Iberian Atlantic” is not so precise as “Cantabro-Atlantic”.
-The former could include southern Portugal (Algarve or Vicentine Coast), which is not at all the case.
-To use the term Cantabrian or Cantabrico is particularly precise to indicate the geographic position of this territory.
-Additionally, it would be necessary to make a more formal description of the “Iberian Atlantic ecoregion” and discuss the differences from this ecoregion with the “Cantabro-Atlantic subprovince”.
+*Hemicryptophytes are the most numerous group of species in the dataset.
+This fact is repeatedly mentioned in the text of results and discussion.
+However, hemicryptophytes are not shown in Fig.
+4.
+It is confusing and not logical.
+I would strongly suggest redrawing the figure and replacing the geophytes with hemicryptophytes (or merging them into one category of perennial herbs).
+/ p.
+12, l.
+20.
+I mentioned above, this is not good decision.
+It is shown that the hemicryptophytes are the most common life form!*
 
-We agree that the definition and name of the studied biogeographical unit requires further clarification First, we must make it clear how we define our study region: "the territories with a temperate climate in the north-western part of the Iberian Peninsula".
-In other words, we are following precisely the definition by Fernández Prieto et al.
-[-@RN5372] in *Naturalia Cantabricae* 8(2): 30-37.
-More specifically, we are following the geographical limits presented in Figure 1 of Fernández Prieto et al.
-[-@RN5372].
+\textcolor{teal}{We have done as requested, and included hemicryptophytes in Figure 4.
+We have kept geophytes as a separate category in Figure 4 because they are an important part of the definition of the alliance Allion triquetri.}
 
-The concept of this biogeographical unit as presented by Fernández Prieto et al.
-[-@RN5372] is essentially equivalent to the *Cantabrian Mixed Forests ecoregion* by Olson et al.
-[-@RN4667].
-The major difference between the two concepts is in the southern borders of the region, which are much more precisely drawn in Fernández Prieto et al.
-[-@RN5372].
+*A flow diagram (or table) showing the translation between alliances obtained by different classification schemes and EUNIS types would be helpful for the readers.*
 
-However, the biogeographical unit does not correspond with Rivas-Martínez's *Cantabro-Atlantic subprovince*, as there are two deviations from that concept.
-First, our study region also includes Rivas-Martínez's *Orocantabrian subprovince*.
-Second, Rivas-Martínez's *Cantabro-Atlantic subprovince* includes a larger portion of French territory.
-In the third (and, to our knowledge, last) edition of the Biogeographic map of Europe [@RN5592], the *Cantabro-Atlantic subprovince* extends in France from the French Basque country in the south to the Cotentin Peninsula in Normandy.
-In sum, the more precise correspondence between our study region and Rivas-Martínez's biogegraphical classification would be "the Iberian part of the *European Atlantic province*".
-We include this equivalence in our methods section, in the subsection "Biogeographical territory under study".
+\textcolor{teal}{The translation between alliances and EUNIS types is provided in Appendix S3, we have revised the text (page 13, line 8) to make this clearer.}
 
-Finally, we agree that, from a geographical point of view, "Iberian Atlantic" can refer to the Iberian coast down to Gibraltar, but the term "Atlantic" is used here in a bioclimatic sense, in accordance with the definition by Fernández Prieto et al.
-[-@RN5372] and the biogeographical unit names by Rivas-Martínez (cf.
-*European Atlantic province* in Rivas-Martínez et al.
-[-@RN5592]) and the European Union's Natura 2000 framework (cf.
-*Atlantic biogeographical region*, https://www.eea.europa.eu/en/analysis/maps-and-charts/biogeographical-regions-in-europe-2).
+*I am not comfortable with the use of Central European historical data for the Iberian peninsula, as described in the introduction.
+A lot of Central European weed species originate in South Europe, not only in the Near East, and the spread of archaeophytes was here probably earlier.
+It should be at least mentioned.
+(p.
+3, l.
+9-16) / p.
+12, l.
+12 Chytrý et al.
+2021?
+Is it an appropriate source for the Iberian peninsula?
+I doubt it.
+The delimitation of alien species in CZ could hardly work for other regions.
+Moreover, the authors of Catalogue of alien species in CZ is Pyšek et al.
+– the newest version was published in Preslia in 2022.*
 
-In any case, to avoid confusion, we have revised the manuscript's section "Biogeographical territory under study" to better explain the study area and its equivalence in different classifications.
-We have also revised the whole manuscript to avoid using the term *Iberian Atlantic ecoregion*, which mixes the concepts *Cantabrian Mixed Forests ecoregion* by Olson et al.
-[-@RN4667] and *Iberian Atlantic territories* by Fernández Prieto et al.
-[-@RN5372].
-Now, we use *Iberian Atlantic territories* in accordance with Fernández Prieto et al.
-[-@RN5372].
-We have also added the reference to Fernández Prieto et al.
-[-@RN5588] which further supports the definition of the studied biogeographical unit.
+\textcolor{teal}{Applying the concept of archaeophyte in southern Europe is problematic.
+However, we wanted to consider this category, solely for the purpose of comparison with similar studies in other European regions.
+Unfortunately, historical vegetation data is not as available in the Iberian Atlantic territories as it is in other parts of Europe, and this is the reason why we used the catalogues for the Czech Republic and Britain.
+We must clarify that we did not take those catalogues at face value.
+We used them to create a starting list of archaeophyte candidates.
+Then, we revised this list using Flora iberica, to remove species that are considered native in the Mediterranean part of the Iberian Peninsula.
+We classified the remaining species as putative archaeophytes in the study region, and only for the purposes of comparison.
+We have revised the text (page 11 line 19 to page 12 line 5) to better explain this, and to further stress that the classification must be interpreted with care (page 12, line 5-10).
+We have also updated the reference for the Czech catalogue.}
 
-> One of the main results of this survey is the selection of 28 alliances (appendix S4) from an initial set of 38 that could be present in the Iberian Atlantic ecoregion.
-So, I understand that there are 10 alliances have not been detected.
-It would be interesting to list them somewhere (in appendix S2, for instance) and make clear that no data from them have been found in the datasets.
+*Neophytes could originate from all the continents, not only from America, Africa, and Asia (p. 3, l. 19)*
 
-The missing 10 alliances are listed in the discussion, section *Vegetation and habitat classification of anthropogenic communities*.
-Some of these alliances were present in the dataset in the sense that the relevés had been assigned to syntaxa belonging to these alliances by the original authors of the relevé.
-However, the alliances were not supported by our numerical classification, i.e.
-their relevés were asigned to another alliance, or to the noise group.
+\textcolor{teal}{We have removed this sentence.}
 
-> Concerning the weakly represented alliances.
-This is inevitable if we start with a definition of the study area prior to surveying its plant communities.
-There will be always some types which are found in the extreme or border of its geographical area or ecological amplitude and are, thus, weakly represented.
-It is an inherent problem of such a territorial planning and has to be conveniently considered.
-Other possibility is to choose the vegetation alliances and study them in all their geographic amplitude.
-This would entail that their geographical territory would be diverse but the survey would bring more accurate results about their character and differences.
+*P5, l. 18, 19 the format of references is not in appropriate format*
 
-We agree that the risk exist, although we believe that this risk is lessened by focusing on a biogeographical unit rather than in e.g.
-a political division.
-Furthermore, there were alliances that are only found in the extremes (geographical or ecological) of the study area, but still they were supported by the numerical classification.
-We believe that the fact that the missing alliances were not supported by the numerical classification indicates that the relevés that had been assigned to these alliances are poor representatives of the vegetation types, and more likely transitional communities at the extreme of the alliance distributions.
+\textcolor{teal}{We are using a reference manager with AVS's official style template.
+After some time trying, we have not been able to fix these specific references.
+We will fix them manually if the manuscript is accepted.}
 
-> In S2 author’s names should be corrected and homogenized e.g.
-Lence Paz is C.
-or Carmen, all the patronimic names such as Pérez, Fernández, etc,, have an accent but have instead a strange signal in the involved letter.
-It is Rivas-Martínez, not Rivas Martínez; Is it Romero or Romero Buján?, etc.
+*p. 6, l. 11-12 the mix of two different taxonomic concepts is not the best option*
 
-In Appendix S2, author names are provided exactly as they are stored in the SIVIM database. This is necessary to ensure cross-referencing between our dataset and the SIVIM database.
+\textcolor{teal}{We agree in principle, however, in this case the taxa revised with POWO were a few non-European introduced taxa, therefore, this has not affected in any way the classification and the results of the article.}
 
-> In the figure 1 the area does not include any French terrotory [...] In the maps of figure 1 there is no French territory included
+*p. 10, l. 9-16. this part of the text belongs to the results and not to the methodological part*
 
-The figure includes a portion of French territory in the Pyrénées-Atlantiques department, following the biogeographical borders shown by Fernández Prieto et al.
-[-@RN5372] in their Figure 1, although our figure may be to small to fully appreciate this.
-We have revised the mention in the abstract to make clear that we are speaking about SW France.
+\textcolor{teal}{We have moved the text to results as requested (page 13 line 10-17).}
 
-> Anthropogenic syntaxa have a larger biogeographical scope than other vegetation types, particularly grasslands and scrub, but they have  narrower distribution than others such as aquatic vegetation.
-On the other hand, they have been described by traditional phytsociology after fine ecological variations.
-If there is inflation it should be somhow documented.
+*p. 14, l. 8. It would be helpful to explain in the methods which taxa are those problematic ones that are merged to some aggregates.*
 
-The inflation is reflected in the 78 associations which had been described in the original relevés as stored in SIVIM (see "Results - Overview of the classification").
-It was not our goal to classify to the level of these associations, but our exploratory analysis indicated that they were not supported by numerical classification (i.e.
-TWINSPAN within the alliances did not result in splits along association lines, and many locally-described associations were only represented by a low number of relevés).
+\textcolor{teal}{Thank you for spotting this.
+The vegetation dataset that we use in this manuscript is part of a wider vegetation dataset for the whole Iberian Atlantic region (including all vegetation types, not just synanthropic).
+When we cleaned and curated the whole dataset, we did create some taxa aggregates.
+However, we have checked the plots and the taxa included in this manuscript and, for this subset, each aggregate can only correspond to one plant species.
+Therefore, all mentions to "taxa aggregates" can be changed to "species".
+We have done so in the text.}
 
-> There are several papers on this topic published in the area by Loidi and other authors, they are quoted in the adjointed sheet
+*Do all of the plots include bryophytes? It is not described in the methods. Bryophytes are highly rare in the annual vegetation of disturbed sites.*
 
-Thank you for these references, we have incorporated them into our manuscript.
+\textcolor{teal}{The SIVIM database does not provide consistent information on whether bryophytes had been recorded or not, but it is very likely that they were not in many cases.
+We have revised the manuscript to include this information (page 16, line 3).
+Bryophytes are rare in most synanthropic vegetation types, but they were relatively frequent and abundant in the vegetation of mesic trampled sites, i.e.
+Saginion procumbentis with Bryum argenteum.}
 
-> classification at the alliance level ..
+*p.
+14, l.
+12-13 The comparison of species numbers per plot of different sizes is inappropriate.
+The authors minimize the variance of the plot sizes, but still, could we be sure that the low number of species is not a result of lower plot sizes?*
 
-We have updated the text as suggested.
+\textcolor{teal}{We have revised the calculation to use only those plots that had the same and most frequent size value (= 20 m2) and recalculated the numbers of species accordingly.}
 
-> In this paper only forest and mantle communities are dealt, no nitrophilous vegetation.
+*References.
+Please fix the correct form of some authors' names in the references.
+/ Němec, R., Lososová, Z., Dřevojan, P & Žáková, K.
+/ Šilc U.
+/ Šilc et al.
+– check all the names / Simonová D.
+/ Tabasevic et al.
+– check / Zaliberova – check.*
 
-We cite this reference as the basis for defining the borders of our study biogeographical territory, we have revised the text to make this clear.
+\textcolor{teal}{Thank you very much for revising the spelling of the names. We have corrected them.}
 
-> In these disturbed communities, geophytes, with some exceptions, use to be  scarce.
-Do you think neccesary to highlight their proportion more than other life-forms?
-It seems to me that the main point is to describe the relative importance of hemicriptophytes in relation to therophytes as they become dominant in their respective alliances.
+*Appendix S5.
+Please add a header to the table with a short description and explanation of what the numbers in the table show.
+Define the phi value for diagnostic taxa and explain the meaning of the different colors used.
+Please note that the phi value is a number between 0 and 1.
+The table probably does not include the complete list of species.
+Some explanation is needed.*
 
-We agree that, in most of the alliances, only the variation of hemicriptophytes and therophytes is relevant.
-However, the presence of geophytes is an important part of the description of some of the alliances (e.g.
-see the description of *Allion triquetri* by Mucina et al.
-[-@RN4698]: *Mesic nitrophilous geophyte-rich fringe vegetation of the Western Mediterranean*).
-Furthermore, the variation in geophytes explained a significant part of the overall variance in a exploratory PCA.
-For these two reasons, we decided to highlight the propotion of geophytes in our manuscript.
+\textcolor{teal}{We have added a header explaining all the requested points.}
 
-> This figure is difficult to read.
-Point clouds overlap and the information is partially not visible.
-I suggest to use two figures for the same PCA, each one representing a half of the alliances selected in form that overlpings are avoided.
-This could also allow to use larger dots which would be more easily identified by the color.
+# Referee: 2
 
-Thank you for your suggestions to improve Figure 7. We have provided a new version of the figure, where each vegetation class is represented separately.
+*I couldn't find a Response Letter.
+If the authors haven't submitted one, I would ask them to do so.
+However, from the track-changes file I see that the authors addressed most of my previous points.
+The only remaining issue is the synoptic table: The authors included such a table as Appendix S5, but it only shows the phi values.
+In my opinion, a synoptic table should show the percentage frequency, and phi values reaching a certain threshold should be indicated by shading.*
 
-> It should be cited as: Castroviejo, S. (coord. gen.). 1986-2020. Flora iberica. Real Jardín Botánico, CSIC, Madrid [and other comments regarding the references]
-
-Thank you for checking our references, we have corrected them as needed.
-
-# Referee 2
-
-> This paper presents a classification and characterisation of the anthropogenic vegetation alliances of the Iberian Atlantic ecoregion in Europe.
-The study area is not very large, but I think that the paper is a valuable contribution to the knowledge of this vegetation type in Europe, and it can be considered suitable for AVS after some revision.
-
-Thank you for your positive evaluation and suggestions.
-
-> The authors claim that (1) there is a need for a numerical classification and synthesis of the full spectrum of anthropogenic vegetation, and (2) an appropriate scale to conduct such study is the ecological region.
-While I strongly agree with point 1, I'm less convinced about point 2.
-Why should a revision at the ecoregion level be superior to a continental pan-European synthesis?
-I think, both is necessary, and the authors should acknowledge this fact.
-
-We fully agree that a pan-European synthesis would be worthwhile goal, and never intended to suggest the contrary.
-In fact, we would be very happy of any research efforts on continental studies of anthropogenic vegetation.
-We have added a new final sentence to our conclusions: *We hope that our synthesis of the anthropogenic communities of the Iberian Atlantic territories serves as a contribution towards a pan-European synthesis of human-made vegetation, a synthesis that could shed light on the coexistence between humans and plants.* We also added a similar closing statement to the abstract.
-We have also modified our text to avoid giving the impression that we are suggesting that the ecoregion (or a similarly ranked biogeographical unit) is the *only* appropriate scale at which one can produce a synthesis of anthropogenic vegetation.
-
-> The title is a bit too general. Please consider to add a subtitle such as "a case-study from SW Europe", or something similar.
-
-We agree and have changed the title to *Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*.
-
-> The merits of the applied two-step approach (first a Twinspan classification, then a "semi-supervised" noise clustering) remain unclear to me.
-More details are needed to fully understand what the authors have done.
-How many Twinspan clusters did you choose, and how did you determine the optimal number of clusters?
-How did you match the individual Twinspan clusters with EVC alliances?
-A crosstab showing the Twinspan clusters and the original alliance assignment should be presented.
-In the noise clustering, no additional clusters were detected, but plots were allowed to be re-assigned, and a considerable part of the data (17%) remained unclassified in a "noise group".
-Still, the "semi-supervised" classification is a numerical cluster algorithm.
-In my opinion, a pure unsupervised classification with subsequent interpretation of the result would have been more straightforward.
-Re-assignment of plots should be done on the basis of well-defined diagnostic species or using formal definitions, not by applying two numerical algorithms one after another.
-Or did I miss something important?**
-
-The main aim of our classification was for it to be consistent with regional phytosociological expertise, and to match that expertise with current EuroVegChecklist alliances [@RN4698], not to create a new classification.
-To achieve this, one advantage of using the SIVIM database is that a majority of the plots had been assigned to associations or alliances by the original authors [@RN5374].
-However, there could be misclassification due to unclear concepts or human error.
-Therefore, we used TWINSPAN to revise the classification by the original authors, to define the number of clusters that better aligned with our syntaxonomical checklist, and to remove outliers or misclassified plots.
-The obtained clusters therefore provided an agreement between phytosociological and numerical classification, and contained only those plots that better represented each alliance.
-These clusters were then used to assign other plots to the most similar cluster.
-We think semi-supervised classification is an optimal procedure to reach an agreement between existing and revised classifications (following Tichý et al.
-[-@RN5593], the aim is "preserving the good old units and searching for new ones").
-We found that this important point was not clearly explained in the text or the abstract, so we revised the text accordingly.
-We have also provided more details on the TWINSPAN procedure as requested by the reviewer.
-
-> The authors conclude that ten alliances previously reported from the study area do not occur there.
-This seems a bit hasty.
-Maybe, these alliances were only represented by a few plots, so they were not detected by the numerical methods.
-
-We note that some alliances were supported by the methods, despite being extreme (ecologically or geographically) for the study region and/or being represented by a few plots.
-Therefore, we believe that the fact that the missing alliances were not supported by the numerical classification supports that the relevés that had been assigned to these alliances are poor representatives of the vegetation types, and more likely transitional communities at the extreme of the alliance distributions.
-Nonetheless, we have modified the text to be less conclusive about their absence, by adding the following sentence to the discussion: *Of course, the fact that these ten alliances are not supported by our numerical methods only indicates that they are poorly represented in our dataset, and should not be taken as absolute proof that they are absent from the Iberian Atlantic territories*.
-
-> Most importantly, a synoptic table of the alliances is missing.
-I consider a table showing the constancy of all species in the 28 alliance, with the diagnostic species highlighted, an absolute prerequisite for the paper to be acceptable.
-A shorted table in the main text would be nice, but there should be a least the full table as supplementary information.
-
-**Pendiente de hacer (Víctor?)**
-
-> p5, l11: "non-overlapping definitions" ... different, but certainly overlapping definitions!
-
-You are right, of course. We have change it to *not fully equivalent*.
-
-> p9, l15: Why did you use 15% as cutlevel, and not 5%?
-The latter would correspond to the border between Braun-Blanquet class 1 and 2, while 15 cuts the Braun-Blanquet class 2 in the middle.
-Which cutlevel then corresponds to Braun-Blanquet class 2?**
-
-Our transformation of Braun-Blanquet values follows the standard of assigning intermediate cover values.
-Thus, the BrBl value 2 (representing 5-25% cover) is transformed to 15% and the value 3 (25-50%) is transformed to 37.5%.
-Although we could use more approximate pseudospecies levels, the results of TWINSPAN will be the same when the transformation values are selected within the original cover ranges.
-
-> p10, l5: Please clarify if the same Twinspan settings were used as before.
-
-Yes, and we have added this information to the manuscript.
-
-> p20, l10: "which are mostly differentiated by dominant vegetation height" ... Is this a new result of your study, or has this been mentioned in the literature (if so, please provide a reference)?
-
-Yes, this is a traditional part of the alliance descriptions, and we have added this to our text with a reference.
-
-> p21, l12: "neophytes tended to be in higher proportions in vegetation types associated to wet conditions (Bidentetea)" ...
-I would have expected an even higher portion of neophytes in the Senecionion fluviatilis.
-Maybe, plots of alluvial forb communities dominated by neophytes are underrepresented in your dataset?
-
-Our *Senecionion fluviatilis* plots are certainly dominated by alien species, especifically by *Arundo donax*, which has a mean cover of 74%.
-However, *Arundo donax* is accompanied by many native species (mean = 12 native species per plot) and so the plot-level proportion of neophytes seems low (as it is just the propotion, not weighted by species cover).
-That being said, it might be possible that plots with neophytes are generally underrepresented across the dataset, because of a sampling bias towards "natural" communities.
-But this underrepresentation is not necessarily higher in *Senecionion fluviatilis* vs.
-other alliances.
-
-> References: In several references the page numbers are missing!
-
-We have revised the references.
-
-> Appendix S3: What is the value for dominant species? Constancy? Mean cover value? Something else? Please clarify.
-
-We have added information on how we calculated the values for diagnostic, constant and dominant species.
-
-# References
-
-::: {#refs}
-:::
+\textcolor{teal}{We are sorry that the reviewer couldn't access the response letter. We have revised the synoptic table following the reviewer's specifications.}

### 2024-11-15 doc/manuscript.Rmd
AVS R2

@@
 * *Results* We assigned 2,086 vegetation plots to 25 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*).
-The plots included 1,149 taxa: 78% natives, 15% archaeophytes and 7% neophytes.
+The plots included 1,149 species: 78% natives, 15% archaeophytes and 7% neophytes.
 Vegetation groups were organized along a principal axis of abiotic stress (dry-sunny to moist-shady habitats) and a secondary axis of disturbance.
@@
 Some of these archaeophytes are 'obligate weeds' that do not occur in non-anthropogenic habitats and are assumed to have evolved together with human agriculture [@RN5494].
-The synanthropic flora of Europe has been further enriched by neophytes, alien taxa that arrived after the expansion of intercontinental trade in the Modern Age (after 1.500 CE) [@RN5439].
-Synanthropic neophytes therefore include many American, African and Asian taxa that arrived in Europe because of the global trade networks established by western European colonialism [@RN5451].
+The synanthropic flora of Europe has been further enriched by neophytes, alien taxa that arrived after the expansion of intercontinental trade in the Modern Age (after 1.500 CE) [@RN5439; @RN5451].
 In the last century, the European anthropogenic vegetation has shown a decrease in species richness and diversity [@RN5423], as rare synanthropic species have become rarer and neophytes have increased their abundance at the expense of natives and archaeophytes [@RN5441].
@@
 To circumscribe our study vegetation, we followed the definition of anthropogenic vegetation in the revised classification of the vegetation of Europe by Mucina *et al.* [-@RN4698].
-In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*.
+In our study area, this potentially includes the vegetation classes *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii* (which includes *Galio-Urticetea*).
 For the sake of completeness, we also considered the class *Cymbalario-Parietarietea diffusae* included by Mucina *et al.* [-@RN4698] in the vegetation of rock crevices and screes, since this class encompasses the vegetation of human-made walls.
@@
 Therefore, we used TWINSPAN to revise the classification by the original authors, to define the number of clusters that better aligned with our syntaxonomical checklist and to remove outliers or misclassified plots.
-
 We started this stage of the analyses by keeping only those plots (n = 2,095) that had been assigned by the original authors of the plot to any of the 9 anthropogenic vegetation classes defined in our syntaxonomical checklist (i.e., we removed the plots without syntaxa, that had been retrieved solely by the expert system).
@@
 the plots that had been assigned by the original authors to that alliance were separated in different clusters).
-By the end of this process, of the 34 alliances in our checklist, 25 had been assigned to clusters.
-Nine alliances were not assigned to any cluster.
-Of these, eight had been assigned to a few plots by the original authors: *Onopordion castellani* (n = 6 plots), *Onopordion acanthii* (5), *Euphorbion prostratae* (4), *Caucalidion lappulae* (4), *Roemerion hybridae* (2), *Hordeion murini* (2), *Taeniathero-Aegilopion geniculatae* (1) and *Fragarion vescae* (1).
-The alliance *Aegopodion podagrariae* had been assigned by the original authors to 39 plots, but our TWINSPAN analysis could not separate this alliance from the alliances *Geo urbani-Alliarion officinalis* and *Balloto-Conion maculati*.
-
 This stage of the analyses allowed us to identify those plots in which there was an agreement between (1) our TWINSPAN-based classification (i.e.
@@
 
-Next, we conducted a semi-supervised classification [@RN5593] of the whole dataset (n = 2,508 plots) into the 25 TWINSPAN-validated anthropogenic alliances.
+Next, we conducted a semi-supervised classification [@RN5593] of the whole dataset (n = 2,508 plots) into 25 TWINSPAN-validated anthropogenic alliances.
 Semi-supervised classification uses a training subset of *a priori* classified vegetation plots to classify a secondary subset of unclassified plots [@RN5383].
@@
 
-We classified the species as native, archaeophytes or neophytes in the Iberian Atlantic territories using the information in *Flora iberica* [-@RN3243] and catalogues of archaeophytes for Britain [@RN5490] and the Czech Republic [@RN5495].
-We must stress that identifying archaeophytes in southern Europe is highly problematic [@RN5444] and our classification must be taken as an indication of putative archaeophyte character for the purposes of vegetation description and comparison with other European regions, rather than a definitive classification of the species; as stated by Preston *et al.* [@RN5490], the classification of a species as an archaeophyte should be interpreted as a hypothesis to be tested by further studies.
+We classified the species as native, archaeophytes or neophytes in the Iberian Atlantic territories.
+Information for archaeophyte status in the Iberian Peninsula is scarce and inconsistent.
+Therefore, we used the catalogues of archaeophytes for Britain [@RN5490] and the Czech Republic [@RN5626] to create a starting list of species that are considered archaeophytes in those countries.
+Then, we used *Flora iberica* [-@RN3243] to remove from that list those species that are considered native in the Mediterranean part of the Iberian Peninsula.
+We classified the remaining species as archaeophytes in the Iberian Atlantic territories.
+Nonetheless, we must stress that identifying archaeophytes in southern Europe is highly problematic [@RN5444] and our classification must be taken as an indication of putative archaeophyte character, solely for the purposes of vegetation comparison with other European regions, rather than a definitive classification of the species; as stated by Preston *et al.* [@RN5490], the classification of a species as an archaeophyte should be interpreted as a hypothesis to be tested by further studies.
 
-## Species life form
+## Species life form, height and flowering phenology
 
-Using *FloraVeg.EU* [-@RN5385] we extracted the species’ life forms and kept *therophyte* and *geophyte* for further analysis since these two categories had the largest contribution to variance among alliances, as per an exploratory Principal Component Analysis (PCA) performed using the R package *FactoMineR* [@RN3166].
-
-## Species height and flowering phenology
-
+We extracted the species’ life forms from *FloraVeg.EU* [-@RN5385].
 From *Flora iberica* [-@RN3243] we extracted the maximum height, median month of flowering and length of the flowering period of each species in the dataset.
@@
 The semi-supervised classification resulted in the classification of 2,086 vegetation plots into 25 anthropogenic alliances (**Fig.
-1**, description of the alliances and correspondence with EUNIS habitat types in **Appendix S3**, groups of characteristic species in **Appendix S4**, synoptic table in **Appendix S5**) representing 9 vegetation classes.
-The regional literature included in our dataset recognized 69 associations (**Appendix S6**) within these alliances, but exploratory analysis (not shown) indicated that the separation of these associations within the alliances had a generally weak support based on numerical classification methods.
+1**, description of the alliances and translation into EUNIS habitat types in **Appendix S3**, groups of characteristic species in **Appendix S4**, synoptic table in **Appendix S5**) representing 9 vegetation classes.
+Of the 34 alliances in our initial syntaxonomical checklist, nine alliances were not assigned to any cluster.
+Of these, eight had been assigned to a few plots by the original authors: *Onopordion castellani* (n = 6 plots), *Onopordion acanthii* (5), *Euphorbion prostratae* (4), *Caucalidion lappulae* (4), *Roemerion hybridae* (2), *Hordeion murini* (2), *Taeniathero-Aegilopion geniculatae* (1) and *Fragarion vescae* (1).
+The alliance *Aegopodion podagrariae* had been assigned by the original authors to 39 plots, but our TWINSPAN analysis could not separate this alliance from the alliances *Geo urbani-Alliarion officinalis* and *Balloto-Conion maculati*.
+The regional literature included in our dataset recognized 69 associations (**Appendix S6**) within the final 25 alliances, but exploratory analysis (not shown) indicated that the separation of these associations within the alliances had a generally weak support based on numerical classification methods.
 
-The classified plots included 1,149 taxa or taxa aggregates.
+The classified plots included 1,149 species.
 The 10 most frequent species were *Ochlopoa annua* (625 occurrences), *Urtica dioica* (588), *Sonchus oleraceus* (548), *Stellaria media* (520), *Capsella bursa-pastoris* (396), *Polygonum aviculare* (373), *Dactylis glomerata* (369), *Senecio vulgaris* (336) and *Anisantha sterilis* (310).
-Considering only those plots with the most frequent plot size range (10-30 m^2^, n = 867 plots), the average species richness per plot was 16 (minimum = 3, maximum = 48).
+Considering only those plots with the most frequent plot size range (20 m^2^, n = 260 plots), the average species richness per plot was 16 (minimum = 3, maximum = 37).
 The class with the richest species pool was *Epilobietea angustifolii* (n = 605) and the poorest was *Bidentetea* (n = 132).
@@
 
-The most frequent life form in the species pool (n = 1,149 species) were hemicryptophytes (44%), followed by therophytes (41%), geophytes (8%), chamaephytes (6%), phanerophytes (5%), bryophytes (1%) and hydrophytes (1%).
+The most frequent life form in the species pool (n = 1,149 species) were hemicryptophytes (44%), followed by therophytes (41%), geophytes (8%), chamaephytes (6%), phanerophytes (5%), bryophytes (1%, not consistently recorded in all plots) and hydrophytes (1%).
 The proportion of therophytes across classes and alliances (**Fig.
@@
 
-The anthropogenic vegetation of the Iberian Atlantic territories is home to more than a thousand plant taxa.
+The anthropogenic vegetation of the Iberian Atlantic territories is home to more than a thousand plant species.
 This is approximately one third of the whole regional species pool represented in the SIVIM database [@RN5374] and one fifth of the Iberian flora [@RN5503].
 More than half of this synanthropic diversity (c.
-600 taxa) occurs in mesic to wet perennial ruderal vegetation, highlighting the potential of some anthropogenic habitats as a source of biodiversity in human-dominated landscapes [@RN5446; @RN5445].
+600 species) occurs in mesic to wet perennial ruderal vegetation, highlighting the potential of some anthropogenic habitats as a source of biodiversity in human-dominated landscapes [@RN5446; @RN5445].
 These results suggest three main objectives for the sustainable management of human-made habitats: (1) focusing efforts on alien plant management in anthropogenic plant communities of wet habitats; (2) conserving examples of archaeophyte-rich annual weed communities, especially those linked to traditional agricultural practices [@RN5502]; and (3) conserving and promoting mesic ruderal and fringe vegetation as a biodiversity refuge in urban and peri-urban areas.
@@
 
-**Appendix S5** Synoptic table of the fidelity (phi) of the diagnostic species of the 25 anthropogenic alliances.
+**Appendix S5** Synoptic table of the anthropogenic communities found in NW Iberia.
 
@@
 
-```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of therophytes, geophytes and other life forms (hemicryptophytes, chamaephytes, phanerophytes, bryophytes, hydrophytes) in the anthropogenic plant communities of the Iberian Atlantic territories.
-Each bar represents the plot-level average proportion of therophytes (dark shade), geophytes (medium shade) and other life forms (light shade) in each vegetation alliance.
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of therophytes, hemicryptophytes, geophytes and other life forms (chamaephytes, phanerophytes, bryophytes, hydrophytes) in the anthropogenic plant communities of the Iberian Atlantic territories.
+Each bar represents the plot-level average proportion of therophytes (dark shade),hemicryptophytes (medium-dark shade), geophytes (medium-light shade) and other life forms (light shade) in each vegetation alliance.
 Colors indicate the vegetation class."}

### 2024-11-15 src/11-lifeforms.R

@@ -12,6 +12,7 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
   filter(Value == 1) %>%
   mutate(origin = ifelse(Trait == "therophyte", "Therophyte", "Other")) %>%
   mutate(origin = ifelse(Trait == "geophyte", "Geophyte", origin)) %>%
+    mutate(origin = ifelse(Trait == "hemicryptophyte", "Hemicryptophyte", origin)) %>%
   merge(read.csv("data/species.csv", fileEncoding = "latin1"), all.y = TRUE) %>%
   group_by(SIVIMID, origin) %>%
   summarise(y = length(Analysis.Names)) %>%
@@ -26,6 +27,7 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
   mutate(origin = fct_relevel(origin, 
                               "Other",
                               "Geophyte",
+                              "Hemicryptophyte",
                               "Therophyte")) %>%
   mutate(origin = fct_recode(origin, 
                              "Natives" = "native",
@@ -93,9 +95,9 @@ read.csv("data/traits.csv", fileEncoding = "latin1") %>%
                                "darkmagenta", 
                                "limegreen")) +
   ggthemes::theme_tufte() +
-  xlab("Alliance") + ylab("Proportion of therophytes, geophytes and other life forms at the plot level") +
+  xlab("Alliance") + ylab("Proportion of therophytes, hemicryptophytes, geophytes and other life forms at the plot level") +
   coord_flip() +
-  scale_alpha_manual(name = "Species origin",values = c(.35, .65, 1), guide = "none") +
+  scale_alpha_manual(name = "Species origin",values = c(0.1, .35, .65, 1), guide = "none") +
   theme(text = element_text(family = "sans"),
         strip.background = element_blank(),
         legend.position = "right", 

### 2024-11-15 src/15-aggregates.R

@@ -1,0 +1,16 @@
+library(tidyverse)
+
+read.csv("data/species.csv", fileEncoding = "latin1") %>%
+  dplyr::filter(grepl("aggr.", Analysis.Names)) %>%
+  select(Analysis.Names) %>%
+  unique -> aggr
+
+read.csv("../canteunis/results/SIVIM/species2.3.csv", fileEncoding = "latin1") %>%
+  select(Analysis.Names, Expert.Revision) %>%
+  unique %>%
+  merge(aggr, by = "Analysis.Names") %>%
+  mutate(Expert.Revision = gsub(" aggr.", "", Expert.Revision)) %>%
+  unique %>%
+  group_by(Analysis.Names) %>%
+  filter(length(Expert.Revision) > 1) %>%
+  data.frame

### 2024-11-15 src/7-spp-numbers.R

@@ -24,9 +24,10 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
   group_by(Class) %>% tally %>% arrange(-n)
 
 read.csv("data/header.csv", fileEncoding = "latin1") %>% 
-  filter(Alliance != "Noise") %>% 
-  filter(Area >= 10) %>%
-  filter(Area <= 30) %>%
+  filter(Alliance != "Noise") %>%
+  # filter(Area >= 10) %>%
+  # filter(Area <= 30) %>%
+  filter(Area == 20) %>%
   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
   group_by(SIVIMID) %>%
   summarise(l = length(Analysis.Names)) %>%
@@ -34,8 +35,9 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
 
 read.csv("data/header.csv", fileEncoding = "latin1") %>% 
   filter(Alliance != "Noise") %>% 
-  filter(Area >= 10) %>%
-  filter(Area <= 30) %>%
+  # filter(Area >= 10) %>%
+  # filter(Area <= 30) %>%
+  filter(Area == 20) %>%
   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
   group_by(SIVIMID) %>%
   summarise(l = length(Analysis.Names)) %>%
@@ -43,8 +45,9 @@ read.csv("data/header.csv", fileEncoding = "latin1") %>%
 
 read.csv("data/header.csv", fileEncoding = "latin1") %>% 
   filter(Alliance != "Noise") %>% 
-  filter(Area >= 10) %>%
-  filter(Area <= 30) %>%
+  # filter(Area >= 10) %>%
+  # filter(Area <= 30) %>%
+  filter(Area == 20) %>%
   merge(read.csv("data/species.csv", fileEncoding = "latin1")) %>%
   group_by(SIVIMID) %>%
   summarise(l = length(Analysis.Names)) %>%

### 2024-11-27 README.Rmd
Zenodo link ⛓️
updated README and GitHub manuscript 📖

--- 2024-04-26 README.Rmd
@@
 
-[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.6077618.svg)](https://doi.org/10.5281/zenodo.6077618)
+[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.11613571.svg)](https://zenodo.org/doi/10.5281/zenodo.11613571)
 
-![Antigua fábrica de La Cerámica Gijonesa](ceramica.JPG)
+![](ceramica.JPG)
 
-# Anthropogenic plant communities of the Iberian Atlantic ecoregion
+# Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula
 
-This repository stores all the information related to the manuscript [*Anthropogenic plant communities of the Iberian Atlantic ecoregion*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), including the raw datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
+This repository stores all the information related to the manuscript [*Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), including the datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
 
@@
 
-* *Questions* Which is the diversity of anthropogenic plant communities at the ecoregional level in a transitional territory between temperate and Mediterranean Europe?
-How are these communities composed by native plants, archaeophytes and neophytes?
-What are the main environmental and disturbance drivers of the diversity in anthropogenic vegetation?
+* *Questions* Can we reconcile regional and European classifications of anthropogenic plant communities at the biogeographical scale?
+How are these communities characterized by species origins, traits and ecological preferences?
 
-* *Location* Iberian Atlantic ecoregion, Iberian Peninsula, south-western Europe.
+* *Location* Atlantic territories in the NW Iberian Peninsula (a.k.a. Cantabrian Mixed Forests ecoregion); south-western Europe.
 
-* *Methods* We compiled a vegetation database of 2,508 synanthropic plots and revised the regional checklist of anthropogenic alliances.
-We used modified TWINSPAN and semi-supervised classification to classify the plots into the revised alliances.
+* *Methods* We classified 2,508 plots with the aim of being consistent with regional phytosociological expertise, while matching that expertise with current EuroVegChecklist alliances.
+We used modified TWINSPAN to revise the original phytosociological classification, followed by semi-supervised re-classification of the whole dataset.
 We determined the proportion of natives, archaeophytes and neophytes.
-We described the alliances in terms of vegetation height, flowering phenology and ecological requirements for temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.
+We also described the alliances in terms of species traits (lifeforms, height and flowering phenology) and ecological requirements (temperature, moisture, light, nutrients, soil reaction, disturbance frequency and severity).
 
-* *Results* We classified 2,081 vegetation plots into 28 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*).
-The plots included 1,162 plant taxa: 78% natives, 15% archaeophytes and 7% neophytes.
-The main ecological gradients behind this diversity went from dry-sunny to moist-shady habitats, followed by a gradient of low to high disturbance frequency and severity.
+* *Results* We assigned 2,086 vegetation plots to 25 anthropogenic alliances representing 9 vegetation classes (*Cymbalario-Parietarietea diffusae*, *Polygono-Poetea annuae*, *Papaveretea rhoeadis*, *Digitario sanguinalis-Eragrostietea minoris*, *Chenopodietea*, *Sisymbrietea*, *Bidentetea*, *Artemisietea vulgaris* and *Epilobietea angustifolii*).
+The plots included 1,149 species: 78% natives, 15% archaeophytes and 7% neophytes.
+Vegetation groups were organized along a principal axis of abiotic stress (dry-sunny to moist-shady habitats) and a secondary axis of disturbance.
 
-* *Conclusions* The human-made habitats of the Iberian Atlantic ecoregion host one third of the ecoregional species pool and one fifth of the Iberian flora.
-Wet anthropogenic habitats are most affected by neophyte invasions.
-Annual vegetation of arable lands is a refuge for archaeophytes.
-Mesic perennial ruderal vegetation is specially rich in native species and can be a biodiversity asset in urban and peri-urban settings.
+* *Conclusions* In the Iberian Atlantic territories, anthropogenic habitats host one third of the regional plant species pool and one fifth of the Iberian flora.
+Mesic perennial ruderal vegetation is especially rich in native species and can be a biodiversity asset in urban landscapes.
+Our biogeographical-level synthesis can improve the management of anthropogenic plant communities and contribute towards a European-level synthesis of human-made vegetation.
 
@@
 > Fernández-Pascual E, González-García, Ivesdal G, Lázaro-Lobo A, Jiménez-Alfaro B.
-[*Anthropogenic plant communities of the Iberian Atlantic ecoregion*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), *in preparation*.
+[*Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), *in preparation*.

### 2025-01-14 README.Rmd
published in issue

@@
 
-This repository stores all the information related to the manuscript [*Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), including the datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
+This repository stores all the information related to the manuscript [*Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*](https://doi.org/10.1111/avsc.70010), including the datasets, the scripts to perform data cleaning and analysis, and the `Rmarkdown` files to create the manuscript.
 
@@
 > Fernández-Pascual E, González-García, Ivesdal G, Lázaro-Lobo A, Jiménez-Alfaro B.
-[*Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*](https://github.com/efernandezpascual/manmade/blob/master/doc/manuscript.md), *in preparation*.
+[*Classification and characterization of anthropogenic plant communities in the northwestern Iberian Peninsula*](https://doi.org/10.1111/avsc.70010).
+*Applied Vegetation Science* 28, e70010.

### 2025-01-14 doc/manuscript.Rmd
published in issue

@@
 ```
+
+Eduardo Fernández-Pascual^1,a^ (<https://orcid.org/0000-0002-4743-9577>), Víctor González-García^1^ (<https://orcid.org/0000-0002-8949-7943>), Greta Ivesdal^2^, Adrián Lázaro-Lobo^1^ (<https://orcid.org/0000-0002-0509-2085>), Borja Jiménez-Alfaro^1^ (<https://orcid.org/0000-0001-6601-9597>)
+
+^1^ Biodiversity Research Institute (IMIB), University of Oviedo - CSIC - Principality of Asturias, Mieres, Spain; ^2^ Jardín Botánico Atlántico, Gijón/Xixón, Spain
+
+Correspondence: Eduardo Fernández-Pascual, Instituto Mixto de Investigación en Biodiversidad, Campus de Mieres, Edificio de Investigación, 5ª planta, c/ Gonzalo Gutiérrez Quirós s/n, E-33600 Mieres, Spain.
+Email: [fernandezpeduardo\@uniovi.es](mailto:fernandezpeduardo@uniovi.es){.email}.
+Telephone: +34985104781.
 
@@
 
+```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Anthropogenic vegetation alliances of the Iberian Atlantic territories. Each dot is a vegetation plot. Dot colors indicate the vegetation class."}
+knitr::include_graphics("../results/figures/F1 - maps.png")
+```
+
 The classified plots included 1,149 species.
@@
 Separate PCAs within the major vegetation groups indicated a coherent separation between most of the alliances, but with some cases of relatively high overlap, especially between *Galio valantiae-Parietarion judaicae* and *Cymbalario-Asplenion*.
+
+```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Floristic diversity of the anthropogenic plant communities of the Iberian Atlantic territories.
+Biplots produced by Principal Component Analysis (PCA) with Hellinger transformation.
+PCAs were conducted for the whole vegetation (A) and for specific subsets (B-F).
+Colors indicate the vegetation class."}
+knitr::include_graphics("../results/figures/F2 - pca.png")
+```
 
@@
 
+```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of neophytes, archaeophytes and native species in the anthropogenic plant communities of the Iberian Atlantic territories.
+Each bar represents the plot-level average proportion of neophytes (dark shade), archaeophytes (medium shade) and natives (light shade) in each vegetation alliance.
+Colors indicate the vegetation class."}
+knitr::include_graphics("../results/figures/F3 - origins.png")
+```
+
 ## Community-level traits
@@
 
+```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of therophytes, hemicryptophytes, geophytes and other life forms (chamaephytes, phanerophytes, bryophytes, hydrophytes) in the anthropogenic plant communities of the Iberian Atlantic territories.
+Each bar represents the plot-level average proportion of therophytes (dark shade),hemicryptophytes (medium-dark shade), geophytes (medium-light shade) and other life forms (light shade) in each vegetation alliance.
+Colors indicate the vegetation class."}
+knitr::include_graphics("../results/figures/F4 - lifeforms.png")
+```
+
 Community-weighted means for plant height (**Fig.
@@
 5C**) was generally high, from 5 months (*Artemisietea vulgaris*, *Epilobietea angustifolii*) to 6 (*Chenopodietea*, *Sisymbrietea*), 7 (*Bidentetea*, *Digitario sanguinalis-Eragrostietea minoris*, *Papaveretea rhoeadis*), 8 (*Polygono-Poetea annuae*) or 9 (*Cymbalario-Parietarietea diffusae*).
+
+```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Vegetation height and flowering phenology of the anthropogenic plant communities of the Iberian Atlantic territories.
+Boxplots show the community-weighted mean for plant height, median month of flowering and length of the flowering season for each vegetation alliance.
+Colors indicate the vegetation class.
+Alliances as ordered as in Figures 1 and 3-4 and alliance numbers are the same as listed in Appendix S3."}
+knitr::include_graphics("../results/figures/F5 - height-flowering.png")
+```
 
@@
 The major contributors to axis 2 (23% variance explained) were disturbance severity and frequency; this axis separated communities preferring less severe and less frequent disturbances (classes *Cymbalario-Parietarietea diffusae*, *Epilobietea angustifolii*) from communities adapted to more severe and more frequent disturbances (classes *Papaveretea rhoeadis*, *Polygono-Poetea annuae*, *Sisymbrietea* and *Digitario sanguinalis-Eragrostietea minoris*).
+
+```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic territories.
+Boxplots show the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity, calculated for each vegetation alliance.
+Colors indicate the vegetation class.
+Alliances as ordered as in Figures 1 and 3-4 and alliance numbers are the same as listed in Appendix S3."}
+knitr::include_graphics("../results/figures/F6 - indicators.png")
+```
+
+```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic territories.
+Biplot produced by Principal Component Analysis (PCA) of the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.
+Each dot is a plot.
+In the left panel, labels and arrows indicate the contribution of each indicator to the first and second principal components.
+The right panels show the space occupied by each vegetation class (colored dots) with respect to the whole anthropogenic vegetation (grey dots)."}
+knitr::include_graphics("../results/figures/F7 - pca.png")
+```
 
@@
 
-# Figures
+# Author contributions
 
-```{r fig1, echo = FALSE, fig.pos = "H", fig.cap = "Anthropogenic vegetation alliances of the Iberian Atlantic territories. Each dot is a vegetation plot. Dot colors indicate the vegetation class."}
-knitr::include_graphics("../results/figures/F1 - maps.png")
-```
+Eduardo Fernández-Pascual: Conceptualization; Methodology; Data Curation; Formal Analysis; Visualization; Writing – Original Draft Preparation; Writing – Review & Editing.
+Víctor González-García: Methodology; Investigation; Formal Analysis; Writing – Review & Editing.
+Greta Ivesdal: Investigation; Formal Analysis; Writing – Review & Editing.
+Adrián Lázaro-Lobo: Investigation; Writing – Review & Editing.
+Borja Jiménez-Alfaro: Methodology; Writing – Review & Editing.
 
-```{r fig2, echo = FALSE, fig.pos = "H", fig.cap = "Floristic diversity of the anthropogenic plant communities of the Iberian Atlantic territories.
-Biplots produced by Principal Component Analysis (PCA) with Hellinger transformation.
-PCAs were conducted for the whole vegetation (A) and for specific subsets (B-F).
-Colors indicate the vegetation class."}
-knitr::include_graphics("../results/figures/F2 - pca.png")
-```
+# Funding
 
-```{r fig3, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of neophytes, archaeophytes and native species in the anthropogenic plant communities of the Iberian Atlantic territories.
-Each bar represents the plot-level average proportion of neophytes (dark shade), archaeophytes (medium shade) and natives (light shade) in each vegetation alliance.
-Colors indicate the vegetation class."}
-knitr::include_graphics("../results/figures/F3 - origins.png")
-```
+Ayuntamiento de Gijón/Xixón, Jardín Botánico Atlántico (SV-23-GIJON-JBA).
 
-```{r fig4, echo = FALSE, fig.pos = "H", fig.cap = "Proportion of therophytes, hemicryptophytes, geophytes and other life forms (chamaephytes, phanerophytes, bryophytes, hydrophytes) in the anthropogenic plant communities of the Iberian Atlantic territories.
-Each bar represents the plot-level average proportion of therophytes (dark shade),hemicryptophytes (medium-dark shade), geophytes (medium-light shade) and other life forms (light shade) in each vegetation alliance.
-Colors indicate the vegetation class."}
-knitr::include_graphics("../results/figures/F4 - lifeforms.png")
-```
+# Conflict of interest statement
 
-```{r fig5, echo = FALSE, fig.pos = "H", fig.cap = "Vegetation height and flowering phenology of the anthropogenic plant communities of the Iberian Atlantic territories.
-Boxplots show the community-weighted mean for plant height, median month of flowering and length of the flowering season for each vegetation alliance.
-Colors indicate the vegetation class.
-Alliances as ordered as in Figures 1 and 3-4 and alliance numbers are the same as listed in Appendix S3."}
-knitr::include_graphics("../results/figures/F5 - height-flowering.png")
-```
+None declared.
 
-```{r fig6, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic territories.
-Boxplots show the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity, calculated for each vegetation alliance.
-Colors indicate the vegetation class.
-Alliances as ordered as in Figures 1 and 3-4 and alliance numbers are the same as listed in Appendix S3."}
-knitr::include_graphics("../results/figures/F6 - indicators.png")
-```
+# Data availability statement
 
-```{r fig7, echo = FALSE, fig.pos = "H", fig.cap = "Ecological requirements of the anthropogenic plant communities of the Iberian Atlantic territories.
-Biplot produced by Principal Component Analysis (PCA) of the plot-level means for the ecological indicator values of temperature, moisture, light, nutrients, soil reaction, disturbance frequency and disturbance severity.
-Each dot is a plot.
-In the left panel, labels and arrows indicate the contribution of each indicator to the first and second principal components.
-The right panels show the space occupied by each vegetation class (colored dots) with respect to the whole anthropogenic vegetation (grey dots)."}
-knitr::include_graphics("../results/figures/F7 - pca.png")
-```
+Original datasets, R code for analysis and Rmarkdown code for creation of the manuscript can be accessed at the GitHub repository <https://github.com/efernandezpascual/anthropogenic> stored at https://zenodo.org/doi/10.5281/zenodo.11613571.
